### 2017-05-27 paper.tex
Moved to Latex, used IEEE style.

\documentclass[10pt, conference, compsocconf]{IEEEtran}

% packages
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{color}
\usepackage[pdftex]{graphicx}
\usepackage{hyperref}

% new commands
\newcommand{\todo}[1]{
  \color{red}\emph{[#1]}
  \color{black}
}

\begin{document}

\title{Sequential algorithms to split and merge large 3D images}

\maketitle

\begin{abstract}
\todo{The abstract goes here. DO NOT USE SPECIAL CHARACTERS, SYMBOLS, OR MATH IN YOUR TITLE OR ABSTRACT.}
\end{abstract}

\section{Introduction}

Three-dimensional images that cannot be loaded in memory are
increasingly found in a variety of disciplines. BigBrain, for
instance, is a 3D histological image of the human brain that
represents 1~TB of data organized in 3600 planes at full resolution
(1$\mu$m in-plane), and 76~GB at a 40$\mu$m isotropic resolution
commonly used in neurosciences~\cite{amunts2013bigbrain}. Other
examples found in medical imaging, our primary domain of interest,
include \todo{electromicroscopy,spectroscopy-francoise-peyrin}. As
such images would typically be processed on a computing cluster,
possibly using locality-aware file systems such as HDFS, software
libraries are needed to split and merged such images efficiently. We
introduce and compare a family of algorithms to address this problem
in a sequential environment made of a single storage disk. We are
aware of the need for parallel algorithms that would split and merge
images to and from an array of disks but we plan to describe such
algorithms in a subsequent paper based on the results of the present
one.

A disk is modeled by its read and write rates ($\alpha_r$ and
$\alpha_w$, in MB/s), and its seek speed ($\beta$, also in
MB/s).

Images are split into chunks representing 3D blocks or 2D slices. We
assume that a block or slice fits in memory. The BigBrain would
perhaps be split into 125 chunks of 600~MB. The decision to split an
image into slices or blocks, and the size of the resulting chunks, may
be constrained by the application. Some applications, for instance
spatial filtering, would commonly require blocks while other ones such
as acquisition artifact removal would rather work on
slices. Flexibility is thus required in the splitting
scheme. Applications processing voxels individually, for instance
histogram computation or k-means clustering, could work on either
slices or blocks.

The main problem encountered in the context of this paper is the fact
that, if not ordered properly, data reads and writes might result in
extensive seek times that, as we illustrate below, drastically limit
the I/O performance. This problem is obviously related to the ordering
of data bytes in image files, which we assume is arbitrary but known
to the algorithm. The problem is also related to the geometry of the
chunks to extract or merge, which we again assume arbitrary but known
to the algorithm. In a nutshell, we are seeking for algorithms to
reduce seek times while allowing for arbitrary chunk geometries and
arbitrary byte orders in data files. The main idea of the variations
presented in the remainder of the paper is to convert data orderings
in memory before writing them on disk.

The literature looks remarkably scarce on the problem described
above. Parallel processing of 3D images has obviously been extensively
studied, but methods have focused more on geometrical approaches to
partition images and on load-balancing and task scheduling techniques
while we aim at algorithms to efficiently split or merge images
regardless of the geometry of the chunks \todo{do a serious literature
review on this topic}. Perhaps this lack of related works could be
attributed to the fact that disk I/O weren't a bottleneck until image
resolution reached a certain threshold. A relevant reference is the
description of the Open Connectome Data Cluster~\cite{burns2013open},
a data warehouse system that allows users to retrieve specific chunks
of large datasets. In this work, the problem of long seek times is
addressed by designing a specific file format based on space-filling
curves, that elegantly preserves spatial proximity on disk. However,
addressing the problem using a specific data organization optimizes
the system for specific chunk geometries, volumetric in this case,
while we aim at algorithms that support multiple types of geometries.

\section{Algorithms}

In our context, split and merge relate to the same dual
problem \todo{not so sure about this actually}. Without loss of
generality, we focus here on merging for the sake of concision. Our
goal then is to merge a set of $N$ chunks into a single reconstructed
3D image with $F$ voxels of size $u$. We assume that all chunks are of
identical size and that slices are squares and blocks are cubes.

Without loss of generality, we consider a file format where voxels are
written after alphanumerical sorting of their (x,y,z)
representation. That is, x slices are written one after the other, and
in every x slice, y rows are written one after the other. NifTI, the
file format defined by the Neuroimaging Informatics Technology
Initiative Nifti\footnote{\url{https://nifti.nimh.nih.gov}} is such a
format, used by the vast majority of the neuroimaging community. In
the remainder, we use the term "slice" for "x slice".

\subsection{Slices vs blocks}

Algorithms~\ref{algo:naive-slices} and~\ref{algo:naive-blocks} are the
naive solutions to reconstruct large images from slices or blocks.
\begin{algorithm}{tbh}
\caption{Naive merging of slices.}
\label{algo:naive-slices} 
\begin{algorithmic}
  \FOR{each slice}
    \STATE read slice
    \STATE write slice in reconstructed image
  \ENDFOR      
\end{algorithmic}
\end{algorithm}

\begin{algorithm}{tbh}
\caption{Naive merging of blocks.}
\label{algo:naive-blocks}
\begin{algorithmic}
  \FOR{each block}
    \STATE read block
    \STATE write block in reconstructed image
  \ENDFOR 
\end{algorithmic}
\end{algorithm}

Although simplistic, Algorithms~\ref{algo:naive-slices}
and~\ref{algo:naive-blocks} actually exhibit very different
performance due to the amount of seeks performed by Algorithm 1.b. The
number of seeks performed in Algo 1.b for each block is indeed the
number of rows in a slice, that is, $(F/N)^{1/3}$. The total number of
seeks, considering $N$ blocks is thus $N^{2/3}.F^{1/3}$. In practice,
this could lead to a tremendous slowdown.

\subsection{Sorted blocks}

Sorting blocks alphanumerically by (x,y,z) as shown on Algorithm 2 reduces the seek time
between blocks.

Algo 2: reconstruction from sorted blocks.
* Sort blocks according to (x,y,z)
* For each block in sorted list:
  - read block
  - write block in reconstructed image

The number of seeks is not reduced but the length of seeks is.

\subsection{Cluster reads}

We fill memory with as many blocks as possible to write in slices. A
given block is read only once. Slices may be written partially.

Algo 3: reconstruction from blocks using cluster reads and sorted blocks
* Sort blocks according to (x,y,z)
* While there are unread blocks:
  * While there is memory available:
    - read next block in sorted list
  * (Partially) reconstruct slices from blocks (in memory).
  * For each partially reconstructed slice:
    - Write in reconstructed image

\subsection{Multiple reads}

We fill memory with as many voxels as possible to write groups of
entire slices. A given block may be read multiple times.

Algo 4: reconstruction from blocks using multiple reads
* While there are empty slices in the reconstructed image:
  * Determine $n$ so that $n$ slices fit in memory
  * Reconstruct next slices $s_1$, ...,  $s_n$:
     - list the blocks contributing to $s_1$, $s_n$
     - for each block in the list, read the voxels of interest
     - reconstruct $s_1$, ..., $s_n$ in memory
     - write $s_1$, ..., $s_n$

\subsection{Benchmark}

As a running example to benchmark our algorithms, we use the 76~GB
BigBrain image split in 125 chunks. The data is available [here]. We
perform experiments using the two disks below:
- HD ...
- SSD ...
Our algorithms are implemented in Python using the Nibabel library.

\section{Results}

-- Splitting, 125 blocks, HD vs SSD.
Figure 1: on HD, 5 algorithms, 10 repetitions per algorithm.
Figure 2: on SSD, 5 algorithms, 10 repetitions per algorithm.

-- Merging, 125 blocks, HD vs SSD.
Figure 3: on HD, 5 algorithms, 10 repetitions per algorithm.
Figure 4: on SSD, 5 algorithms, 10 repetitions per algorithm.

-- HD, 250 blocks vs 125 blocks. 
Figure 5: different block sizes

-- HD, 125 blocks, F vs F/2
- Figure 6: Different images sizes, with constant memory.

\section{Discussion}

What is the best algorithm (between 3 and 4), in which context. Did we
solve the seek problem or is there any issue remaining?

File formats:
\begin{itemize}
\item would MINC~\cite{vincent2016minc} support such algorithms? See on-going email discussion
with P. Bellec.
\item some formats optimize the storage for particular split shapes. See
  ndstore~\cite{burns2013open}. Such formats would behave equally
  poorly with splits that do not comply to this geometry. Our
  algorithms would work equally well with such formats.
\end{itemize}

Reconstructing (therefore splitting) a complete 3D image may not
always be necessary. Re-splitting, for instance, could be done from
existing splits.

\section{Conclusion}

Future work:
- obliques
- parallel algorithms
- re-splitting
- model

\section*{Acknowledgment}

(If not added as co-authors): Greg Kiar, Pierre Bellec.

\bibliographystyle{IEEEtran}
\bibliography{IEEEabrv,biblio.bib}

\end{document}

### 2017-05-29 paper.tex
Note on compression and memo on authors

@@
 \begin{abstract}
-\todo{The abstract goes here. DO NOT USE SPECIAL CHARACTERS, SYMBOLS, OR MATH IN YOUR TITLE OR ABSTRACT.}
+  \todo{The abstract goes here. DO NOT USE SPECIAL CHARACTERS,
+    SYMBOLS, OR MATH IN YOUR TITLE OR ABSTRACT.}  Authors (final order
+  to be discussed when the paper shapes up): Yongping/Val\'erie,
+  Yuhong Yan, Tristan. Perhaps Pierre Bellec.
 \end{abstract}
@@
 
-\subsection{Benchmark}
+\subsection{Lossless compression}
+
+Lossless compression might reduce overall read and write times by
+reducing seek time and data sizes. However, it needs to be implemented
+on-the-fly, i.e., in chunks. See the work in~\cite{rajna2015speeding}
+and \href{https://github.com/pauldmccarthy/indexed\_gzip}{corresponding
+  implementation}.
+
+\subsection{Implementation and Benchmark}
 

### 2017-05-29 paper.tex
Added comment on compression, following discussion with @ValHayot on Slack.

@@
 on-the-fly, i.e., in chunks. See the work in~\cite{rajna2015speeding}
-and \href{https://github.com/pauldmccarthy/indexed\_gzip}{corresponding
-  implementation}.
+and
+\href{https://github.com/pauldmccarthy/indexed\_gzip}{corresponding
+  implementation}. This is straightforward to apply for random reads
+but not for random writes. Multiple reads is the only merging strategy
+that works with literally no seek in the reconstructed image which
+could therefore potentially use in-memory compression.
 

### 2017-05-31 function.py
Model implemented

@@ -1,0 +1,73 @@
+import math
+import matplotlib.pyplot as plt
+
+def cubic_root(x):
+  return x**(1.0/3)
+
+# returns (t2-t1)/t1, where t1 is the sequential time ("read slices, write slices")
+# and t2 is the "random time" (read blocks, write blocks)
+#
+# F: file size in bytes. Also file size in voxels. Typical value: 10**9 bytes (1 GB).
+# N: number of blocks and number of slices. Must be lower than F**(1/3). Typical value: 100.
+# alpha_r: data read rate. Typical value: 100 MB/s
+# alpha_w: data write rate. Typical value: 100 MB/s (?)
+# beta: disk access time. Typical value: 5 ms
+# m: amount of available memory, in bytes
+# b: number of bytes per voxel
+
+def t_slices(N,F,alpha_r,alpha_w,beta,b):
+  return b*F/alpha_r+b*F/alpha_w+2*beta*N
+
+def t_buff_slices(N,F,alpha_r,alpha_w,beta,m,b):
+  return b*F/alpha_r+b*F/alpha_w+beta*(N+math.ceil(b*F/m))
+
+def t_blocks(N,F,alpha_r,alpha_w,beta,b):
+    return F*b/alpha_r+F*b/alpha_w+beta*N*(1+cubic_root(F/N))
+
+def t_multiple_reads(N,F,alpha_r,alpha_w,beta,m,b):
+    return F*b/alpha_r+F*b/alpha_w+beta*math.ceil(F*b/m)*(1+math.ceil(m*N/(F*b)))
+
+def t_cluster_reads(N,F,alpha_r,alpha_w,beta,m,b):
+    a = m/(b*F)*cubic_root(N)
+    return b*F/alpha_r+b*F/alpha_w+beta*(N+math.ceil(F*b/m)*(cubic_root(F/N)*(math.ceil(a)-math.floor(a))+1))
+  
+N=125.0
+b=8.0
+F=76000000000.0/b
+alpha_r=145000000.0
+alpha_w=alpha_r
+beta=20*10**(-3)
+m=F/125.0 
+print "Slices",t_slices(N,F,alpha_r,alpha_w,beta,b)
+print "Buffered slices",t_buff_slices(N,F,alpha_r,alpha_w,beta,m,b)
+print "Blocks",t_blocks(N,F,alpha_r,alpha_w,beta,b)
+print "Multiple reads",t_multiple_reads(N,F,alpha_r,alpha_w,beta,m,b)
[…]

### 2017-05-31 paper.tex
Added models

@@
 \usepackage{hyperref}
+\usepackage{mathtools}
+\DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}
+\DeclarePairedDelimiter{\floor}{\lfloor}{\rfloor}
 
@@
 
-\section{Algorithms}
+\section{Algorithms and Models}
 
@@
 the remainder, we use the term "slice" for "x slice".
+
+\subsection{Notations}
+
+We adopt the following notations:
+\begin{itemize}
+\item $R$: number of voxels in the reconstructed images.
+\item $b$: bytes per voxel in the reconstructed images and splits (in B).
+\item $n$: number of splits (blocks or slices).
+\item $\alpha_r$: data read rate from disk (in B/s).
+\item $\alpha_w$: data write rate to disk (in B/s).
+\item $\beta$: disk access time (in s).
+\item $m$: amount of RAM available (in B).
+\end{itemize}
+
+The following properties are straightforward applications of these
+definitions:
+\begin{itemize}
+\item bytes per chunk: $\frac{Rb}{n}$
+\item number of chunks that can fit in memory: $\frac{m.n}{Rb}$
+\end{itemize}
+For the sake of mathematical simplicity, the models below assume that
+the reconstructed image and blocks are cubic. Under this assumption,
+we have the following definitions and properties:
+\begin{itemize}
+\item Number of slices in a block: $\sqrt[3]{\frac{R}{n}}$
+\item Number of blocks required to fully cover a slice in the
+  reconstructed image: $\sqrt[3]{n}^2=n^{\frac{2}{3}}$
+  \item Number of full slices that can be reconstructed from $k$
+    blocks: $\floor*{\frac{k}{n^{\frac{2}{3}}}}\sqrt[3]{\frac{R}{n}}$
+\end{itemize}
+\todo{Make a figure to illustrate the notations.}
 
@@
 number of seeks performed in Algo 1.b for each block is indeed the
-number of rows in a slice, that is, $(F/N)^{1/3}$. The total number of
-seeks, considering $N$ blocks is thus $N^{2/3}.F^{1/3}$. In practice,
+number of rows in a slice, that is, $(R/n)^{1/3}$. The total number of
+seeks, considering $n$ blocks is thus $N^{2/3}.F^{1/3}$. In practice,
 this could lead to a tremendous slowdown.
+
+For slices, the total merge time is expressed as follows:
+\begin{equation}
+T_\mathrm{slices} = \frac{bR}{\alpha_r}+\frac{bR}{\alpha_w}+2\beta n
+\end{equation}
+and for blocks:
+\begin{equation}
+  T_\mathrm{blocks} = \frac{bR}{\alpha_r}+\frac{bR}{\alpha_w}+\beta n\left(1+\sqrt[3]{\frac{R}{n}}\right)
+\end{equation}
 
@@
 
-\subsection{Cluster reads}
+\subsection{Buffered slices}
+
+\begin{equation}
+T_\mathrm{buff\_slices} = \frac{bR}{\alpha_r}+\frac{bR}{\alpha_w}+\beta\left( n + \ceil*{\frac{bR}{m}}\right)
+\end{equation}
+
+\subsection{Buffered blocks: Cluster reads}
 
@@
 
-\subsection{Multiple reads}
+    \begin{equation}
+      \begin{multlined}
+  T_\mathrm{cluster\_reads} = \frac{bR}{\alpha_r}+\frac{bR}{\alpha_w} +\\
+  \beta
+  \left[n+ 
+  \ceil*{\frac{bR}{m}} \left[
+           \sqrt[3]{\frac{R}{n}} \left(
+              \ceil*{\frac{m\sqrt[3]{n}}{bR}} - \floor*{\frac{m\sqrt[3]{n}}{bR}}
+           \right)+1
+             \right]
+             \right]
+             \end{multlined}
+\end{equation}
+
+\subsection{Buffered blocks: Multiple reads}
 
@@
 
+     \begin{equation}
+       T_\mathrm{multiple\_reads}= \frac{bR}{\alpha_r}+\frac{bR}{\alpha_w} + \beta\ceil*{\frac{bR}{m}}\left(1+ \ceil*{\frac{m.n}{bR}}\right)\\
+\end{equation}
+
 \subsection{Lossless compression}
@@
 
-\subsection{Implementation and Benchmark}
+\subsection{Benchmark}
+\label{sec:benchmark}
 
+\subsubsection{Data}
 As a running example to benchmark our algorithms, we use the 76~GB
-BigBrain image split in 125 chunks. The data is available [here]. We
+BigBrain image split in 125 chunks. The data is available
+\todo{[here]}.
+
+\subsubsection{Infrastructure}
+ We
 perform experiments using the two disks below:
@@
 - SSD ...
-Our algorithms are implemented in Python using the Nibabel library.
+
+\todo{Benchmark the disks to measure $\alpha_r$, $\alpha_w$ and $\beta$.}
+
+\subsubsection{Software implementation}
+
+Our algorithms are implemented in Python using the Nibabel
+library. \todo{Software release.}
 
@@
 
--- Splitting, 125 blocks, HD vs SSD.
-Figure 1: on HD, 5 algorithms, 10 repetitions per algorithm.
-Figure 2: on SSD, 5 algorithms, 10 repetitions per algorithm.
+\subsection{Slices vs blocks}
 
--- Merging, 125 blocks, HD vs SSD.
-Figure 3: on HD, 5 algorithms, 10 repetitions per algorithm.
-Figure 4: on SSD, 5 algorithms, 10 repetitions per algorithm.
+\todo{Valerie's results at PERFORM + model.}
 
--- HD, 250 blocks vs 125 blocks. 
-Figure 5: different block sizes
+\subsection{Sorted blocks vs blocks}
 
--- HD, 125 blocks, F vs F/2
-- Figure 6: Different images sizes, with constant memory.
+\subsection{Buffered slices vs slices}
+
+\subsection{Cluster reads vs Multiple reads vs Buffered slices}
+
+\subsection{Model-based comparisons}
+
+Figure~\ref{fig:models} plots the total merge time obtained with the
+different algorithms, using the values in~\ref{sec:benchmark} and
+different amounts of available memory. Multiple reads behave
+consistently better regardless of the amount of available
+memory. Cluster reads reach similar performance as available memory
+increases.
+\begin{figure}
+  \includegraphics[width=\columnwidth]{figures/model.pdf}
+  % see function.py
+  \caption{Model comparison}
+  \label{fig:models}
+\end{figure}
+\todo{redraw the figure. Benchmark $\alpha$ and $\beta$ more
+  accurately on the HDD or SSD. In the previous sections, compare the
+  model with real experiments, to validate the model.}
+
+\subsection{Lossless Compression}
+
+\subsection{Splitting}
 

### 2017-06-01 scripts/disk-benchmark.py
A script to measure read/write rates and access times.

@@ -1,0 +1,43 @@
+#!/usr/bin/env python
+
+import time
+import random
+import string
+import tempfile
+import os
+
+# A 256 MB random string
+print "Creating random string of 256MB..."
+s= ''.join(random.choice(string.ascii_uppercase + string.digits) for _ in range(256*1024*1024))
+    
+while(True):
+    # Picks a random file size between 256 MB and 6 GB.
+    # size is expressed in multiples of 256MB
+    size = random.randint(1,24)
+    
+    # Generates the string we will write to disk
+    buff = ''
+    n = 0
+    while ( n < size ):
+        buff += s
+        n += 1
+
+    # Creates temp file and writes buffer
+    fd, filename = tempfile.mkstemp(dir=os.getcwd())
+    f = open(filename,"w")
+    write_start = time.time()
+    f.write(buff)
+    write_end = time.time()
+    f.close()
+    f = open(filename,"r")
+    read_start = time.time()
+    f.read(len(buff))
+    read_end = time.time()
+    f.close()
+    os.remove(filename)
+    
+    # Print result
+    print len(buff),filename,write_end-write_start,read_end-read_start
+    # Just in case...
+    time.sleep(1)
+    

### 2017-06-01 scripts/disk-benchmark.py
Updates in HDD benchmark

@@ -37,7 +37,9 @@ while(True):
     os.remove(filename)
     
     # Print result
-    print len(buff),filename,write_end-write_start,read_end-read_start
+    f = open("benchmark.csv","a")
+    f.write(''+str(len(buff))+","+str(write_end-write_start)+","+str(read_end-read_start)+"\n")
+    f.close()
     # Just in case...
     time.sleep(1)
     

### 2017-06-03 paper.tex
Added a README
Updates

@@
 
+\newcommand{\note}[1]{
+  \color{blue}\emph{[Note: #1]}
+  \color{black}
+}
+
 \begin{document}
 
-\title{Sequential algorithms to split and merge large 3D images}
+\title{Sequential algorithms to split and merge terabyte-scale 3D images}
 
@@
   to be discussed when the paper shapes up): Yongping/Val\'erie,
-  Yuhong Yan, Tristan. Perhaps Pierre Bellec.
+  Yuhong Yan, Tristan. Perhaps Pierre Bellec. \note{This is a note.}
 \end{abstract}
@@
 
-Three-dimensional images that cannot be loaded in memory are
+Three-dimensional images that exceed typical memory size are
 increasingly found in a variety of disciplines. BigBrain, for
@@
 A disk is modeled by its read and write rates ($\alpha_r$ and
-$\alpha_w$, in MB/s), and its seek speed ($\beta$, also in
-MB/s).
+$\alpha_w$, in MB/s), and its access time ($\beta$, in seconds).
+%also in
+%MB/s).
 
@@
 
+This paper makes the following contributions:
+\begin{itemize}
+  \item We propose a set of algorithms to split and merge 3D images
+    that do not fit in memory.
+  \item We model the performance (execution time) of these algorithms,
+    based on a simple but realistic disk I/O model.
+  \item We evaluate our model and algorithms on a representative
+    use-case involving a 1~TB image a hard drive and a solid-state
+    drive.
+  \item We compare the algorithms in a variety of conditions based on the validated model.
+\end{itemize}
+Section~\ref{sec:algos} presents our algorithms and models,
+Section~\ref{sec:results} details the results and
+Sections~\ref{sec:discussion} discusses them.
+
 \section{Algorithms and Models}
+\label{sec:algos}
 
@@
 generality, we focus here on merging for the sake of concision. Our
-goal then is to merge a set of $N$ chunks into a single reconstructed
-3D image with $F$ voxels of size $u$. We assume that all chunks are of
+goal then is to merge a set of $n$ chunks into a single reconstructed
+3D image with $R$ voxels of size $b$. We assume that all chunks are of
 identical size and that slices are squares and blocks are cubes.
@@
 format, used by the vast majority of the neuroimaging community. In
-the remainder, we use the term "slice" for "x slice".
+the remainder, we use the term ``slice" for ``x slice".
 
@@
 \begin{itemize}
-\item $R$: number of voxels in the reconstructed images.
-\item $b$: bytes per voxel in the reconstructed images and splits (in B).
-\item $n$: number of splits (blocks or slices).
+\item $R$: number of voxels in the reconstructed image.
+\item $b$: number of bytes per voxel in the reconstructed image and chunks (in B).
+\item $n$: number of chunks (blocks or slices).
 \item $\alpha_r$: data read rate from disk (in B/s).
@@
 \item $\beta$: disk access time (in s).
-\item $m$: amount of RAM available (in B).
+\item $m$: amount of available RAM (in B).
 \end{itemize}
@@
 \begin{itemize}
-\item bytes per chunk: $\frac{Rb}{n}$
+\item number of bytes per chunk: $\frac{Rb}{n}$
 \item number of chunks that can fit in memory: $\frac{m.n}{Rb}$
@@
 Algorithms~\ref{algo:naive-slices} and~\ref{algo:naive-blocks} are the
-naive solutions to reconstruct large images from slices or blocks.
-\begin{algorithm}{tbh}
+naive merging methods for slices and blocks.
+\begin{algorithm}[h]
 \caption{Naive merging of slices.}
@@
 
-\begin{algorithm}{tbh}
+\begin{algorithm}[h]
 \caption{Naive merging of blocks.}
@@
 
-Although simplistic, Algorithms~\ref{algo:naive-slices}
-and~\ref{algo:naive-blocks} actually exhibit very different
-performance due to the amount of seeks performed by Algorithm 1.b. The
-number of seeks performed in Algo 1.b for each block is indeed the
-number of rows in a slice, that is, $(R/n)^{1/3}$. The total number of
-seeks, considering $n$ blocks is thus $N^{2/3}.F^{1/3}$. In practice,
-this could lead to a tremendous slowdown.
-
-For slices, the total merge time is expressed as follows:
+Algorithms~\ref{algo:naive-slices} and~\ref{algo:naive-blocks}
+actually have very different performance even though blocks and slices
+have identical sizes. This is due to the number of seeks performed by
+Algorithm~\ref{algo:naive-blocks} to write a block, which equals the
+total number of rows in all slices of a block, that is (since the
+blocks are cubic): $\sqrt[3]{\frac{R}{n}}^2$. \note{Note the square
+  which is omitted in equation 2 below and in model.py. Adding the
+  square to the model just makes the runtime aberrant. This is because
+  not all seeks are equal: only ``long'' seeks between block slices
+  are significant; seeks between rows of a given block slice are much
+  shorter. The way to model this is to replace $\beta$ by $\beta +
+  \frac{s}{\gamma}$ where $\beta$ is in seconds, $\gamma$ is in B/s
+  and $s$ is the seek ``distance'' in bytes. I bet $\gamma$ is very
+  high on the SSD. } In practice, this could lead to a tremendous
+slowdown as we will show later. For slices, the total merge time is
+modeled as follows:
 \begin{equation}
@@
 \section{Results}
+\label{sec:results}
 
@@
 \section{Discussion}
+\label{sec:discussion}
 
@@
   ndstore~\cite{burns2013open}. Such formats would behave equally
-  poorly with splits that do not comply to this geometry. Our
+  poorly with chunks that do not comply to this geometry. Our
   algorithms would work equally well with such formats.
@@
 
+The performance model might be dependent on the particular
+infrastructure (disks, OS and filesystem) used in the experiments. In
+particular, caching might behave differently in a different setup.
+
 Reconstructing (therefore splitting) a complete 3D image may not
 always be necessary. Re-splitting, for instance, could be done from
-existing splits.
+existing chunks.
 

### 2017-06-03 scripts/model.py

@@ -32,11 +32,12 @@ def t_cluster_reads(N,F,alpha_r,alpha_w,beta,m,b):
     return b*F/alpha_r+b*F/alpha_w+beta*(N+math.ceil(F*b/m)*(cubic_root(F/N)*(math.ceil(a)-math.floor(a))+1))
   
 N=125.0
-b=8.0
+b=2.0
 F=76000000000.0/b
-alpha_r=145000000.0
-alpha_w=alpha_r
-beta=20*10**(-3)
+alpha_w=68*1024*1024 #109506075.558291 # according to benchmark on May 31st
+alpha_r=alpha_w
+alpha_w=1000000000000000000000000000000000000
+beta=15*10**(-3)
 m=F/125.0 
 print "Slices",t_slices(N,F,alpha_r,alpha_w,beta,b)
 print "Buffered slices",t_buff_slices(N,F,alpha_r,alpha_w,beta,m,b)

### 2017-06-05 scripts/benchmark/seek.py
More benchmkar

@@ -1,0 +1,30 @@
+#!/usr/bin/env python
+
+import time
+import random
+import string
+import tempfile
+import os
+
+filename="file.dat"
+
+
+old_offset = 0
+while(True):
+    # Picks a random file size between 256 MB and 4718440448 (size of file)
+    offset = random.randint(0,22548578304)
+
+    f = open(filename,"w")
+    seek_start = time.time()
+    f.seek(offset)
+    seek_end = time.time()
+    f.close()
+       
+    # Print result
+    out = open("benchmark-seek.csv","a")
+    out.write(''+str(offset-old_offset)+","+str(seek_end-seek_start)+"\n")
+    out.close()
+    old_offset=offset
+    # Just in case...
+    time.sleep(1)
+    

### 2017-06-05 scripts/benchmark/disk-benchmark.py
More data

@@ -22,23 +22,32 @@ while(True):
         buff += s
         n += 1
 
-    # Creates temp file and writes buffer
+    # Creates temp file
     fd, filename = tempfile.mkstemp(dir=os.getcwd())
+    # write file
     f = open(filename,"w")
     write_start = time.time()
     f.write(buff)
     write_end = time.time()
     f.close()
+    # read file
     f = open(filename,"r")
     read_start = time.time()
     f.read(len(buff))
     read_end = time.time()
     f.close()
+    # open file and seek to the end
+    f = open(filename,"w")
+    seek_start = time.time()
+    f.seek(len(buff))
+    seek_end = time.time()
+    f.close()
+    
     os.remove(filename)
     
     # Print result
     f = open("benchmark.csv","a")
-    f.write(''+str(len(buff))+","+str(write_end-write_start)+","+str(read_end-read_start)+"\n")
+    f.write(''+str(len(buff))+","+str(write_end-write_start)+","+str(read_end-read_start)+","+str(seek_end-seek_start)+"\n")
     f.close()
     # Just in case...
     time.sleep(1)

### 2017-06-05 scripts/benchmark/seek.py

@@ -6,23 +6,28 @@ import string
 import tempfile
 import os
 
-filename="file.dat"
+filename="/data/bigbrain_40microns.nii"
 
 
 old_offset = 0
+f = open(filename,"r")
 while(True):
     # Picks a random file size between 256 MB and 4718440448 (size of file)
-    offset = random.randint(0,22548578304)
+    offset = random.randint(0,81523750352)
+    small_distance = random.randint(-1024**3,1024**3)
 
-    f = open(filename,"w")
     seek_start = time.time()
     f.seek(offset)
     seek_end = time.time()
-    f.close()
-       
+    small_start = time.time()
+    f.seek(offset+small_distance)
+    small_end = time.time()  
+
+   
     # Print result
-    out = open("benchmark-seek.csv","a")
+    out = open("benchmark-seek-hdd-1.csv","a")
     out.write(''+str(offset-old_offset)+","+str(seek_end-seek_start)+"\n")
+    out.write(''+str(small_distance)+","+str(small_end-small_start)+"\n")
     out.close()
     old_offset=offset
     # Just in case...

### 2017-06-07 paper.tex
Wrote remaining algorithms

@@
 \begin{algorithm}[h]
-\caption{Naive merging of slices.}
+\caption{Naive merging from slices.}
 \label{algo:naive-slices} 
@@
 \begin{algorithm}[h]
-\caption{Naive merging of blocks.}
+\caption{Naive merging from blocks.}
 \label{algo:naive-blocks}
@@
 
-Algo 2: reconstruction from sorted blocks.
-* Sort blocks according to (x,y,z)
-* For each block in sorted list:
-  - read block
-  - write block in reconstructed image
+\begin{algorithm}[h]
+  \caption{Merging from sorted blocks.}
+  \label{algo:sorted-blocks}
+  \begin{algorithmic}
+    \STATE list = sort blocks according to (x,y,z)
+    \FOR{each block in list}
+      \STATE read block
+      \STATE write block in reconstructed image
+    \ENDFOR
+  \end{algorithmic}
+  \end{algorithm}
 
@@
 \subsection{Buffered slices}
+
+\begin{algorithm}[h]
+  \caption{Buffered merging from slices}
+  \label{algo:buffered-slices}
+  \begin{algorithmic}
+    \STATE list = sort slices by increasing x values
+    \STATE initialize buffer
+    \STATE index = 0
+    \WHILE{index\textless n}
+      \WHILE{sizeof(buffer)\textless m}
+        \STATE slice = list[index]
+        \STATE read slice and append it to buffer
+        \STATE index += 1
+       \ENDWHILE
+       \STATE write buffer in reconstructed image
+       \STATE clear buffer
+    \ENDWHILE
+  \end{algorithmic}
+\end{algorithm}
 
@@
 
-We fill memory with as many blocks as possible to write in slices. A
-given block is read only once. Slices may be written partially.
-
-Algo 3: reconstruction from blocks using cluster reads and sorted blocks
-* Sort blocks according to (x,y,z)
-* While there are unread blocks:
-  * While there is memory available:
-    - read next block in sorted list
-  * (Partially) reconstruct slices from blocks (in memory).
-  * For each partially reconstructed slice:
-    - Write in reconstructed image
+See Algorithm~\ref{algo:cluster-reads}. As many blocks as possible are
+loaded in memory, concatenated in a buffer and written as sequentially
+as possible from the buffer. Although the Algorithm looks very similar
+to Algorithm~\ref{algo:buffered-slices}, it should be noted that in
+Algorithm~\ref{algo:cluster-reads} the buffer should actually be a
+more complex data structure capable of storing multiple disjoint
+sequences of contiguous bytes without having to allocate memory to
+store the ``holes'' between such sequences \todo{a figure showing the
+  buffer and how it fillus up would help}.  \note{More on
+  defragmentation might be needed here.}
+\begin{algorithm}[h]
+  \caption{Buffered merging from blocks using cluster reads}
+  \label{algo:cluster-reads}
+  \begin{algorithmic}
+    \STATE list = sort blocks by increasing (x,y,z) values
+    \STATE initialize buffer
+    \STATE index = 0
+    \WHILE{index\textless n}
+      \WHILE{sizeof(buffer)\textless m}
+        \STATE block = list[index]
+        \STATE read block and append it to buffer
+        \STATE index += 1
+        \ENDWHILE
+       \STATE write buffer in reconstructed image
+       \STATE clear buffer
+    \ENDWHILE
+  \end{algorithmic}
+\end{algorithm}
 
@@
 
-We fill memory with as many voxels as possible to write groups of
-entire slices. A given block may be read multiple times.
+In multiple reads, blocks are read partially to ensure that the memory
+buffer only contains contiguous bytes, at any time. That is, a given
+block will be read multiple times.
 
-Algo 4: reconstruction from blocks using multiple reads
-* While there are empty slices in the reconstructed image:
-  * Determine $n$ so that $n$ slices fit in memory
-  * Reconstruct next slices $s_1$, ...,  $s_n$:
-     - list the blocks contributing to $s_1$, $s_n$
-     - for each block in the list, read the voxels of interest
-     - reconstruct $s_1$, ..., $s_n$ in memory
-     - write $s_1$, ..., $s_n$
+\begin{algorithm}[h]
+  \caption{Buffered merging from blocks using multiple reads
+    \todo{improve notations and algo}.}
+  \label{algo:multiple-reads}
+  \begin{algorithmic}
+    \STATE list = sort blocks by increasing (x,y,z) values
+    \STATE initialize buffer
+    \STATE pointer = 0, max\_index=(m-1)
+    \WHILE{max\_index \textless sizeof(reconstructed\_image)}
+    \WHILE{pointer \textless max\_index}
+        \STATE block = read block that contains voxel index\_to\_voxel(pointer)
+        \STATE insert in buffer all voxels (x,y,z) of block where voxel\_to\_index(x,y,z) \textless max\_index
+        \STATE update pointer to point to the next unassigned element in buffer
+        \ENDWHILE
+       \STATE write buffer in reconstructed image
+       \STATE clear buffer
+       \STATE pointer = max\_index + 1
+       \STATE max\_index += m
+    \ENDWHILE
+  \end{algorithmic}
+\end{algorithm}
 

### 2017-06-10 paper.tex
Changed model to count number of seeks instead of measuring total time

@@
 \usepackage{algorithmic}
+\usepackage{amsfonts} % for R symbol (the set of real numbers)
 \usepackage{color}
@@
 
-A disk is modeled by its read and write rates ($\alpha_r$ and
-$\alpha_w$, in MB/s), and its access time ($\beta$, in seconds).
 %also in
@@
 \begin{itemize}
-\item $R$: number of voxels in the reconstructed image.
+\item $R=D^3$: number of voxels in the reconstructed image.
 \item $b$: number of bytes per voxel in the reconstructed image and chunks (in B).
@@
 \item $\beta$: disk access time (in s).
+\item $\sigma: \mathbb{R}^{+} \rightarrow \mathbb{R}^{+}$: seek time (in s), function of the seek
+  distance (in B).
 \item $m$: amount of available RAM (in B).
@@
 \begin{itemize}
-\item Number of slices in a block: $\sqrt[3]{\frac{R}{n}}$
+\item Number of slices in a block: $\sqrt[3]{\frac{R}{n}}=d$. This is
+  also the number of rows and columns in a slice and the number of
+  voxel in a row or column.
 \item Number of blocks required to fully cover a slice in the
@@
 \todo{Make a figure to illustrate the notations.}
+
+\subsection{Disk model}
+
+A disk is modeled by its read and write rates ($\alpha_r$ and
+$\alpha_w$, in B/s), its access time $\beta$ (in s) and its seek time,
+a function $\sigma$ of the seek distance (in B). More precisely,
+$\sigma$ would typically be in the order of $\beta$ when an opened
+file is seeked to a random position, but it would be much smaller when
+a file is seeked to a nearby position.
+
+The disk model encompasses file systems, kernel modules and disk
+controllers required to access the disk from a program run in user
+space. In modern systems, this often includes caching at various
+levels \todo{\cite{x}}, which we will model by zeroing $\alpha_r$ or
+$\alpha_w$ values depending on the application.
 
@@
 actually have very different performance even though blocks and slices
-have identical sizes. This is due to the number of seeks performed by
-Algorithm~\ref{algo:naive-blocks} to write a block, which equals the
-total number of rows in all slices of a block, that is (since the
-blocks are cubic): $\sqrt[3]{\frac{R}{n}}^2$. \note{Note the square
-  which is omitted in equation 2 below and in model.py. Adding the
-  square to the model just makes the runtime aberrant. This is because
-  not all seeks are equal: only ``long'' seeks between block slices
-  are significant; seeks between rows of a given block slice are much
-  shorter. The way to model this is to replace $\beta$ by $\beta +
-  \frac{s}{\gamma}$ where $\beta$ is in seconds, $\gamma$ is in B/s
-  and $s$ is the seek ``distance'' in bytes. I bet $\gamma$ is very
-  high on the SSD. } In practice, this could lead to a tremendous
-slowdown as we will show later. For slices, the total merge time is
-modeled as follows:
+have identical sizes. Since slices are stored contiguously in the
+reconstructed image, the number of seeks in
+Algorithm~\ref{algo:naive-slices} is only $2n$:
 \begin{equation}
-T_\mathrm{slices} = \frac{bR}{\alpha_r}+\frac{bR}{\alpha_w}+2\beta n
+  N_\mathrm{slices} = 2n 
 \end{equation}
-and for blocks:
+However,
+Algorithm~\ref{algo:naive-blocks} has to do extra seeks for each row
+in each slice of each block. That is, the total number of seeks
+performed by Algorithm~\ref{algo:naive-blocks} is:
 \begin{equation}
-  T_\mathrm{blocks} = \frac{bR}{\alpha_r}+\frac{bR}{\alpha_w}+\beta n\left(1+\sqrt[3]{\frac{R}{n}}\right)
+N_\mathrm{blocks} = 2n+nd^2 = n\left(1+\sqrt[3]{\frac{R}{n}}^2\right)
 \end{equation}
+In
+practice, this difference could lead to a tremendous slowdown as we will show
+later.
+
+%%  For slices, the total merge time is modeled as follows:
+%% \begin{equation}
+%% T_\mathrm{slices} = \frac{bR}{\alpha_r}+\frac{bR}{\alpha_w}+2\beta n
+%% \end{equation}
+%% and for blocks:
+%% \begin{equation}
+%%   \begin{multlined}
+%%     T_\mathrm{blocks} = \frac{bR}{\alpha_r}+\frac{bR}{\alpha_w}+ 2\beta n + \\
+%%     n\left({\sigma(bR/2)} + d(d-1)\sigma\left((D-d)b\right)+d\sigma\left((D^2-d^2)b\right) \right)
+%%     \end{multlined}
+%% \end{equation}
 
@@
 
-Sorting blocks alphanumerically by (x,y,z) as shown on Algorithm 2 reduces the seek time
-between blocks.
+Sorting blocks alphanumerically by (x,y,z) as shown on Algorithm 2
+reduces the seek time between blocks.
 
@@
 
-The number of seeks is not reduced but the length of seeks is.
+\todo{Not sure what we are measuring here. The number of seeks will be identical. The difference in seek time is supposed to be very minimal.}
 
 \subsection{Buffered slices}
+
+The algorithms presented up to now are in fact particular cases of
+memory buffering where the amount of available memory equals the
+maximum size of a chunk.
 
@@
 \begin{equation}
-T_\mathrm{buff\_slices} = \frac{bR}{\alpha_r}+\frac{bR}{\alpha_w}+\beta\left( n + \ceil*{\frac{bR}{m}}\right)
+N_\mathrm{buff\_slices} =  n + \ceil*{\frac{bR}{m}}
 \end{equation}
@@
         \STATE block = list[index]
-        \STATE read block and append it to buffer
+        \STATE read block and insert it in buffer
         \STATE index += 1
@@
 \begin{equation}
-  \begin{multlined}
-    T_\mathrm{cluster\_reads} = \frac{bR}{\alpha_r}+\frac{bR}{\alpha_w} +\\
-    \beta
-    \left[n+ 
-      \ceil*{\frac{bR}{m}} \left[
-        \sqrt[3]{\frac{R}{n}} \left(
-        \ceil*{\frac{m\sqrt[3]{n}}{bR}} - \floor*{\frac{m\sqrt[3]{n}}{bR}}
-        \right)+1
-        \right]
-      \right]
-  \end{multlined}
+  N_\mathrm{cluster\_reads}= n + \ceil*{\frac{bR}{m}}f(m,n,R)
+\end{equation}
+where:
+\begin{equation}
+  f(m,n,R)=
+  \begin{cases}
+    d^2 & $if there is an incomplete line$\\
+    d   & $if there is an incomplete slice$\\
+    1   & $otherwise$
+\end{cases}
 \end{equation}
@@
 \begin{equation}
-       T_\mathrm{multiple\_reads}= \frac{bR}{\alpha_r}+\frac{bR}{\alpha_w} + \beta\ceil*{\frac{bR}{m}}\left(1+ \ceil*{\frac{m.n}{bR}}\right)\\
+  N_\mathrm{multiple\_reads}= n \max\left(1,\ceil*{\frac{d}{\ceil*{\frac{m}{nd^2b}}}}\right)+ \ceil*{\frac{bR}{m}}
 \end{equation}

### 2017-06-10 paper.tex
Better description of cluster reads

@@
 chunks to extract or merge, which we again assume arbitrary but known
-to the algorithm. In a nutshell, we are seeking for algorithms to
+to the algorithm. In a nutshell, we are looking for algorithms to
 reduce seek times while allowing for arbitrary chunk geometries and
@@
 
-A disk is modeled by its read and write rates ($\alpha_r$ and
-$\alpha_w$, in B/s), its access time $\beta$ (in s) and its seek time,
-a function $\sigma$ of the seek distance (in B). More precisely,
-$\sigma$ would typically be in the order of $\beta$ when an opened
-file is seeked to a random position, but it would be much smaller when
-a file is seeked to a nearby position.
+We use a simplistic disk model where a disk is characterized by its
+read and write rates ($\alpha_r$ and $\alpha_w$, in B/s) and its
+access time $\beta$ (in s). In particular, we assume that seeks
+require a constant amount of time, regardless of the position seeked
+to. In practice, large variations would be expected, but modeling such
+variations would inevitably lead to models specific to the hardware,
+file system or operating system, which we intentionally avoid
+here. Thus, our goal is to find algorithms that minimize the
+\emph{number} of seek operations. Likewise, in modern systems, read
+and write times are greatly impacted by caching occurring at several
+levels, which we do not model here.
 
-The disk model encompasses file systems, kernel modules and disk
-controllers required to access the disk from a program run in user
-space. In modern systems, this often includes caching at various
-levels \todo{\cite{x}}, which we will model by zeroing $\alpha_r$ or
-$\alpha_w$ values depending on the application.
+%% and its seek time,
+%% a function $\sigma$ of the seek distance (in B). More precisely,
+%% $\sigma$ would typically be in the order of $\beta$ when an opened
+%% file is seeked to a random position, but it would be much smaller when
+%% a file is seeked to a nearby position.
+%% The disk model encompasses file systems, kernel modules and disk
+%% controllers required to access the disk from a program run in user
+%% space. In modern systems, this often includes caching at various
+%% levels \todo{\cite{x}}, which we will model by zeroing $\alpha_r$ or
+%% $\alpha_w$ values depending on the application.
 
@@
 
-See Algorithm~\ref{algo:cluster-reads}. As many blocks as possible are
-loaded in memory, concatenated in a buffer and written as sequentially
-as possible from the buffer. Although the Algorithm looks very similar
-to Algorithm~\ref{algo:buffered-slices}, it should be noted that in
-Algorithm~\ref{algo:cluster-reads} the buffer should actually be a
-more complex data structure capable of storing multiple disjoint
-sequences of contiguous bytes without having to allocate memory to
-store the ``holes'' between such sequences \todo{a figure showing the
-  buffer and how it fillus up would help}.  \note{More on
-  defragmentation might be needed here.}
+As detailed in Algorithm~\ref{algo:cluster-reads}, cluster reads load
+multiple blocks in memory, concatenate them in a buffer and write the
+buffer as sequentially as possible. The buffer is a slightly complex
+data structure, for instance an associative array or Python
+dictionary, capable of storing multiple disjoint sequences of
+contiguous bytes without having to allocate memory to store the
+``holes'' between such sequences \todo{a figure showing the buffer and
+  how it fillus up would help}.  \note{More on defragmentation might
+  be needed here.}
 \begin{algorithm}[h]

### 2017-06-11 paper.tex
Explained nseeks for cluster reads

@@
 buffer as sequentially as possible. The buffer is a slightly complex
-data structure, for instance an associative array or Python
+data structure, for instance an associative array or a Python
 dictionary, capable of storing multiple disjoint sequences of
-contiguous bytes without having to allocate memory to store the
-``holes'' between such sequences \todo{a figure showing the buffer and
-  how it fillus up would help}.  \note{More on defragmentation might
-  be needed here.}
+contiguous bytes without having to allocate memory for the bytes
+between such sequences. Implementation details about the buffer are
+given later in the document.
+
+\todo{a figure showing the buffer and how it
+  fillus up would help}.
+%\note{More on defragmentation might be
+%  needed here.}
 \begin{algorithm}[h]
@@
 
+The number of seeks performed by cluster reads depends on the way that
+blocks loaded in memory arrange in the reconstructed image. In the
+best case, blocks in memory completely cover slices in the
+reconstructed image. In this case, slices of the reconstructed images
+can be written without seeking at all. If the blocks in memory
+completely cover rows in the reconstructed image but do not cover
+entire slices, then the blocks are written with $d$ seeks, one per
+slice. And if the blocks do not cover entire rows, then they are
+written with $d^2$ seeks, one per row and per slice.  In conclusion:
 \begin{equation}
@@
   \begin{cases}
-    d^2 & $if there is an incomplete line$\\
-    d   & $if there is an incomplete slice$\\
+    d^2 & $if there is an incomplete row$\\
+    d   & $if there is an incomplete slice but not incomplete row$\\
     1   & $otherwise$

### 2017-06-12 figures/export.sh
Figures for cluster reads

@@ -1,0 +1,9 @@
+#!/bin/bash
+
+for name in d2-seeks d-seeks no-seek 
+do
+  echo "Exporting ${name}.svg"
+  inkscape -D -z --file=figures/${name}.svg --export-pdf=figures/${name}.pdf --export-latex
+  sed s,${name}\.pdf,figures/${name}\.pdf,g figures/${name}.pdf_tex > figures/${name}.pdf_tex.new
+  \mv figures/${name}.pdf_tex.new figures/${name}.pdf_tex
+done

### 2017-06-12 paper.tex
gitignore and Figure 1

@@
 multiple blocks in memory, concatenate them in a buffer and write the
-buffer as sequentially as possible. The buffer is a slightly complex
+buffer sequentially in the reconstructed image. The buffer is a slightly complex
 data structure, for instance an associative array or a Python
@@
 
+\begin{figure}
+\centering
+\def\svgwidth{0.3\columnwidth}
+\input{figures/no-seek.pdf_tex}
+\def\svgwidth{0.3\columnwidth}
+\input{figures/d-seeks.pdf_tex}
+\def\svgwidth{0.3\columnwidth}
+\input{figures/d2-seeks.pdf_tex}
+\caption{Left: complete slices and rows are reconstructed in memory:
+  no seeking is required. Center: complete rows but incomplete slices
+  are reconstructed: $d$ seeks are required. Right: incomplete rows
+  are reconstructed: $d^2$ seeks are required.}
+\label{fig:cluster-reads}
+\end{figure}
+
 \todo{a figure showing the buffer and how it
-  fillus up would help}.
+  fill up would help}.
 %\note{More on defragmentation might be
@@
 \begin{algorithm}[h]
-  \caption{Buffered merging from blocks using cluster reads}
+  \caption{Buffered merging of blocks with cluster reads}
   \label{algo:cluster-reads}
@@
 
-The number of seeks performed by cluster reads depends on the way that
-blocks loaded in memory arrange in the reconstructed image. In the
-best case, blocks in memory completely cover slices in the
-reconstructed image. In this case, slices of the reconstructed images
-can be written without seeking at all. If the blocks in memory
-completely cover rows in the reconstructed image but do not cover
-entire slices, then the blocks are written with $d$ seeks, one per
-slice. And if the blocks do not cover entire rows, then they are
-written with $d^2$ seeks, one per row and per slice.  In conclusion:
+The number of seeks performed by cluster reads depends on how blocks
+loaded in memory arrange in the reconstructed image. In the best case,
+complete contiguous slices of the reconstructed image can be assembled
+in memory and written without seeking at all. In the worst case, some
+rows in the reconstructed image are not covered by blocks loaded in
+memory: $d^2$ seeks are then required, one for every partial row in
+every partial slice. In the intermediary case, rows are complete but
+some slices can only be partially reconstructed: $d$ seeks are then
+required. In summary:
 \begin{equation}

### 2017-06-12 paper.tex
Added figure on buffer, added variants of cluster reads

@@
 
+\todo{Overlapping blocks. non-cubic blocks.}
+
 In our context, split and merge relate to the same dual
@@
   voxel in a row or column.
-\item Number of blocks required to fully cover a slice in the
-  reconstructed image: $\sqrt[3]{n}^2=n^{\frac{2}{3}}$
-  \item Number of full slices that can be reconstructed from $k$
-    blocks: $\floor*{\frac{k}{n^{\frac{2}{3}}}}\sqrt[3]{\frac{R}{n}}$
+\item Number of blocks required to fill a row of the reconstructed
+  image: $\sqrt[3]{n}=\nu$.
 \end{itemize}
@@
 \begin{equation}
-N_\mathrm{blocks} = 2n+nd^2 = n\left(1+\sqrt[3]{\frac{R}{n}}^2\right)
+N_\mathrm{blocks} = n+nd^2 
 \end{equation}
@@
 multiple blocks in memory, concatenate them in a buffer and write the
-buffer sequentially in the reconstructed image. The buffer is a slightly complex
-data structure, for instance an associative array or a Python
-dictionary, capable of storing multiple disjoint sequences of
+buffer sequentially in the reconstructed image. The buffer is a
+slightly complex data structure, for instance an associative array or
+a Python dictionary, capable of storing multiple disjoint sequences of
 contiguous bytes without having to allocate memory for the bytes
-between such sequences. Implementation details about the buffer are
-given later in the document.
-
+between such sequences. Figure~\ref{fig:cluster-reads-buffer}
+illustrates how the buffer would fill up when two blocks of one slice
+and three rows per slice are loaded. Implementation details about the
+buffer are given later in the document.
 \begin{figure}
 \centering
-\def\svgwidth{0.3\columnwidth}
-\input{figures/no-seek.pdf_tex}
-\def\svgwidth{0.3\columnwidth}
-\input{figures/d-seeks.pdf_tex}
-\def\svgwidth{0.3\columnwidth}
-\input{figures/d2-seeks.pdf_tex}
-\caption{Left: complete slices and rows are reconstructed in memory:
-  no seeking is required. Center: complete rows but incomplete slices
-  are reconstructed: $d$ seeks are required. Right: incomplete rows
-  are reconstructed: $d^2$ seeks are required.}
-\label{fig:cluster-reads}
+\def\svgwidth{\columnwidth}
+\input{figures/buffer-cluster-reads.pdf_tex}
+\caption{Illustration of the buffer used in cluster reads. White
+  portions are not allocated. [a,b] refers to block a, row b.}
+\label{fig:cluster-reads-buffer}
 \end{figure}
-
-\todo{a figure showing the buffer and how it
-  fill up would help}.
-%\note{More on defragmentation might be
-%  needed here.}
 \begin{algorithm}[h]
@@
 some slices can only be partially reconstructed: $d$ seeks are then
-required. In summary:
+required. See also Figure~\ref{fig:cluster-reads}. In summary:
 \begin{equation}
@@
     d^2 & $if there is an incomplete row$\\
-    d   & $if there is an incomplete slice but not incomplete row$\\
+    d   & $if there is an incomplete slice$\\
+        & $but not incomplete row$\\
     1   & $otherwise$
@@
 \end{equation}
+\begin{figure}
+\centering
+\def\svgwidth{0.3\columnwidth}
+\input{figures/no-seek.pdf_tex}
+\def\svgwidth{0.3\columnwidth}
+\input{figures/d-seeks.pdf_tex}
+\def\svgwidth{0.3\columnwidth}
+\input{figures/d2-seeks.pdf_tex}
+\caption{Left: complete slices and rows are reconstructed in memory:
+  no seeking is required. Center: complete rows but incomplete slices
+  are reconstructed: $d$ seeks are required. Right: incomplete rows
+  are reconstructed: $d^2$ seeks are required.}
+\label{fig:cluster-reads}
+\end{figure}
+
+If $\frac{Rb}{\sqrt[3]{n}}>m\geq\frac{Rb}{\sqrt[3]{n}^2}$, i.e., a
+full row of blocks fits in memory but a full slice does not, then a
+variant of cluster reads consists in adjusting the memory load to the
+highest value that makes sure that only full rows are written. In this
+case:
+\begin{equation}
+  N_\mathrm{cluster\_reads\_v1}= n + \frac{bR}{m|}d
+\end{equation}
+where $m|=\frac{Rb}{\sqrt[3]{n}^2}\floor*{\frac{m\sqrt[3]{n}^2}{Rb}}$
+
+And if $m\geq\frac{Rb}{\sqrt[3]{n}}$, i.e., a full slice of blocks
+fits in memory, then a second variant adjusts
+the memory load to the highest value that makes sure that only full
+slices are written. In this case:
+\begin{equation}
+  N_\mathrm{cluster\_reads\_v2}= n + \frac{bR}{m||}d
+\end{equation}
+where $m||=\frac{Rb}{\sqrt[3]{n}}\floor*{\frac{m\sqrt[3]{n}}{Rb}}$
 
@@
 could therefore potentially use in-memory compression.
+
+\todo{could we merge without de-compressing?}
 

### 2017-06-12 figures/export.sh

@@ -1,6 +1,6 @@
 #!/bin/bash
 
-for name in d2-seeks d-seeks no-seek 
+for name in d2-seeks d-seeks no-seek buffer-cluster-reads
 do
   echo "Exporting ${name}.svg"
   inkscape -D -z --file=figures/${name}.svg --export-pdf=figures/${name}.pdf --export-latex

### 2017-06-20 paper.tex
Updates in cluster reads following discussion with @ValHayot (still work in progress)

@@
 \def\svgwidth{0.3\columnwidth}
-\input{figures/no-seek.pdf_tex}
+\input{figures/case1.pdf_tex}
 \def\svgwidth{0.3\columnwidth}
-\input{figures/d-seeks.pdf_tex}
+\input{figures/case2.pdf_tex}
 \def\svgwidth{0.3\columnwidth}
-\input{figures/d2-seeks.pdf_tex}
-\caption{Left: complete slices and rows are reconstructed in memory:
-  no seeking is required. Center: complete rows but incomplete slices
-  are reconstructed: $d$ seeks are required. Right: incomplete rows
-  are reconstructed: $d^2$ seeks are required.}
+\input{figures/case3-a.pdf_tex}
+\def\svgwidth{0.3\columnwidth}
+\input{figures/case3-b.pdf_tex}
+\def\svgwidth{0.3\columnwidth}
+\input{figures/case4.pdf_tex}
+\def\svgwidth{0.3\columnwidth}
+\input{figures/case5.pdf_tex}
+\def\svgwidth{0.3\columnwidth}
+\input{figures/case6-a.pdf_tex}
+\def\svgwidth{0.3\columnwidth}
+\input{figures/case6-b.pdf_tex}
+\def\svgwidth{0.3\columnwidth}
+\input{figures/case7.pdf_tex}
+
+\caption{Cluster reads configurations leading to different number of
+  seeks.}
 \label{fig:cluster-reads}
+% See source files in
+% https://drive.google.com/drive/folders/0BzT7pbvPDUdQRXppVlhfdS0ycGc?usp=sharing
 \end{figure}

### 2017-06-20 figures/export.sh

@@ -1,6 +1,6 @@
 #!/bin/bash
 
-for name in d2-seeks d-seeks no-seek buffer-cluster-reads
+for name in buffer-cluster-reads case1 case2 case3-a case3-b case4 case5 case6-a case6-b case7
 do
   echo "Exporting ${name}.svg"
   inkscape -D -z --file=figures/${name}.svg --export-pdf=figures/${name}.pdf --export-latex

### 2017-06-21 paper.tex
Updates in section II.F

@@
 \usepackage{mathtools}
+\usepackage{multirow}
 \DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}
@@
 and three rows per slice are loaded. Implementation details about the
-buffer are given later in the document.
+buffer are given later.
 \begin{figure}
@@
 complete contiguous slices of the reconstructed image can be assembled
-in memory and written without seeking at all. In the worst case, some
+in memory and written without any seeking. In the worst case, some
 rows in the reconstructed image are not covered by blocks loaded in
-memory: $d^2$ seeks are then required, one for every partial row in
-every partial slice. In the intermediary case, rows are complete but
-some slices can only be partially reconstructed: $d$ seeks are then
-required. See also Figure~\ref{fig:cluster-reads}. In summary:
+memory: O($d^2$) seeks are then required to write the memory load, one
+for every partial row in every partial slice. In the intermediary
+case, rows are complete but some slices can only be partially
+reconstructed: O($d$) seeks are then
+required. We have:
 \begin{equation}
-  N_\mathrm{cluster\_reads}= n + \ceil*{\frac{bR}{m}}f(m,n,R)
+  N_\mathrm{cluster\_reads}= n + \sum_{i=1}^{\ceil*{\frac{bR}{m}}}f_i
 \end{equation}
-where:
+where $f_i$ is the number of seeks done in the $i^{th}$ memory
+load. Figure~\ref{fig:cluster-reads} lists the possible block
+arrangements and the corresponding values of $f_i$. \todo{Val\'erie: Check that
+  ordering is consistent with Nifti's.}
+%% In case (1) on
+%% this Figure, the memory load completely fills rows and slices in the
+%% reconstructed image. Thus it can be written with a single seek,
+%% $f_i$=1. In case (2), all the rows are complete but $d$ slices are
+%% incomplete; one seek is required for each slices, $f_i=d$. In case
+%% (3), $d$ rows of $d$ slices are incomplete, $f_i=d^2$. In case (4),
+%% $2d$ rows of $d$ slices are incomplete, $f_i=2d^2-d$. In case (5),
+%% $2d$ slices are incomplete but no row is incomplete, $f_i=2d$. In case
+%% (6), $d$ rows of $d$ slices are incomplete, and $d$ additional slices
+%% are incomplete, $f_i$=$d^2+d-1$. And in case (7), $d$ rows of $2d$
+%% slices are incomplete, $f_i$=$2d^2-1$.
+In general:
 \begin{equation}
-  f(m,n,R)=
-  \begin{cases}
-    d^2 & $if there is an incomplete row$\\
-    d   & $if there is an incomplete slice$\\
-        & $but not incomplete row$\\
-    1   & $otherwise$
-\end{cases}
+  N_\mathrm{cluster\_reads}= n + \ceil*{\frac{bR}{m}}O(d^2)
 \end{equation}
+
 \begin{figure}
@@
 
-\caption{Cluster reads configurations leading to different number of
-  seeks.}
+\caption{In-memory block arrangements leading to different number of
+  seeks in cluster reads. Blue blocks need seeking before each of
+  their slices ($d$ seeks in total). Red blocks need seeking before
+  each of their rows ($d^2$ seeks). Purple blocks require seeking
+  before each row and each slice ($d^2$ seeks). Green blocks are
+  written with no seeking.}
 \label{fig:cluster-reads}
@@
 
-If $\frac{Rb}{\sqrt[3]{n}}>m\geq\frac{Rb}{\sqrt[3]{n}^2}$, i.e., a
-full row of blocks fits in memory but a full slice does not, then a
-variant of cluster reads consists in adjusting the memory load to the
-highest value that makes sure that only full rows are written. In this
-case:
+If $m\geq\frac{Rb}{\sqrt[3]{n}^2}$, i.e., a
+full row of blocks fits in memory, then a
+variant adjusts the memory load to the
+highest value for which only full rows are written. In this
+case, $f_i \in \{1,d,2d\}$ and:
 \begin{equation}
-  N_\mathrm{cluster\_reads\_v1}= n + \frac{bR}{m|}d
+  N_\mathrm{cluster\_reads\_v1}= n + \frac{bR}{m|}O(d)
 \end{equation}
-where $m|=\frac{Rb}{\sqrt[3]{n}^2}\floor*{\frac{m\sqrt[3]{n}^2}{Rb}}$
+where
+$m|=\frac{Rb}{\sqrt[3]{n}^2}\floor*{\frac{m\sqrt[3]{n}^2}{Rb}}$. It
+should be noted that $m|$ is lower than $m$, therefore
+$N_\mathrm{cluster\_reads\_v1}$ might be greater than
+$N_\mathrm{cluster\_reads}$ for small values of $d$.
 
-And if $m\geq\frac{Rb}{\sqrt[3]{n}}$, i.e., a full slice of blocks
+If $m\geq\frac{Rb}{\sqrt[3]{n}}$, i.e., a full slice of blocks
 fits in memory, then a second variant adjusts
-the memory load to the highest value that makes sure that only full
+the memory load to the highest value for which only full
 slices are written. In this case:
 \begin{equation}
-  N_\mathrm{cluster\_reads\_v2}= n + \frac{bR}{m||}d
+  N_\mathrm{cluster\_reads\_v2}= n + \frac{bR}{m||}
 \end{equation}
-where $m||=\frac{Rb}{\sqrt[3]{n}}\floor*{\frac{m\sqrt[3]{n}}{Rb}}$
+where $m||=\frac{Rb}{\sqrt[3]{n}}\floor*{\frac{m\sqrt[3]{n}}{Rb}}$.
 

### 2017-06-21 paper.tex
Remove model-based comparisons (now pointless)

@@
 
-\subsection{Model-based comparisons}
-
-Figure~\ref{fig:models} plots the total merge time obtained with the
-different algorithms, using the values in~\ref{sec:benchmark} and
-different amounts of available memory. Multiple reads behave
-consistently better regardless of the amount of available
-memory. Cluster reads reach similar performance as available memory
-increases.
-\begin{figure}
-  \includegraphics[width=\columnwidth]{figures/model.pdf}
-  % see function.py
-  \caption{Model comparison}
-  \label{fig:models}
-\end{figure}
-\todo{redraw the figure. Benchmark $\alpha$ and $\beta$ more
-  accurately on the HDD or SSD. In the previous sections, compare the
-  model with real experiments, to validate the model.}
-
 \subsection{Lossless Compression}

### 2017-06-21 paper.tex
Updates about disks

@@
 \end{itemize}
-Section~\ref{sec:algos} presents our algorithms and models,
+Section~\ref{sec:algos} presents our algorithms and their complexity,
 Section~\ref{sec:results} details the results and
@@
 
-\section{Algorithms and Models}
+\section{Algorithms}
 \label{sec:algos}
@@
 \end{itemize}
-For the sake of mathematical simplicity, the models below assume that
+For the sake of mathematical simplicity, we assume that
 the reconstructed image and blocks are cubic. Under this assumption,
@@
 
-\subsection{Benchmark}
+\section{Benchmark}
 \label{sec:benchmark}
 
+\subsubsection{Implementation}
+
+Our algorithms are implemented in Python using the Nibabel
+library. \todo{Val\'erie and Yongping: (1) extract ImageUtils from
+  mapreduce repository (preserve git history; just copy the repo and
+  remove the mapreduce stuff in the copy). Link to the github
+  repo. (2) Describe the code briefly, in particular the
+  implementation tricks used for each implementation.}
+
 \subsubsection{Data}
-As a running example to benchmark our algorithms, we use the 76~GB
-BigBrain image split in 125 chunks. The data is available
-\todo{[here]}.
+To benchmark our library we use the 76~GB BigBrain image split in 125
+non-overlapping chunks. The data is available \todo{[here]}.
 
 \subsubsection{Infrastructure}
- We
-perform experiments using the two disks below:
-- HD ...
-- SSD ...
 
-\todo{Benchmark the disks to measure $\alpha_r$, $\alpha_w$ and $\beta$.}
-
-\subsubsection{Software implementation}
-
-Our algorithms are implemented in Python using the Nibabel
-library. \todo{Software release.}
+We run experiments using a Dell Precision Tower 3620 with CentOS Linux release
+7.3.1611, 32~GB of RAM and two disks:
+\begin{itemize}
+  \item Hard disk drive (HDD): HGST Travelstar 7K1000, 7200~rpm, 931GiB (1TB), firmware version JB0OA3W0.
+    \item Solid-state drive (SSD): SanDisk X400 2.5, 238GiB (256GB),
+      firmware version X4130012.
+\end{itemize}
+Both drives used 512-byte logical sectors, 4096-byte physical sectors,
+SATA >3.1 (6.0 Gb/s) and were accessed through the XFS file system.
+%smartctl --xall /dev/sda
 

### 2017-07-04 paper.tex
Fixed number of seeks for multiple reads
change mreads algo pseudocode

@@
 Sections~\ref{sec:discussion} discusses them.
+
+%\href{http://dl.acm.org/citation.cfm?id=218395}
+
+\newpage
 
@@
   \begin{algorithmic}
-    \STATE list = sort blocks by increasing (x,y,z) values
-    \STATE initialize buffer
-    \STATE pointer = 0, max\_index=(m-1)
-    \WHILE{max\_index \textless sizeof(reconstructed\_image)}
-    \WHILE{pointer \textless max\_index}
-        \STATE block = read block that contains voxel index\_to\_voxel(pointer)
-        \STATE insert in buffer all voxels (x,y,z) of block where voxel\_to\_index(x,y,z) \textless max\_index
-        \STATE update pointer to point to the next unassigned element in buffer
+  \STATE sorted\_split\_name\_list = sort blocks by increasing (x,y,z) values 
+  \STATE start\_index = 0, end\_index=(m-1)
+  \WHILE{end\_index \textless sizeof(reconstructed\_image)}
+    \STATE initialize data\_buffer
+    \FOR{split\_name in sorted\_split\_name\_list}
+      \IF{indexes of split\_name is in next\_write\_range}
+        \STATE read split;
+        \STATE extract all indexes in next write range;
+        \STATE assgin the corresponding data to the data\_buffer
+      \ENDIF
+    \ENDFOR
+    \STATE write data\_buffer in reconstructed\_image 
+    \STATE clear data\_buffer
+    \STATE start\_index = end\_index + 1
+    \STATE end\_index += mem
   \ENDWHILE
-       \STATE write buffer in reconstructed image
-       \STATE clear buffer
-       \STATE pointer = max\_index + 1
-       \STATE max\_index += m
-    \ENDWHILE
+
   \end{algorithmic}
@@
 \begin{equation}
-  N_\mathrm{multiple\_reads}= n \max\left(1,\ceil*{\frac{d}{\ceil*{\frac{m}{nd^2b}}}}\right)+ \ceil*{\frac{bR}{m}}
+  N_\mathrm{multiple\_reads}= n\left( \ceil*{\frac{\left((d-1)D^2+dD\right)b}{m}}\right) + 2\ceil*{\frac{bR}{m}}
 \end{equation}
@@
 
+\newpage
+
 \section{Discussion}

### 2017-07-24 scripts/benchmark/seek.py
Updates in figures and model

@@ -6,7 +6,7 @@ import string
 import tempfile
 import os
 
-filename="/data/bigbrain_40microns.nii"
+filename="/home/glatard/code/mapreduce/data/recon.nii"
 
 
 old_offset = 0

### 2017-07-24 scripts/model.py

@@ -1,74 +1,105 @@
+#!/usr/bin/env python
+
 import math
 import matplotlib.pyplot as plt
 
 def cubic_root(x):
   return x**(1.0/3)
 
-# returns (t2-t1)/t1, where t1 is the sequential time ("read slices, write slices")
-# and t2 is the "random time" (read blocks, write blocks)
-#
-# F: file size in bytes. Also file size in voxels. Typical value: 10**9 bytes (1 GB).
-# N: number of blocks and number of slices. Must be lower than F**(1/3). Typical value: 100.
-# alpha_r: data read rate. Typical value: 100 MB/s
-# alpha_w: data write rate. Typical value: 100 MB/s (?)
-# beta: disk access time. Typical value: 5 ms
-# m: amount of available memory, in bytes
-# b: number of bytes per voxel
 
-def t_slices(N,F,alpha_r,alpha_w,beta,b):
-  return b*F/alpha_r+b*F/alpha_w+2*beta*N
+# Seek function, typically a step function
+def sigma(distance,beta):
+  return beta
+  if distance > 10**9:
+    return beta # 50ms, on HDD
+  else:
+    return beta/4.0 # 7ms - consistent with fastest seek time in
+  # https://en.wikipedia.org/wiki/Hard_disk_drive_performance_characteristics#SEEKTIME
 
-def t_buff_slices(N,F,alpha_r,alpha_w,beta,m,b):
-  return b*F/alpha_r+b*F/alpha_w+beta*(N+math.ceil(b*F/m))
-
-def t_blocks(N,F,alpha_r,alpha_w,beta,b):
-    return F*b/alpha_r+F*b/alpha_w+beta*N*(1+cubic_root(F/N))
-
-def t_multiple_reads(N,F,alpha_r,alpha_w,beta,m,b):
-    return F*b/alpha_r+F*b/alpha_w+beta*math.ceil(F*b/m)*(1+math.ceil(m*N/(F*b)))
-
-def t_cluster_reads(N,F,alpha_r,alpha_w,beta,m,b):
-    a = m/(b*F)*cubic_root(N)
-    return b*F/alpha_r+b*F/alpha_w+beta*(N+math.ceil(F*b/m)*(cubic_root(F/N)*(math.ceil(a)-math.floor(a))+1))
+# See notations in paper, section II.A
   
-N=125.0
[…]

### 2017-07-24 paper.tex
Started explaning equation for multiple reads. Stopped because equation seems to be incorrect.
Updates in description of cluster reads

@@
 
-\title{Sequential algorithms to split and merge terabyte-scale 3D images}
+\title{Sequential algorithms to split and merge ultra-high resolution 3D images}
 
@@
 possibly using locality-aware file systems such as HDFS, software
-libraries are needed to split and merged such images efficiently. We
+libraries are needed to split and merge such images efficiently. We
 introduce and compare a family of algorithms to address this problem
@@
 \item $m$: amount of available RAM (in B).
+  \item $m'$: amount of used RAM (in B). $m'\leq m$.
 \end{itemize}
@@
 
-As detailed in Algorithm~\ref{algo:cluster-reads}, cluster reads load
-multiple blocks in memory, concatenate them in a buffer and write the
-buffer sequentially in the reconstructed image. The buffer is a
+The main idea of cluster reads is to load multiple blocks in memory,
+concatenate them in a buffer and write the buffer in the reconstructed
+image. Seeking is reduced compared to naive block merging since
+contiguous parts of the buffer will be written without seeking. In
+cluster reads, a given block is accessed only once. That is, the
+amount of used memory $m'$ is such that $m'=kd^3b, k \in \mathbb{N}$.
+
+The buffer is a
 slightly complex data structure, for instance an associative array or
@@
 \end{figure}
+
+The number of seeks performed by cluster reads depends on how blocks
+loaded in memory arrange in the reconstructed image. In the best case,
+complete contiguous slices of the reconstructed image can be assembled
+in memory and written without any seeking. In the worst case, some
+rows in the reconstructed image are not covered by blocks loaded in
+memory: O($d^2$) seeks are then required to write the memory load, one
+for every partial row in every partial slice. In the intermediary
+case, rows are complete but some slices can only be partially
+reconstructed: O($d$) seeks are then
+required. In the general case, we have:
+\begin{equation}
+  N_\mathrm{cluster\_reads}= n + \sum_{i=1}^{\ceil*{\frac{bR}{m}}}f_i
+\end{equation}
+where $f_i$ is the number of seeks done in the $i^{th}$ memory
+load. Figure~\ref{fig:cluster-reads} lists the possible block
+arrangements and the corresponding values of $f_i$. \todo{Val\'erie:
+  Check that ordering in the Figure is consistent with Nifti's.}  From
+this first analysis, we conjecture that incomplete block rows and
+incomplete block slices should be avoided whenever possible. That is,
+our implementation of cluster reads will consider only three cases:
+\begin{itemize}
+  \item Case 1: $m<d^3\sqrt[3]{n}b=\frac{D^3}{\sqrt[3]{n}^2b}$, that is, a
+    row of blocks does not fit in memory. Incomplete rows of
+    blocks will be merged, with complexity O($d^2$).
+  \item Case 2: $\frac{D^3}{\sqrt[3]{n}^2}b\leq m<\frac{D^3}{\sqrt[3]{n}}b$,
+    that is, a row of blocks fits in memory but a slice of blocks does
+    not. Only complete rows of blocks will be merged, with
+    complexity O(d). We have $m'=\floor*{\frac{m}{d^3\sqrt[3]{n}b}}$.
+    \item Case 3: $\frac{D^3}{\sqrt[3]{n}}b \leq m < D^3b$, that is, a slice
+      of blocks fits in memory. Only complete slices of blocks
+      will be merged, with complexity O(1). We have
+      $m'=\floor*{\frac{m}{d^3\sqrt[3]{n}^2b}}$.
+\end{itemize}
+
+Algorithm~\ref{algo:cluster-reads} shows the resulting cluster reads
+method. Function \texttt{adjust\_memory} (line 4) returns $m'$
+according to the three cases above. Function \texttt{sizeof} (line 6)
+returns the memory used by the buffer, including only its allocated
+segments. It should be noted that writing the buffer to the
+reconstructed image (line 11) might involve some manipulation in the
+buffer to merge its continuous segments. 
 \begin{algorithm}[h]
@@
   \label{algo:cluster-reads}
-  \begin{algorithmic}
+  \begin{algorithmic}[1]
     \STATE list = sort blocks by increasing (x,y,z) values
@@
     \STATE index = 0
+    \STATE m'=adjust\_memory(m)    
     \WHILE{index\textless n}
-      \WHILE{sizeof(buffer)\textless m}
+      \WHILE{sizeof(buffer)\textless m'}
         \STATE block = list[index]
@@
 
-The number of seeks performed by cluster reads depends on how blocks
-loaded in memory arrange in the reconstructed image. In the best case,
-complete contiguous slices of the reconstructed image can be assembled
-in memory and written without any seeking. In the worst case, some
-rows in the reconstructed image are not covered by blocks loaded in
-memory: O($d^2$) seeks are then required to write the memory load, one
-for every partial row in every partial slice. In the intermediary
-case, rows are complete but some slices can only be partially
-reconstructed: O($d$) seeks are then
-required. We have:
-\begin{equation}
-  N_\mathrm{cluster\_reads}= n + \sum_{i=1}^{\ceil*{\frac{bR}{m}}}f_i
-\end{equation}
-where $f_i$ is the number of seeks done in the $i^{th}$ memory
-load. Figure~\ref{fig:cluster-reads} lists the possible block
-arrangements and the corresponding values of $f_i$. \todo{Val\'erie: Check that
-  ordering is consistent with Nifti's.}
 %% In case (1) on
@@
 %% slices are incomplete, $f_i$=$2d^2-1$.
-In general:
-\begin{equation}
-  N_\mathrm{cluster\_reads}= n + \ceil*{\frac{bR}{m}}O(d^2)
-\end{equation}
 
@@
 
-If $m\geq\frac{Rb}{\sqrt[3]{n}^2}$, i.e., a
-full row of blocks fits in memory, then a
-variant adjusts the memory load to the
-highest value for which only full rows are written. In this
-case, $f_i \in \{1,d,2d\}$ and:
+The number of seeks performed by cluster reads depends on the three
+cases listed above. In case 1, $f_i$ might be (a) $d^2$
+ or (b) $2d^2$
+depending whether the memory load
+overlaps on multiple block slices, which will happen
+$\alpha$
+times where:
+\begin{equation*}
+  \alpha=\sqrt[3]{n}-1-\floor*{\frac{n-\sqrt[3]{n}^2}{LCM\left(\frac{m'}{d^3},\sqrt[3]{n}^2\right)}}
+\end{equation*}
+ We have:
 \begin{equation}
-  N_\mathrm{cluster\_reads\_v1}= n + \frac{bR}{m|}O(d)
+  N_\mathrm{cluster\_reads\_v1}= n + \frac{bR}{m'}d^2+d^2\alpha
 \end{equation}
-where
-$m|=\frac{Rb}{\sqrt[3]{n}^2}\floor*{\frac{m\sqrt[3]{n}^2}{Rb}}$. It
-should be noted that $m|$ is lower than $m$, therefore
-$N_\mathrm{cluster\_reads\_v1}$ might be greater than
-$N_\mathrm{cluster\_reads}$ for small values of $d$.
-
-If $m\geq\frac{Rb}{\sqrt[3]{n}}$, i.e., a full slice of blocks
-fits in memory, then a second variant adjusts
-the memory load to the highest value for which only full
-slices are written. In this case:
+Similarly, in Case 2 we have:
 \begin{equation}
-  N_\mathrm{cluster\_reads\_v2}= n + \frac{bR}{m||}
+  N_\mathrm{cluster\_reads\_v2}= n + \frac{bR}{m'}d+d\alpha
 \end{equation}
-where $m||=\frac{Rb}{\sqrt[3]{n}}\floor*{\frac{m\sqrt[3]{n}}{Rb}}$.
+And in Case 3:
+\begin{equation}
+  N_\mathrm{cluster\_reads\_v3}= n + \frac{bR}{m'}
+\end{equation}
 
@@
         \STATE extract all indexes in next write range;
-        \STATE assgin the corresponding data to the data\_buffer
+        \STATE write the corresponding data to the data\_buffer
       \ENDIF
@@
 
+In multiple reads, a given chunk will be accessed $k$ times and we have: 
+\begin{equation*}
+   N_\mathrm{multiple\_reads}= nk + \ceil*{\frac{bR}{m}}
+\end{equation*}
+$k$ depends on the span of memory offsets covered by a chunk in the
+reconstructed image, that is, $\left[(d-1)D^2+dD\right]b$.
+
+Finally:
 \begin{equation}

### 2017-07-24 figures/export.sh
Updated figure 2 to reflect the new cases

@@ -1,6 +1,6 @@
 #!/bin/bash
 
-for name in buffer-cluster-reads case1 case2 case3-a case3-b case4 case5 case6-a case6-b case7
+for name in buffer-cluster-reads case1-a case1-b case1-c case2-a case2-b case3
 do
   echo "Exporting ${name}.svg"
   inkscape -D -z --file=figures/${name}.svg --export-pdf=figures/${name}.pdf --export-latex

### 2017-07-24 paper.tex
Updated figure 2 to reflect the new cases
Fixed figure 2 (again)

@@
 
-The buffer is a
+The buffer might be a
 slightly complex data structure, for instance an associative array or
@@
 complete contiguous slices of the reconstructed image can be assembled
-in memory and written without any seeking. In the worst case, some
-rows in the reconstructed image are not covered by blocks loaded in
-memory: O($d^2$) seeks are then required to write the memory load, one
-for every partial row in every partial slice. In the intermediary
-case, rows are complete but some slices can only be partially
-reconstructed: O($d$) seeks are then
-required. In the general case, we have:
-\begin{equation}
-  N_\mathrm{cluster\_reads}= n + \sum_{i=1}^{\ceil*{\frac{bR}{m}}}f_i
-\end{equation}
-where $f_i$ is the number of seeks done in the $i^{th}$ memory
-load. Figure~\ref{fig:cluster-reads} lists the possible block
-arrangements and the corresponding values of $f_i$. \todo{Val\'erie:
-  Check that ordering in the Figure is consistent with Nifti's.}  From
-this first analysis, we conjecture that incomplete block rows and
-incomplete block slices should be avoided whenever possible. That is,
-our implementation of cluster reads will consider only three cases:
+in memory and written without any seeking. In the worst case, the
+memory load only partially covers rows in the reconstructed image:
+O($d^2$) seeks are then required during writing, one for every partial
+row in every partial slice. In the intermediary case, rows are
+complete but some slices can only be partially reconstructed: O($d$)
+seeks are then required.
+
+\todo{Check that ordering in the Figure is consistent with Nifti's.}
+
+Incomplete block rows and incomplete block slices lead to extra
+seeking and should be avoided as much as possible.  Thus, cluster
+reads consider only three cases, represented in
+Figure~\ref{fig:cluster-reads}:
 \begin{itemize}
@@
     row of blocks does not fit in memory. Incomplete rows of
-    blocks will be merged, with complexity O($d^2$).
+    blocks will be merged, with $d^2$ seeks per memory load.
   \item Case 2: $\frac{D^3}{\sqrt[3]{n}^2}b\leq m<\frac{D^3}{\sqrt[3]{n}}b$,
@@
     not. Only complete rows of blocks will be merged, with
-    complexity O(d). We have $m'=\floor*{\frac{m}{d^3\sqrt[3]{n}b}}$.
+    $d$ seeks per memory load. We have $m'=\floor*{\frac{m}{d^3\sqrt[3]{n}b}}$.
     \item Case 3: $\frac{D^3}{\sqrt[3]{n}}b \leq m < D^3b$, that is, a slice
       of blocks fits in memory. Only complete slices of blocks
-      will be merged, with complexity O(1). We have
+      will be merged, a single seek per memory load. We have
       $m'=\floor*{\frac{m}{d^3\sqrt[3]{n}^2b}}$.
@@
 \def\svgwidth{0.3\columnwidth}
-\input{figures/case1.pdf_tex}
+\input{figures/case1-a.pdf_tex}
 \def\svgwidth{0.3\columnwidth}
-\input{figures/case2.pdf_tex}
+\input{figures/case2-a.pdf_tex}
 \def\svgwidth{0.3\columnwidth}
-\input{figures/case3-a.pdf_tex}
-\def\svgwidth{0.3\columnwidth}
-\input{figures/case3-b.pdf_tex}
-\def\svgwidth{0.3\columnwidth}
-\input{figures/case4.pdf_tex}
-\def\svgwidth{0.3\columnwidth}
-\input{figures/case5.pdf_tex}
-\def\svgwidth{0.3\columnwidth}
-\input{figures/case6-a.pdf_tex}
-\def\svgwidth{0.3\columnwidth}
-\input{figures/case6-b.pdf_tex}
-\def\svgwidth{0.3\columnwidth}
-\input{figures/case7.pdf_tex}
+\input{figures/case3.pdf_tex}
 
-\caption{In-memory block arrangements leading to different number of
-  seeks in cluster reads. Blue blocks need seeking before each of
-  their slices ($d$ seeks in total). Red blocks need seeking before
-  each of their rows ($d^2$ seeks). Purple blocks require seeking
-  before each row and each slice ($d^2$ seeks). Green blocks are
-  written with no seeking.}
+\caption{Memory-load configurations in cluster reads, leading to
+  different number of seeks. Red blocks need seeking before each of
+  their rows ($d^2$ seeks). Blue blocks need seeking before each of
+  their slices ($d$ seeks in total). Green blocks need only a single
+  seek. Grey, dashed, transparent blocks represent the contiguous
+  memory loads and are added for the sake of visualization.}
 \label{fig:cluster-reads}
@@
 
-The number of seeks performed by cluster reads depends on the three
-cases listed above. In case 1, $f_i$ might be (a) $d^2$
- or (b) $2d^2$
-depending whether the memory load
-overlaps on multiple block slices, which will happen
-$\alpha$
-times where:
-\begin{equation*}
-  \alpha=\sqrt[3]{n}-1-\floor*{\frac{n-\sqrt[3]{n}^2}{LCM\left(\frac{m'}{d^3},\sqrt[3]{n}^2\right)}}
-\end{equation*}
- We have:
+The total number of seeks performed by cluster reads depends on the three
+cases listed above and on the overall number of memory loads.
 \begin{equation}

### 2017-07-24 figures/export.sh

@@ -1,6 +1,6 @@
 #!/bin/bash
 
-for name in buffer-cluster-reads case1-a case1-b case1-c case2-a case2-b case3
+for name in buffer-cluster-reads case1-a case2-a case3
 do
   echo "Exporting ${name}.svg"
   inkscape -D -z --file=figures/${name}.svg --export-pdf=figures/${name}.pdf --export-latex

### 2017-07-25 paper.tex
Updated methods sub-section on cluster reads

@@
 % new commands
-\newcommand{\todo}[1]{
-  \color{red}\emph{[#1]}
-  \color{black}
-}
+\newcommand{\todo}[1]{\marginpar{\parbox{18mm}{\flushleft\tiny\color{red}\textbf{TODO}:
+      #1}}}
 
@@
 representation. That is, x slices are written one after the other, and
-in every x slice, y rows are written one after the other. NifTI, the
+in every x slice, y rows are written one after the other. NIfTI, the
 file format defined by the Neuroimaging Informatics Technology
-Initiative Nifti\footnote{\url{https://nifti.nimh.nih.gov}} is such a
+Initiative NIfTI\footnote{\url{https://nifti.nimh.nih.gov}} is such a
 format, used by the vast majority of the neuroimaging community. In
@@
 contiguous parts of the buffer will be written without seeking. In
-cluster reads, a given block is accessed only once. That is, the
-amount of used memory $m'$ is such that $m'=kd^3b, k \in \mathbb{N}$.
+cluster reads, a given block is accessed only once.
 
@@
 between such sequences. Figure~\ref{fig:cluster-reads-buffer}
-illustrates how the buffer would fill up when two blocks of one slice
-and three rows per slice are loaded. Implementation details about the
+illustrates how the buffer would fill up for the two first blocks in a reconstructed image.Implementation details about the
 buffer are given later.
@@
 \centering
-\def\svgwidth{\columnwidth}
-\input{figures/buffer-cluster-reads.pdf_tex}
-\caption{Illustration of the buffer used in cluster reads. White
-  portions are not allocated. [a,b] refers to block a, row b.}
+\def\svgwidth{0.3\columnwidth}
+\input{figures/buffer.pdf_tex}
+\caption{Illustration of the buffer used in cluster reads, showing how
+  the two first 4x4 blocks of a reconstructed image would be arranged
+  in the buffer.  White portions in the buffer are not allocated.  The
+  column-major order is used, similar to the NIfTI format: i is the
+  fastest changing dimension, followed by j, followed by k.}
 \label{fig:cluster-reads-buffer}
@@
 
-\todo{Check that ordering in the Figure is consistent with Nifti's.}
+\begin{figure}
+\centering
+\def\svgwidth{0.3\columnwidth}
+\input{figures/case1-a.pdf_tex}
+\def\svgwidth{0.3\columnwidth}
+\input{figures/case2-a.pdf_tex}
+\def\svgwidth{0.3\columnwidth}
+\input{figures/case3.pdf_tex}
+\caption{Memory-load configurations in cluster reads, leading to
+  different number of seeks. Red blocks need seeking before each of
+  their rows ($d^2$ seeks). Blue blocks need seeking before each of
+  their slices ($d$ seeks in total). Green blocks need only a single
+  seek. Grey, dashed, transparent blocks represent the contiguous
+  memory loads and are added for the sake of visualization.}
+\label{fig:cluster-reads}
+% See source files in
+% https://drive.google.com/drive/folders/0BzT7pbvPDUdQRXppVlhfdS0ycGc?usp=sharing
+\end{figure}
+Our cluster reads algorithm focuses on the three memory load
+configurations represented in Figure~\ref{fig:cluster-reads}, that is,
+the amount of memory $m'$ used by the algorithm is rounded down to the
+closest number of complete blocks (case 1), complete block rows (case
+2) or complete block slices (case 3). This is in general reasonable
+since adding an incomplete row to a set of complete ones multiplies
+the number of required seeks by $d$, as illustrated in
+Figure~\ref{fig:avoided-configurations}-Left. In some cases though,
+rounding $m$ down to $m'$ might increase the number of required
+memory load to a point that the overall number of seeks also
+increases. However, such cases are a bit pathological and their
+complete description requires extensive calculations involving modulo
+arithmetic, which we felt were unwieldy to report here.
+\begin{figure}[b]
+  \centering
+\def\svgwidth{0.3\columnwidth}
+\input{figures/incomplete-rows.pdf_tex}
+\hfill
+\def\svgwidth{0.3\columnwidth}
+\input{figures/overlap.pdf_tex}
+\caption{Configurations that increase the number of seeks per
+  memory-load and are thus deliberately avoided by cluster
+  reads. Left: configuration with incomplete block rows (multiplies
+  number of seeks by $d$). Right: configuration with block rows that
+  overlap multiple block slices (multiplies number of seeks by $2$).}
+\label{fig:avoided-configurations}
+\end{figure}
+To summarize, the amount of memory used $m'$ is set as follows in each of the 3 cases:
+%% \begin{equation*}
+%%   m_{case 1}' = \floor*{\frac{m}{d^3b}}; m_{case 2}' = \floor*{\frac{m}{d^3b\sqrt[3]{n}}}; m_{case 3}' = \floor*{\frac{m}{d^3b\sqrt[3]{n}^2}}
+%% \end{equation*}
+%% which gives:
+\begin{equation*}
+  m_1' = \floor*{\frac{mn}{Rb}};
+  m_2' = \floor*{\frac{m\sqrt[3]{n}^2}{Rb}};
+  m_3' = \floor*{\frac{m\sqrt[3]{n}}{Rb}}
+\end{equation*}
 
-Incomplete block rows and incomplete block slices lead to extra
-seeking and should be avoided as much as possible.  Thus, cluster
-reads consider only three cases, represented in
-Figure~\ref{fig:cluster-reads}:
-\begin{itemize}
-  \item Case 1: $m<d^3\sqrt[3]{n}b=\frac{D^3}{\sqrt[3]{n}^2b}$, that is, a
-    row of blocks does not fit in memory. Incomplete rows of
-    blocks will be merged, with $d^2$ seeks per memory load.
-  \item Case 2: $\frac{D^3}{\sqrt[3]{n}^2}b\leq m<\frac{D^3}{\sqrt[3]{n}}b$,
-    that is, a row of blocks fits in memory but a slice of blocks does
-    not. Only complete rows of blocks will be merged, with
-    $d$ seeks per memory load. We have $m'=\floor*{\frac{m}{d^3\sqrt[3]{n}b}}$.
-    \item Case 3: $\frac{D^3}{\sqrt[3]{n}}b \leq m < D^3b$, that is, a slice
-      of blocks fits in memory. Only complete slices of blocks
-      will be merged, a single seek per memory load. We have
-      $m'=\floor*{\frac{m}{d^3\sqrt[3]{n}^2b}}$.
-\end{itemize}
+In addition, our algorithm avoids configurations where the memory load
+overlaps multiple block slices in case 2 or multiple block rows in
+case 1, as such overlaps multiply the number of required seeks
+(see Figure~\ref{fig:avoided-configurations}-Right). Thus, the total
+number of memory loads $x$ required to reconstruct the image is:
+\begin{equation*}
+  x_1 = \ceil*{\frac{Rb}{\sqrt[3]{n}^2m_1'}}\sqrt[3]{n}^2;
+  x_2 = \ceil*{\frac{Rb}{\sqrt[3]{n}m_2'}}\sqrt[3]{n};
+  x_3 = \ceil*{\frac{Rb}{m_3'}}
+\end{equation*}
 
-Algorithm~\ref{algo:cluster-reads} shows the resulting cluster reads
-method. Function \texttt{adjust\_memory} (line 4) returns $m'$
-according to the three cases above. Function \texttt{sizeof} (line 6)
-returns the memory used by the buffer, including only its allocated
-segments. It should be noted that writing the buffer to the
-reconstructed image (line 11) might involve some manipulation in the
-buffer to merge its continuous segments. 
+Finally, the total number of seeks performed by cluster reads to reconstruct the image is:
+%% \begin{equation*}
+%% N_\mathrm{cr} =
+%% \begin{cases}
+%%   n + \sqrt[3]{n}^2 \ceil*{\frac{Rb}{\sqrt[3]{n}^2m_1'}}d^2 & \text{if } \frac{m}{d^3b} < \sqrt[3]{n}\\
+%%   n + \sqrt[3]{n} \ceil*{\frac{Rb}{\sqrt[3]{n}m_2'}}d      & \text{if } \sqrt[3]{n} \leq \frac{m}{d^3b} < \sqrt[3]{n}^2\\
+%%   n + \ceil*{\frac{Rb}{m_3'}}                              & \text{if } \sqrt[3]{n}^2 \leq \frac{m}{d^3b} < n\\
+%%   n + 1                                                          & \text{otherwise}
+%% \end{cases} \label{eq:seeks-cluster-reads}
+%% \end{equation*}
+%% And since $d=\frac{D}{\sqrt[3]{n}}$ and $R=D^3$, we finally have:
+\begin{equation}
+N_\mathrm{cr} =
+\begin{cases}
+  n + \ceil*{\frac{Rb}{\sqrt[3]{n}^2m_1'}}\sqrt[3]{R}^2     & \text{if } \frac{mn}{Rb} < \sqrt[3]{n}\\
+  n + \ceil*{\frac{Rb}{\sqrt[3]{n}m_2'}}\sqrt[3]{R}        & \text{if } \sqrt[3]{n} \leq \frac{mn}{Rb} < \sqrt[3]{n}^2\\
+  n + \ceil*{\frac{Rb}{m_3'}}                              & \text{if } \sqrt[3]{n}^2 \leq \frac{mn}{Rb} < n\\
+  n + 1                                                          & \text{otherwise}
+\end{cases} \label{eq:seeks-cluster-reads-1}
+\end{equation}
+
+Cluster reads are described in
+Algorithm~\ref{algo:cluster-reads}. Function \texttt{switch} (line 4)
+selects one of the three cases based on the amount of available memory
+and the number of blocks. It returns \texttt{m'} and \texttt{case},
+the identifier of the selected case. Function \texttt{check\_overlap}
+(line 7) determines whether two blocks overlap multiple block slices
+(case 2) or multiple block rows (case 1). For case 3 it always returns
+true.  Function \texttt{sizeof} (line 8) returns the actual memory
+used by the buffer, including only its allocated segments. It should
+be noted that writing the buffer to the reconstructed image (line 11)
+might involve some manipulation in the buffer to merge its continuous
+segments. \todo{Precise this based on Valerie's final code.}
 \begin{algorithm}[h]
@@
   \begin{algorithmic}[1]
-    \STATE list = sort blocks by increasing (x,y,z) values
+    \STATE list = sort blocks by increasing (z,y,x) values
     \STATE initialize buffer
-    \STATE index = 0
-    \STATE m'=adjust\_memory(m)    
-    \WHILE{index\textless n}
-      \WHILE{sizeof(buffer)\textless m'}
-        \STATE block = list[index]
+    \STATE (m',case)=\texttt{switch}(m,n,R,b)
+    \STATE old\_block = list[0]
+    \FOR{i = 0 ; i\textless n ; i+=1}
+      \STATE block = list[i]
+      \STATE overlap = \texttt{check\_overlap}(block,old\_block,case)
+      \IF{\texttt{sizeof}(buffer) \textless m' \textbf{and} overlap=false}
       \STATE read block and insert it in buffer
-        \STATE index += 1
-        \ENDWHILE
+      \ELSE
       \STATE write buffer in reconstructed image
       \STATE clear buffer
-    \ENDWHILE
+      \STATE overlap = false
+      \ENDIF
+    \ENDFOR
   \end{algorithmic}
@@
 
-\begin{figure}
-\centering
-\def\svgwidth{0.3\columnwidth}
-\input{figures/case1-a.pdf_tex}
-\def\svgwidth{0.3\columnwidth}
-\input{figures/case2-a.pdf_tex}
-\def\svgwidth{0.3\columnwidth}
-\input{figures/case3.pdf_tex}
-
-\caption{Memory-load configurations in cluster reads, leading to
-  different number of seeks. Red blocks need seeking before each of
-  their rows ($d^2$ seeks). Blue blocks need seeking before each of
-  their slices ($d$ seeks in total). Green blocks need only a single
-  seek. Grey, dashed, transparent blocks represent the contiguous
-  memory loads and are added for the sake of visualization.}
-\label{fig:cluster-reads}
-% See source files in
-% https://drive.google.com/drive/folders/0BzT7pbvPDUdQRXppVlhfdS0ycGc?usp=sharing
-\end{figure}
-
-The total number of seeks performed by cluster reads depends on the three
-cases listed above and on the overall number of memory loads.
-\begin{equation}
-  N_\mathrm{cluster\_reads\_v1}= n + \frac{bR}{m'}d^2+d^2\alpha
-\end{equation}
-Similarly, in Case 2 we have:
-\begin{equation}
-  N_\mathrm{cluster\_reads\_v2}= n + \frac{bR}{m'}d+d\alpha
-\end{equation}
-And in Case 3:
-\begin{equation}
-  N_\mathrm{cluster\_reads\_v3}= n + \frac{bR}{m'}
-\end{equation}
-
 \subsection{Buffered blocks: Multiple reads}
@@
 buffer only contains contiguous bytes, at any time. That is, a given
-block will be read multiple times.
+block will be read multiple times.     \todo{improve notations and algo in algorithm}
 
 \begin{algorithm}[h]
-  \caption{Buffered merging from blocks using multiple reads
-    \todo{improve notations and algo}.}
+  \caption{Buffered merging from blocks using multiple reads.}
   \label{algo:multiple-reads}

### 2017-07-25 figures/export.sh

@@ -1,6 +1,6 @@
 #!/bin/bash
 
-for name in buffer-cluster-reads case1-a case2-a case3
+for name in buffer case1-a case2-a case3 incomplete-rows overlap
 do
   echo "Exporting ${name}.svg"
   inkscape -D -z --file=figures/${name}.svg --export-pdf=figures/${name}.pdf --export-latex

### 2017-07-25 scripts/model.py
Reorganized data and scripts directories

@@ -1,105 +1,52 @@
 #!/usr/bin/env python
 
 import math
-import matplotlib.pyplot as plt
+from decimal import Decimal as dec
 
 def cubic_root(x):
   return x**(1.0/3)
 
-
-# Seek function, typically a step function
-def sigma(distance,beta):
-  return beta
-  if distance > 10**9:
-    return beta # 50ms, on HDD
-  else:
-    return beta/4.0 # 7ms - consistent with fastest seek time in
-  # https://en.wikipedia.org/wiki/Hard_disk_drive_performance_characteristics#SEEKTIME
-
 # See notations in paper, section II.A
   
-def t_slices(b,R,alpha_r,alpha_w,beta,n):
-  read_time = b*R/alpha_r + beta*n
-  write_time = +b*R/alpha_w+beta*n
-  seek_time = 0
-  total_time = read_time + write_time + seek_time
-  return (read_time,write_time,seek_time,total_time)
+def t_slices(b,R,n):
+  return 2*n
 
-#def t_buff_slices(N,F,alpha_r,alpha_w,beta,m,b):
-#  return b*F/alpha_r+b*F/alpha_w+beta*(N+math.ceil(b*F/m))
+def t_buff_slices(b,R,n,m):
+  return n+math.ceil(b*R/m)
 
-def t_blocks(b,R,alpha_r,alpha_w,beta,n):
+def t_blocks(b,R,n):
     d = cubic_root(R/n)
-    D = cubic_root(R)
-    read_time = R*b/alpha_r + beta*n
-    write_time = R*b/alpha_w + beta*n
-    seek_time_block = n*(sigma(b*R/2,beta))
-    seek_time_rows = n*(d*(d-1)*sigma(b*(D-d),beta))
-    seek_time_slices = n*(d*sigma(b*(D**2-d**2),beta))
-    total_time = read_time + write_time + seek_time_block+seek_time_rows+seek_time_slices
-    return (read_time,write_time,seek_time_block,seek_time_rows,seek_time_slices,total_time)
+    return n+n*d**2
     
-
-def t_sorted_blocks(b,R,alpha_r,alpha_w,beta,n):
-    (read,write,seek,total) = t_blocks(b,R,alpha_r,alpha_w,beta,n)
-    write_ = write
[…]

### 2017-07-25 paper.tex
Typo in cluster reads
Added figures and description for slices and blocks.

@@
 \label{fig:cluster-reads-buffer}
+% Data file: path
+% Gnuplot file: path
+% Script file: path
 \end{figure}
@@
 \label{fig:cluster-reads}
-% See source files in
-% https://drive.google.com/drive/folders/0BzT7pbvPDUdQRXppVlhfdS0ycGc?usp=sharing
 \end{figure}
@@
 %% \begin{equation*}
-%%   m_{case 1}' = \floor*{\frac{m}{d^3b}}; m_{case 2}' = \floor*{\frac{m}{d^3b\sqrt[3]{n}}}; m_{case 3}' = \floor*{\frac{m}{d^3b\sqrt[3]{n}^2}}
+%%   m_{case 1}' = bd^3\floor*{\frac{m}{d^3b}}; m_{case 2}' = bd^3\sqrt[3]{n}\floor*{\frac{m}{d^3b\sqrt[3]{n}}}; m_{case 3}' = bd^3\sqrt[3]{n}^2\floor*{\frac{m}{d^3b\sqrt[3]{n}^2}}
 %% \end{equation*}
@@
 \begin{equation*}
-  m_1' = \floor*{\frac{mn}{Rb}};
-  m_2' = \floor*{\frac{m\sqrt[3]{n}^2}{Rb}};
-  m_3' = \floor*{\frac{m\sqrt[3]{n}}{Rb}}
+  m_1' = \frac{Rb}{n}\floor*{\frac{mn}{Rb}};
+  m_2' = \frac{Rb}{\sqrt[3]{n}^2}\floor*{\frac{m\sqrt[3]{n}^2}{Rb}};
+  m_3' = \frac{Rb}{\sqrt[3]{n}}\floor*{\frac{m\sqrt[3]{n}}{Rb}}
 \end{equation*}
@@
 
-\subsection{Slices vs blocks}
+\subsection{Slices vs Blocks}
 
-\todo{Valerie's results at PERFORM + model.}
+Figure~\ref{fig:slices-vs-blocks} shows the merging time for Slices vs
+Blocks, on SSD and HDD. Merging with Blocks takes about twice the time
+of Slices, which motivates the need for improved algorithms to merge
+blocks. As expected, seek time is mainly responsible for this
+difference, although differences in read times are also observed,
+perhaps due to a caching effect at the disk or OS level. This behavior is
+reproduced in SSD and HDD, which shows that SSDs might not be the only
+solution to reduce seek time.  \todo{On figure: check that breakdowns
+  add up to total time. Harmonize time formats. Harmonize colors (all
+  slices in purple, all blocks in green). If time permits, try
+  histogram for breakdown, on same figure than total time.}
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=0.45\columnwidth]{figures/blocks-slices/totalmergetimeSSD1.pdf}
+  \hfill
+  \includegraphics[width=0.45\columnwidth]{figures/blocks-slices/totalmergetimeHDD1.pdf}\\
+  \includegraphics[width=0.45\columnwidth]{figures/blocks-slices/hddBreakDown1.pdf}
+  \hfill
+  \includegraphics[width=0.45\columnwidth]{figures/blocks-slices/ssdBreakDown1.pdf}
+  \caption{Merging time for Slices and Blocks. Left column: SSD. Right column: HDD. Top row: overall merging time. Bottom row: breakdown by read, write and seek time.}
+  % Gnuplot file: scripts/blocks-slices/createboxplot.gnplt
+\label{fig:slices-vs-blocks}
+\end{figure}
 
-\subsection{Sorted blocks vs blocks}
+\subsection{Sorted blocks vs blocks?}
 
@@
 
-\subsection{Cluster reads vs Multiple reads vs Buffered slices}
+\subsection{Cluster reads vs blocks and slices}
+
+\subsection{Multiple reads vs blocks and slices}
+
+\subsection{Cluster reads vs Multiple reads}
 

### 2017-07-25 paper.tex
Removed sorted blocks. Updates in algorithms.

@@
 \item $n$: number of chunks (blocks or slices).
-\item $\alpha_r$: data read rate from disk (in B/s).
-\item $\alpha_w$: data write rate to disk (in B/s).
-\item $\beta$: disk access time (in s).
-\item $\sigma: \mathbb{R}^{+} \rightarrow \mathbb{R}^{+}$: seek time (in s), function of the seek
-  distance (in B).
 \item $m$: amount of available RAM (in B).
@@
 
-The following properties are straightforward applications of these
+The following relations are straightforward applications of these
 definitions:
@@
 \end{itemize}
-\todo{Make a figure to illustrate the notations.}
+%\todo{Make a figure to illustrate the notations.}
 
@@
 We use a simplistic disk model where a disk is characterized by its
-read and write rates ($\alpha_r$ and $\alpha_w$, in B/s) and its
-access time $\beta$ (in s). In particular, we assume that seeks
-require a constant amount of time, regardless of the position seeked
-to. In practice, large variations would be expected, but modeling such
-variations would inevitably lead to models specific to the hardware,
-file system or operating system, which we intentionally avoid
-here. Thus, our goal is to find algorithms that minimize the
-\emph{number} of seek operations. Likewise, in modern systems, read
-and write times are greatly impacted by caching occurring at several
-levels, which we do not model here.
+read and write rates and its access time. In particular, we assume
+that seeks require a constant amount of time, regardless of the
+position seeked to. In practice, large variations would be expected,
+but modeling such variations would inevitably lead to models specific
+to the hardware, file system or operating system, which we
+intentionally avoid here. Thus, our goal is to find algorithms that
+minimize the \emph{number} of seek operations. Likewise, in modern
+systems, read and write times are greatly impacted by caching
+occurring at several levels, which we do not model here.
 
@@
 
-\subsection{Sorted blocks}
-
-Sorting blocks alphanumerically by (x,y,z) as shown on Algorithm 2
-reduces the seek time between blocks.
-
-\begin{algorithm}[h]
-  \caption{Merging from sorted blocks.}
-  \label{algo:sorted-blocks}
-  \begin{algorithmic}
-    \STATE list = sort blocks according to (x,y,z)
-    \FOR{each block in list}
-      \STATE read block
-      \STATE write block in reconstructed image
-    \ENDFOR
-  \end{algorithmic}
-  \end{algorithm}
-
-\todo{Not sure what we are measuring here. The number of seeks will be identical. The difference in seek time is supposed to be very minimal.}
-
 \subsection{Buffered slices}
@@
   \label{algo:buffered-slices}
-  \begin{algorithmic}
+  \begin{algorithmic}[1]
     \STATE list = sort slices by increasing x values
     \STATE initialize buffer
-    \STATE index = 0
-    \WHILE{index\textless n}
-      \WHILE{sizeof(buffer)\textless m}
+    \FOR{i = 0 ; i \textless n ; i+=1 }
       \STATE slice = list[index]
-        \STATE read slice and append it to buffer
-        \STATE index += 1
-       \ENDWHILE
+      \IF{\texttt{sizeof}(buffer)+\texttt{sizeof}(slice) $\geq$ m}
         \STATE write buffer in reconstructed image
         \STATE clear buffer
-    \ENDWHILE
+      \ENDIF
+      \STATE read slice and append it to buffer
+    \ENDFOR
   \end{algorithmic}
@@
 Cluster reads are described in
-Algorithm~\ref{algo:cluster-reads}. Function \texttt{switch} (line 4)
+Algorithm~\ref{algo:cluster-reads}. Function \texttt{switch} (line 3)
 selects one of the three cases based on the amount of available memory
@@
 true.  Function \texttt{sizeof} (line 8) returns the actual memory
-used by the buffer, including only its allocated segments. It should
-be noted that writing the buffer to the reconstructed image (line 11)
-might involve some manipulation in the buffer to merge its continuous
-segments. \todo{Precise this based on Valerie's final code.}
+used by its argument, including only its allocated segments in case
+the argument is a buffer. It should be noted that writing the buffer
+to the reconstructed image (line 11) might involve some manipulation
+in the buffer to merge its continuous segments. \todo{Precise this
+  based on Valerie's final code.}
 \begin{algorithm}[h]
@@
       \STATE overlap = \texttt{check\_overlap}(block,old\_block,case)
-      \IF{\texttt{sizeof}(buffer) \textless m' \textbf{and} overlap=false}
-      \STATE read block and insert it in buffer
-      \ELSE
+      \IF{\texttt{sizeof}(buffer)+\texttt{sizeof}(block) $\geq$ m' \textbf{or} overlap=true}
       \STATE write buffer in reconstructed image
@@
       \ENDIF
+      \STATE read block and insert it in buffer
       \ENDFOR
@@
   \label{algo:multiple-reads}
-  \begin{algorithmic}
-  \STATE sorted\_split\_name\_list = sort blocks by increasing (x,y,z) values 
+  \begin{algorithmic}[1]
+  \STATE sorted\_split\_name\_list = sort blocks by increasing (z,y,x) values 
   \STATE start\_index = 0, end\_index=(m-1)
@@
 \end{algorithm}
-
-In multiple reads, a given chunk will be accessed $k$ times and we have: 
-\begin{equation*}
-   N_\mathrm{multiple\_reads}= nk + \ceil*{\frac{bR}{m}}
-\end{equation*}
-$k$ depends on the span of memory offsets covered by a chunk in the
-reconstructed image, that is, $\left[(d-1)D^2+dD\right]b$.
-
-Finally:
-\begin{equation}
-   N_\mathrm{multiple\_reads}= n\left( \ceil*{\frac{\left((d-1)D^2+dD\right)b}{m}}\right) + 2\ceil*{\frac{bR}{m}}
-\end{equation}
 
@@
 
-\subsection{Sorted blocks vs blocks?}
-
 \subsection{Buffered slices vs slices}

### 2017-07-26 paper.tex
Updates in Multiple read model

@@
 \usepackage{multirow}
+\usepackage{stmaryrd} % for llbracket and rrbracket
 \DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}
@@
 \item $n$: number of chunks (blocks or slices).
-\item $m$: amount of available RAM (in B).
-\item $m'$: amount of used RAM (in B). $m'\leq m$.
+\item $m$: amount of available memory (in B).
+\item $m'$: amount of used memory (in B), $m'\leq m$.
 \end{itemize}
 
-The following relations are straightforward applications of these
-definitions:
-\begin{itemize}
-\item number of bytes per chunk: $\frac{Rb}{n}$
-\item number of chunks that can fit in memory: $\frac{m.n}{Rb}$
-\end{itemize}
 For the sake of mathematical simplicity, we assume that
 the reconstructed image and blocks are cubic. Under this assumption,
-we have the following definitions and properties:
+we also have the following relations:
 \begin{itemize}
@@
 \item Number of blocks required to fill a row of the reconstructed
-  image: $\sqrt[3]{n}=\nu$.
+  image: $\sqrt[3]{n}$.
 \end{itemize}
@@
 arithmetic, which we felt were unwieldy to report here.
-\begin{figure}[b]
+\begin{figure}
   \centering
@@
 \end{equation}
+\todo{Is $N_{cr}$ a continuous function of m?}
 
@@
 \begin{algorithm}[h]
-  \caption{Buffered merging from blocks using multiple reads.}
+  \caption{Buffered merging of blocks with multiple reads}
   \label{algo:multiple-reads}
@@
       \IF{indexes of split\_name is in next\_write\_range}
-        \STATE read split;
-        \STATE extract all indexes in next write range;
+        \STATE read split
+        \STATE extract all indexes in next write range
         \STATE write the corresponding data to the data\_buffer
@@
 \end{algorithm}
+
+For simplicity, we assume that $m'$ represents an integer number $k$
+of block rows (case 1, $k<\sqrt[3]{n}$), of complete rows (case 2,
+$k<d$), of tile rows (case 3, $k<\sqrt[3]{n}$), of slices (case 4,
+$k<d$) or of block slices (case 5, $k<\sqrt[3]{n}$) \todo{make a
+  figure}. In each of these 5 cases, we define $v$ as follows:
+\begin{equation*}
+  v_1=db \text{ ; }  v_2=Db \text{ ; } v_3=Ddb \text{ ; } v_4=D^2b \text{ ; } v_5=D^2db
+\end{equation*}
+so that we have:
+\begin{equation*}
+k_i=\floor*{\frac{m}{v_i}} \quad \mathrm{and} \quad m_i'=k_iv_i, \quad i \in \llbracket 1,5 \rrbracket
+\end{equation*}
+Using these notations, the number of seeks performed by multiple reads is:
+\begin{equation*}
+  N_\mathrm{MR} =
+\begin{cases}
+  \floor*{\frac{Rb}{m_1'}}(k_1+1)+1
+  + n \text{ mod } k_1& \text{if } d \leq \frac{m}{b} < D\\
+  \ceil*{\frac{Rb}{m_2'}}(\sqrt[3]{n}+1)   & \text{if } D \leq \frac{m}{b} < Dd\\
+  \floor*{\frac{Rb}{m_3'}}(k_3\sqrt[3]{n}+1)+1+\\
+  \quad \quad \sqrt[3]{n}\left(\frac{n}{\sqrt[3]{n}} \text{ mod } k_3\right)&\
+  \text{if } Dd \leq \frac{m}{b} < D^2\\
+  \ceil*{\frac{Rb}{m_4'}}(\sqrt[3]{n}^2+1) & \text{if } D^2 \leq \frac{m}{b} < D^2d\\
+  \floor*{\frac{Rb}{m_5'}}(k_5\sqrt[3]{n}^2+1)+1+&\\
+  \quad \quad \sqrt[3]{n}^2\left(\frac{n}{\sqrt[3]{n}^2} \text{ mod } k_5\right)&\
+  \text{if } D^2d \leq \frac{m}{b} < R
+\end{cases} 
+\end{equation*}
+
+which gives, using $R$ and $n$ as main variables:
+\begin{equation}
+N_\mathrm{MR} =
+\begin{cases}
+  \floor*{\frac{Rb}{m_1'}}\left(\frac{m_1'\sqrt[3]{n}}{\sqrt[3]{R}b}+1\right)+1+\\
+  \quad \quad \quad n \text{ mod }\floor*{\frac{m\sqrt[3]{n}}{\sqrt[3]{R}b}}\
+  & \text{if } \sqrt[3]{\frac{R}{n}} \leq \frac{m}{b} < \sqrt[3]{R}\\
+
+  \ceil*{\frac{Rb}{m_2'}}\left(\sqrt[3]{n}+1\right)\
+  & \text{if } \sqrt[3]{R} \leq \frac{m}{b} < \frac{\sqrt[3]{R}^2}{\sqrt[3]{n}}\\
+
+  \floor*{\frac{Rb}{m_3'}}\left(\frac{m'_3\sqrt[3]{n}^2}{\sqrt[3]{R}^2b}+1\right)+1+\\
+  \sqrt[3]{n}\left( \frac{n}{\sqrt[3]{n}} \text{ mod } \floor*{\frac{m\sqrt[3]{n}}{\sqrt[3]{R}^2b}} \right)
+  & \text{if } \frac{\sqrt[3]{R}^2}{\sqrt[3]{n}} \leq \frac{m}{b} < \sqrt[3]{R}^2\\
+
+  \ceil*{\frac{Rb}{m_4'}}\left(\sqrt[3]{n}^2+1\right)\
+  & \text{if } \sqrt[3]{R}^2 \leq \frac{m}{b} < \frac{R}{\sqrt[3]{n}}\\
+
+  \ceil*{\frac{Rb}{m_5'}}\left(\frac{m_5'n}{Rb}+1\right)+1+\\
+  \sqrt[3]{n}^2\left( \frac{n}{\sqrt[3]{n}^2} \text{ mod } \floor*{\frac{m\sqrt[3]{n}}{Rb}}\right)
+  & \text{if } \frac{R}{\sqrt[3]{n}} \leq \frac{m}{b} < R
+\end{cases} \label{eq:seeks-multiple-reads}
+\end{equation}
 
@@
 
+\subsubsection{Simulations}
+After running a few simulations, we saw that the comparison between
+Cluster reads and Multiple reads essentially depends on the three
+Cluster reads configurations.  When Cluster reads merge incomplete
+block rows (case 1), they are always outperformed by Multiple reads.
+When Cluster reads merge complete block rows (case 2), the comparison
+between Cluster reads and Multiple reads depends on the values of $D$
+and $n$. Figure~\ref{fig:model-comparison} shows two situations where
+Cluster reads and Multiple reads are alternatively better than each
+other. When Cluster reads merge complete block slices (case 3),
+Cluster reads and Multiple reads are equivalent.
+
+\begin{figure}
+  \includegraphics[width=0.45\columnwidth]{figures/model-big-brain.pdf}
+  \includegraphics[width=0.45\columnwidth]{figures/model-big-brain-rescan.pdf}
+  \caption{Number of seeks for Cluster reads vs Multiple reads when
+    Cluster reads merge complete block rows (case 2), for D=3000 and
+    b=2. Left: n=125; Right: n=64,000.}
+  % Gnuplot file: scripts/model/model.gnplt
+  \label{fig:model-comparison}
+\end{figure}
+
+\subsubsection{Experiments}
+
 \subsection{Lossless Compression}

### 2017-07-26 scripts/mreads/benchmark_mreads.py
add breakdown figure for mreads

@@ -39,12 +39,12 @@ def benchmark_mreads(mem, reconstructed, legend):
     total_time = time() - s_time
     return (total_read_time, total_write_time, total_seek_time, total_seek_number, total_time)
 
-def write_to_csv(data_dict, csv_file):
+def write_to_file(data_dict, csv_file):
     # (total_read_time, total_write_time, total_seek_time, total_seek_number, total_time)
     with open(csv_file, "a") as f:
         for k in sorted(data_dict.keys()):
             for e in data_dict[k]:
-                f.write(str(e) + ",")
+                f.write(str(e) + " ")
         f.write("\n")
 
 
@@ -69,7 +69,7 @@ def main():
             os.system("rm {}".format(files[disk][0]))
             data = benchmark_mreads(mem=mem, reconstructed=files[disk][0], legend=files[disk][1])
             data_dict[mem] = data
-        write_to_csv(data_dict, files[disk][2])
+        write_to_file(data_dict, files[disk][2])
 
 if __name__ == '__main__':
     main()

### 2017-07-26 scripts/mreads/generate_avg_data.py

@@ -1,0 +1,156 @@
+#!/usr/bin/env python
+
+input_file_mreads_ssd = "../../data/mreads/mreads_ssd.dat"
+output_file_mreads_ssd = "../../data/mreads/mreads_ssd_avg.dat"
+
+input_file_mreads_hdd = "../../data/mreads/mreads_hdd.dat"
+output_file_mreads_hdd = "../../data/mreads/mreads_hdd_avg.dat"
+
+input_file_naive_ssd = "../../data/blocks-slices/totaltime.dat"
+output_file_naive_ssd = "../../data/blocks-slices/blocks_slices_avg_ssd.dat"
+
+input_file_naive_hdd = "../../data/blocks-slices/totaltime.dat"
+output_file_naive_hdd = "../../data/blocks-slices/blocks_slices_avg_hdd.dat"
+
+def avg(l):
+    return sum(l) / float(len(l))
+
+def get_avg_mreads(input_file, output_file):
+    read_times_3g = []
+    write_times_3g = []
+    seek_times_3g = []
+    calculation_times_3g = []
+
+    read_times_6g = []
+    write_times_6g = []
+    seek_times_6g = []
+    calculation_times_6g = []
+
+    read_times_9g = []
+    write_times_9g = []
+    seek_times_9g = []
+    calculation_times_9g = []
+
+    read_times_12g = []
+    write_times_12g = []
+    seek_times_12g = []
+    calculation_times_12g = []
+
+    with open(input_file, 'r') as f:
+        for line in f:
+            if '#' not in line:
+                items = line.split(' ')
+
+                read_time = float(items[0])
+                write_time = float(items[1])
+                seek_time = float(items[2])
+                total_time = float(items[4])
+                calculation_time = total_time - seek_time - write_time - read_time
+
+                read_times_3g.append(read_time)
+                write_times_3g.append(write_time)
+                seek_times_3g.append(seek_time)
+                calculation_times_3g.append(calculation_time)
+
+                read_time = float(items[5])
+                write_time = float(items[6])
+                seek_time = float(items[7])
+                total_time = float(items[9])
+                calculation_time = total_time - seek_time - write_time - read_time
+
+                read_times_6g.append(read_time)
+                write_times_6g.append(write_time)
+                seek_times_6g.append(seek_time)
+                calculation_times_6g.append(calculation_time)
+
+                read_time = float(items[10])
+                write_time = float(items[11])
+                seek_time = float(items[12])
[…]

### 2017-07-26 paper.tex
add mwrites scipts, data

--- 2017-07-25 paper.tex
@@
 
-\subsection{Slices vs Blocks}
+\subsection{Slices vs blocks}
 
-Figure~\ref{fig:slices-vs-blocks} shows the merging time for Slices vs
-Blocks, on SSD and HDD. Merging with Blocks takes about twice the time
-of Slices, which motivates the need for improved algorithms to merge
-blocks. As expected, seek time is mainly responsible for this
-difference, although differences in read times are also observed,
-perhaps due to a caching effect at the disk or OS level. This behavior is
-reproduced in SSD and HDD, which shows that SSDs might not be the only
-solution to reduce seek time.  \todo{On figure: check that breakdowns
-  add up to total time. Harmonize time formats. Harmonize colors (all
-  slices in purple, all blocks in green). If time permits, try
-  histogram for breakdown, on same figure than total time.}
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=0.45\columnwidth]{figures/blocks-slices/totalmergetimeSSD1.pdf}
-  \hfill
-  \includegraphics[width=0.45\columnwidth]{figures/blocks-slices/totalmergetimeHDD1.pdf}\\
-  \includegraphics[width=0.45\columnwidth]{figures/blocks-slices/hddBreakDown1.pdf}
-  \hfill
-  \includegraphics[width=0.45\columnwidth]{figures/blocks-slices/ssdBreakDown1.pdf}
-  \caption{Merging time for Slices and Blocks. Left column: SSD. Right column: HDD. Top row: overall merging time. Bottom row: breakdown by read, write and seek time.}
-  % Gnuplot file: scripts/blocks-slices/createboxplot.gnplt
-\label{fig:slices-vs-blocks}
-\end{figure}
+\todo{Valerie's results at PERFORM + model.}
+% will be used as reference
 

### 2017-07-26 scripts/mwrites/generate_avg_data.py

@@ -1,0 +1,156 @@
+#!/usr/bin/env python
+
+input_file_mwrites_ssd = "../../data/mwrites/mwrites_ssd.dat"
+output_file_mwrites_ssd = "../../data/mwrites/mwrites_ssd_avg.dat"
+
+input_file_mwrites_hdd = "../../data/mwrites/mwrites_hdd.dat"
+output_file_mwrites_hdd = "../../data/mwrites/mwrites_hdd_avg.dat"
+
+input_file_naive_ssd = "../../data/blocks-slices/totaltime.dat"
+output_file_naive_ssd = "../../data/blocks-slices/blocks_slices_avg_ssd.dat"
+
+input_file_naive_hdd = "../../data/blocks-slices/totaltime.dat"
+output_file_naive_hdd = "../../data/blocks-slices/blocks_slices_avg_hdd.dat"
+
+def avg(l):
+    return sum(l) / float(len(l))
+
+def get_avg_mwrites(input_file, output_file):
+    read_times_3g = []
+    write_times_3g = []
+    seek_times_3g = []
+    calculation_times_3g = []
+
+    read_times_6g = []
+    write_times_6g = []
+    seek_times_6g = []
+    calculation_times_6g = []
+
+    read_times_9g = []
+    write_times_9g = []
+    seek_times_9g = []
+    calculation_times_9g = []
+
+    read_times_12g = []
+    write_times_12g = []
+    seek_times_12g = []
+    calculation_times_12g = []
+
+    with open(input_file, 'r') as f:
+        for line in f:
+            if '#' not in line:
+                items = line.split(' ')
+
+                read_time = float(items[0])
+                write_time = float(items[1])
+                seek_time = float(items[2])
+                total_time = float(items[4])
+                calculation_time = total_time - seek_time - write_time - read_time
+
+                read_times_3g.append(read_time)
+                write_times_3g.append(write_time)
+                seek_times_3g.append(seek_time)
+                calculation_times_3g.append(calculation_time)
+
+                read_time = float(items[5])
+                write_time = float(items[6])
+                seek_time = float(items[7])
+                total_time = float(items[9])
+                calculation_time = total_time - seek_time - write_time - read_time
+
+                read_times_6g.append(read_time)
+                write_times_6g.append(write_time)
+                seek_times_6g.append(seek_time)
+                calculation_times_6g.append(calculation_time)
+
+                read_time = float(items[10])
+                write_time = float(items[11])
+                seek_time = float(items[12])
[…]

### 2017-07-26 scripts/model.py
minor fix

@@ -1,52 +1,0 @@
-#!/usr/bin/env python
-
-import math
-from decimal import Decimal as dec
-
-def cubic_root(x):
-  return x**(1.0/3)
-
-# See notations in paper, section II.A
-  
-def t_slices(b,R,n):
-  return 2*n
-
-def t_buff_slices(b,R,n,m):
-  return n+math.ceil(b*R/m)
-
-def t_blocks(b,R,n):
-    d = cubic_root(R/n)
-    return n+n*d**2
-    
-def t_cluster_reads(b,R,n,m):
-    a=(m*n)/(R*b)
-    nu = cubic_root(n)
-    d = cubic_root(R/n)
-    if a < nu: # case 1
-      m_prime = (R*b)/(n)*math.floor(a)
-     # print "R=",R
-    #  print "b=",b
-   #   print "n=",n
-  #    print "m_prime=",m_prime
- #     print "nu=",nu
-      x=math.ceil(R*b)/(nu**2*m_prime)*nu**2
-      return n + math.ceil((R*b)/(nu**2*m_prime))*cubic_root(R)**2
-    if a < nu**2: # case 2
-      m_prime = (R*b)/(nu**2)*math.floor((m*nu**2)/(R*b))
-      return n + math.ceil((R*b)/(nu*m_prime))*cubic_root(R)
-    if a < n: #case 3
-      m_prime = (R*b)/(nu)*math.floor((m*nu)/(R*b))
-      return n + math.ceil((R*b)/(m_prime))
-    return n + 1 # memory is larger than image size
-    
-n=125.0
-b=2.0
-R=76000000000.0/b
-mem_values=[10**9,2*10**9,4*10**9,6*10**9,8*10**9,10*10**9,12*10**9,14*10**9,16*10**9]
-print "#Memory Slices Blocks Buff Slices Cluster Reads"
-for m in mem_values:
-  
-  print  m, t_slices(b,R,n), t_blocks(b,R,n), t_buff_slices(b,R,n,m), t_cluster_reads(b,R,n,m)
-
-
-

### 2017-07-26 scripts/mreads/generate_avg_data.py
fix avg generatation bug

@@ -112,30 +112,56 @@ def get_avg_naive(input_file, output_file):
     calculation_times_blocks = []
 
     with open(input_file, 'r') as f:
-        for line in f:
-            if '#' not in line:
-                items = line.split(' ')
-                read_time = float(items[9])
-                write_time = float(items[10])
-                seek_time = 0
-                total_time = float(items[5])
-                calculation_time = total_time - seek_time - write_time - read_time
+        if "hdd" in output_file:
+            for line in f:
+                if '#' not in line:
+                    items = line.split(' ')
+                    read_time = float(items[9])
+                    write_time = float(items[10])
+                    seek_time = 0
+                    total_time = float(items[5])
+                    calculation_time = total_time - seek_time - write_time - read_time
 
-                read_times_slices.append(read_time)
-                write_times_slices.append(write_time)
-                seek_times_slices.append(seek_time)
-                calculation_times_slices.append(calculation_time)
+                    read_times_slices.append(read_time)
+                    write_times_slices.append(write_time)
+                    seek_times_slices.append(seek_time)
+                    calculation_times_slices.append(calculation_time)
 
-                read_time = float(items[14])
-                write_time = float(items[16])
-                seek_time = float(items[15])
-                total_time = float(items[6])
-                calculation_time = total_time - seek_time - write_time - read_time
+                    read_time = float(items[14])
+                    write_time = float(items[16])
+                    seek_time = float(items[15])
+                    total_time = float(items[6])
+                    calculation_time = total_time - seek_time - write_time - read_time
 
-                read_times_blocks.append(read_time)
-                write_times_blocks.append(write_time)
-                seek_times_blocks.append(seek_time)
-                calculation_times_blocks.append(calculation_time)
+                    read_times_blocks.append(read_time)
+                    write_times_blocks.append(write_time)
+                    seek_times_blocks.append(seek_time)
+                    calculation_times_blocks.append(calculation_time)
+        else:
+            for line in f:
+                if '#' not in line:
+                    items = line.split(' ')
+                    read_time = float(items[7])
+                    write_time = float(items[8])
+                    seek_time = 0
+                    total_time = float(items[0])
+                    calculation_time = total_time - seek_time - write_time - read_time
+
+                    read_times_slices.append(read_time)
+                    write_times_slices.append(write_time)
[…]

### 2017-07-26 generate_figures.sh
Removed figure pdfs. Added all gnuplot scripts to generate_figures.sh.

@@ -1,0 +1,15 @@
+#!/bin/bash
+
+set -e
+set -u
+
+figures/svg/export_svg_files.sh
+scripts/blocks-slices/createboxplot.gnplt
+scripts/model/model.gnplt
+scripts/mwrites/compare_total_time.gnuplot
+scripts/mwrites/comparasion_break_down.gnuplot
+scripts/mwrites/generate_avg_data.py
+scripts/mreads/compare_total_time.gnuplot
+scripts/mreads/comparasion_break_down.gnuplot
+scripts/mreads/generate_avg_data.py
+#./scripts/mreads/number_of_seeks.gnuplot # crashes for now

### 2017-07-26 scripts/mwrites/generate_avg_data.py

@@ -1,16 +1,16 @@
 #!/usr/bin/env python
 
-input_file_mwrites_ssd = "../../data/mwrites/mwrites_ssd.dat"
-output_file_mwrites_ssd = "../../data/mwrites/mwrites_ssd_avg.dat"
+input_file_mwrites_ssd = "./data/mwrites/mwrites_ssd.dat"
+output_file_mwrites_ssd = "./data/mwrites/mwrites_ssd_avg.dat"
 
-input_file_mwrites_hdd = "../../data/mwrites/mwrites_hdd.dat"
-output_file_mwrites_hdd = "../../data/mwrites/mwrites_hdd_avg.dat"
+input_file_mwrites_hdd = "./data/mwrites/mwrites_hdd.dat"
+output_file_mwrites_hdd = "./data/mwrites/mwrites_hdd_avg.dat"
 
-input_file_naive_ssd = "../../data/blocks-slices/totaltime.dat"
-output_file_naive_ssd = "../../data/blocks-slices/blocks_slices_avg_ssd.dat"
+input_file_naive_ssd = "./data/blocks-slices/totaltime.dat"
+output_file_naive_ssd = "./data/blocks-slices/blocks_slices_avg_ssd.dat"
 
-input_file_naive_hdd = "../../data/blocks-slices/totaltime.dat"
-output_file_naive_hdd = "../../data/blocks-slices/blocks_slices_avg_hdd.dat"
+input_file_naive_hdd = "./data/blocks-slices/totaltime.dat"
+output_file_naive_hdd = "./data/blocks-slices/blocks_slices_avg_hdd.dat"
 
 def avg(l):
     return sum(l) / float(len(l))

### 2017-07-27 paper.tex
clean
Small updates in outline
Restored files permissions
Removed figure pdfs. Added all gnuplot scripts to generate_figures.sh.

--- 2017-07-26 paper.tex
@@
 \def\svgwidth{0.3\columnwidth}
-\input{figures/buffer.pdf_tex}
+\input{figures/svg/buffer.pdf_tex}
 \caption{Illustration of the buffer used in cluster reads, showing how
@@
 \def\svgwidth{0.3\columnwidth}
-\input{figures/case1-a.pdf_tex}
+\input{figures/svg/case1-a.pdf_tex}
 \def\svgwidth{0.3\columnwidth}
-\input{figures/case2-a.pdf_tex}
+\input{figures/svg/case2-a.pdf_tex}
 \def\svgwidth{0.3\columnwidth}
-\input{figures/case3.pdf_tex}
+\input{figures/svg/case3.pdf_tex}
 \caption{Memory-load configurations in cluster reads, leading to
@@
 \def\svgwidth{0.3\columnwidth}
-\input{figures/incomplete-rows.pdf_tex}
+\input{figures/svg/incomplete-rows.pdf_tex}
 \hfill
 \def\svgwidth{0.3\columnwidth}
-\input{figures/overlap.pdf_tex}
+\input{figures/svg/overlap.pdf_tex}
 \caption{Configurations that increase the number of seeks per
@@
 
-\section{Benchmark}
-\label{sec:benchmark}
-
-\subsubsection{Implementation}
+\subsection{Implementation}
 
@@
 
-\subsubsection{Data}
+\section{Benchmark}
+\label{sec:benchmark}
+
+\subsection{Data}
 To benchmark our library we use the 76~GB BigBrain image split in 125
@@
 
-\subsubsection{Infrastructure}
+\subsection{Infrastructure}
 
@@
 \subsection{Multiple reads vs blocks and slices}
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=0.45\columnwidth]{figures/benchmark-mreads/mreads-comparision-ssd.pdf}
+  \hfill
+  \includegraphics[width=0.45\columnwidth]{figures/benchmark-mreads/mreads-comparision-hdd.pdf}\\
+  \includegraphics[width=0.45\columnwidth]{figures/benchmark-mreads/mreads-breakdown-ssd.pdf}
+  \hfill
+    \includegraphics[width=0.45\columnwidth]{figures/benchmark-mreads/mreads-breakdown-hdd.pdf}
+  \caption{Comparison between Multiple reads, blocks and slices. Left column: SSD. Right column: HDD. Top row: overall merging time. Bottom row: breakdown by read, write and seek time.}
+  % Gnuplot file: scripts/blocks-slices/createboxplot.gnplt
+\label{fig:multiple-reads}
+\end{figure}
 

### 2017-07-27 paper.tex
Integrated multiple writes in paper

@@
 
+\subsubsection{Multiple writes}
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=0.45\columnwidth]{figures/benchmark-mwrites/mwrites-comparision-ssd.pdf}
+  \hfill
+  \includegraphics[width=0.45\columnwidth]{figures/benchmark-mwrites/mwrites-comparision-hdd.pdf}\\
+  \includegraphics[width=0.45\columnwidth]{figures/benchmark-mwrites/mwrites-breakdown-ssd.pdf}
+  \hfill
+    \includegraphics[width=0.45\columnwidth]{figures/benchmark-mwrites/mwrites-breakdown-hdd.pdf}
+  \caption{Comparison between Multiple writes, blocks and slices. Left column: SSD. Right column: HDD. Top row: overall merging time. Bottom row: breakdown by read, write and seek time.}
+  % Gnuplot file: scripts/blocks-slices/createboxplot.gnplt
+\label{fig:multiple-writes}
+\end{figure}
+
 \newpage

### 2017-07-27 paper.tex
Minor edits

@@
 
+\author{Final ordering
+  to be discussed when the paper: Yongping/Val\'erie,
+  Yuhong Yan, Tristan.\\
+  Department of Computer Science and Software Engineering, Concordia University, Montreal, Quebec, Canada
+}
+
 \maketitle
@@
   \todo{The abstract goes here. DO NOT USE SPECIAL CHARACTERS,
-    SYMBOLS, OR MATH IN YOUR TITLE OR ABSTRACT.}  Authors (final order
-  to be discussed when the paper shapes up): Yongping/Val\'erie,
-  Yuhong Yan, Tristan. Perhaps Pierre Bellec. \note{This is a note.}
+    SYMBOLS, OR MATH IN YOUR TITLE OR ABSTRACT.} 
 \end{abstract}
@@
 \end{equation*}
+\todo{add one more preliminary step to explain how these equations were derived. Then jump to the final ones.}
 Using these notations, the number of seeks performed by multiple reads is:
@@
 
-\subsection{Buffered slices vs slices}
+\subsection{Buffered slices}
 
-\subsection{Cluster reads vs blocks and slices}
+\subsection{Cluster reads}
 
-\subsection{Multiple reads vs blocks and slices}
+\subsection{Multiple reads}
 

### 2017-07-27 generate_figures.sh
add error bars and fix number_of_seeks.gnuplot

@@ -6,10 +6,15 @@ set -u
 figures/svg/export_svg_files.sh
 scripts/blocks-slices/createboxplot.gnplt
 scripts/model/model.gnplt
-scripts/mwrites/compare_total_time.gnuplot
-scripts/mwrites/comparasion_break_down.gnuplot
-scripts/mwrites/generate_avg_data.py
+scripts/mreads/number_of_seeks.gnuplot
+scripts/mreads/generate_avg_var_data.py
 scripts/mreads/compare_total_time.gnuplot
-scripts/mreads/comparasion_break_down.gnuplot
-scripts/mreads/generate_avg_data.py
-#./scripts/mreads/number_of_seeks.gnuplot # crashes for now
+scripts/mreads/compare_breakdown_errbar.gnuplot
+scripts/mwrites/generate_avg_data.py
+scripts/mwrites/compare_total_time.gnuplot
+scripts/mwrites/compare_breakdown_errbar.gnuplot
+
+
+
+
+

### 2017-07-27 scripts/mwrites/generate_avg_data.py

@@ -1,156 +1,0 @@
-#!/usr/bin/env python
-
-input_file_mwrites_ssd = "./data/mwrites/mwrites_ssd.dat"
-output_file_mwrites_ssd = "./data/mwrites/mwrites_ssd_avg.dat"
-
-input_file_mwrites_hdd = "./data/mwrites/mwrites_hdd.dat"
-output_file_mwrites_hdd = "./data/mwrites/mwrites_hdd_avg.dat"
-
-input_file_naive_ssd = "./data/blocks-slices/totaltime.dat"
-output_file_naive_ssd = "./data/blocks-slices/blocks_slices_avg_ssd.dat"
-
-input_file_naive_hdd = "./data/blocks-slices/totaltime.dat"
-output_file_naive_hdd = "./data/blocks-slices/blocks_slices_avg_hdd.dat"
-
-def avg(l):
-    return sum(l) / float(len(l))
-
-def get_avg_mwrites(input_file, output_file):
-    read_times_3g = []
-    write_times_3g = []
-    seek_times_3g = []
-    calculation_times_3g = []
-
-    read_times_6g = []
-    write_times_6g = []
-    seek_times_6g = []
-    calculation_times_6g = []
-
-    read_times_9g = []
-    write_times_9g = []
-    seek_times_9g = []
-    calculation_times_9g = []
-
-    read_times_12g = []
-    write_times_12g = []
-    seek_times_12g = []
-    calculation_times_12g = []
-
-    with open(input_file, 'r') as f:
-        for line in f:
-            if '#' not in line:
-                items = line.split(' ')
-
-                read_time = float(items[0])
-                write_time = float(items[1])
-                seek_time = float(items[2])
-                total_time = float(items[4])
-                calculation_time = total_time - seek_time - write_time - read_time
-
-                read_times_3g.append(read_time)
-                write_times_3g.append(write_time)
-                seek_times_3g.append(seek_time)
-                calculation_times_3g.append(calculation_time)
-
-                read_time = float(items[5])
-                write_time = float(items[6])
-                seek_time = float(items[7])
-                total_time = float(items[9])
-                calculation_time = total_time - seek_time - write_time - read_time
-
-                read_times_6g.append(read_time)
-                write_times_6g.append(write_time)
-                seek_times_6g.append(seek_time)
-                calculation_times_6g.append(calculation_time)
-
-                read_time = float(items[10])
-                write_time = float(items[11])
-                seek_time = float(items[12])
[…]

### 2017-07-28 paper.tex
Updates following discussion with Yongping

@@
 algorithms in a subsequent paper based on the results of the present
-one.
+one. \todo{
+There is BigBrain-1 rescanned at 1 micron in the plane. So resolution
+is 1x1x20, 7404 slices. If raw data are about 1Tb, then multiply by
+100 for an estimation of the size of the new image. }
 
@@
 
+\subsubsection{Cluster reads}
+
+\subsubsection{Multiple reads}
+
+\todo{Yongping: among other implementation techniques (e.g. dictionary
+  vs list, etc), explain clearly and concisely what the calculation
+  time covers.}
+
 \section{Benchmark}
@@
 \end{figure}
+
+Results discussion:
+\begin{itemize}
+\item Comment on calculation time.
+\item Write time is higher than read time because use nibabel for writing but not reading.
+\item It seems that the merge time is quite stable w.r.t the amount of available memory. This is good news for multiple reads. 
+\end{itemize}
 
@@
 
+% Ideas for the future: measure the actual amount of consumed memory.
+
 \section*{Acknowledgment}

### 2017-07-29 paper.tex
Wrote some discussion. Fixed references

@@
 \author{Final ordering
-  to be discussed when the paper: Yongping/Val\'erie,
+  to be discussed: Yongping/Val\'erie,
   Yuhong Yan, Tristan.\\
@@
 \begin{abstract}
-  \todo{The abstract goes here. DO NOT USE SPECIAL CHARACTERS,
-    SYMBOLS, OR MATH IN YOUR TITLE OR ABSTRACT.} 
+  Splitting and merging data is a requirement for most parallel
+    or distributed processing operations, however, naive algorithms
+    perform very poorly on ultra-high resolution 3D images due to
+    important seek times. To address this issue, we created two
+    algorithms based on in-memory reconstruction of image segments and
+    analyzed their complexity in terms of number of seeks. Evaluation
+    on a 3000x3000x3000 image of the brain shows that both our
+    algorithms can split or merge blocks with a performance similar to
+    slices, which solves our initial problem. We conclude that
+    splitting and merging large 3D images can be done efficiently
+    based on simple data formats.
 \end{abstract}
@@
 examples found in medical imaging, our primary domain of interest,
-include \todo{electromicroscopy,spectroscopy-francoise-peyrin}. As
-such images would typically be processed on a computing cluster,
-possibly using locality-aware file systems such as HDFS, software
-libraries are needed to split and merge such images efficiently. We
-introduce and compare a family of algorithms to address this problem
-in a sequential environment made of a single storage disk. We are
-aware of the need for parallel algorithms that would split and merge
-images to and from an array of disks but we plan to describe such
-algorithms in a subsequent paper based on the results of the present
-one. \todo{
-There is BigBrain-1 rescanned at 1 micron in the plane. So resolution
-is 1x1x20, 7404 slices. If raw data are about 1Tb, then multiply by
-100 for an estimation of the size of the new image. }
+include high-resolution 3D electron microscopy (see, e.g.,
+~\cite{bock2011network}) or micro and
+nano-tomography~\cite{10.1371/journal.pone.0035691}. As such images
+would typically be processed on a computing cluster, possibly using
+locality-aware file systems such as HDFS, software libraries are
+needed to split and merge such images efficiently. We introduce and
+compare a family of algorithms to address this problem in a sequential
+environment made of a single storage disk. We are aware of the need
+for parallel algorithms that would split and merge images to and from
+an array of disks but we plan to describe such algorithms in a
+subsequent paper based on the results of the present one. \todo{ There
+  is BigBrain-1 rescanned at 1 micron in the plane. So resolution is
+  1x1x20, 7404 slices. If raw data are about 1Tb, then multiply by 100
+  for an estimation of the size of the new image. }
 
@@
 
-The literature looks remarkably scarce on the problem described
+The literature is remarkably scarce on the problem described
 above. Parallel processing of 3D images has obviously been extensively
@@
     that do not fit in memory.
-  \item We model the performance (execution time) of these algorithms,
+  \item We model the performance (number of seeks) of these algorithms,
     based on a simple but realistic disk I/O model.
@@
     drive.
-  \item We compare the algorithms in a variety of conditions based on the validated model.
 \end{itemize}
@@
 \subsection{Implementation}
+\label{sec:implementation}
 
@@
 Both drives used 512-byte logical sectors, 4096-byte physical sectors,
-SATA >3.1 (6.0 Gb/s) and were accessed through the XFS file system.
+SATA >3.1 (6.0 Gb/s) and were accessed through the XFS file system
+v4.5.0.
 %smartctl --xall /dev/sda
+
+\subsection{Execution conditions}
+
+We used release xxx of our library.
+
+\todo{Yongping: mention the cache cleaning thing that you run before every experiment.}
+
+Experiment scripts are available at xxx.
 
@@
 
-What is the best algorithm (between 3 and 4), in which context. Did we
-solve the seek problem or is there any issue remaining?
+% Problem solved!
+Cluster reads and multiple reads are able to reduce to a negligible
+amount the overall seek time required to split or merge 3D blocks in a
+high-resolution image where data is stored linearly. Both algorithms
+performed equivalently on our benchmark and compared to the reference
+configuration where slices are merged or split without seeking. Our
+initial problem is solved.
 
-File formats:
-\begin{itemize}
-\item would MINC~\cite{vincent2016minc} support such algorithms? See on-going email discussion
-with P. Bellec.
-\item some formats optimize the storage for particular split shapes. See
-  ndstore~\cite{burns2013open}. Such formats would behave equally
-  poorly with chunks that do not comply to this geometry. Our
-  algorithms would work equally well with such formats.
-\end{itemize}
+% Parallel algorithms
+In a real system, high-resolution images are likely to be processed on
+computing clusters, for instance using software from the Hadoop
+project, in particular the Hadoop Distributed File
+System~\cite{shvachko2010hadoop}. In this context parallel split-and-merge algorithms
+will be required, since the various splits of a large image would be
+uploaded to different disks concurrently. In the same vein,
+``re-spliting'' algorithms would have to be devised in case a split
+image need to be split in a different geometry. Cluster reads and
+multiple reads could be used as starting points for such algorithms.
 
-The performance model might be dependent on the particular
-infrastructure (disks, OS and filesystem) used in the experiments. In
-particular, caching might behave differently in a different setup.
+% Several other factors might influence performance
+I/O optimization is a holistic problem that is in practice highly
+dependent on the hardware used, firmware, operating system, file
+system and programming language. Caching occurs at various levels and
+might always influence performance, potentially differently depending
+on the split or merge algorithm used. In some disks, seek time greatly
+varies with the seek distance, which would open the door to additional
+opportunities for I/O optimization. Interactions between those
+components might also result in performance differences particular to
+a particular algorithm. To ensure the portability of our library
+across systems and configurations, we focused on reducing the overall
+number of seeks and ignored specific system configurations. We
+demonstrated the performance of our methods on both an HDD and an SSD
+disk, using state-of-the-art and widely used versions of Linux
+(CentOS7) and file systems (XFS v4.5.0).
 
-Reconstructing (therefore splitting) a complete 3D image may not
-always be necessary. Re-splitting, for instance, could be done from
-existing chunks.
-
-\section{Conclusion}
-
-Future work:
-- obliques
-- parallel algorithms
-- re-splitting
-- model
+% File formats
+In terms of file formats, our results demonstrate that simple imaging
+formats might be used for the split-and-merge problem without
+performance loss compared to more complex formats that try to preserve
+spatial locality on disk, for instance MINC 2.0~\cite{vincent2016minc}
+or the format based on space-filling curves mentioned
+in~\cite{burns2013open}. This is of major interest in the current
+open-science context since simpler formats favor data-sharing and
+interoperability.  Moreover, our algorithms could potentially be
+adapted to any split geometry, even though we demonstrated them on
+slices and blocks only, while file formats inevitably assume a
+particular geometry. For instance, the format in~\cite{burns2013open}
+wouldn't perform nicely with a naive splitting algorithm if slices had
+to be extracted instead of blocks.  Two points need to be made to be
+fair though. First, we aimed at extracting all the blocks while the
+problem targeted by the format in~\cite{burns2013open} was to extract
+a single block from a large image. Second, we have not considered
+on-the-fly data compression yet, which is a unique feature of MINC
+2.0.
 

### 2017-07-29 generate_figures.sh
remove all average files and pdfs of figures generated by scripts

@@ -4,17 +4,10 @@ set -e
 set -u
 
 figures/svg/export_svg_files.sh
-scripts/blocks-slices/createboxplot.gnplt
 scripts/model/model.gnplt
 scripts/mreads/number_of_seeks.gnuplot
-scripts/mreads/generate_avg_var_data.py
-scripts/mreads/compare_total_time.gnuplot
 scripts/mreads/compare_breakdown_errbar.gnuplot
-scripts/mwrites/generate_avg_var_data.py
-scripts/mwrites/compare_total_time.gnuplot
 scripts/mwrites/compare_breakdown_errbar.gnuplot
-
-
-
-
+#crash
+#scripts/blocks-slices/createboxplot.gnplt
 

### 2017-07-29 paper.tex
Minor updates following Yongping's edits

@@
   \centering
-  \includegraphics[width=0.45\columnwidth]{figures/benchmark-mreads/mreads-comparision-ssd.pdf}
-  \hfill
-  \includegraphics[width=0.45\columnwidth]{figures/benchmark-mreads/mreads-comparision-hdd.pdf}\\
   \includegraphics[width=0.45\columnwidth]{figures/benchmark-mreads/mreads-breakdown-ssd.pdf}
@@
     \includegraphics[width=0.45\columnwidth]{figures/benchmark-mreads/mreads-breakdown-hdd.pdf}
-  \caption{Comparison between Multiple reads, blocks and slices. Left column: SSD. Right column: HDD. Top row: overall merging time. Bottom row: breakdown by read, write and seek time.}
+  \caption{Comparison between Multiple reads, blocks and slices. Left: SSD. Right: HDD.}
   % Gnuplot file: scripts/blocks-slices/createboxplot.gnplt
@@
   \centering
-  \includegraphics[width=0.45\columnwidth]{figures/benchmark-mwrites/mwrites-comparision-ssd.pdf}
-  \hfill
-  \includegraphics[width=0.45\columnwidth]{figures/benchmark-mwrites/mwrites-comparision-hdd.pdf}\\
   \includegraphics[width=0.45\columnwidth]{figures/benchmark-mwrites/mwrites-breakdown-ssd.pdf}
@@
     \includegraphics[width=0.45\columnwidth]{figures/benchmark-mwrites/mwrites-breakdown-hdd.pdf}
-  \caption{Comparison between Multiple writes, blocks and slices. Left column: SSD. Right column: HDD. Top row: overall merging time. Bottom row: breakdown by read, write and seek time.}
+  \caption{Comparison between Multiple writes, blocks and slices. Left: SSD. Right: HDD.}
   % Gnuplot file: scripts/blocks-slices/createboxplot.gnplt

### 2017-07-29 paper.tex
Moved simulation section

@@
 
+\begin{itemize}
+  \item In MR, no seeking withing the splits.
+\end{itemize}
+
 \subsubsection{Cluster reads}
@@
   time covers.}
+
+\section{Simulations}
+
+\todo{Make more simulations to support that claim.}
+After running a few simulations, we saw that the comparison between
+Cluster reads and Multiple reads essentially depends on the three
+Cluster reads configurations.  When Cluster reads merge incomplete
+block rows (case 1), they are always outperformed by Multiple reads.
+When Cluster reads merge complete block rows (case 2), the comparison
+between Cluster reads and Multiple reads depends on the values of $D$
+and $n$. Figure~\ref{fig:model-comparison} shows two situations where
+Cluster reads and Multiple reads are alternatively better than each
+other. When Cluster reads merge complete block slices (case 3),
+Cluster reads and Multiple reads are equivalent.
+
+\begin{figure}
+  \includegraphics[width=0.45\columnwidth]{figures/model-big-brain.pdf}
+  \includegraphics[width=0.45\columnwidth]{figures/model-big-brain-rescan.pdf}
+  \caption{Number of seeks for Cluster reads vs Multiple reads when
+    Cluster reads merge complete block rows (case 2), for D=3000 and
+    b=2. Left: n=125; Right: n=64,000.}
+  % Gnuplot file: scripts/model/model.gnplt
+  \label{fig:model-comparison}
+\end{figure}
 
@@
 \subsection{Cluster reads vs Multiple reads}
-
-\subsubsection{Simulations}
-After running a few simulations, we saw that the comparison between
-Cluster reads and Multiple reads essentially depends on the three
-Cluster reads configurations.  When Cluster reads merge incomplete
-block rows (case 1), they are always outperformed by Multiple reads.
-When Cluster reads merge complete block rows (case 2), the comparison
-between Cluster reads and Multiple reads depends on the values of $D$
-and $n$. Figure~\ref{fig:model-comparison} shows two situations where
-Cluster reads and Multiple reads are alternatively better than each
-other. When Cluster reads merge complete block slices (case 3),
-Cluster reads and Multiple reads are equivalent.
-
-\begin{figure}
-  \includegraphics[width=0.45\columnwidth]{figures/model-big-brain.pdf}
-  \includegraphics[width=0.45\columnwidth]{figures/model-big-brain-rescan.pdf}
-  \caption{Number of seeks for Cluster reads vs Multiple reads when
-    Cluster reads merge complete block rows (case 2), for D=3000 and
-    b=2. Left: n=125; Right: n=64,000.}
-  % Gnuplot file: scripts/model/model.gnplt
-  \label{fig:model-comparison}
-\end{figure}
-
-\subsubsection{Experiments}
 

### 2017-07-30 figures/svg/export_svg_files.sh
add implementation of mreads in paper

@@ -1,4 +1,4 @@
-#!/bin/bash
+#!/usr/bin/env bash
 
 for name in buffer case1-a case2-a case3 incomplete-rows overlap
 do

### 2017-07-30 paper.tex

@@
   \begin{algorithmic}[1]
-  \STATE sorted\_split\_name\_list = sort blocks by increasing (z,y,x) values
+  \STATE sorted\_split\_list = sort splits based on its postion in increasing (z,y,x) values
   \STATE start\_index = 0, end\_index=(m-1)
+  \STATE write\_range = (start\_index, end\_index)
   \WHILE{end\_index \textless sizeof(reconstructed\_image)}
     \STATE initialize data\_buffer
-    \FOR{split\_name in sorted\_split\_name\_list}
-      \IF{indexes of split\_name is in next\_write\_range}
+    \FOR{split in sorted\_split\_list}
+      \IF{split is in write\_range}
         \STATE read split
-        \STATE extract all indexes in next write range
-        \STATE write the corresponding data to the data\_buffer
+        \STATE extract the rows in the write\_range
+        \STATE write rows to the data\_buffer
       \ENDIF
     \ENDFOR
-    \STATE write data\_buffer in reconstructed\_image
+    \STATE write data\_buffer continuously to reconstructed\_image
     \STATE clear data\_buffer
@@
 
+For faster reading all the splits, before the merging process start,
+we read all the split’s headers to get their position information and
+calculate each row’s index in the reconstructed image. We only read the
+split when there is a row in that writing range (available memory),
+otherwise, we skip it.
+
+Initially, we use a numpy array as our data buffer, after the split is read,
+we assign the rows (a numpy array of unsigned integers) into this
+buffer to the right position. However, after some tests, we found that
+there is some overhead. Each time we assign the data to the corresponding
+position of the buffer, the system will copy these integers first.
+This is neither time-efficient nor space-efficient.
+
+We then think of a better solution which we use the dictionary in python
+as our data buffer. We use the row’s index as a key, the data (in bytes)
+as the value. As for the multiple reads, the data buffer only contains
+contiguous bytes, we can just seek once (the first key of the sorted keys
+in dictionary), then write all the bytes continuously to the file.
+There is no copy in the memory, which proves a bit faster than using
+a numpy array as the buffer.
+
 \todo{Yongping: among other implementation techniques (e.g. dictionary
@@
 Experiment scripts are available at xxx.
+As our experiments results are heavily rely on IO performance of the disk,
+we drop the system cache to get more precise result before each experiment.
 

### 2017-07-30 paper.tex
Updates

--- 2017-07-29 paper.tex
@@
 \author{Final ordering
-  to be discussed: Yongping/Val\'erie,
-  Yuhong Yan, Tristan.\\
-  Department of Computer Science and Software Engineering, Concordia University, Montreal, Quebec, Canada
+  to be discussed: Yongping Gao/Val\'erie Hayot-Sasson,
+  Yuhong Yan, Tristan Glatard\\
+  Department of Computer Science and Software Engineering\\ Concordia University, Montreal, Quebec, Canada\\
+  {first.last}@concordia.ca
 }
@@
 %\href{http://dl.acm.org/citation.cfm?id=218395}
-
-\newpage
 
@@
 \item In MR, no seeking withing the splits.
+\item Avoid in-memory copies.
 \end{itemize}

### 2017-07-31 paper.tex
Fixed (minor) error in multiple reads formulas. Explained them better.

@@
   \begin{algorithmic}[1]
-    \STATE list = sort blocks by increasing (z,y,x) values
+    \STATE sorted\_splits = sort splits by increasing (z,y,x) values
     \STATE initialize buffer
@@
 
-In multiple reads, blocks are read partially to ensure that the memory
-buffer only contains contiguous bytes, at any time. That is, a given
-block will be read multiple times.     \todo{improve notations and algo in algorithm}
-
+Multiple reads are shown in Algorithm~\ref{algo:multiple-reads}.  The
+main idea of this algorithm is that blocks are read partially (line 9)
+to ensure that the memory buffer only contains contiguous bytes, at
+any time. Therefore, the buffer can be written continuously to the
+reconstructed image, without seeking (line 13).  It should be noted
+that a given block might be read multiple times, in different memory
+loads. 
 \begin{algorithm}[h]
@@
   \begin{algorithmic}[1]
-  \STATE sorted\_split\_list = sort splits based on its postion in increasing (z,y,x) values
-  \STATE start\_index = 0, end\_index=(m-1)
+  \STATE sorted\_splits = sort splits by increasing (z,y,x) values
+  \STATE start\_index = 0 ; end\_index=(m-1)
   \STATE write\_range = (start\_index, end\_index)
-  \WHILE{end\_index \textless sizeof(reconstructed\_image)}
-    \STATE initialize data\_buffer
-    \FOR{split in sorted\_split\_list}
-      \IF{split is in write\_range}
-        \STATE read split
-        \STATE extract the rows in the write\_range
-        \STATE write rows to the data\_buffer
+  \WHILE{end\_index \textless Rb}
+    \STATE initialize buffer
+    \FOR{split in sorted\_splits}
+      \IF{split has voxels in write\_range}
+        \STATE split\_data = read split
+        \STATE in split\_data, extract the rows in write\_range
+        \STATE write rows to buffer
       \ENDIF
     \ENDFOR
-    \STATE write data\_buffer continuously to reconstructed\_image
-    \STATE clear data\_buffer
+    \STATE write buffer to reconstructed\_image
     \STATE start\_index = end\_index + 1
-    \STATE end\_index += mem
+    \STATE end\_index += m
   \ENDWHILE
@@
 \end{equation*}
-\todo{add one more preliminary step to explain how these equations were derived. Then jump to the final ones.}
-Using these notations, the number of seeks performed by multiple reads is:
+
+The total number of seeks performed by multiple reads in case $i$ is:
+\begin{equation*}
+  N^i_{\mathrm{MR}} = x_i + (x_i-1)b_i +b_i', \quad i \in \llbracket 1,5 \rrbracket
+\end{equation*}
+where $x_i$ is the total number of required memory loads, $b_i$ is the
+number of blocks accessed by the first ($x_i$-1) memory loads and
+$b_i'$ is the number of blocks access by the last memory load. We have:
+\begin{equation*}
+  x_i = \ceil*{\frac{Rb}{m_i'}} \quad ; \quad x_i -1 = \floor*{\frac{Rb}{m_i'}}, \quad i \in \llbracket 1,5 \rrbracket
+\end{equation*}
+and:
+\begin{equation*}
+b_1=k_1 ; \quad b_2=\sqrt[3]{n} ; \quad b_3 = k_3\sqrt[3]{n} ; \quad b_4=\sqrt[3]{n}^2 ; \quad b_5=k_5\sqrt[3]{n}^2
+\end{equation*}
+and:
+\begin{eqnarray*}
+b_1'=\sqrt[3]{n}D^2\text{ mod }k_1; &b_2'=b_2;   & b_3' = \sqrt[3]{n}\left( \sqrt[3]{n}D\text{ mod }k_3\right)\\
+                                   &b_4'=b_4;  & b_5'=\sqrt[3]{n}^2\left( \sqrt[3]{n}\text{ mod }k_5\right) 
+\end{eqnarray*}
+It gives the following expression for $N_\mathrm{MR}$:
 \begin{equation*}
@@
   \floor*{\frac{Rb}{m_1'}}(k_1+1)+1
-  + n \text{ mod } k_1& \text{if } d \leq \frac{m}{b} < D\\
+  + \sqrt[3]{n}D^2 \text{ mod } k_1& \text{if } d \leq \frac{m}{b} < D\\
   \ceil*{\frac{Rb}{m_2'}}(\sqrt[3]{n}+1)   & \text{if } D \leq \frac{m}{b} < Dd\\
   \floor*{\frac{Rb}{m_3'}}(k_3\sqrt[3]{n}+1)+1+\\
-  \quad \quad \sqrt[3]{n}\left(\frac{n}{\sqrt[3]{n}} \text{ mod } k_3\right)&\
+  \quad \quad \sqrt[3]{n}\left(\sqrt[3]{n}D\text{ mod } k_3\right)&\
   \text{if } Dd \leq \frac{m}{b} < D^2\\
@@
   \floor*{\frac{Rb}{m_5'}}(k_5\sqrt[3]{n}^2+1)+1+&\\
-  \quad \quad \sqrt[3]{n}^2\left(\frac{n}{\sqrt[3]{n}^2} \text{ mod } k_5\right)&\
+  \quad \quad \sqrt[3]{n}^2\left(\sqrt[3]{n} \text{ mod } k_5\right)&\
   \text{if } D^2d \leq \frac{m}{b} < R
@@
 \end{equation*}
-
-which gives, using $R$ and $n$ as main variables:
+And finally, using $R$ and $n$ as main variables:
 \begin{equation}
@@
   \floor*{\frac{Rb}{m_1'}}\left(\frac{m_1'\sqrt[3]{n}}{\sqrt[3]{R}b}+1\right)+1+\\
-  \quad \quad \quad n \text{ mod }\floor*{\frac{m\sqrt[3]{n}}{\sqrt[3]{R}b}}\
+  \quad \quad \quad \sqrt[3]{n}\sqrt[3]{R}^2 \text{ mod }\floor*{\frac{m\sqrt[3]{n}}{\sqrt[3]{R}b}}\
   & \text{if } \sqrt[3]{\frac{R}{n}} \leq \frac{m}{b} < \sqrt[3]{R}\\
@@
   \floor*{\frac{Rb}{m_3'}}\left(\frac{m'_3\sqrt[3]{n}^2}{\sqrt[3]{R}^2b}+1\right)+1+\\
-  \sqrt[3]{n}\left( \frac{n}{\sqrt[3]{n}} \text{ mod } \floor*{\frac{m\sqrt[3]{n}}{\sqrt[3]{R}^2b}} \right)
+  \sqrt[3]{n}\left( \sqrt[3]{nR} \text{ mod } \floor*{\frac{m\sqrt[3]{n}}{\sqrt[3]{R}^2b}} \right)
   & \text{if } \frac{\sqrt[3]{R}^2}{\sqrt[3]{n}} \leq \frac{m}{b} < \sqrt[3]{R}^2\\
@@
   \ceil*{\frac{Rb}{m_5'}}\left(\frac{m_5'n}{Rb}+1\right)+1+\\
-  \sqrt[3]{n}^2\left( \frac{n}{\sqrt[3]{n}^2} \text{ mod } \floor*{\frac{m\sqrt[3]{n}}{Rb}}\right)
+  \sqrt[3]{n}^2\left( \sqrt[3]{n} \text{ mod } \floor*{\frac{m\sqrt[3]{n}}{Rb}}\right)
   & \text{if } \frac{R}{\sqrt[3]{n}} \leq \frac{m}{b} < R

### 2017-07-31 paper.tex
Updates in simulations, still work in progress

@@
 
-\section{Simulations}
-
-\todo{Make more simulations to support that claim.}
-After running a few simulations, we saw that the comparison between
-Cluster reads and Multiple reads essentially depends on the three
-Cluster reads configurations.  When Cluster reads merge incomplete
-block rows (case 1), they are always outperformed by Multiple reads.
-When Cluster reads merge complete block rows (case 2), the comparison
-between Cluster reads and Multiple reads depends on the values of $D$
-and $n$. Figure~\ref{fig:model-comparison} shows two situations where
-Cluster reads and Multiple reads are alternatively better than each
-other. When Cluster reads merge complete block slices (case 3),
-Cluster reads and Multiple reads are equivalent.
-
-\begin{figure}
-  \includegraphics[width=0.45\columnwidth]{figures/model-big-brain.pdf}
-  \includegraphics[width=0.45\columnwidth]{figures/model-big-brain-rescan.pdf}
-  \caption{Number of seeks for Cluster reads vs Multiple reads when
-    Cluster reads merge complete block rows (case 2), for D=3000 and
-    b=2. Left: n=125; Right: n=64,000.}
-  % Gnuplot file: scripts/model/model.gnplt
-  \label{fig:model-comparison}
-\end{figure}
-
 \section{Benchmark}
@@
 
+\section{Simulations}
+
+\todo{Make more simulations to support that claim.}
+
+After running a few simulations, we saw that the comparison between
+Cluster reads and Multiple reads essentially depends on the three
+Cluster reads configurations.  When Cluster reads merge incomplete
+block rows (case 1), they are always outperformed by Multiple reads.
+When Cluster reads merge complete block rows (case 2), the comparison
+between Cluster reads and Multiple reads depends on the values of $D$
+and $n$. Figure~\ref{fig:model-comparison} shows two situations where
+Cluster reads and Multiple reads are alternatively better than each
+other. When Cluster reads merge complete block slices (case 3),
+Cluster reads and Multiple reads are equivalent.
+
+\begin{figure}
+  \includegraphics[width=0.45\columnwidth]{figures/model-big-brain-creads.pdf}
+  \includegraphics[width=0.45\columnwidth]{figures/model-big-brain-mreads.pdf}
+  \caption{Number of seeks for Cluster reads (Left) and Multiple reads (Right), for
+    D=3000, b=2 and n=125.}
+  % Gnuplot file: scripts/model/model.gnplt
+  \label{fig:model-individual}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=0.45\columnwidth]{figures/model-big-brain-comparison.pdf}
+  \includegraphics[width=0.45\columnwidth]{figures/model-big-brain-comparison-moreblocks.pdf}
+  \caption{Number of seeks for Cluster reads vs Multiple reads when
+    Cluster reads merge complete block rows (case 2), for D=3000 and
+    b=2. Left: n=125; Right: n=64,000.}
+  % Gnuplot file: scripts/model/model.gnplt
+  \label{fig:model-comparison}
+\end{figure}
+
+\section{Experiments}
+\label{sec:results}
+
 \subsection{Execution conditions}
@@
 we drop the system cache to get more precise result before each experiment.
-
-\section{Results}
-\label{sec:results}
 

### 2017-07-31 figures/svg/export_svg_files.sh
add mreads-case figures

@@ -1,6 +1,7 @@
 #!/usr/bin/env bash
 
-for name in buffer case1-a case2-a case3 incomplete-rows overlap
+for name in buffer case1-a case2-a case3 incomplete-rows overlap mreads-case1 \
+        mreads-case2 mreads-case3 mreads-case4 mreads-case5
 do
   #echo "Exporting ${name}.svg"
   inkscape -D -z --file=figures/svg/${name}.svg --export-pdf=figures/svg/${name}.pdf --export-latex

### 2017-07-31 paper.tex

--- 2017-07-31 paper.tex
@@
 loads.
+
+\begin{figure}
+\centering
+\def\svgwidth{0.3\columnwidth}
+\input{figures/svg/mreads-case1.pdf_tex}
+\def\svgwidth{0.3\columnwidth}
+\input{figures/svg/mreads-case2.pdf_tex}
+\def\svgwidth{0.3\columnwidth}
+\input{figures/svg/mreads-case3.pdf_tex}
+
+\medskip
+
+\def\svgwidth{0.3\columnwidth}
+\input{figures/svg/mreads-case4.pdf_tex}
+\def\svgwidth{0.3\columnwidth}
+\input{figures/svg/mreads-case5.pdf_tex}
+\caption{Memory-load configurations in multiple reads, leading to
+  different number of seeks. Red regions shows different memory
+  load in each case. Small dots are used to represent block frontiers,
+  long dashes used to represent rows/slices in blocks.}
+\label{fig:multiple-reads-cases}
+\end{figure}
+
 \begin{algorithm}[h]
@@
 $k<d$) or of block slices (case 5, $k<\sqrt[3]{n}$) \todo{make a
-  figure}. In each of these 5 cases, we define $v$ as follows:
+  figure}. In each of these 5 cases in Figure~\ref{fig:multiple-reads-cases}, we define $v$ as follows:
 \begin{equation*}

### 2017-08-01 paper.tex
Updates in mreads model

--- 2017-07-31 paper.tex
@@
 loads.
+
+\begin{figure}
+\centering
+\def\svgwidth{0.3\columnwidth}
+\input{figures/svg/mreads-case1.pdf_tex}
+\def\svgwidth{0.3\columnwidth}
+\input{figures/svg/mreads-case2.pdf_tex}
+\def\svgwidth{0.3\columnwidth}
+\input{figures/svg/mreads-case3.pdf_tex}
+
+\medskip
+
+\def\svgwidth{0.3\columnwidth}
+\input{figures/svg/mreads-case4.pdf_tex}
+\def\svgwidth{0.3\columnwidth}
+\input{figures/svg/mreads-case5.pdf_tex}
+\caption{Memory-load configurations in multiple reads (d=4, D=16,
+  n=64, $k_1=k_2=k_3=k_4=k_5=1$). Red colour shows the content of the
+  memory load. Small dots depict block frontiers and long dashes
+  depict rows and slices within blocks.}
+\label{fig:multiple-reads-cases}
+\end{figure}
+
 \begin{algorithm}[h]
@@
 
-For simplicity, we assume that $m'$ represents an integer number $k$
+As illustrated in Figure~\ref{fig:multiple-reads-cases}, we assume that $m'$ represents an integer number $k$
 of block rows (case 1, $k<\sqrt[3]{n}$), of complete rows (case 2,
 $k<d$), of tile rows (case 3, $k<\sqrt[3]{n}$), of slices (case 4,
-$k<d$) or of block slices (case 5, $k<\sqrt[3]{n}$) \todo{make a
-  figure}. In each of these 5 cases, we define $v$ as follows:
+$k<d$) or of block slices (case 5, $k<\sqrt[3]{n}$). In each of these 5 cases, we define $v$ as follows:
 \begin{equation*}
@@
 The total number of seeks performed by multiple reads in case $i$ is:
-\begin{equation*}
-  N^i_{\mathrm{MR}} = x_i + (x_i-1)b_i +b_i', \quad i \in \llbracket 1,5 \rrbracket
-\end{equation*}
+\begin{eqnarray*}
+  N^i_{\mathrm{MR}} &=& x_i + (x_i-1)b_i +b_i', \quad i \in \llbracket 1,5 \rrbracket\\
+  &=& x_i \left(1+b_i\right)-b_i+b_i'
+\end{eqnarray*}
 where $x_i$ is the total number of required memory loads, $b_i$ is the
@@
 \begin{equation*}
-  x_i = \ceil*{\frac{Rb}{m_i'}} \quad ; \quad x_i -1 = \floor*{\frac{Rb}{m_i'}}, \quad i \in \llbracket 1,5 \rrbracket
+  x_i = \ceil*{\frac{Rb}{m_i'}}, \quad i \in \llbracket 1,5 \rrbracket
 \end{equation*}
@@
 \begin{cases}
-  \floor*{\frac{Rb}{m_1'}}(k_1+1)+1
-  + \sqrt[3]{n}D^2 \text{ mod } k_1& \text{if } d \leq \frac{m}{b} < D\\
-  \ceil*{\frac{Rb}{m_2'}}(\sqrt[3]{n}+1)   & \text{if } D \leq \frac{m}{b} < Dd\\
-  \floor*{\frac{Rb}{m_3'}}(k_3\sqrt[3]{n}+1)+1+\\
-  \quad \quad \sqrt[3]{n}\left(\sqrt[3]{n}D\text{ mod } k_3\right)&\
-  \text{if } Dd \leq \frac{m}{b} < D^2\\
-  \ceil*{\frac{Rb}{m_4'}}(\sqrt[3]{n}^2+1) & \text{if } D^2 \leq \frac{m}{b} < D^2d\\
-  \floor*{\frac{Rb}{m_5'}}(k_5\sqrt[3]{n}^2+1)+1+&\\
+  \ceil*{\frac{Rb}{m_1'}}(k_1+1)-k_1
+  + \left( \sqrt[3]{n}D^2 \text{ mod } k_1 \right)& \text{\hspace*{-0.9em}if } d \leq \frac{m}{b} < D\\[0.4em]
+
+  \ceil*{\frac{Rb}{m_2'}}(\sqrt[3]{n}+1)   & \text{\hspace*{-0.9em}if } D \leq \frac{m}{b} < Dd\\[0.4em]
+
+  \ceil*{\frac{Rb}{m_3'}}(k_3\sqrt[3]{n}+1)-k_3\sqrt[3]{n}+\\[0.4em]
+  \quad \quad \quad \quad \sqrt[3]{n}\left(\sqrt[3]{n}D\text{ mod } k_3\right)& \text{\hspace*{-0.9em}if } Dd \leq \frac{m}{b} < D^2\\[0.4em]
+  \ceil*{\frac{Rb}{m_4'}}(\sqrt[3]{n}^2+1) & \text{\hspace*{-0.9em}if } D^2 \leq \frac{m}{b} < D^2d\\[0.4em]
+  \ceil*{\frac{Rb}{m_5'}}(k_5\sqrt[3]{n}^2+1)-k_5\sqrt[3]{n}^2+&\\[0.4em]
   \quad \quad \sqrt[3]{n}^2\left(\sqrt[3]{n} \text{ mod } k_5\right)&\
-  \text{if } D^2d \leq \frac{m}{b} < R
+  \text{\hspace*{-0.9em}if } D^2d \leq \frac{m}{b} < R
 \end{cases}
@@
 \begin{cases}
-  \floor*{\frac{Rb}{m_1'}}\left(\frac{m_1'\sqrt[3]{n}}{\sqrt[3]{R}b}+1\right)+1+\\
-  \quad \quad \quad \sqrt[3]{n}\sqrt[3]{R}^2 \text{ mod }\floor*{\frac{m\sqrt[3]{n}}{\sqrt[3]{R}b}}\
-  & \text{if } \sqrt[3]{\frac{R}{n}} \leq \frac{m}{b} < \sqrt[3]{R}\\
+  \ceil*{\frac{Rb}{m_1'}}\left(\frac{m_1'\sqrt[3]{n}}{\sqrt[3]{R}b}+1\right)-\frac{m_1'\sqrt[3]{n}}{\sqrt[3]{R}b}\\[0.4em]
+   +\left( \sqrt[3]{n}\sqrt[3]{R}^2 \text{ mod }\floor*{\frac{m\sqrt[3]{n}}{\sqrt[3]{R}b}}\right)\
+  & \text{\hspace*{-0.9em}if } \sqrt[3]{\frac{R}{n}} \leq \frac{m}{b} < \sqrt[3]{R}\\[0.4em]
 
   \ceil*{\frac{Rb}{m_2'}}\left(\sqrt[3]{n}+1\right)\
-  & \text{if } \sqrt[3]{R} \leq \frac{m}{b} < \frac{\sqrt[3]{R}^2}{\sqrt[3]{n}}\\
+  & \text{\hspace*{-0.9em}if } \sqrt[3]{R} \leq \frac{m}{b} < \frac{\sqrt[3]{R}^2}{\sqrt[3]{n}}\\[0.4em]
 
-  \floor*{\frac{Rb}{m_3'}}\left(\frac{m'_3\sqrt[3]{n}^2}{\sqrt[3]{R}^2b}+1\right)+1+\\
-  \sqrt[3]{n}\left( \sqrt[3]{nR} \text{ mod } \floor*{\frac{m\sqrt[3]{n}}{\sqrt[3]{R}^2b}} \right)
-  & \text{if } \frac{\sqrt[3]{R}^2}{\sqrt[3]{n}} \leq \frac{m}{b} < \sqrt[3]{R}^2\\
+  \ceil*{\frac{Rb}{m_3'}}\left(\frac{m'_3\sqrt[3]{n}^2}{\sqrt[3]{R}^2b}+1\right)-\frac{m'_3\sqrt[3]{n}^2}{\sqrt[3]{R}^2b}\\[0.4em]
+  +\sqrt[3]{n}\left( \sqrt[3]{nR} \text{ mod } \floor*{\frac{m\sqrt[3]{n}}{\sqrt[3]{R}^2b}} \right)
+  & \text{\hspace*{-0.9em}if } \frac{\sqrt[3]{R}^2}{\sqrt[3]{n}} \leq \frac{m}{b} < \sqrt[3]{R}^2\\[0.4em]
 
   \ceil*{\frac{Rb}{m_4'}}\left(\sqrt[3]{n}^2+1\right)\
-  & \text{if } \sqrt[3]{R}^2 \leq \frac{m}{b} < \frac{R}{\sqrt[3]{n}}\\
+  & \text{\hspace*{-0.9em}if } \sqrt[3]{R}^2 \leq \frac{m}{b} < \frac{R}{\sqrt[3]{n}}\\[0.4em]
 
-  \ceil*{\frac{Rb}{m_5'}}\left(\frac{m_5'n}{Rb}+1\right)+1+\\
-  \sqrt[3]{n}^2\left( \sqrt[3]{n} \text{ mod } \floor*{\frac{m\sqrt[3]{n}}{Rb}}\right)
-  & \text{if } \frac{R}{\sqrt[3]{n}} \leq \frac{m}{b} < R
+  \ceil*{\frac{Rb}{m_5'}}\left(\frac{m_5'n}{Rb}+1\right)-\frac{m_5'n}{Rb}\\[0.4em]
+  +\sqrt[3]{n}^2\left( \sqrt[3]{n} \text{ mod } \floor*{\frac{m\sqrt[3]{n}}{Rb}}\right)
+  & \text{\hspace*{-0.9em}if } \frac{R}{\sqrt[3]{n}} \leq \frac{m}{b} < R
 \end{cases} \label{eq:seeks-multiple-reads}

### 2017-08-01 paper.tex
Fixed model for cluster reads

@@
 
-The buffer might be a
-slightly complex data structure, for instance an associative array or
-a Python dictionary, capable of storing multiple disjoint sequences of
-contiguous bytes without having to allocate memory for the bytes
-between such sequences. Figure~\ref{fig:cluster-reads-buffer}
-illustrates how the buffer would fill up for the two first blocks in a reconstructed image.Implementation details about the
-buffer are given later.
+The buffer might be a slightly complex data structure, for instance an
+associative array or a Python dictionary, capable of storing multiple
+disjoint sequences of contiguous bytes without having to allocate
+memory for the bytes between such
+sequences. Figure~\ref{fig:cluster-reads-buffer} illustrates how the
+buffer would fill up for the two first blocks in a reconstructed
+image. Implementation details about the buffer are given later.
 \begin{figure}
@@
 complete contiguous slices of the reconstructed image can be assembled
-in memory and written without any seeking. In the worst case, the
+in memory and written in a single seek. In the worst case, the
 memory load only partially covers rows in the reconstructed image:
@@
 the amount of memory $m'$ used by the algorithm is rounded down to the
-closest number of complete blocks (case 1), complete block rows (case
-2) or complete block slices (case 3). This is in general reasonable
+closest number of complete blocks (case 1), of complete block rows (case
+2) or of complete block slices (case 3). This is in general reasonable
 since adding an incomplete row to a set of complete ones multiplies
@@
 memory load to a point that the overall number of seeks also
-increases. However, such cases are a bit pathological and their
+increases. Such cases are, however, slightly unusual and their
 complete description requires extensive calculations involving modulo
@@
 \end{figure}
-To summarize, the amount of memory used $m'$ is set as follows in each of the 3 cases:
-%% \begin{equation*}
-%%   m_{case 1}' = bd^3\floor*{\frac{m}{d^3b}}; m_{case 2}' = bd^3\sqrt[3]{n}\floor*{\frac{m}{d^3b\sqrt[3]{n}}}; m_{case 3}' = bd^3\sqrt[3]{n}^2\floor*{\frac{m}{d^3b\sqrt[3]{n}^2}}
-%% \end{equation*}
-%% which gives:
-\begin{equation*}
-  m_1' = \frac{Rb}{n}\floor*{\frac{mn}{Rb}};
-  m_2' = \frac{Rb}{\sqrt[3]{n}^2}\floor*{\frac{m\sqrt[3]{n}^2}{Rb}};
-  m_3' = \frac{Rb}{\sqrt[3]{n}}\floor*{\frac{m\sqrt[3]{n}}{Rb}}
-\end{equation*}
-
-In addition, our algorithm avoids configurations where the memory load
-overlaps multiple block slices in case 2 or multiple block rows in
-case 1, as such overlaps multiply the number of required seeks
-(see Figure~\ref{fig:avoided-configurations}-Right). Thus, the total
-number of memory loads $x$ required to reconstruct the image is:
-\begin{equation*}
-  x_1 = \ceil*{\frac{Rb}{\sqrt[3]{n}^2m_1'}}\sqrt[3]{n}^2;
-  x_2 = \ceil*{\frac{Rb}{\sqrt[3]{n}m_2'}}\sqrt[3]{n};
-  x_3 = \ceil*{\frac{Rb}{m_3'}}
-\end{equation*}
-
-Finally, the total number of seeks performed by cluster reads to reconstruct the image is:
-%% \begin{equation*}
-%% N_\mathrm{cr} =
-%% \begin{cases}
-%%   n + \sqrt[3]{n}^2 \ceil*{\frac{Rb}{\sqrt[3]{n}^2m_1'}}d^2 & \text{if } \frac{m}{d^3b} < \sqrt[3]{n}\\
-%%   n + \sqrt[3]{n} \ceil*{\frac{Rb}{\sqrt[3]{n}m_2'}}d      & \text{if } \sqrt[3]{n} \leq \frac{m}{d^3b} < \sqrt[3]{n}^2\\
-%%   n + \ceil*{\frac{Rb}{m_3'}}                              & \text{if } \sqrt[3]{n}^2 \leq \frac{m}{d^3b} < n\\
-%%   n + 1                                                          & \text{otherwise}
-%% \end{cases} \label{eq:seeks-cluster-reads}
-%% \end{equation*}
-%% And since $d=\frac{D}{\sqrt[3]{n}}$ and $R=D^3$, we finally have:
-\begin{equation}
-N_\mathrm{cr} =
-\begin{cases}
-  n + \ceil*{\frac{Rb}{\sqrt[3]{n}^2m_1'}}\sqrt[3]{R}^2     & \text{if } \frac{mn}{Rb} < \sqrt[3]{n}\\
-  n + \ceil*{\frac{Rb}{\sqrt[3]{n}m_2'}}\sqrt[3]{R}        & \text{if } \sqrt[3]{n} \leq \frac{mn}{Rb} < \sqrt[3]{n}^2\\
-  n + \ceil*{\frac{Rb}{m_3'}}                              & \text{if } \sqrt[3]{n}^2 \leq \frac{mn}{Rb} < n\\
-  n + 1                                                          & \text{otherwise}
-\end{cases} \label{eq:seeks-cluster-reads-1}
-\end{equation}
-\todo{Is $N_{cr}$ a continuous function of m?}
 
@@
     \STATE (m',case)=\texttt{switch}(m,n,R,b)
-    \STATE old\_block = list[0]
+    \STATE old\_block = sorted\_splits[0]
     \FOR{i = 0 ; i\textless n ; i+=1}
-      \STATE block = list[i]
+      \STATE block = sorted\_splits[i]
       \STATE overlap = \texttt{check\_overlap}(block,old\_block,case)
@@
 \end{algorithm}
+
+The amount of memory used $m'$ is set as follows in each of the 3 cases:
+\begin{equation*}
+  m_1' = \frac{Rb}{n}\floor*{\frac{mn}{Rb}};
+  m_2' = \frac{Rb}{\sqrt[3]{n}^2}\floor*{\frac{m\sqrt[3]{n}^2}{Rb}};
+  m_3' = \frac{Rb}{\sqrt[3]{n}}\floor*{\frac{m\sqrt[3]{n}}{Rb}}
+\end{equation*}
+The number of seeks performed by cluster reads in each of the three cases is:
+\begin{equation*}
+  N^i_{\mathrm{CR}} = n + x_ib_i, \quad i \in \llbracket 1, 3\rrbracket,
+\end{equation*}
+where $x_i$ is the number of memory loads required to reconstruct the
+image and $b_i$ is the number of seeks required to write a the memory
+load. The first $n$ seeks in the equation are required to read all the blocks
+once. According to Figure~\ref{fig:cluster-reads}, we have:
+\begin{equation*}
+  b_1=d^2=\sqrt[3]{\frac{R}{n}}^2 \quad ; \quad b_2=d=\sqrt[3]{\frac{R}{n}} \quad ; \quad b_3=1
+\end{equation*}
+
+Our algorithm avoids configurations where the memory load
+overlaps multiple block slices in case 2 or multiple block rows in
+case 1, as such overlaps multiply the number of required seeks
+(see Figure~\ref{fig:avoided-configurations}-Right). Thus, the
+numbers of memory loads required to reconstruct the image are:
+\begin{equation*}
+  x_1 = \ceil*{\frac{Rb}{\sqrt[3]{n}^2m_1'}}\sqrt[3]{n}^2;
+  x_2 = \ceil*{\frac{Rb}{\sqrt[3]{n}m_2'}}\sqrt[3]{n};
+  x_3 = \ceil*{\frac{Rb}{m_3'}}
+\end{equation*}
+
+Finally, the total number of seeks performed by cluster reads to reconstruct the image is:
+\begin{equation}
+N_\mathrm{CR} =
+\begin{cases}
+  n + \ceil*{\frac{Rb}{\sqrt[3]{n}^2m_1'}}\sqrt[3]{R}^2     & \text{if } \frac{mn}{Rb} < \sqrt[3]{n}\\[0.4em]
+  n + \ceil*{\frac{Rb}{\sqrt[3]{n}m_2'}}\sqrt[3]{R}        & \text{if } \sqrt[3]{n} \leq \frac{mn}{Rb} < \sqrt[3]{n}^2\\[0.4em]
+  n + \ceil*{\frac{Rb}{m_3'}}                              & \text{if } \sqrt[3]{n}^2 \leq \frac{mn}{Rb} < n\\[0.4em]
+  n + 1                                                          & \text{otherwise}
+\end{cases} \label{eq:seeks-cluster-reads-1}
+\end{equation}
+It should be noted that $N_{CR}$ is not a continuous function of m,
+due to the differences among $b_i$ values.
 
@@
         \STATE in split\_data, extract the rows in write\_range
-        \STATE write rows to buffer
+        \STATE insert rows in buffer
       \ENDIF
@@
     \STATE write buffer to reconstructed\_image
-    \STATE start\_index = end\_index + 1
-    \STATE end\_index += m
+    \STATE start\_index = end\_index + 1 ; end\_index += m
   \ENDWHILE
@@
 number of blocks accessed by the first ($x_i$-1) memory loads and
-$b_i'$ is the number of blocks access by the last memory load. We have:
+$b_i'$ is the number of blocks access by the last memory load. The
+first $x_i$ seeks in the equation are required to write each memory
+load (1 seek per memory load). We have:
 \begin{equation*}
@@
 \begin{cases}
-  \ceil*{\frac{Rb}{m_1'}}(k_1+1)-k_1
+  \ceil*{\frac{Rb}{m_1'}}(k_1+1)-k_1 \\
   + \left( \sqrt[3]{n}D^2 \text{ mod } k_1 \right)& \text{\hspace*{-0.9em}if } d \leq \frac{m}{b} < D\\[0.4em]
@@
 
-\section{Benchmark}
-\label{sec:benchmark}
-
-\subsection{Data}
-To benchmark our library we use the 76~GB BigBrain image split in 125
-non-overlapping chunks. The data is available \todo{[here]}.
-
-\subsection{Infrastructure}
-
-We run experiments using a Dell Precision Tower 3620 with CentOS Linux release
-7.3.1611, 32~GB of RAM and two disks:
-\begin{itemize}
-  \item Hard disk drive (HDD): HGST Travelstar 7K1000, 7200~rpm, 931GiB (1TB), firmware version JB0OA3W0.
-    \item Solid-state drive (SSD): SanDisk X400 2.5, 238GiB (256GB),
-      firmware version X4130012.
-\end{itemize}
-Both drives used 512-byte logical sectors, 4096-byte physical sectors,
-SATA >3.1 (6.0 Gb/s) and were accessed through the XFS file system
-v4.5.0.
-%smartctl --xall /dev/sda
-
 \section{Simulations}
@@
 \end{figure}
+
+\section{Benchmark}
+\label{sec:benchmark}
+
+\subsection{Data}
+To benchmark our library we use the 76~GB BigBrain image split in 125
+non-overlapping chunks. The data is available \todo{[here]}.
+
+\subsection{Infrastructure}
+
+We run experiments using a Dell Precision Tower 3620 with CentOS Linux release
+7.3.1611, 32~GB of RAM and two disks:
+\begin{itemize}
+  \item Hard disk drive (HDD): HGST Travelstar 7K1000, 7200~rpm, 931GiB (1TB), firmware version JB0OA3W0.
+    \item Solid-state drive (SSD): SanDisk X400 2.5, 238GiB (256GB),
+      firmware version X4130012.
+\end{itemize}
+Both drives used 512-byte logical sectors, 4096-byte physical sectors,
+SATA >3.1 (6.0 Gb/s) and were accessed through the XFS file system
+v4.5.0.
+%smartctl --xall /dev/sda
 

### 2017-08-01 paper.tex
Added figures for notations

@@
 
-In our context, split and merge relate to the same dual
-problem \todo{not so sure about this actually}. Without loss of
-generality, we focus here on merging for the sake of concision. Our
-goal then is to merge a set of $n$ chunks into a single reconstructed
-3D image with $R$ voxels of size $b$. We assume that all chunks are of
+In our context, split and merge relate to the same dual problem . We
+focus here on merging for the sake of concision. Splitting algorithms
+can be derived from merging by swapping reads and writes. Our goal
+then is to merge a set of $n$ chunks into a single reconstructed 3D
+image with $R$ voxels of size $b$. We assume that all chunks are of
 identical size and that slices are squares and blocks are cubes.
@@
 
-We adopt the following notations:
+We adopt the following notations (see Figure\ref{fig:notations}):
 \begin{itemize}
@@
 
+\begin{figure}
+  \centering
+  \begin{minipage}[b]{0.3\columnwidth}
+    \def\svgwidth{\columnwidth}
+    \input{figures/svg/Notations.pdf_tex}
+    \caption{Notations.}
+    \label{fig:notations}
+  \end{minipage}
+  \quad \quad \quad
+  \begin{minipage}[b]{0.3\columnwidth}
+    \def\svgwidth{\columnwidth}
+    \input{figures/svg/buffer.pdf_tex}
+    \caption{Buffer used in cluster reads.}
+    \label{fig:cluster-reads-buffer}
+  \end{minipage}
+\end{figure}
+
 \subsection{Buffered slices}
@@
 buffer would fill up for the two first blocks in a reconstructed
-image. Implementation details about the buffer are given later.
-\begin{figure}
-\centering
-\def\svgwidth{0.3\columnwidth}
-\input{figures/svg/buffer.pdf_tex}
-\caption{Illustration of the buffer used in cluster reads, showing how
-  the two first 4x4 blocks of a reconstructed image would be arranged
-  in the buffer.  White portions in the buffer are not allocated.  The
-  column-major order is used, similar to the NIfTI format: i is the
-  fastest changing dimension, followed by j, followed by k.}
-\label{fig:cluster-reads-buffer}
-% Data file: path
-% Gnuplot file: path
-% Script file: path
-\end{figure}
+image. It shows how the two first 4x4 blocks of a reconstructed image
+would be arranged in the buffer.  White portions in the buffer are not
+allocated.  The column-major order is used, similar to the NIfTI
+format: i is the fastest changing dimension, followed by j, followed
+by k.
 
@@
 \begin{cases}
-  n + \ceil*{\frac{Rb}{\sqrt[3]{n}^2m_1'}}\sqrt[3]{R}^2     & \text{if } \frac{mn}{Rb} < \sqrt[3]{n}\\[0.4em]
-  n + \ceil*{\frac{Rb}{\sqrt[3]{n}m_2'}}\sqrt[3]{R}        & \text{if } \sqrt[3]{n} \leq \frac{mn}{Rb} < \sqrt[3]{n}^2\\[0.4em]
-  n + \ceil*{\frac{Rb}{m_3'}}                              & \text{if } \sqrt[3]{n}^2 \leq \frac{mn}{Rb} < n\\[0.4em]
+  n + \ceil*{\frac{Rb}{\sqrt[3]{n}^2m_1'}}\sqrt[3]{R}^2     & \text{if } m < \frac{Rb}{\sqrt[3]{n}^2}\\[0.4em]
+  n + \ceil*{\frac{Rb}{\sqrt[3]{n}m_2'}}\sqrt[3]{R}        & \text{if } \frac{Rb}{\sqrt[3]{n}^2} \leq m < \frac{Rb}{\sqrt[3]{n}}\\[0.4em]
+  n + \ceil*{\frac{Rb}{m_3'}}                              & \text{if } \frac{Rb}{\sqrt[3]{n}} \leq Rb < n\\[0.4em]
   n + 1                                                          & \text{otherwise}
@@
 
-\todo{Yongping: mention the cache cleaning thing that you run before every experiment.}
-
 Experiment scripts are available at xxx.
@@
 
-(If not added as co-authors): Greg Kiar, Pierre Bellec.
+(If not added as co-authors): Greg Kiar, Pierre Bellec, Claude Lepage, Lindsay B. Lewis.
 

### 2017-08-01 figures/svg/export_svg_files.sh

@@ -1,7 +1,7 @@
 #!/usr/bin/env bash
 
 for name in buffer case1-a case2-a case3 incomplete-rows overlap mreads-case1 \
-        mreads-case2 mreads-case3 mreads-case4 mreads-case5
+        mreads-case2 mreads-case3 mreads-case4 mreads-case5 Notations
 do
   #echo "Exporting ${name}.svg"
   inkscape -D -z --file=figures/svg/${name}.svg --export-pdf=figures/svg/${name}.pdf --export-latex

### 2017-08-01 paper.tex
Minor updates
add number of seeks figure and minor fix

@@
 \input{figures/svg/incomplete-rows.pdf_tex}
-\hfill
+\quad \quad \quad
 \def\svgwidth{0.3\columnwidth}
@@
 used by its argument, including only its allocated segments in case
-the argument is a buffer. It should be noted that writing the buffer
-to the reconstructed image (line 11) might involve some manipulation
-in the buffer to merge its continuous segments. \todo{Precise this
-  based on Valerie's final code.}
+the argument is a buffer. 
 \begin{algorithm}[h]
@@
   \caption{Comparison between Multiple reads, blocks and slices. Left: SSD. Right: HDD.}
-  % Gnuplot file: scripts/blocks-slices/createboxplot.gnplt
+  % Python file: scripts/mreads/generate_avg_var_data.py
+  % Gnuplot file: scripts/mreads/compare_breakdown_errbar.gnuplot
 \label{fig:multiple-reads}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=0.70\columnwidth]{figures/benchmark-mreads/mreads-number-of-seeks.pdf}
+  \hfill
+  \caption{Number of seeks in Multiple reads}
+  % Gnuplot file: scripts/mreads/number_of_seeks.gnuplot
+\label{fig:multiple-reads-number-of-seek}
 \end{figure}

### 2017-08-01 paper.tex
A better intro, still not perfect

--- 2017-08-01 paper.tex
@@
 \begin{abstract}
-  Splitting and merging data is a requirement for most parallel
-    or distributed processing operations, however, naive algorithms
-    perform very poorly on ultra-high resolution 3D images due to
-    important seek times. To address this issue, we created two
-    algorithms based on in-memory reconstruction of image segments and
-    analyzed their complexity in terms of number of seeks. Evaluation
-    on a 3000x3000x3000 image of the brain shows that both our
-    algorithms can split or merge blocks with a performance similar to
-    slices, which solves our initial problem. We conclude that
-    splitting and merging large 3D images can be done efficiently
-    based on simple data formats.
+  Splitting and merging data is a requirement for most parallel or
+  distributed processing operations. Naive algorithms, however,
+  perform very poorly to split or merge 3D blocks from ultra-high
+  resolution images, due to important seek times. We propose and
+  analyze two algorithms based on in-memory reconstruction of image
+  parts. Cluster reads access image splits only once while Multiple
+  reads minimize the accesses to the image. Evaluation on a
+  3000x3000x3000 brain image shows that our algorithms perform
+  similarly to the optimal configuration (slice merging) provided that
+  enough memory is available, which solves our initial problem.  We
+  conclude that splitting and merging large 3D images can be done
+  efficiently based on simple data formats.
 \end{abstract}
@@
 
+% Problem definition
 Three-dimensional images that exceed typical memory size are
-increasingly found in a variety of disciplines. BigBrain, for
+increasingly found in a variety of disciplines. Big~Brain, for
 instance, is a 3D histological image of the human brain that
@@
 (1$\mu$m in-plane), and 76~GB at a 40$\mu$m isotropic resolution
-commonly used in neurosciences~\cite{amunts2013bigbrain}. Other
-examples found in medical imaging, our primary domain of interest,
-include high-resolution 3D electron microscopy (see, e.g.,
-~\cite{bock2011network}) or micro and
+commonly used in neurosciences~\cite{amunts2013bigbrain}. This dataset
+is being re-scanned at an even higher resolution, leading to 100~TB of
+raw data. Other examples found in medical imaging, our primary domain
+of interest, include high-resolution 3D electron microscopy (see,
+e.g., ~\cite{bock2011network}) or micro- and
 nano-tomography~\cite{10.1371/journal.pone.0035691}. As such images
 would typically be processed on a computing cluster, possibly using
-locality-aware file systems such as HDFS, software libraries are
-needed to split and merge such images efficiently. We introduce and
-compare a family of algorithms to address this problem in a sequential
-environment made of a single storage disk. We are aware of the need
-for parallel algorithms that would split and merge images to and from
-an array of disks but we plan to describe such algorithms in a
-subsequent paper based on the results of the present one. \todo{ There
-  is BigBrain-1 rescanned at 1 micron in the plane. So resolution is
-  1x1x20, 7404 slices. If raw data are about 1Tb, then multiply by 100
-  for an estimation of the size of the new image. }
+locality-aware file systems such as HDFS~\cite{shvachko2010hadoop},
+efficient software libraries are needed to split and merge them
+efficiently. In particular, data reads and writes might result in
+extensive seek times that drastically limit I/O performance.  In this
+paper we introduce and compare a family of sequential algorithms to
+split and merge images with minimal seek time.
 
-%also in
-%MB/s).
+Images are split into chunks representing 3D blocks or 2D slices,
+which we assume fit in memory. A dataset such as Big~Brain would
+perhaps be split into 125 chunks of 600~MB. The decision to split an
+image into slices or blocks, and the size of the resulting chunks, is
+usually done by the application. Some applications would commonly
+require blocks, for instance spatial filtering, while other ones such
+as acquisition artifact removal would rather work on
+slices. Applications processing voxels individually, for instance
+histogram computation or k-means clustering, could work on either
+slices or blocks. Flexibility is thus required in the splitting
+scheme.
 
-Images are split into chunks representing 3D blocks or 2D slices. We
-assume that a block or slice fits in memory. The BigBrain would
-perhaps be split into 125 chunks of 600~MB. The decision to split an
-image into slices or blocks, and the size of the resulting chunks, may
-be constrained by the application. Some applications, for instance
-spatial filtering, would commonly require blocks while other ones such
-as acquisition artifact removal would rather work on
-slices. Flexibility is thus required in the splitting
-scheme. Applications processing voxels individually, for instance
-histogram computation or k-means clustering, could work on either
-slices or blocks.
+Seek time is highly dependent on the byte ordering in image files,
+which we assume is arbitrary but known to the algorithm. The geometry
+of the chunks to split or merge (slices or blocks) and their sizes are
+also critical, which we again assume arbitrary but known to the
+algorithm.  The main idea of our algorithms will be to convert data
+orderings in memory before writing them to disk.
 
-The main problem encountered in the context of this paper is the fact
-that, if not ordered properly, data reads and writes might result in
-extensive seek times that, as we illustrate below, drastically limit
-the I/O performance. This problem is obviously related to the ordering
-of data bytes in image files, which we assume is arbitrary but known
-to the algorithm. The problem is also related to the geometry of the
-chunks to extract or merge, which we again assume arbitrary but known
-to the algorithm. In a nutshell, we are looking for algorithms to
-reduce seek times while allowing for arbitrary chunk geometries and
-arbitrary byte orders in data files. The main idea of the variations
-presented in the remainder of the paper is to convert data orderings
-in memory before writing them on disk.
-
-The literature is remarkably scarce on the problem described
-above. Parallel processing of 3D images has obviously been extensively
-studied, but methods have focused more on geometrical approaches to
-partition images and on load-balancing and task scheduling techniques
-while we aim at algorithms to efficiently split or merge images
-regardless of the geometry of the chunks \todo{do a serious literature
-review on this topic}. Perhaps this lack of related works could be
-attributed to the fact that disk I/O weren't a bottleneck until image
-resolution reached a certain threshold. A relevant reference is the
-description of the Open Connectome Data Cluster~\cite{burns2013open},
-a data warehouse system that allows users to retrieve specific chunks
-of large datasets. In this work, the problem of long seek times is
-addressed by designing a specific file format based on space-filling
-curves, that elegantly preserves spatial proximity on disk. However,
-addressing the problem using a specific data organization optimizes
-the system for specific chunk geometries, volumetric in this case,
-while we aim at algorithms that support multiple types of geometries.
+The literature on this problem is remarkably scarce. Parallel and
+distributed image processing has obviously been extensively studied
+and used in various
+platforms~\cite{miguet1991elastic,tang2007eman2,yang2008parallel,braunl2013parallel,moise2013terabyte,bajcsy2013terabyte},
+but methods have focused more on geometrical approaches to partition
+images and on load-balancing and task scheduling techniques while we
+aim at algorithms to efficiently split or merge images regardless of
+the geometry of the chunks. Seek times are, however, often identified
+as an issue but the preferred solution is usually to optimize data
+storage formats for a certain application while we are searching for
+algorithms that would reduce the seek time regardless of the data
+format. For instance, the Open Connectome Data
+Cluster~\cite{burns2013open} is a data warehouse system that allows
+users to retrieve specific chunks of large image datasets. It
+addresses seek times by designing a specific file format based on
+space-filling curves, which elegantly preserves spatial proximity on
+disk. Such specific data formats optimize the system for particular
+chunk geometries, blocks in this case, while we aim at algorithms to
+support multiple types of geometries.
+%http://dl.acm.org/citation.cfm?id=218395 also minimizes seek time using a specific format.
 
@@
 \begin{itemize}
-  \item We propose a set of algorithms to split and merge 3D images
-    that do not fit in memory.
-  \item We model the performance (number of seeks) of these algorithms,
-    based on a simple but realistic disk I/O model.
+  \item We propose a set of algorithms to split and merge
+    terabyte-scale 3D images from blocks or slices.
+  \item We model the performance (number of seeks) of these algorithms.
   \item We evaluate our model and algorithms on a representative
@@
 Section~\ref{sec:results} details the results and
-Sections~\ref{sec:discussion} discusses them.
+Sections~\ref{sec:discussion} discusses them.\todo{Revise when outline is final.}
 
@@
 \label{sec:algos}
-
-\todo{Overlapping blocks. non-cubic blocks.}
 
@@
 Without loss of generality, we consider a file format where voxels are
-written after alphanumerical sorting of their (x,y,z)
-representation. That is, x slices are written one after the other, and
-in every x slice, y rows are written one after the other. NIfTI, the
-file format defined by the Neuroimaging Informatics Technology
-Initiative NIfTI\footnote{\url{https://nifti.nimh.nih.gov}} is such a
-format, used by the vast majority of the neuroimaging community. In
-the remainder, we use the term ``slice" for ``x slice".
+written after alphanumerical sorting of their (z,y,x)
+representation. That is, z slices are written one after the other, and
+in every z slice, y rows are written one after the other. x is the
+fastest-changing dimention.  NIfTI, the file format defined by the
+Neuroimaging Informatics Technology Initiative
+NIfTI\footnote{\url{https://nifti.nimh.nih.gov}} is such a format,
+used by the vast majority of the neuroimaging community. In the
+remainder, we use the term ``slice" for ``z slice".
 
@@
 
-For the sake of mathematical simplicity, we assume that
-the reconstructed image and blocks are cubic. Under this assumption,
-we also have the following relations:
+For the sake of mathematical simplicity, we assume that the
+reconstructed image and blocks are cubic and that they do not
+overlap. Under this assumption, we also have the following relations:
 \begin{itemize}
@@
 \end{itemize}
+
 %\todo{Make a figure to illustrate the notations.}
@@
 We use a simplistic disk model where a disk is characterized by its
-read and write rates and its access time. In particular, we assume
-that seeks require a constant amount of time, regardless of the
-position seeked to. In practice, large variations would be expected,
-but modeling such variations would inevitably lead to models specific
-to the hardware, file system or operating system, which we
-intentionally avoid here. Thus, our goal is to find algorithms that
-minimize the \emph{number} of seek operations. Likewise, in modern
-systems, read and write times are greatly impacted by caching
-occurring at several levels, which we do not model here.
+read and write rates and its seek time. We do not distinguish between
+access time and seek time. We also assume that seeks require a
+constant amount of time, regardless of the position seeked to. In
+practice, large variations would be expected, but modeling such
+variations would inevitably lead to models specific to the hardware,
+file system or operating system, which we intentionally avoid
+here. Thus, our goal is to find algorithms that minimize the
+\emph{number} of seek operations. Likewise, in modern systems, read
+and write times are greatly impacted by caching occurring at several
+levels, which we do not model here.
 
@@
 \end{algorithm}
-
 \begin{algorithm}[h]
@@
 \end{algorithm}
-
 Algorithms~\ref{algo:naive-slices} and~\ref{algo:naive-blocks}
@@
 reconstructed image, the number of seeks in
-Algorithm~\ref{algo:naive-slices} is only $2n$:
+Algorithm~\ref{algo:naive-slices} is only $2n$ as $n$ seeks are
+required to read the slices and $n$ seeks are required to write them:
 \begin{equation}
@@
 performed by Algorithm~\ref{algo:naive-blocks} is:
+\begin{equation*}
+N_\mathrm{blocks} = n+nd^2 
+\end{equation*}
+or, using $R$ and $n$ as our main variables:
 \begin{equation}
-N_\mathrm{blocks} = n+nd^2 
+N_\mathrm{blocks} = n+n\left(\sqrt[3]{\frac{R}{n}}\right)^2
 \end{equation}
@@
 
-The algorithms presented up to now are in fact particular cases of
-memory buffering where the amount of available memory equals the
-maximum size of a chunk.
-
+Algorithm~\ref{algo:naive-slices} is a particular case of memory
+buffering where the amount of available memory equals the maximum size
+of a chunk. More buffering can be achieved when the amount of
+available memory increases, as shown in
+Algorithm~\ref{algo:buffered-slices}.
 \begin{algorithm}[h]
@@
   \begin{algorithmic}[1]
-    \STATE list = sort slices by increasing x values
+    \STATE sorted\_slices = sort slices by increasing z values
     \STATE initialize buffer
     \FOR{i = 0 ; i \textless n ; i+=1 }
-      \STATE slice = list[index]
+      \STATE slice = sorted\_slices[i]
       \IF{\texttt{sizeof}(buffer)+\texttt{sizeof}(slice) $\geq$ m}
@@
 \end{algorithm}
-
+The number of seeks obtained with buffered slices is:
 \begin{equation}
@@
 \end{equation}
+The next two algorithms attempt to generalize buffered slices to blocks.
 
@@
 \subsection{Data}
-To benchmark our library we use the 76~GB BigBrain image split in 125
+To benchmark our library we use the 76~GB Big~Brain image split in 125
 non-overlapping chunks. The data is available \todo{[here]}.

### 2017-08-01 paper.tex
Revised up until II.F

@@
   analyze two algorithms based on in-memory reconstruction of image
-  parts. Cluster reads access image splits only once while Multiple
+  parts. Cluster reads access image chunks only once while Multiple
   reads minimize the accesses to the image. Evaluation on a
@@
 
-In our context, split and merge relate to the same dual problem . We
+Split and merge relate to the same dual problem in our context. We
 focus here on merging for the sake of concision. Splitting algorithms
-can be derived from merging by swapping reads and writes. Our goal
-then is to merge a set of $n$ chunks into a single reconstructed 3D
-image with $R$ voxels of size $b$. We assume that all chunks are of
-identical size and that slices are squares and blocks are cubes.
+can be derived from merging ones by swapping reads and writes. Our
+goal then is to merge a set of $n$ chunks into a single reconstructed
+3D image with $R$ voxels of size $b$. For simplicity, we assume that
+all chunks are of identical size, that they do not overlap, and that
+slices are squares and blocks are cubes.
 
-Without loss of generality, we consider a file format where voxels are
-written after alphanumerical sorting of their (z,y,x)
-representation. That is, z slices are written one after the other, and
-in every z slice, y rows are written one after the other. x is the
-fastest-changing dimention.  NIfTI, the file format defined by the
+We consider a file format where voxels are written in column-major
+order, that is, for an image of dimensions (i, j, k), i is the fastest
+changing dimension, followed by dimension j, followed by dimension
+k. All voxels in a \emph{slice} have a the same $k$ and all voxels in
+a \emph{row} have the same $j$.  NIfTI, the file format defined by the
 Neuroimaging Informatics Technology Initiative
 NIfTI\footnote{\url{https://nifti.nimh.nih.gov}} is such a format,
-used by the vast majority of the neuroimaging community. In the
-remainder, we use the term ``slice" for ``z slice".
+used by the vast majority of the neuroimaging community.
 
@@
 \item $R=D^3$: number of voxels in the reconstructed image.
-\item $b$: number of bytes per voxel in the reconstructed image and chunks (in B).
+\item $b$: number of bytes per voxel (in B).
 \item $n$: number of chunks (blocks or slices).
@@
 \end{itemize}
-
-For the sake of mathematical simplicity, we assume that the
-reconstructed image and blocks are cubic and that they do not
-overlap. Under this assumption, we also have the following relations:
+We also have the following relations:
 \begin{itemize}
-\item Number of slices in a block: $\sqrt[3]{\frac{R}{n}}=d$. This is
-  also the number of rows and columns in a slice and the number of
-  voxel in a row or column.
-\item Number of blocks required to fill a row of the reconstructed
-  image: $\sqrt[3]{n}$.
+\item Number of slices, rows, columns in a block: $\sqrt[3]{\frac{R}{n}}=d$.
+\item Number of blocks in a block row: $\sqrt[3]{n}$.
 \end{itemize}
-
-%\todo{Make a figure to illustrate the notations.}
 
@@
 
-We use a simplistic disk model where a disk is characterized by its
-read and write rates and its seek time. We do not distinguish between
-access time and seek time. We also assume that seeks require a
-constant amount of time, regardless of the position seeked to. In
-practice, large variations would be expected, but modeling such
-variations would inevitably lead to models specific to the hardware,
-file system or operating system, which we intentionally avoid
-here. Thus, our goal is to find algorithms that minimize the
-\emph{number} of seek operations. Likewise, in modern systems, read
-and write times are greatly impacted by caching occurring at several
-levels, which we do not model here.
+A disk is characterized by its read and write rates and its seek
+time. We do not distinguish between access time and seek time. We also
+assume that seeks require a constant amount of time, regardless of the
+position seeked to. In practice, large variations would be expected,
+but modeling such variations would inevitably lead to models specific
+to the hardware, file system or operating system, which we
+intentionally avoid here. Likewise, in modern systems, read and write
+times are greatly impacted by caching occurring at several levels,
+which we do not model here. Thus, our goal is to find algorithms that
+minimize the \emph{number} of seek operations.
 
-%% and its seek time,
-%% a function $\sigma$ of the seek distance (in B). More precisely,
-%% $\sigma$ would typically be in the order of $\beta$ when an opened
-%% file is seeked to a random position, but it would be much smaller when
-%% a file is seeked to a nearby position.
-%% The disk model encompasses file systems, kernel modules and disk
-%% controllers required to access the disk from a program run in user
-%% space. In modern systems, this often includes caching at various
-%% levels \todo{\cite{x}}, which we will model by zeroing $\alpha_r$ or
-%% $\alpha_w$ values depending on the application.
+\todo{Add ballpark values of seek time (5ms) and read/write rates.}
 
@@
 
-Algorithms~\ref{algo:naive-slices} and~\ref{algo:naive-blocks} are the
-naive merging methods for slices and blocks.
-\begin{algorithm}[h]
-\caption{Naive merging from slices.}
+\begin{figure}
+  \centering
+  \begin{minipage}[b]{0.42\columnwidth}
+    \def\svgwidth{0.8\columnwidth}
+    \input{figures/svg/Notations.pdf_tex}
+    \caption{Notations. A \emph{block row} is shown in red. A
+      \emph{block slice} is shown in blue.}
+    \label{fig:notations}
+  \end{minipage}
+  \quad \quad \quad
+  \begin{minipage}[b]{0.42\columnwidth}
+    \def\svgwidth{0.8\columnwidth}
+    \input{figures/svg/buffer.pdf_tex}
+    \caption{Buffer used in cluster reads (d=4).  White portions in
+      the buffer are not allocated. }
+    \label{fig:cluster-reads-buffer}
+  \end{minipage}
+\end{figure}
+Algorithms~\ref{algo:naive-slices} and~\ref{algo:naive-blocks} show
+the naive merging methods for slices and blocks. These algorithms
+actually have very different complexities even though blocks and
+slices have identical sizes. Since slices are stored contiguously in
+the reconstructed image, the number of seeks in
+Algorithm~\ref{algo:naive-slices} is only $2n$ as $n$ seeks are
+required to read the slices and $n$ seeks are required to write them:
+\begin{equation}
+  N_\mathrm{slices} = 2n 
+\end{equation}
+However,
+Algorithm~\ref{algo:naive-blocks} has to do extra seeks for each row
+in each slice of each block:
+\begin{equation*}
+N_\mathrm{blocks} = n+nd^2 
+\end{equation*}
+or, using $R$ and $n$ as main variables:
+\begin{equation}
+N_\mathrm{blocks} = n+n\left(\sqrt[3]{\frac{R}{n}}\right)^2
+\end{equation}
+In practice, this difference could lead to a tremendous slowdown, as
+we will show later.
+
+\begin{algorithm}
+\caption{Naive merging from slices}
 \label{algo:naive-slices} 
@@
 \begin{algorithm}[h]
-\caption{Naive merging from blocks.}
+\caption{Naive merging from blocks}
 \label{algo:naive-blocks}
@@
 \end{algorithm}
-Algorithms~\ref{algo:naive-slices} and~\ref{algo:naive-blocks}
-actually have very different performance even though blocks and slices
-have identical sizes. Since slices are stored contiguously in the
-reconstructed image, the number of seeks in
-Algorithm~\ref{algo:naive-slices} is only $2n$ as $n$ seeks are
-required to read the slices and $n$ seeks are required to write them:
-\begin{equation}
-  N_\mathrm{slices} = 2n 
-\end{equation}
-However,
-Algorithm~\ref{algo:naive-blocks} has to do extra seeks for each row
-in each slice of each block. That is, the total number of seeks
-performed by Algorithm~\ref{algo:naive-blocks} is:
-\begin{equation*}
-N_\mathrm{blocks} = n+nd^2 
-\end{equation*}
-or, using $R$ and $n$ as our main variables:
-\begin{equation}
-N_\mathrm{blocks} = n+n\left(\sqrt[3]{\frac{R}{n}}\right)^2
-\end{equation}
-In
-practice, this difference could lead to a tremendous slowdown as we will show
-later.
-
 %%  For slices, the total merge time is modeled as follows:
@@
 
-\begin{figure}
-  \centering
-  \begin{minipage}[b]{0.3\columnwidth}
-    \def\svgwidth{\columnwidth}
-    \input{figures/svg/Notations.pdf_tex}
-    \caption{Notations.}
-    \label{fig:notations}
-  \end{minipage}
-  \quad \quad \quad
-  \begin{minipage}[b]{0.3\columnwidth}
-    \def\svgwidth{\columnwidth}
-    \input{figures/svg/buffer.pdf_tex}
-    \caption{Buffer used in cluster reads.}
-    \label{fig:cluster-reads-buffer}
-  \end{minipage}
-\end{figure}
-
 \subsection{Buffered slices}
@@
   \begin{algorithmic}[1]
-    \STATE sorted\_slices = sort slices by increasing z values
+    \STATE sorted\_slices = sort slices by increasing k values
     \STATE initialize buffer
@@
 \end{algorithm}
-The number of seeks obtained with buffered slices is:
+This algorithm writes in the reconstructed image using a single seek
+per memory load. Therefore:
 \begin{equation}
@@
 \end{equation}
-The next two algorithms attempt to generalize buffered slices to blocks.
+Buffered slices are straightforward to implement, however, their
+extension to block merging is not easy. The remainder of this Section
+presents our attempts for such a generalization.
 
@@
 
-The main idea of cluster reads is to load multiple blocks in memory,
-concatenate them in a buffer and write the buffer in the reconstructed
-image. Seeking is reduced compared to naive block merging since
-contiguous parts of the buffer will be written without seeking. In
-cluster reads, a given block is accessed only once.
+Cluster reads are the more direct extension of buffered slices to
+blocks: they load multiple blocks in memory, concatenate them in a
+buffer and write the buffer in the reconstructed image. Seeking is
+reduced compared to naive block merging since contiguous parts of the
+buffer will be written without seeking. A given
+block is accessed only once.
 
@@
 buffer would fill up for the two first blocks in a reconstructed
-image. It shows how the two first 4x4 blocks of a reconstructed image
-would be arranged in the buffer.  White portions in the buffer are not
-allocated.  The column-major order is used, similar to the NIfTI
-format: i is the fastest changing dimension, followed by j, followed
-by k.
+image, assuming that blocks are of size 4x4x4.
 
@@
 rounding $m$ down to $m'$ might increase the number of required
-memory load to a point that the overall number of seeks also
+memory loads to a point that the overall number of seeks also
 increases. Such cases are, however, slightly unusual and their
@@
 
+Our algorithm also avoids configurations where the memory load
+overlaps multiple block slices in case 2 or multiple block rows in
+case 1, as such overlaps multiply the number of required seeks (see
+Figure~\ref{fig:avoided-configurations}-Right).
+
 Cluster reads are described in
@@
 (case 2) or multiple block rows (case 1). For case 3 it always returns
-true.  Function \texttt{sizeof} (line 8) returns the actual memory
+false.  Function \texttt{sizeof} (line 8) returns the actual memory
 used by its argument, including only its allocated segments in case
@@
   \begin{algorithmic}[1]
-    \STATE sorted\_splits = sort splits by increasing (z,y,x) values
+    \STATE sorted\_blocks = sort blocks by increasing (k,j,i)
     \STATE initialize buffer
-    \STATE (m',case)=\texttt{switch}(m,n,R,b)
-    \STATE old\_block = sorted\_splits[0]
+    \STATE (m',case)=switch(m,n,R,b)
+    \STATE old\_block = sorted\_blocks[0]
     \FOR{i = 0 ; i\textless n ; i+=1}
-      \STATE block = sorted\_splits[i]
-      \STATE overlap = \texttt{check\_overlap}(block,old\_block,case)
-      \IF{\texttt{sizeof}(buffer)+\texttt{sizeof}(block) $\geq$ m' \textbf{or} overlap=true}
+      \STATE block = sorted\_blocks[i]
+      \STATE overlap = check\_overlap(block,old\_block,case)
+      \IF{sizeof(buffer)+sizeof(block) $\geq$ m' \textbf{or} overlap=true}
       \STATE write buffer in reconstructed image
@@
 where $x_i$ is the number of memory loads required to reconstruct the
-image and $b_i$ is the number of seeks required to write a the memory
-load. The first $n$ seeks in the equation are required to read all the blocks
-once. According to Figure~\ref{fig:cluster-reads}, we have:
+image and $b_i$ is the number of seeks required to write the memory
+load. The first $n$ seeks in the equation are required to read all the
+blocks once. According to Figure~\ref{fig:cluster-reads}, we have:
 \begin{equation*}
@@
 \end{equation*}
-
-Our algorithm avoids configurations where the memory load
-overlaps multiple block slices in case 2 or multiple block rows in
-case 1, as such overlaps multiply the number of required seeks
-(see Figure~\ref{fig:avoided-configurations}-Right). Thus, the
+The
 numbers of memory loads required to reconstruct the image are:
@@
 \end{equation*}
+Because our algorithm avoids overlapping configurations, $x_1$ is
+proportional to the total number of block rows in the image
+($\sqrt[3]{n}^2$) and $x_2$ is proportional to the total number of
+block slices ($\sqrt[3]{n}$).
 
-Finally, the total number of seeks performed by cluster reads to reconstruct the image is:
+Finally, the total number of seeks
+performed by cluster reads to reconstruct the image is:
 \begin{equation}
@@
 main idea of this algorithm is that blocks are read partially (line 9)
-to ensure that the memory buffer only contains contiguous bytes, at
-any time. Therefore, the buffer can be written continuously to the
-reconstructed image, without seeking (line 13).  It should be noted
-that a given block might be read multiple times, in different memory
-loads.
+to ensure that the memory buffer only contains contiguous
+bytes. Therefore, the buffer can be written continuously to the
+reconstructed image, without seeking (line 13).  However, a given
+block might be read multiple times, in different memory loads.
 
@@
   \begin{algorithmic}[1]
-  \STATE sorted\_splits = sort splits by increasing (z,y,x) values
+  \STATE sorted\_blocks = sort blocks by increasing (k,j,i)
   \STATE start\_index = 0 ; end\_index=(m-1)
@@
     \STATE initialize buffer
-    \FOR{split in sorted\_splits}
+    \FOR{split in sorted\_blocks}
       \IF{split has voxels in write\_range}
@@
 
-As illustrated in Figure~\ref{fig:multiple-reads-cases}, we assume that $m'$ represents an integer number $k$
-of block rows (case 1, $k<\sqrt[3]{n}$), of complete rows (case 2,
-$k<d$), of tile rows (case 3, $k<\sqrt[3]{n}$), of slices (case 4,
-$k<d$) or of block slices (case 5, $k<\sqrt[3]{n}$). In each of these 5 cases, we define $v$ as follows:
+As illustrated in Figure~\ref{fig:multiple-reads-cases}, we assume
+that $m'$ represents an integer number $k$ of sub-rows (case 1,
+$k_1<\sqrt[3]{n}$), of complete rows (case 2, $k_2<d$), of tile rows
+(case 3, $k_3<\sqrt[3]{n}$), of slices (case 4, $k_4<d$) or of block
+slices (case 5, $k_5<\sqrt[3]{n}$). In each of these 5 cases, we
+define $v_i$ as follows:
 \begin{equation*}
@@
 \end{eqnarray*}
-where $x_i$ is the total number of required memory loads, $b_i$ is the
+where $x_i$ is the total number of memory loads, $b_i$ is the
 number of blocks accessed by the first ($x_i$-1) memory loads and
@@
 \begin{itemize}
-\item In MR, no seeking withing the splits.
+\item In MR, no seeking withing the blocks.
 \item Avoid in-memory copies.
@@
 
-For faster reading all the splits, before the merging process start,
+For faster reading all the blocks, before the merging process start,
 we read all the split’s headers to get their position information and
@@
 System~\cite{shvachko2010hadoop}. In this context parallel split-and-merge algorithms
-will be required, since the various splits of a large image would be
+will be required, since the various blocks of a large image would be
 uploaded to different disks concurrently. In the same vein,
@@
 
+% Library is available. We recommend using xxx for most cases. Same
+% algo works for blocks or slices.
+
 % Ideas for the future: measure the actual amount of consumed memory.

### 2017-08-01 paper.tex
Updates in analysis and implementation

@@
 \usepackage{hyperref}
+\hypersetup{colorlinks=true,linkcolor=black,citecolor=black,filecolor=black,urlcolor=blue}
 \usepackage{mathtools}
@@
 
+\subsection{Analysis}
+
+\todo{Make more simulations to support that claim.}
+
+Based on a few simulations of Equations~\ref{eq:seeks-cluster-reads-1}
+and~\ref{eq:seeks-multiple-reads}, we conjectured that the comparison
+between Cluster reads and Multiple reads essentially depends on the
+three Cluster reads cases. When Cluster reads merge incomplete block
+rows (case 1), they are always outperformed by Multiple reads
+(\textbf{prop 1}).  When Cluster reads merge complete block rows (case
+2), the comparison between Cluster reads and Multiple reads depends on
+the values of $D$ and $n$ (\textbf{prop 2}). When Cluster reads merge complete
+block slices (case 3), Cluster reads and Multiple reads are equivalent
+(\textbf{prop 3}).
+
+\textbf{prop 1} \todo{do it!}
+
+ \textbf{prop 2} is proven by Figure~\ref{fig:model-comparison} that
+ shows two situations where Cluster reads and Multiple reads are
+ alternatively better than each other.
+
+ \textbf{prop 3} is easy to infer from
+ equations~\ref{eq:seeks-cluster-reads-1}
+ and~\ref{eq:seeks-multiple-reads} when assuming that $Rb$ is an exact
+ multiple of $m$. In this case, both equations boils down to
+ $n+\frac{Rb}{m}$.
+
+\begin{figure}
+  \includegraphics[width=0.45\columnwidth]{figures/model-big-brain-comparison.pdf}
+  \includegraphics[width=0.45\columnwidth]{figures/model-big-brain-comparison-moreblocks.pdf}
+  \caption{Number of seeks for Cluster reads vs Multiple reads when
+    Cluster reads merge complete block rows (case 2), for D=3000 and
+    b=2. Left: n=125; Right: n=64,000.}
+  % Gnuplot file: scripts/model/model.gnplt
+  \label{fig:model-comparison}
+\end{figure}
+
+\section{Implementation}
+\label{sec:implementation}
+
+The 5 algorithms presented earlier are implemented in a Python library
+called \texttt{sam} (for ``split and merge''), available at
+\url{https://github.com/big-data-lab-team/-bin-sam} under MIT
+license. It uses Nibabel~\cite{matthew_brett_2016_60808} for image I/O
+and NumPy for array manipulations.
+
+% buffer implementation
+The data buffer used in Cluster reads and Multiple reads is
+implemented as a Python dictionary where the keys are offsets in the
+reconstructed image and the values are NumPy arrays containing the
+data starting at this offset. When the memory load is complete,
+dictionary entries are written sequentially to the reconstructed
+image. In Cluster reads, some seeking might be required between
+writes. We implemented a defragmentation procedure for the dictionary
+that merges contiguous dictionary entries in a single one, but we
+abandoned it as it proved more time-consuming than going through all
+the initial entries, due to the overhead of resizing NumPy arrays to
+merge entries. In Multiple reads, dictionary entries are always
+contiguous in the reconstructed image.  We tried to use a single NumPy
+array as a buffer, but we finally abandoned it as inserting data at a
+specific position in a NumPy array copies the data in memory, which
+increases the runtime and the peak memory consumption.
+
+% Split methods
+The implementation of splitting algorithms was greatly facilitated by
+the availability of so-called array proxies in Nibabel, which help to
+read specific sub-parts of large images. Nibabel's array proxies
+essentially provided the buffer implementation for splitting
+algorithms. Unfortunately, they are not available to write
+data. \todo{Yongping, Valerie: any other comment on split
+  implementations?}
+
+% Additional optimizations.
+In Multiple reads, for faster reading all the blocks, before the
+merging process start, we read all the block headers to get their
+position information and calculate each corresponding row index in the
+reconstructed image. We only read the block when there is a row in
+that writing range (available memory), otherwise, we skip it.
+\todo{Valerie, Yongping, any other implementation trick to report?}
+
 \subsection{Lossless compression}
@@
 
-\subsection{Implementation}
-\label{sec:implementation}
-
-Our algorithms are implemented in Python using the Nibabel
-library. \todo{Val\'erie and Yongping: (1) extract ImageUtils from
-  mapreduce repository (preserve git history; just copy the repo and
-  remove the mapreduce stuff in the copy). Link to the github
-  repo. (2) Describe the code briefly, in particular the
-  implementation tricks used for each implementation.}
-
-\begin{itemize}
-\item In MR, no seeking withing the blocks.
-\item Avoid in-memory copies.
-\end{itemize}
-
-\subsubsection{Cluster reads}
-
-\subsubsection{Multiple reads}
-
-For faster reading all the blocks, before the merging process start,
-we read all the split’s headers to get their position information and
-calculate each row’s index in the reconstructed image. We only read the
-split when there is a row in that writing range (available memory),
-otherwise, we skip it.
-
-Initially, we use a numpy array as our data buffer, after the split is read,
-we assign the rows (a numpy array of unsigned integers) into this
-buffer to the right position. However, after some tests, we found that
-there is some overhead. Each time we assign the data to the corresponding
-position of the buffer, the system will copy these integers first.
-This is neither time-efficient nor space-efficient.
-
-We then think of a better solution which we use the dictionary in python
-as our data buffer. We use the row’s index as a key, the data (in bytes)
-as the value. As for the multiple reads, the data buffer only contains
-contiguous bytes, we can just seek once (the first key of the sorted keys
-in dictionary), then write all the bytes continuously to the file.
-There is no copy in the memory, which proves a bit faster than using
-a numpy array as the buffer.
-
-\todo{Yongping: among other implementation techniques (e.g. dictionary
-  vs list, etc), explain clearly and concisely what the calculation
-  time covers.}
-
-\section{Simulations}
-
-\todo{Make more simulations to support that claim.}
-
-After running a few simulations, we saw that the comparison between
-Cluster reads and Multiple reads essentially depends on the three
-Cluster reads configurations.  When Cluster reads merge incomplete
-block rows (case 1), they are always outperformed by Multiple reads.
-When Cluster reads merge complete block rows (case 2), the comparison
-between Cluster reads and Multiple reads depends on the values of $D$
-and $n$. Figure~\ref{fig:model-comparison} shows two situations where
-Cluster reads and Multiple reads are alternatively better than each
-other. When Cluster reads merge complete block slices (case 3),
-Cluster reads and Multiple reads are equivalent.
-
-\begin{figure}
-  \includegraphics[width=0.45\columnwidth]{figures/model-big-brain-creads.pdf}
-  \includegraphics[width=0.45\columnwidth]{figures/model-big-brain-mreads.pdf}
-  \caption{Number of seeks for Cluster reads (Left) and Multiple reads (Right), for
-    D=3000, b=2 and n=125.}
-  % Gnuplot file: scripts/model/model.gnplt
-  \label{fig:model-individual}
-\end{figure}
-
-\begin{figure}
-  \includegraphics[width=0.45\columnwidth]{figures/model-big-brain-comparison.pdf}
-  \includegraphics[width=0.45\columnwidth]{figures/model-big-brain-comparison-moreblocks.pdf}
-  \caption{Number of seeks for Cluster reads vs Multiple reads when
-    Cluster reads merge complete block rows (case 2), for D=3000 and
-    b=2. Left: n=125; Right: n=64,000.}
-  % Gnuplot file: scripts/model/model.gnplt
-  \label{fig:model-comparison}
-\end{figure}
-
-\section{Benchmark}
-\label{sec:benchmark}
+\section{Experiments}
+\label{sec:results}
 
@@
 %smartctl --xall /dev/sda
-
-\section{Experiments}
-\label{sec:results}
 

### 2017-08-02 paper.tex
Inserted @ValHayot's results on SSD. Updates in text.

@@
   reads minimize the accesses to the image. Evaluation on a
-  3000x3000x3000 brain image shows that our algorithms perform
+  3850x3025x3500 brain image shows that our algorithms perform
   similarly to the optimal configuration (slice merging) provided that
@@
 
-\todo{Make more simulations to support that claim.}
-
 Based on a few simulations of Equations~\ref{eq:seeks-cluster-reads-1}
@@
   \caption{Number of seeks for Cluster reads vs Multiple reads when
-    Cluster reads merge complete block rows (case 2), for D=3000 and
+    Cluster reads merge complete block rows (case 2), for D=3458 and
     b=2. Left: n=125; Right: n=64,000.}
@@
 
-\subsection{Infrastructure}
+\subsection{Hardware}
 
@@
   histogram for breakdown, on same figure than total time.}
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=0.45\columnwidth]{figures/blocks-slices/totalmergetimeSSD1.pdf}
-  \hfill
-  \includegraphics[width=0.45\columnwidth]{figures/blocks-slices/totalmergetimeHDD1.pdf}\\
-  \includegraphics[width=0.45\columnwidth]{figures/blocks-slices/hddBreakDown1.pdf}
-  \hfill
-  \includegraphics[width=0.45\columnwidth]{figures/blocks-slices/ssdBreakDown1.pdf}
-  \caption{Merging time for Slices and Blocks. Left column: SSD. Right column: HDD. Top row: overall merging time. Bottom row: breakdown by read, write and seek time.}
-  % Gnuplot file: scripts/blocks-slices/createboxplot.gnplt
-\label{fig:slices-vs-blocks}
-\end{figure}
 
@@
 
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=0.45\columnwidth]{figures/benchmark-buff-slices/buff-slices-breakdown-ssd.pdf}
+  \hfill
+    \includegraphics[width=0.45\columnwidth]{figures/benchmark--buff-slices/buff-slices-breakdown-hdd.pdf}
+  \caption{Comparison between Bufered slices, naive blocks and naive slices. Left: SSD. Right: HDD.}
+  % Python file: scripts/experiment/generate_avg_var_data.py
+  % Gnuplot file: scripts/experiment/compare_breakdown_errbar.gnuplot
+\label{fig:buffered-slices}
+\end{figure}
+
 \subsection{Cluster reads}
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=0.45\columnwidth]{figures/benchmark-creads/creads-breakdown-ssd.pdf}
+  \hfill
+    \includegraphics[width=0.45\columnwidth]{figures/benchmark-creads/creads-breakdown-hdd.pdf}
+  \caption{Comparison between Cluster reads, naive blocks and naive slices. Left: SSD. Right: HDD.}
+  % Python file: scripts/experiment/generate_avg_var_data.py
+  % Gnuplot file: scripts/experiment/compare_breakdown_errbar.gnuplot
+\label{fig:cluster-reads}
+\end{figure}
 
@@
     \includegraphics[width=0.45\columnwidth]{figures/benchmark-mreads/mreads-breakdown-hdd.pdf}
-  \caption{Comparison between Multiple reads, blocks and slices. Left: SSD. Right: HDD.}
-  % Python file: scripts/mreads/generate_avg_var_data.py
-  % Gnuplot file: scripts/mreads/compare_breakdown_errbar.gnuplot
+  \caption{Comparison between Multiple reads, naive blocks and naive slices. Left: SSD. Right: HDD.}
+  % Python file: scripts/experiment/generate_avg_var_data.py
+  % Gnuplot file: scripts/experiment/compare_breakdown_errbar.gnuplot
 \label{fig:multiple-reads}
@@
   \centering
-  \includegraphics[width=0.70\columnwidth]{figures/benchmark-mreads/mreads-number-of-seeks.pdf}
+  \includegraphics[width=0.70\columnwidth]{figures/number-of-seeks.pdf}
   \hfill
-  \caption{Number of seeks in Multiple reads}
-  % Gnuplot file: scripts/mreads/number_of_seeks.gnuplot
-\label{fig:multiple-reads-number-of-seek}
+  \caption{Number of seeks for all algorithms, comparing modeled and
+    experimental values.}
+  % Gnuplot file: scripts/number_of_seeks.gnuplot
+\label{fig:number-of-seeks}
 \end{figure}

### 2017-08-02 generate_figures.sh

@@ -3,10 +3,11 @@
 set -e
 set -u
 
+scripts/experiment/generate_avg_var_data.py
 figures/svg/export_svg_files.sh
-scripts/model/model.gnplt
-scripts/mreads/number_of_seeks.gnuplot
-scripts/mreads/compare_breakdown_errbar.gnuplot
+scripts/model/model.gnplt &> data/seeks-model.dat
+scripts/experiment/number_of_seeks.gnuplot # don't run this before the model as the model generates data to plot
+scripts/experiment/compare_breakdown_errbar.gnuplot
 scripts/mwrites/compare_breakdown_errbar.gnuplot
 #crash
 #scripts/blocks-slices/createboxplot.gnplt

### 2017-08-02 scripts/experiment/generate_avg_var_data.py

@@ -1,12 +1,20 @@
 #!/usr/bin/env python
 
 input_file_mreads_ssd = "./data/mreads/mreads_ssd.dat"
-input_file_naive_ssd = "./data/blocks-slices/totaltime.dat"
 output_file_mreads_ssd = "./data/mreads/mreads_ssd_avg_var.dat"
-
 input_file_mreads_hdd = "./data/mreads/mreads_hdd.dat"
-input_file_naive_hdd = "./data/blocks-slices/totaltime.dat"
 output_file_mreads_hdd = "./data/mreads/mreads_hdd_avg_var.dat"
+
+input_file_creads_ssd = "./data/creads/creads_ssd.dat"
+output_file_creads_ssd = "./data/creads/creads_ssd_avg_var.dat"
+input_file_creads_hdd = "./data/creads/creads_hdd.dat"
+output_file_creads_hdd = "./data/creads/creads_hdd_avg_var.dat"
+
+input_file_buff_slices_ssd = "./data/buff-slices/buff-slices_ssd.dat"
+output_file_buff_slices_ssd = "./data/buff-slices/buff-slices_ssd_avg_var.dat"
+input_file_buff_slices_hdd = "./data/buff-slices/buff-slices_hdd.dat"
+output_file_buff_slices_hdd = "./data/buff-slices/buff-slices_hdd_avg_var.dat"
+
 
 def avg(l):
     return sum(l) / float(len(l))
@@ -19,7 +27,7 @@ def var(l):
         s2 += i
     return (float(s1) / len(l) - (float(s2) / len(l)) ** 2) ** 0.5
 
-def get_avg_var_mreads_naive(input_file_mreads, input_file_naive, output_file, disk):
+def get_avg_var_naive(input_file, output_file, input_file_naive_blocks, input_file_naive_slices, hasnaive, disk):
     read_times_3g = []
     write_times_3g = []
     seek_times_3g = []
@@ -56,15 +64,19 @@ def get_avg_var_mreads_naive(input_file_mreads, input_file_naive, output_file, d
     calculation_times_blocks = []
     total_time_blocks = []
 
-    with open(input_file_mreads, 'r') as f:
+    offset=0
+    if hasnaive:
+        offset=5
+        
+    with open(input_file, 'r') as f:
         for line in f:
             if '#' not in line:
                 items = line.split(' ')
                 # 3g
-                read_time = float(items[0])
-                write_time = float(items[1])
-                seek_time = float(items[2])
[…]

### 2017-08-02 paper.tex
Minor updates

@@
   \centering
-  \includegraphics[width=0.70\columnwidth]{figures/number-of-seeks.pdf}
+  \includegraphics[width=\columnwidth]{figures/number-of-seeks.pdf}
   \hfill
-  \caption{Number of seeks for all algorithms, comparing modeled and
-    experimental values.}
+  \caption{Number of seeks for all algorithms. y scale is logarithmic. Each algorithm is
+    represented with a different color. Dark color is experimental
+    value; bright color is model.}
   % Gnuplot file: scripts/number_of_seeks.gnuplot

### 2017-08-02 paper.tex
Wrote in Section IV (Experiments)

@@
 increasingly found in a variety of disciplines. Big~Brain, for
-instance, is a 3D histological image of the human brain that
-represents 1~TB of data organized in 3600 planes at full resolution
-(1$\mu$m in-plane), and 76~GB at a 40$\mu$m isotropic resolution
+\todo{summarize sizes} instance, is a 3D histological image of the human brain that
+represents 1~TB of raw data organized in 3600 planes at full resolution
+and 76~GB at a 40-micrometer isotropic resolution
 commonly used in neurosciences~\cite{amunts2013bigbrain}. This dataset
@@
 Neuroimaging Informatics Technology Initiative
-NIfTI\footnote{\url{https://nifti.nimh.nih.gov}} is such a format,
+NIfTI (\url{https://nifti.nimh.nih.gov}) is such a format,
 used by the vast majority of the neuroimaging community.
@@
 \begin{equation}
-  N_\mathrm{slices} = 2n 
+  N_\mathrm{slices} = 2n \label{eq:naive-blocks}
 \end{equation}
@@
 \begin{equation}
-N_\mathrm{blocks} = n+n\left(\sqrt[3]{\frac{R}{n}}\right)^2
+N_\mathrm{blocks} = n+n\left(\sqrt[3]{\frac{R}{n}}\right)^2 \label{eq:naive-slices}
 \end{equation}
@@
 \begin{equation}
-N_\mathrm{buff\_slices} =  n + \ceil*{\frac{bR}{m}}
+N_\mathrm{buff\_slices} =  n + \ceil*{\frac{bR}{m}} \label{eq:buff-slices}
 \end{equation}
@@
 \subsection{Analysis}
-
+\label{sec:analysis}
 Based on a few simulations of Equations~\ref{eq:seeks-cluster-reads-1}
@@
 The 5 algorithms presented earlier are implemented in a Python library
-called \texttt{sam} (for ``split and merge''), available at
-\url{https://github.com/big-data-lab-team/-bin-sam} under MIT
-license. It uses Nibabel~\cite{matthew_brett_2016_60808} for image I/O
+called \texttt{sam} (for ``split and merge''). It uses Nibabel~\cite{matthew_brett_2016_60808} for image I/O
 and NumPy for array manipulations.
@@
 \section{Experiments}
-\label{sec:results}
+\label{sec:experiments}
 
 \subsection{Data}
-To benchmark our library we use the 76~GB Big~Brain image split in 125
-non-overlapping chunks. The data is available \todo{[here]}.
+We used the 3850x3025x3500 Big~Brain image split in 125
+non-overlapping chunks of size 770x605x700 with 2 bytes per voxel
+(total size, uncompressed is 75.92~GB). Big~Brain is a reference brain
+based on the reconstruction of 7404 histological sections at nearly
+cellular resolution of 20 micrometers~\cite{amunts2013bigbrain}. It is
+a freely, publicly available tool with numerous applications in
+neurosciences and neurosurgery.
+
+We used the blocks of the 2015 Big~Brain release with 40-micrometer
+isotropic resolution available at
+\url{ftp://bigbrain.loris.ca/BigBrainRelease.2015/3D\_Blocks/40um}. We
+converted them to NiFTI 1.0 using Nibabel and left them uncompressed.
+
+\todo{add slices}
 
@@
 
-We run experiments using a Dell Precision Tower 3620 with CentOS Linux release
-7.3.1611, 32~GB of RAM and two disks:
-\begin{itemize}
-  \item Hard disk drive (HDD): HGST Travelstar 7K1000, 7200~rpm, 931GiB (1TB), firmware version JB0OA3W0.
-    \item Solid-state drive (SSD): SanDisk X400 2.5, 238GiB (256GB),
-      firmware version X4130012.
-\end{itemize}
-Both drives used 512-byte logical sectors, 4096-byte physical sectors,
-SATA >3.1 (6.0 Gb/s) and were accessed through the XFS file system
-v4.5.0.
+We used a Dell Precision Tower 3620 workstation with CentOS Linux
+release 7.3.1611, 32~GB of RAM and two disks. (1) a Hard disk drive
+(HDD): HGST Travelstar 7K1000, 7200~rpm, 931GiB (1TB), firmware
+version JB0OA3W0; (2) a Solid-state drive (SSD): SanDisk X400 2.5,
+238GiB (256GB), firmware version X4130012.  Both drives used 512-byte
+logical sectors, 4096-byte physical sectors, SATA \textgreater 3.1 (6.0 Gb/s) and
+were accessed through the XFS file system v4.5.0. We used
+\texttt{iotop} (\url{http://guichaz.free.fr/iotop}) to monitor I/Os on
+the workstation and make sure that no other process was compromising
+our measures.
 %smartctl --xall /dev/sda
@@
 
-We used release xxx of our library.
+We used Git tag 0.1 of our \texttt{sam} library. Our experiment
+scripts are available at
+\url{https://github.com/big-data-lab-team/paper-sequential-split-merge/blob/master/scripts/experiment}. We
+used them to split and merge using Buffered slices, Cluster reads and
+Multiple reads, with 3GB, 6GB, 9GB, 12GB and 16GB of memory. We also
+did 0GB of memory for Buffered slices and Cluster reads, which
+triggered naive slice and block merging. We did 5 repetitions for each
+memory value. Memory values were shuffled in each repetition, to
+correct for potential biases coming from ordering, such as caching
+effects. To improve reproducibility, we dropped the kernel page,
+dentry and inode caches before each run (\texttt{echo 3 | sudo tee
+  /proc/sys/vm/drop\_caches}).
 
-Experiment scripts are available at xxx.
-As our experiments results are heavily rely on IO performance of the disk,
-we drop the system cache to get more precise result before each experiment.
+\subsection{Results}
 
-\subsection{Slices vs Blocks}
+\subsubsection{Merging time}
 
+% For each paragraph below, comment on overal trend and breakdown.
+
+% Slices vs blocks
 Figure~\ref{fig:slices-vs-blocks} shows the merging time for Slices vs
@@
 
-\subsection{Buffered slices}
+% Buffered slices
+
+% Cluster reads
+%% ultimately fine but not stable due to discontinuities
+At 3GB, Cluster reads were in case 1, which explains that more than 10
+million seeks were done, almost equivalent to naive blocks. At 6, 9
+and 12~GB, Cluster reads were in case 2 and at 16~GB they were in case
+3. This confirms that Cluster reads are beneficial in case 2 and 3
+only, consistently with the analysis in Section~\ref{sec:analysis} .
+
+% Multiple reads
+%% more stable
+
+\begin{itemize}
+\item Comment on calculation time.
+\item Write time is higher than read time because use nibabel for writing but not reading.
+\item It seems that the merge time is quite stable w.r.t the amount of available memory. This is good news for multiple reads. 
+\end{itemize}
 
@@
 
-\subsection{Cluster reads}
-
 \begin{figure}[h]
@@
 \end{figure}
-
-\subsection{Multiple reads}
 
@@
 
+\subsubsection{Seeks}
+
+The number of seeks is reported in Figure~\ref{fig:number-of-seeks},
+for all algorithms and the corresponding models
+(Equations~\ref{eq:naive-blocks}
+to~\ref{eq:seeks-multiple-reads}). Note the logarithmic y scale. Error
+bars are not reported as numbers of seeks were constant across all
+repetitions. The mean model error is \todo{x\%}, explained by the fact
+that the model assumes cubic blocks while we used non-cubic ones in
+the experiment. Overall, the model is validated as it correctly
+explains the observations.   The measured seek times correlated with
+the measured merging times ($\rho$=\todo{x}), which suggest that the
+observed time reductions were indeed caused by reduced seek times.
 \begin{figure}[h]
@@
 
-Results discussion:
-\begin{itemize}
-\item Comment on calculation time.
-\item Write time is higher than read time because use nibabel for writing but not reading.
-\item It seems that the merge time is quite stable w.r.t the amount of available memory. This is good news for multiple reads. 
-\end{itemize}
+% correlated to merging time
 
-\subsection{Cluster reads vs Multiple reads}
-
-\subsection{Lossless Compression}
+% model error
 
@@
 
-\subsubsection{Multiple writes}
+% Time
+
+% Number of seeks
+
 \begin{figure}[h]
@@
 initial problem is solved.
+
+% Library is available. 
+% overall, multiple reads is recommended as default parameters for the
+% library. Cluster reads could be useful in some cases.
+
+%Same algo works for blocks or slices.
 
@@
 
-% Library is available. We recommend using xxx for most cases. Same
-% algo works for blocks or slices.
+The complete Latex sources of this paper, including the collected data
+and Gnuplot scripts used to generate the results figures are available
+at \url{https://github.com/big-data-lab-team/paper-sequential-split-merge} under
+GPLv3 license. Our \texttt{sam} library is at
+\url{https://github.com/big-data-lab-team/sam}, under MIT license.
 

### 2017-08-03 generate_figures.sh
merge total time figures to one figure

@@ -4,9 +4,11 @@ set -e
 set -u
 
 scripts/experiment/generate_avg_var_data.py
+scripts/experiment/generate_total_avg_var_data.py # for total merge time figure
 figures/svg/export_svg_files.sh
 scripts/model/model.gnplt &> data/seeks-model.dat
 scripts/experiment/number_of_seeks.gnuplot # don't run this before the model as the model generates data to plot
+scripts/experiment/merge_total_time.gnuplot
 scripts/experiment/compare_breakdown_errbar.gnuplot
 scripts/mwrites/compare_breakdown_errbar.gnuplot
 #crash

### 2017-08-03 paper.tex
Worked on results

@@
 \usepackage{stmaryrd} % for llbracket and rrbracket
+\usepackage{subcaption}
 \DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}
@@
 Section~\ref{sec:algos} presents our algorithms and their complexity,
-Section~\ref{sec:results} details the results and
+Section~\ref{sec:experiments} details the results and
 Sections~\ref{sec:discussion} discusses them.\todo{Revise when outline is final.}
@@
 
-\subsection{Lossless compression}
+%% \subsection{Lossless compression}
 
-Lossless compression might reduce overall read and write times by
-reducing seek time and data sizes. However, it needs to be implemented
-on-the-fly, i.e., in chunks. See the work in~\cite{rajna2015speeding}
-and
-\href{https://github.com/pauldmccarthy/indexed\_gzip}{corresponding
-  implementation}. This is straightforward to apply for random reads
-but not for random writes. Multiple reads is the only merging strategy
-that works with literally no seek in the reconstructed image which
-could therefore potentially use in-memory compression.
+%% Lossless compression might reduce overall read and write times by
+%% reducing seek time and data sizes. However, it needs to be implemented
+%% on-the-fly, i.e., in chunks. See the work in~\cite{rajna2015speeding}
+%% and
+%% \href{https://github.com/pauldmccarthy/indexed\_gzip}{corresponding
+%%   implementation}. This is straightforward to apply for random reads
+%% but not for random writes. Multiple reads is the only merging strategy
+%% that works with literally no seek in the reconstructed image which
+%% could therefore potentially use in-memory compression.
 
-\todo{could we merge without de-compressing?}
+%% \todo{could we merge without de-compressing?}
 
@@
 converted them to NiFTI 1.0 using Nibabel and left them uncompressed.
-
-\todo{add slices}
+We generated NiFTI slices of size 3850x3025x28 from the reconstructed
+image, using our naive split algorithm. Slices are available at
+\todo{\url{x}}.
 
@@
 used them to split and merge using Buffered slices, Cluster reads and
-Multiple reads, with 3GB, 6GB, 9GB, 12GB and 16GB of memory. We also
-did 0GB of memory for Buffered slices and Cluster reads, which
-triggered naive slice and block merging. We did 5 repetitions for each
-memory value. Memory values were shuffled in each repetition, to
-correct for potential biases coming from ordering, such as caching
-effects. To improve reproducibility, we dropped the kernel page,
-dentry and inode caches before each run (\texttt{echo 3 | sudo tee
-  /proc/sys/vm/drop\_caches}).
+Multiple reads, with 3~GB, 6~GB, 9~GB, 12~GB and 16~GB of
+memory. Table~\ref{table:configs} shows the configurations of Cluster
+reads and Multiple reads for each memory value, according to
+Equations~\ref{eq:seeks-cluster-reads-1}
+and~\ref{eq:seeks-multiple-reads}. For instance, Cluster reads are in
+case 1 for 3~GB. We also did 0~GB of memory for Buffered slices and
+Cluster reads, which triggered naive slice and block merging. We did 5
+repetitions for each memory value. Memory values were shuffled in each
+repetition, to correct for potential biases coming from ordering, such
+as caching effects. To improve reproducibility, we dropped the kernel
+page, dentry and inode caches before each run (\texttt{echo 3 | sudo
+  tee /proc/sys/vm/drop\_caches}). We measured the cumulative read,
+write and seek time in each run, as well as the overhead time defined
+as the total time minus the sum of all other times.
+\begin{table}
+  \centering
+\begin{footnotesize}
+\begin{tabular}{|c|ccccc|}
+  \hline
+                 & 3~GB & 6~GB & 9~GB & 12~GB & 16~GB\\
+  \hline
+  Cluster  reads &  1 &  2 &  2 &  2 &  3 \\
+  Multiple reads &  4 &  4 &  4 &  4 &  5\\
+  \hline
+\end{tabular}
+\end{footnotesize}
+\caption{Algorithm configurations by memory values.}
+\label{table:configs}
+\end{table}
 
 \subsection{Results}
-
-\subsubsection{Merging time}
-
-% For each paragraph below, comment on overal trend and breakdown.
-
-% Slices vs blocks
-Figure~\ref{fig:slices-vs-blocks} shows the merging time for Slices vs
-Blocks, on SSD and HDD. Merging with Blocks takes about twice the time
-of Slices, which motivates the need for improved algorithms to merge
-blocks. As expected, seek time is mainly responsible for this
-difference, although differences in read times are also observed,
-perhaps due to a caching effect at the disk or OS level. This behavior is
-reproduced in SSD and HDD, which shows that SSDs might not be the only
-solution to reduce seek time.  \todo{On figure: check that breakdowns
-  add up to total time. Harmonize time formats. Harmonize colors (all
-  slices in purple, all blocks in green). If time permits, try
-  histogram for breakdown, on same figure than total time.}
-
-% Buffered slices
-
-% Cluster reads
-%% ultimately fine but not stable due to discontinuities
-At 3GB, Cluster reads were in case 1, which explains that more than 10
-million seeks were done, almost equivalent to naive blocks. At 6, 9
-and 12~GB, Cluster reads were in case 2 and at 16~GB they were in case
-3. This confirms that Cluster reads are beneficial in case 2 and 3
-only, consistently with the analysis in Section~\ref{sec:analysis} .
-
-% Multiple reads
-%% more stable
-
-\begin{itemize}
-\item Comment on calculation time.
-\item Write time is higher than read time because use nibabel for writing but not reading.
-\item It seems that the merge time is quite stable w.r.t the amount of available memory. This is good news for multiple reads. 
-\end{itemize}
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=0.45\columnwidth]{figures/benchmark-buff-slices/buff-slices-breakdown-ssd.pdf}
-  \hfill
-    \includegraphics[width=0.45\columnwidth]{figures/benchmark--buff-slices/buff-slices-breakdown-hdd.pdf}
-  \caption{Comparison between Bufered slices, naive blocks and naive slices. Left: SSD. Right: HDD.}
-  % Python file: scripts/experiment/generate_avg_var_data.py
-  % Gnuplot file: scripts/experiment/compare_breakdown_errbar.gnuplot
-\label{fig:buffered-slices}
-\end{figure}
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=0.45\columnwidth]{figures/benchmark-creads/creads-breakdown-ssd.pdf}
-  \hfill
-    \includegraphics[width=0.45\columnwidth]{figures/benchmark-creads/creads-breakdown-hdd.pdf}
-  \caption{Comparison between Cluster reads, naive blocks and naive slices. Left: SSD. Right: HDD.}
-  % Python file: scripts/experiment/generate_avg_var_data.py
-  % Gnuplot file: scripts/experiment/compare_breakdown_errbar.gnuplot
-\label{fig:cluster-reads}
-\end{figure}
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=0.45\columnwidth]{figures/benchmark-mreads/mreads-breakdown-ssd.pdf}
-  \hfill
-    \includegraphics[width=0.45\columnwidth]{figures/benchmark-mreads/mreads-breakdown-hdd.pdf}
-  \caption{Comparison between Multiple reads, naive blocks and naive slices. Left: SSD. Right: HDD.}
-  % Python file: scripts/experiment/generate_avg_var_data.py
-  % Gnuplot file: scripts/experiment/compare_breakdown_errbar.gnuplot
-\label{fig:multiple-reads}
-\end{figure}
 
@@
 the experiment. Overall, the model is validated as it correctly
-explains the observations.   The measured seek times correlated with
-the measured merging times ($\rho$=\todo{x}), which suggest that the
-observed time reductions were indeed caused by reduced seek times.
+explains the observations.  
 \begin{figure}[h]
@@
 
-% correlated to merging time
+As expected, the difference between naive blocks and naive slices is
+tremendous, in the order of 50 million seeks. For 3~GB, the number of
+seeks in Cluster reads is 4 orders of magnitude higher than for
+Multiple reads, and 5 orders of magnitude higher than naive
+slices. This huge difference comes from the fact that at 3~GB Cluster
+reads are in case 1. For 6~GB, 9~GB and 12~GB, Cluster reads are in
+case 2 and the difference with Multiple reads and Buffered slices
+reduces. At 16~GB, all algorithms perform the same.
 
-% model error
+\subsubsection{Merge time}
 
-\subsection{Splitting}
+\todo{ The measured seek times correlated with
+the measured merging times ($\rho$=x), which suggest that the
+observed time reductions were indeed caused by reduced seek times.}
+
+% Slices and blocks
+Figure~\ref{fig:merge-time} shows the merge time for all algorithms by
+memory values. Naive blocks take about 5 times longer than naive
+slices to merge the data, which quantifies the effect of the problem
+targeted by our algorithms. This effect appears both on HDD and SSD,
+which shows that SSDs might not be the only solution to reduce seek
+time. In the remainder we use naive blocks and naive slices and
+references to evaluate our algorithms.
+
+% Buffered slices
+Buffered slices provide a negligible speed-up compared to naive
+slices. In most cases, their memory overhead would not be worth the
+time gain.
+
+% Cluster reads
+Cluster reads provide important speed ups compared to naive blocks,
+both on HDD and on SSD. On SSD, they are up to 5 times faster than
+naive blocks at 16~GB, and 3 on average \todo{compute the actual
+  values}. \todo{On HDD, \ldots} Surprisingly, they perform
+substantially faster than naive blocks even at 3~GB, while in case
+1. This may be explained by the fact that the seeks required to write
+incomplete block rows to the reconstructed image are shorter than the
+ones for naive blocks.
+
+% Multiple reads
+Multiple reads are even faster than Cluster reads. On SSD, they are 5
+times faster than naive blocks on average \todo{compute the actual
+  values} ( \todo{On HDD, \ldots}). At 12~GB, the total merging time
+is only 1.2 times slower than naive slices, which solves our initial
+problem to a large extent. Even at 3~GB the speed-up is considerable
+compared to naive blocks.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=\columnwidth]{figures/times-hdd.pdf}\\
+  \includegraphics[width=\columnwidth]{figures/times-ssd.pdf}
+  \hfill
+  \caption{Merge time by algorithm. Top: HDD. Bottom: SSD. Each
+    algorithm is represented with a different color. Averages over 5
+    repetitions. Error bars show $\pm$ 1 standard deviation. }
+  % Gnuplot file: scripts/experiment/number_of_seeks.gnuplot
+\label{fig:merge-time}
+\end{figure}
+
+\subsubsection{Merge time breakdown}
+Figure~\ref{fig:breakdowns-ssd} 
+\begin{figure}
+  \begin{subfigure}[b]{\columnwidth}
+    \includegraphics[width=0.45\columnwidth]{figures/benchmark-buff-slices/buff-slices-breakdown-ssd.pdf}
+    \includegraphics[width=0.45\columnwidth]{figures/benchmark-buff-slices/buff-slices-breakdown-hdd.pdf}
+    \caption{Buffered slices}
+  \end{subfigure}
+  \begin{subfigure}[b]{\columnwidth}
+    \includegraphics[width=0.45\columnwidth]{figures/benchmark-creads/creads-breakdown-ssd.pdf}
+    \includegraphics[width=0.45\columnwidth]{figures/benchmark-creads/creads-breakdown-hdd.pdf}
+    \caption{Cluster reads}
+  \end{subfigure}
+  \begin{subfigure}[b]{\columnwidth}
+    \includegraphics[width=0.45\columnwidth]{figures/benchmark-mreads/mreads-breakdown-ssd.pdf}
+    \includegraphics[width=0.45\columnwidth]{figures/benchmark-mreads/mreads-breakdown-hdd.pdf}
+    \caption{Multiple reads}
+  \end{subfigure}
+  \caption{Breakdown of total merge times. Left column: SSD. Right column: HDD.}
+  \label{fig:breakdowns-ssd}
+  % Gnuplot file: scripts/experiment/compare_breakdown_errbar.gnuplot
+\end{figure}
+
+\subsubsection{Split time}
 
@@
   \centering
-  \includegraphics[width=0.45\columnwidth]{figures/benchmark-mwrites/mwrites-breakdown-ssd.pdf}
+  \includegraphics[width=\columnwidth]{figures/number-of-seeks.pdf}\\
+  \includegraphics[width=\columnwidth]{figures/number-of-seeks.pdf}
   \hfill
-    \includegraphics[width=0.45\columnwidth]{figures/benchmark-mwrites/mwrites-breakdown-hdd.pdf}
-  \caption{Comparison between Multiple writes, blocks and slices. Left: SSD. Right: HDD.}
-  % Gnuplot file: scripts/blocks-slices/createboxplot.gnplt
-\label{fig:multiple-writes}
+  \caption{Split time by algorithm. Top: HDD. Bottom: SSD. Each
+    algorithm is represented with a different color. Averages over 5
+    repetitions. Error bars show $\pm$ 1 standard deviation. }
+  % Gnuplot file: ???
+\label{fig:split-time}
 \end{figure}
-
-\newpage
 
@@
 
-The complete Latex sources of this paper, including the collected data
-and Gnuplot scripts used to generate the results figures are available
-at \url{https://github.com/big-data-lab-team/paper-sequential-split-merge} under
-GPLv3 license. Our \texttt{sam} library is at
+For more investigation and reproducibility purposes, the complete \LaTeX
+sources of this paper, including the collected data and Gnuplot
+scripts used to generate the results figures are available at
+\url{https://github.com/big-data-lab-team/paper-sequential-split-merge}
+under GPLv3 license. Our \texttt{sam} library is at
 \url{https://github.com/big-data-lab-team/sam}, under MIT license.

### 2017-08-03 scripts/experiment/generate_avg_var_data.py

@@ -17,9 +17,13 @@ output_file_buff_slices_hdd = "./data/buff-slices/buff-slices_hdd_avg_var.dat"
 
 
 def avg(l):
+    if len(l) == 0:
+        return 0
     return sum(l) / float(len(l))
-
+    
 def var(l):
+    if len(l) == 0:
+        return 0
     s1 = 0
     s2 = 0
     for i in l:
@@ -177,11 +181,13 @@ def get_avg_var_naive(input_file, output_file, input_file_naive_blocks, input_fi
 def main():
     # mreads
     get_avg_var_naive(input_file_mreads_ssd, output_file_mreads_ssd, input_file_creads_ssd, input_file_buff_slices_ssd, hasnaive=False, disk="ssd")
-    #get_avg_var_naive(input_file_mreads_hdd, output_file_mreads_hdd, input_file_creads_hdd, input_file_buff_slices_hdd, hasnaive=False, disk="hdd")
+    get_avg_var_naive(input_file_mreads_hdd, output_file_mreads_hdd, input_file_creads_hdd, input_file_buff_slices_hdd, hasnaive=False, disk="hdd")
     # creads
     get_avg_var_naive(input_file_creads_ssd, output_file_creads_ssd, input_file_creads_ssd, input_file_buff_slices_ssd, hasnaive=True, disk="ssd")
+    get_avg_var_naive(input_file_creads_hdd, output_file_creads_hdd, input_file_creads_hdd, input_file_buff_slices_hdd, hasnaive=True, disk="hdd")
     # buffered slicse
     get_avg_var_naive(input_file_buff_slices_ssd, output_file_buff_slices_ssd, input_file_creads_ssd, input_file_buff_slices_ssd, hasnaive=True, disk="ssd")
+    get_avg_var_naive(input_file_buff_slices_hdd, output_file_buff_slices_hdd, input_file_creads_hdd, input_file_buff_slices_hdd, hasnaive=True, disk="hdd")
 
 if __name__ == '__main__':
     main()

### 2017-08-03 scripts/experiment/generate_total_avg_var_data.py
Added total merge time for HDD

@@ -1,8 +1,12 @@
 #!/usr/bin/env python
 def avg(l):
+    if len(l) == 0:
+        return 0
     return sum(l) / float(len(l))
 
 def var(l):
+    if len(l) == 0:
+        return 0
     s1 = 0
     s2 = 0
     for i in l:
@@ -102,6 +106,9 @@ def generate(input_file_mreads, input_file_creads, input_file_buff_slices, outpu
 def main():
     output_file = "./data/total-merge-time-ssd.dat"
     generate(input_file_mreads_ssd, input_file_creads_ssd, input_file_buff_slices_ssd, output_file=output_file)
+    output_file = "./data/total-merge-time-hdd.dat"
+    generate(input_file_mreads_hdd, input_file_creads_hdd, input_file_buff_slices_hdd, output_file=output_file)
 
+    
 if __name__ == '__main__':
     main()

### 2017-08-03 paper.tex
Added total merge time for HDD
Worked on split results

@@
   \centering
-  \includegraphics[width=\columnwidth]{figures/times-hdd.pdf}\\
-  \includegraphics[width=\columnwidth]{figures/times-ssd.pdf}
+  \includegraphics[width=\columnwidth]{figures/total-merge-time-hdd.pdf}\\
+  \includegraphics[width=\columnwidth]{figures/total-merge-time-ssd.pdf}
   \hfill
@@
     repetitions. Error bars show $\pm$ 1 standard deviation. }
-  % Gnuplot file: scripts/experiment/number_of_seeks.gnuplot
+  % Gnuplot file: scripts/experiment/split_and_merge_total_time.gnuplot
 \label{fig:merge-time}
@@
 \subsubsection{Merge time breakdown}
-Figure~\ref{fig:breakdowns-ssd} 
+Figure~\ref{fig:breakdowns-ssd} shows how the total merge time breaks
+down to read, write, seek and overhead time for our algorithms. Naive
+blocks and naive slices are shown as references. The huge difference
+between naive blocks and naive slices is coming from both the seek
+time and the write time, which suggests that seeking degrades the
+write rate importantly in addition to introducing extra delays.
+Cluster reads reduce the seek time and thus the read time very
+substantially. Multiple reads almost anihilate the seek time and bring
+the read time to a value comparable to naive slices.  The same
+behavior is observed on HDD and on SSD, although the effect of seeking
+is slightly lower on SSD, as expected. Read times are consistently and substantially lower than write times\todo{explain why}. The overhead time is small for
+both Cluster reads and Multiple reads, although it is not negligible
+for Multiple reads on HDD \todo{Overhead was not well measured for
+  mreads. Needs to be redone.}.
 \begin{figure}
   \begin{subfigure}[b]{\columnwidth}
+    \includegraphics[width=0.45\columnwidth]{figures/benchmark-buff-slices/buff-slices-breakdown-hdd.pdf}
     \includegraphics[width=0.45\columnwidth]{figures/benchmark-buff-slices/buff-slices-breakdown-ssd.pdf}
-    \includegraphics[width=0.45\columnwidth]{figures/benchmark-buff-slices/buff-slices-breakdown-hdd.pdf}
     \caption{Buffered slices}
@@
   \begin{subfigure}[b]{\columnwidth}
+    \includegraphics[width=0.45\columnwidth]{figures/benchmark-creads/creads-breakdown-hdd.pdf}
     \includegraphics[width=0.45\columnwidth]{figures/benchmark-creads/creads-breakdown-ssd.pdf}
-    \includegraphics[width=0.45\columnwidth]{figures/benchmark-creads/creads-breakdown-hdd.pdf}
     \caption{Cluster reads}
@@
   \begin{subfigure}[b]{\columnwidth}
+    \includegraphics[width=0.45\columnwidth]{figures/benchmark-mreads/mreads-breakdown-hdd.pdf}
     \includegraphics[width=0.45\columnwidth]{figures/benchmark-mreads/mreads-breakdown-ssd.pdf}
-    \includegraphics[width=0.45\columnwidth]{figures/benchmark-mreads/mreads-breakdown-hdd.pdf}
     \caption{Multiple reads}
   \end{subfigure}
-  \caption{Breakdown of total merge times. Left column: SSD. Right column: HDD.}
+  \caption{Breakdown of total merge times. Left column: HDD. Right column: SSD.}
   \label{fig:breakdowns-ssd}
@@
 
-% Time
+\todo{define cluster writes and multiple writes somewhere.}
 
-% Number of seeks
+The number of seeks for Cluster writes and Multiple writes were
+identical to the ones in Cluster reads and Multiple reads. \todo{Check
+  that.} The total split time by algorithm is shown in
+Figure~\ref{fig:split-time} and the breakdown by read, write, seek and
+overhead time is in Figure~\ref{fig:breakdowns-ssd-split}. \todo{comment on the figures when results are available.}
 
@@
   \centering
-  \includegraphics[width=\columnwidth]{figures/number-of-seeks.pdf}\\
-  \includegraphics[width=\columnwidth]{figures/number-of-seeks.pdf}
+  \includegraphics[width=\columnwidth]{figures/total-split-time-hdd.pdf}\\
+  \includegraphics[width=\columnwidth]{figures/total-split-time-ssd.pdf}
   \hfill
@@
     repetitions. Error bars show $\pm$ 1 standard deviation. }
-  % Gnuplot file: ???
+  % Gnuplot file: scripts/experiment/split_and_merge_total_time.gnuplot
 \label{fig:split-time}
+\end{figure}
+
+\begin{figure}
+  \begin{subfigure}[b]{\columnwidth}
+    \includegraphics[width=0.45\columnwidth]{figures/benchmark-buff-slices/buff-slices-breakdown-hdd-split.pdf}
+    \includegraphics[width=0.45\columnwidth]{figures/benchmark-buff-slices/buff-slices-breakdown-ssd-split.pdf}
+    \caption{Buffered slices}
+  \end{subfigure}
+  \begin{subfigure}[b]{\columnwidth}
+    \includegraphics[width=0.45\columnwidth]{figures/benchmark-cwrite/cwrites-breakdown-hdd.pdf}
+    \includegraphics[width=0.45\columnwidth]{figures/benchmark-cwrites/cwrites-breakdown-ssd.pdf}
+    \caption{Cluster writes}
+  \end{subfigure}
+  \begin{subfigure}[b]{\columnwidth}
+    \includegraphics[width=0.45\columnwidth]{figures/benchmark-mwrites/mwrites-breakdown-hdd.pdf}
+    \includegraphics[width=0.45\columnwidth]{figures/benchmark-mwrites/mwrites-breakdown-ssd.pdf}
+    \caption{Multiple reads}
+  \end{subfigure}
+  \caption{Breakdown of total split times. Left column: HDD. Right column: SSD.}
+  \label{fig:breakdowns-ssd-split}
+  % Gnuplot file: scripts/experiment/compare_breakdown_errbar.gnuplot
 \end{figure}
@@
 \section*{Acknowledgment}
-
-(If not added as co-authors): Greg Kiar, Pierre Bellec, Claude Lepage, Lindsay B. Lewis.
+We warmly thank Lindsay B. Lewis and Claude Lepage for helping us with
+Big~Brain, Greg Kiar for useful discussion about the Open
+Connectome Data Cluster and Pierre Bellec for discussions about MINC.
 

### 2017-08-03 generate_figures.sh

@@ -8,9 +8,6 @@ scripts/experiment/generate_total_avg_var_data.py # for total merge time figure
 figures/svg/export_svg_files.sh
 scripts/model/model.gnplt &> data/seeks-model.dat
 scripts/experiment/number_of_seeks.gnuplot # don't run this before the model as the model generates data to plot
-scripts/experiment/merge_total_time.gnuplot
+scripts/experiment/split_and_merge_total_time.gnuplot
 scripts/experiment/compare_breakdown_errbar.gnuplot
-scripts/mwrites/compare_breakdown_errbar.gnuplot
-#crash
-#scripts/blocks-slices/createboxplot.gnplt
 

### 2017-08-03 scripts/experiment/generate_total_avg_var_data.py

@@ -14,6 +14,7 @@ def var(l):
         s2 += i
     return (float(s1) / len(l) - (float(s2) / len(l)) ** 2) ** 0.5
 
+# merge
 #ssd
 input_file_mreads_ssd = "./data/mreads/mreads_ssd.dat"
 input_file_creads_ssd = "./data/creads/creads_ssd.dat"
@@ -23,6 +24,17 @@ input_file_buff_slices_ssd = "./data/buff-slices/buff-slices_ssd.dat"
 input_file_mreads_hdd = "./data/mreads/mreads_hdd.dat"
 input_file_creads_hdd = "./data/creads/creads_hdd.dat"
 input_file_buff_slices_hdd = "./data/buff-slices/buff-slices_hdd.dat"
+
+# split
+#ssd
+input_file_mwrites_ssd = "./data/mwrites/mwrites_ssd.dat"
+input_file_cwrites_ssd = "./data/cwrites/cwrites_ssd.dat"
+input_file_buff_slices_split_ssd = "./data/buff-slices/buff-slices_split_ssd.dat"
+
+#hdd
+input_file_mwrites_hdd = "./data/mwrites/mwrites_hdd.dat"
+input_file_cwrites_hdd = "./data/cwrites/cwrites_hdd.dat"
+input_file_buff_slices_split_hdd = "./data/buff-slices/buff-slices_split_hdd.dat"
 
 
 
@@ -108,7 +120,11 @@ def main():
     generate(input_file_mreads_ssd, input_file_creads_ssd, input_file_buff_slices_ssd, output_file=output_file)
     output_file = "./data/total-merge-time-hdd.dat"
     generate(input_file_mreads_hdd, input_file_creads_hdd, input_file_buff_slices_hdd, output_file=output_file)
+    output_file = "./data/total-split-time-ssd.dat"
+    generate(input_file_mwrites_ssd, input_file_cwrites_ssd, input_file_buff_slices_split_ssd, output_file=output_file)
+    output_file = "./data/total-split-time-hdd.dat"
+    generate(input_file_mwrites_hdd, input_file_cwrites_hdd, input_file_buff_slices_split_hdd, output_file=output_file)
 
-    
+  
 if __name__ == '__main__':
     main()

### 2017-08-03 scripts/mwrites/generate_avg_var_data.py

@@ -1,193 +1,0 @@
-#!/usr/bin/env python
-
-input_file_mwrites_ssd = "./data/mwrites/mwrites_ssd.dat"
-# TODO - benchmark naive split
-input_file_naive_ssd = ""
-output_file_mwrites_ssd = "./data/mwrites/mwrites_ssd_avg_var.dat"
-
-input_file_mwrites_hdd = "./data/mwrites/mwrites_hdd.dat"
-# TODO - benchmark naive split
-input_file_naive_hdd = ""
-output_file_mwrites_hdd = "./data/mwrites/mwrites_hdd_avg_var.dat"
-
-def avg(l):
-    return sum(l) / float(len(l))
-
-def var(l):
-    s1 = 0
-    s2 = 0
-    for i in l:
-        s1 += i ** 2
-        s2 += i
-    return (float(s1) / len(l) - (float(s2) / len(l)) ** 2) ** 0.5
-
-def get_avg_var_mwrites_naive(input_file_mwrites, input_file_naive, output_file, disk):
-    read_times_3g = []
-    write_times_3g = []
-    seek_times_3g = []
-    calculation_times_3g = []
-    total_time_3g = []
-
-    read_times_6g = []
-    write_times_6g = []
-    seek_times_6g = []
-    calculation_times_6g = []
-    total_time_6g = []
-
-    read_times_9g = []
-    write_times_9g = []
-    seek_times_9g = []
-    calculation_times_9g = []
-    total_time_9g = []
-
-    read_times_12g = []
-    write_times_12g = []
-    seek_times_12g = []
-    calculation_times_12g = []
-    total_time_12g = []
-
-    read_times_slices = []
-    write_times_slices = []
-    seek_times_slices = []
-    calculation_times_slices = []
-    total_time_slices = []
-
-    read_times_blocks = []
-    write_times_blocks = []
-    seek_times_blocks = []
-    calculation_times_blocks = []
-    total_time_blocks = []
-
-    with open(input_file_mwrites, 'r') as f:
-        for line in f:
-            if '#' not in line:
-                items = line.split(' ')
-                # 3g
-                read_time = float(items[0])
-                write_time = float(items[1])
-                seek_time = float(items[2])
-                total_time = float(items[4])
-                calculation_time = total_time - seek_time - write_time - read_time
-                read_times_3g.append(read_time)
-                write_times_3g.append(write_time)
-                seek_times_3g.append(seek_time)
[…]

### 2017-08-06 scripts/experiment/benchmark_merge.py
rerun mreads benchmark - fix overhead/readtime and add mem=16g

@@ -1,6 +1,6 @@
 #!/usr/bin/env python
 # Ref: imageutils.py from
-# https://github.com/big-data-lab-team/mapreduce/blob/master/python/imageutils.py
+# https://github.com/big-data-lab-team/sam/blob/master/imageutils.py
 import imageutils as img_utils
 import numpy as np
 from time import time
@@ -9,8 +9,8 @@ import random
 import os
 
 # example
-# ./benchmark_mreads.py -m 3221225472 9663676416 -r 5 -d ssd
-# ./benchmark_mreads.py -m 3221225472 9663676416 -r 5 -d hdd
+# ./benchmark_merge.py -m 3221225472 9663676416 -r 5 -d ssd
+# ./benchmark_merge.py -m 3221225472 9663676416 -r 5 -d hdd
 
 
 # on the consider
@@ -18,11 +18,11 @@ reconstructed_hdd = "/data/gao/new_image.nii"
 legend_hdd = "/data/gao/blocks125/legend.txt"
 
 reconstructed_ssd = "/home/gao/new_image.nii"
-legend_ssd = "/home/v_hayots/Documents/Thesis/mapreduce/python/bigbrain_blocks2/legend.txt"
+legend_ssd = "/home/gao/blocks125/legend.txt"
 
 
-csv_file_hdd = "./hdd.csv"
-csv_file_ssd = "./ssd.csv"
+csv_file_hdd = "./hdd.dat"
+csv_file_ssd = "./ssd.dat"
 
 
 first_dim=3850
@@ -37,13 +37,13 @@ files = {
 def benchmark_mreads(mem, reconstructed, legend):
     img = img_utils.ImageUtils(reconstructed, first_dim, second_dim, third_dim, np.uint16)
     s_time = time()
-    total_read_time, total_write_time, total_seek_time, total_seek_number = img.reconstruct_img(legend, "multiple", mem)
+    total_read_time, total_write_time, total_seek_time, total_seek_number = img.reconstruct_img(legend, "multiple", mem, benchmark=True)
     total_time = time() - s_time
     return (total_read_time, total_write_time, total_seek_time, total_seek_number, total_time)
 
-def write_to_file(data_dict, csv_file):
+def write_to_file(data_dict, dat_file):
     # (total_read_time, total_write_time, total_seek_time, total_seek_number, total_time)
-    with open(csv_file, "a") as f:
+    with open(dat_file, "a") as f:
         for k in sorted(data_dict.keys()):
[…]

### 2017-08-06 scripts/experiment/benchmark_split.py

@@ -1,6 +1,6 @@
 #!/usr/bin/env python
 # Ref: imageutils.py from
-# https://github.com/big-data-lab-team/mapreduce/blob/master/python/imageutils.py
+# https://github.com/big-data-lab-team/sam/blob/master/imageutils.py
 import imageutils as img_utils
 import numpy as np
 from time import time
@@ -9,15 +9,15 @@ import random
 import os
 
 # example
-# ./benchmark_mwrites.py -m 3221225472 9663676416 6442450944 13043807040 -r 5 -d ssd
-# ./benchmark_mwrites.py -m 3221225472 9663676416 6442450944 13043807040 -r 5 -d hdd
+# ./benchmark_split.py -m 3221225472 9663676416 6442450944 13043807040 -r 5 -d ssd
+# ./benchmark_split.py -m 3221225472 9663676416 6442450944 13043807040 -r 5 -d hdd
 
 
 # on the consider
 ori_image_hdd = "/data/bigbrain_40microns.nii"
 out_dir_hdd = "/data/gao/blocks_split"
 
-ori_image_ssd = "/home/gao/reconstructed.nii"
+ori_image_ssd = "/home/gao/new_image.nii"
 out_dir_ssd = "/home/gao/blocks125"
 
 csv_file_hdd = "./hdd_mwrites.csv"
@@ -43,11 +43,11 @@ def benchmark_mwrites(mem, ori_image, out_dir):
 
     return (total_read_time, total_write_time, total_seek_time, total_seek_number, total_time)
 
-def write_to_csv(data_dict, csv_file):
-    with open(csv_file, "a") as f:
+def write_to_csv(data_dict, dat_file):
+    with open(dat_file, "a") as f:
         for k in sorted(data_dict.keys()):
             for e in data_dict[k]:
-                f.write(str(e) + ",")
+                f.write(str(e) + " ")
         f.write("\n")
 
 

### 2017-08-06 scripts/experiment/generate_total_avg_var_data.py

@@ -41,7 +41,6 @@ input_file_buff_slices_split_hdd = "./data/buff-slices/buff-slices_split_hdd.dat
 
 
 def generate(input_file_mreads, input_file_creads, input_file_buff_slices, output_file):
-    row_06g, row_3g, row_6g, row_9g, row_12g, row_16g = [],[],[],[],[],[]
     total_time_06g, total_time_3g, total_time_6g, total_time_9g, total_time_12g, total_time_16g = [], [], [], [], [], []
 
     with open(input_file_creads, 'r') as f:
@@ -79,9 +78,11 @@ def generate(input_file_mreads, input_file_creads, input_file_buff_slices, outpu
                 total_time_9g.append(float(items[14]))
                 # 12g - mr
                 total_time_12g.append(float(items[19]))
+                # 16g - mr
+                total_time_16g.append(float(items[24]))
 
-        mr_total = [0, avg(total_time_3g), avg(total_time_6g), avg(total_time_9g), avg(total_time_12g), 0]
-        mr_total_err = [0, var(total_time_3g), var(total_time_6g), var(total_time_9g), var(total_time_12g), 0]
+        mr_total = [0, avg(total_time_3g), avg(total_time_6g), avg(total_time_9g), avg(total_time_12g), avg(total_time_16g)]
+        mr_total_err = [0, var(total_time_3g), var(total_time_6g), var(total_time_9g), var(total_time_12g), var(total_time_16g)]
 
     total_time_06g, total_time_3g, total_time_6g, total_time_9g, total_time_12g, total_time_16g = [], [], [], [], [], []
     with open(input_file_buff_slices, 'r') as f:
@@ -121,10 +122,8 @@ def main():
     output_file = "./data/total-merge-time-hdd.dat"
     generate(input_file_mreads_hdd, input_file_creads_hdd, input_file_buff_slices_hdd, output_file=output_file)
     output_file = "./data/total-split-time-ssd.dat"
-    generate(input_file_mwrites_ssd, input_file_cwrites_ssd, input_file_buff_slices_split_ssd, output_file=output_file)
+    # generate(input_file_mwrites_ssd, input_file_cwrites_ssd, input_file_buff_slices_split_ssd, output_file=output_file)
     output_file = "./data/total-split-time-hdd.dat"
-    generate(input_file_mwrites_hdd, input_file_cwrites_hdd, input_file_buff_slices_split_hdd, output_file=output_file)
-
-  
[…]

### 2017-08-06 scripts/experiment/generate_total_avg_var_data.py
minor fix

@@ -107,8 +107,15 @@ def generate(input_file_mreads, input_file_creads, input_file_buff_slices, outpu
 
     mem = ["0.6", "3", "6", "9", "12", "16"]
     with open(output_file, 'w+') as f:
-        f.write("mem \"Naive blocks\" NBR \"Naive slices\" NSR \"Cluster reads\" CRR \"Multiple reads\" MRR \"Buffered slices \" BSR")
-        f.write('\n')
+        if "merge" in output_file:
+            f.write("mem \"Naive blocks\" NBR \"Naive slices\" NSR \"Cluster reads\" CRR \"Multiple reads\" MRR \"Buffered slices \" BSR")
+            f.write('\n')
+        elif "split" in output_file:
+            f.write("mem \"Naive blocks\" NBR \"Naive slices\" NSR \"Cluster writes\" CRR \"Multiple writes\" MRR \"Buffered slices \" BSR")
+            f.write('\n')
+        else:
+            raise ValueError("output file should contains \"split\" or \"merge\"")
+
         for i in range(0,6):
             f.write(
                 "{0} {1} {2} {3} {4} {5} {6} {7} {8} {9} {10}"
@@ -116,14 +123,17 @@ def generate(input_file_mreads, input_file_creads, input_file_buff_slices, outpu
                             ,mr_total[i], mr_total_err[i], bs_total[i], bs_total_err[i]))
             f.write('\n')
 
+
+
 def main():
     output_file = "./data/total-merge-time-ssd.dat"
     generate(input_file_mreads_ssd, input_file_creads_ssd, input_file_buff_slices_ssd, output_file=output_file)
     output_file = "./data/total-merge-time-hdd.dat"
     generate(input_file_mreads_hdd, input_file_creads_hdd, input_file_buff_slices_hdd, output_file=output_file)
-    output_file = "./data/total-split-time-ssd.dat"
-    # generate(input_file_mwrites_ssd, input_file_cwrites_ssd, input_file_buff_slices_split_ssd, output_file=output_file)
-    output_file = "./data/total-split-time-hdd.dat"
-    # generate(input_file_mwrites_hdd, input_file_cwrites_hdd, input_file_buff_slices_split_hdd, output_file=output_file)
+    #crash - add split data in cwrites/buff-slices
+    #output_file = "./data/total-split-time-ssd.dat"
[…]

### 2017-08-08 scripts/experiment/benchmark_split.py
remeasure mwrites and add mem=16g

@@ -20,8 +20,8 @@ out_dir_hdd = "/data/gao/blocks_split"
 ori_image_ssd = "/home/gao/new_image.nii"
 out_dir_ssd = "/home/gao/blocks125"
 
-csv_file_hdd = "./hdd_mwrites.csv"
-csv_file_ssd = "./ssd_mwrites.csv"
+csv_file_hdd = "./hdd_mwrites.dat"
+csv_file_ssd = "./ssd_mwrites.dat"
 
 
 Y_splits=5

### 2017-08-08 paper.tex
Updates in results and simulations

@@
 \label{sec:analysis}
-Based on a few simulations of Equations~\ref{eq:seeks-cluster-reads-1}
-and~\ref{eq:seeks-multiple-reads}, we conjectured that the comparison
-between Cluster reads and Multiple reads essentially depends on the
-three Cluster reads cases. When Cluster reads merge incomplete block
-rows (case 1), they are always outperformed by Multiple reads
-(\textbf{prop 1}).  When Cluster reads merge complete block rows (case
-2), the comparison between Cluster reads and Multiple reads depends on
-the values of $D$ and $n$ (\textbf{prop 2}). When Cluster reads merge complete
-block slices (case 3), Cluster reads and Multiple reads are equivalent
-(\textbf{prop 3}).
 
-\textbf{prop 1} \todo{do it!}
-
- \textbf{prop 2} is proven by Figure~\ref{fig:model-comparison} that
- shows two situations where Cluster reads and Multiple reads are
- alternatively better than each other.
-
- \textbf{prop 3} is easy to infer from
- equations~\ref{eq:seeks-cluster-reads-1}
- and~\ref{eq:seeks-multiple-reads} when assuming that $Rb$ is an exact
- multiple of $m$. In this case, both equations boils down to
- $n+\frac{Rb}{m}$.
+Figure~\ref{fig:model-comparison} plots
+Equations~\ref{eq:seeks-cluster-reads-1}
+and~\ref{eq:seeks-multiple-reads} for different values of $n$. When
+Cluster reads are in Case 1 or 2, they may outperform Multiple reads
+only for large values of $n$. When Cluster reads are in case 3, they
+are equivalent to Multiple reads; indeed, assuming that $Rb$ is an
+exact multiple of $m$, Equations~\ref{eq:seeks-cluster-reads-1}
+and~\ref{eq:seeks-multiple-reads} both boil down to $n+\frac{Rb}{m}$.
 
@@
   \includegraphics[width=0.45\columnwidth]{figures/model-big-brain-comparison-moreblocks.pdf}
-  \caption{Number of seeks for Cluster reads vs Multiple reads when
-    Cluster reads merge complete block rows (case 2), for D=3458 and
-    b=2. Left: n=125; Right: n=64,000.}
+  \caption{Number of seeks for Cluster reads vs Multiple reads, for
+    D=3458 and b=2. Left: n=125; Right: n=64,000. Case 1 and Case 2
+    denote Cluster read configurations.}
   % Gnuplot file: scripts/model/model.gnplt
@@
 both Cluster reads and Multiple reads, although it is not negligible
-for Multiple reads on HDD \todo{Overhead was not well measured for
-  mreads. Needs to be redone.}.
+for Multiple reads on HDD.
 \begin{figure}
@@
     \includegraphics[width=0.45\columnwidth]{figures/benchmark-mwrites/mwrites-breakdown-ssd.pdf}
-    \caption{Multiple reads}
+    \caption{Multiple writes}
   \end{subfigure}

### 2017-08-08 scripts/experiment/generate_avg_var_data.py

@@ -56,6 +56,12 @@ def get_avg_var_naive(input_file, output_file, input_file_naive_blocks, input_fi
     calculation_times_12g = []
     total_time_12g = []
 
+    read_times_16g = []
+    write_times_16g = []
+    seek_times_16g = []
+    calculation_times_16g = []
+    total_time_16g = []
+    
     read_times_slices = []
     write_times_slices = []
     seek_times_slices = []
@@ -120,6 +126,17 @@ def get_avg_var_naive(input_file, output_file, input_file_naive_blocks, input_fi
                 seek_times_12g.append(seek_time)
                 calculation_times_12g.append(calculation_time)
                 total_time_12g.append(total_time)
+                # 16g
+                read_time = float(items[offset+20])
+                write_time = float(items[offset+21])
+                seek_time = float(items[offset+22])
+                total_time = float(items[offset+24])
+                calculation_time = total_time - seek_time - write_time - read_time
+                read_times_16g.append(read_time)
+                write_times_16g.append(write_time)
+                seek_times_16g.append(seek_time)
+                calculation_times_16g.append(calculation_time)
+                total_time_16g.append(total_time)
 
     # open naive block file:
     with open(input_file_naive_blocks, 'r') as f:
@@ -173,6 +190,9 @@ def get_avg_var_naive(input_file, output_file, input_file_naive_blocks, input_fi
         f.write('\n')
         f.write("12 {} {} {} {} {} ".format(avg(calculation_times_12g), avg(read_times_12g), avg(write_times_12g),
                                              avg(seek_times_12g), var(total_time_12g)))
+        f.write('\n')
+        f.write("16 {} {} {} {} {} ".format(avg(calculation_times_16g), avg(read_times_16g), avg(write_times_16g),
+                                             avg(seek_times_16g), var(total_time_16g)))
         f.write('\n')
         f.write("naive-slice {} {} {} {} {}".format(avg(calculation_times_slices), avg(read_times_slices),
                                                     avg(write_times_slices),

### 2017-08-08 scripts/experiment/generate_total_avg_var_data.py

@@ -99,11 +99,13 @@ def generate(input_file_mreads, input_file_creads, input_file_buff_slices, outpu
                 total_time_9g.append(float(items[19]))
                 # 12g - bs
                 total_time_12g.append(float(items[24]))
+                # 16g - bs
+                total_time_16g.append(float(items[24]))
 
         ns_total = [avg(total_time_06g), 0, 0, 0, 0, 0]
         ns_total_err = [var(total_time_06g), 0, 0, 0, 0, 0]
-        bs_total = [0, avg(total_time_3g), avg(total_time_6g), avg(total_time_9g), avg(total_time_12g), 0]
-        bs_total_err = [0, var(total_time_3g), var(total_time_6g), var(total_time_9g), var(total_time_12g), 0]
+        bs_total = [0, avg(total_time_3g), avg(total_time_6g), avg(total_time_9g), avg(total_time_12g), avg(total_time_16g)]
+        bs_total_err = [0, var(total_time_3g), var(total_time_6g), var(total_time_9g), var(total_time_12g), var(total_time_16g)]
 
     mem = ["0.6", "3", "6", "9", "12", "16"]
     with open(output_file, 'w+') as f:

### 2017-08-10 paper.tex
Updated abstract

@@
   perform very poorly to split or merge 3D blocks from ultra-high
-  resolution images, due to important seek times. We propose and
-  analyze two algorithms based on in-memory reconstruction of image
-  parts. Cluster reads access image chunks only once while Multiple
-  reads minimize the accesses to the image. Evaluation on a
-  3850x3025x3500 brain image shows that our algorithms perform
-  similarly to the optimal configuration (slice merging) provided that
-  enough memory is available, which solves our initial problem.  We
-  conclude that splitting and merging large 3D images can be done
-  efficiently based on simple data formats.
+  resolution images, due to important seek times. We introduce and
+  analyze sequential algorithms that leverage memory buffering, to
+  address this issue. Cluster reads and Cluster writes access image
+  chunks only once, but they have to seek in the reconstructed
+  image. Multiple reads and Multiple writes minimize seeks in the
+  reconstructed image, but they access image chunks multiple
+  times. Evaluation on a 3850x3025x3500 brain image shows that our
+  algorithms perform similarly to the optimal configuration (slice
+  merging) provided that enough memory is available, which solves our
+  initial problem to a large extent.  We conclude that splitting and
+  merging large 3D images can be done efficiently without the need for
+  complex data formats.
 \end{abstract}

### 2017-08-10 paper.tex
Updated introduction

@@
 increasingly found in a variety of disciplines. Big~Brain, for
-\todo{summarize sizes} instance, is a 3D histological image of the human brain that
-represents 1~TB of raw data organized in 3600 planes at full resolution
-and 76~GB at a 40-micrometer isotropic resolution
-commonly used in neurosciences~\cite{amunts2013bigbrain}. This dataset
-is being re-scanned at an even higher resolution, leading to 100~TB of
-raw data. Other examples found in medical imaging, our primary domain
-of interest, include high-resolution 3D electron microscopy (see,
-e.g., ~\cite{bock2011network}) or micro- and
+instance, is a 3D histological image of the
+human brain that represents 1~TB of raw data organized in 3600 planes
+at full resolution and 76~GB at a 40-micrometer isotropic resolution
+commonly used in neurosciences~\cite{amunts2013bigbrain}. Future
+versions of this dataset will be acquired at an even higher resolution
+and will represent 100~TB of raw data. Other examples found in medical
+imaging, our primary domain of interest, include high-resolution 3D
+electron microscopy (see, e.g., ~\cite{bock2011network}) or micro- and
 nano-tomography~\cite{10.1371/journal.pone.0035691}. As such images
 would typically be processed on a computing cluster, possibly using
-locality-aware file systems such as HDFS~\cite{shvachko2010hadoop},
-efficient software libraries are needed to split and merge them
-efficiently. In particular, data reads and writes might result in
-extensive seek times that drastically limit I/O performance.  In this
-paper we introduce and compare a family of sequential algorithms to
-split and merge images with minimal seek time.
+locality-aware file systems such as the Hadoop Distributed File System
+(HDFS~\cite{shvachko2010hadoop}), software libraries are needed to
+split and merge them efficiently, in particular to limit file
+seeks. In this paper we introduce and compare a family of sequential
+algorithms to split and merge images with limited numbers of seeks.
 
-Images are split into chunks representing 3D blocks or 2D slices,
-which we assume fit in memory. A dataset such as Big~Brain would
-perhaps be split into 125 chunks of 600~MB. The decision to split an
-image into slices or blocks, and the size of the resulting chunks, is
-usually done by the application. Some applications would commonly
-require blocks, for instance spatial filtering, while other ones such
-as acquisition artifact removal would rather work on
-slices. Applications processing voxels individually, for instance
+We assume that the high-resolution image is split into chunks
+representing 3D blocks or 2D slices that fit in memory. A dataset such
+as Big~Brain would perhaps be split into 125 chunks of 600~MB. The
+decision to split an image into slices or blocks, and the size of the
+chunks, is up to the application. Some applications, for instance spatial filtering, would
+commonly require blocks while other
+ones, such as acquisition artifact removal, would rather work on
+slices. Applications that process voxels individually, for instance
 histogram computation or k-means clustering, could work on either
@@
 
-Seek time is highly dependent on the byte ordering in image files,
-which we assume is arbitrary but known to the algorithm. The geometry
-of the chunks to split or merge (slices or blocks) and their sizes are
-also critical, which we again assume arbitrary but known to the
-algorithm.  The main idea of our algorithms will be to convert data
-orderings in memory before writing them to disk.
+We also assume that the byte organization in image files is arbitrary
+but known to the algorithm. Some formats, for instance the file format
+defined by the Neuroimaging Informatics Technology Initiative (NIfTI
+-- \url{https://nifti.nimh.nih.gov}) store the complete image in
+column-major order, that is, for an image of dimensions (i, j, k), i
+is the fastest changing dimension, followed by dimension j, followed
+by dimension k. Other formats, for instance HDF5-based MINC
+2.0~\cite{vincent2016minc}, provide more flexibility by allowing data
+to be partitioned in limited-sized chunks, each chunk being stored in
+a specific order. Byte organization is obviously a critical factor of
+seek time. The main idea of our algorithms will be to rearrange
+ordering in memory before or after I/O operations.
 
@@
 platforms~\cite{miguet1991elastic,tang2007eman2,yang2008parallel,braunl2013parallel,moise2013terabyte,bajcsy2013terabyte},
-but methods have focused more on geometrical approaches to partition
-images and on load-balancing and task scheduling techniques while we
+but methods have focused on geometrical approaches to partition images
+and on load-balancing and on task scheduling techniques. Instead, we
 aim at algorithms to efficiently split or merge images regardless of
-the geometry of the chunks. Seek times are, however, often identified
-as an issue but the preferred solution is usually to optimize data
-storage formats for a certain application while we are searching for
-algorithms that would reduce the seek time regardless of the data
-format. For instance, the Open Connectome Data
-Cluster~\cite{burns2013open} is a data warehouse system that allows
-users to retrieve specific chunks of large image datasets. It
-addresses seek times by designing a specific file format based on
-space-filling curves, which elegantly preserves spatial proximity on
-disk. Such specific data formats optimize the system for particular
-chunk geometries, blocks in this case, while we aim at algorithms to
-support multiple types of geometries.
+the geometry of the chunks. Although important seek times are often
+identified as an issue, the preferred solution is usually to optimize
+data storage formats for a certain application.  For instance, the
+Open Connectome Data Cluster~\cite{burns2013open} is a data warehouse
+system that allows users to retrieve specific 3D blocks from large
+image datasets. It reduces seek times through a specific file format
+based on space-filling curves, which elegantly preserves spatial
+proximity on disk. On the contrary, we are searching for algorithms
+that would reduce the seek time regardless of the data format, so that
+applications with flexible splitting schemes can be served
+
 %http://dl.acm.org/citation.cfm?id=218395 also minimizes seek time using a specific format.
+To summarize, our paper makes the following contributions. (1) We
+propose a set of algorithms to split and merge terabyte-scale 3D
+images from 3D blocks or 2D slices. (2) We determine the complexity of
+those algorithms in terms of numbers of seeks, as a function of the image size, splitting scheme and available memory.
+(3)We evaluate our algorithms using the Big~Brain dataset and two storage drives with different characteristics.
 
-This paper makes the following contributions:
-\begin{itemize}
-  \item We propose a set of algorithms to split and merge
-    terabyte-scale 3D images from blocks or slices.
-  \item We model the performance (number of seeks) of these algorithms.
-  \item We evaluate our model and algorithms on a representative
-    use-case involving a 1~TB image a hard drive and a solid-state
-    drive.
-\end{itemize}
-Section~\ref{sec:algos} presents our algorithms and their complexity,
-Section~\ref{sec:experiments} details the results and
-Sections~\ref{sec:discussion} discusses them.\todo{Revise when outline is final.}
+Section~\ref{sec:algos} presents our algorithms,
+Section~\ref{sec:implementation} describes their implementation,
+Section~\ref{sec:experiments} reports experimental results and
+Section~\ref{sec:discussion} concludes the paper.
 
@@
 We consider a file format where voxels are written in column-major
-order, that is, for an image of dimensions (i, j, k), i is the fastest
-changing dimension, followed by dimension j, followed by dimension
-k. All voxels in a \emph{slice} have a the same $k$ and all voxels in
-a \emph{row} have the same $j$.  NIfTI, the file format defined by the
-Neuroimaging Informatics Technology Initiative
-NIfTI (\url{https://nifti.nimh.nih.gov}) is such a format,
+order. All voxels in a \emph{slice} have a the same $k$ and all voxels in
+a \emph{row} have the same $j$.  NIfTI is such a format,
 used by the vast majority of the neuroimaging community.

### 2017-08-10 paper.tex
Updates in II.A and II.B

@@
 3D image with $R$ voxels of size $b$. For simplicity, we assume that
-all chunks are of identical size, that they do not overlap, and that
-slices are squares and blocks are cubes.
+chunks are non-overlapping cuboids that all have the same dimension.
 
-We consider a file format where voxels are written in column-major
-order. All voxels in a \emph{slice} have a the same $k$ and all voxels in
-a \emph{row} have the same $j$.  NIfTI is such a format,
-used by the vast majority of the neuroimaging community.
+Although our algorithms could be applied to any byte organization, we
+consider a file format where voxels are written in column-major
+order. All voxels in a \emph{slice} have the same $k$ and all voxels
+in a \emph{row} have the same $j$.
 
@@
 
-We adopt the following notations (see Figure\ref{fig:notations}):
+We adopt the following notations (see Figure~\ref{fig:notations}):
 \begin{itemize}
@@
 \item $m$: amount of available memory (in B).
-\item $m'$: amount of used memory (in B), $m'\leq m$.
+\item $m'<m$: amount of used memory (in B).
 \end{itemize}
@@
 \begin{itemize}
-\item Number of slices, rows, columns in a block: $\sqrt[3]{\frac{R}{n}}=d$.
+\item Number of slices/rows/columns in a block: $\sqrt[3]{\frac{R}{n}}=d$.
 \item Number of blocks in a block row: $\sqrt[3]{n}$.
@@
 
-A disk is characterized by its read and write rates and its seek
-time. We do not distinguish between access time and seek time. We also
-assume that seeks require a constant amount of time, regardless of the
-position seeked to. In practice, large variations would be expected,
-but modeling such variations would inevitably lead to models specific
-to the hardware, file system or operating system, which we
-intentionally avoid here. Likewise, in modern systems, read and write
-times are greatly impacted by caching occurring at several levels,
-which we do not model here. Thus, our goal is to find algorithms that
-minimize the \emph{number} of seek operations.
+A disk is characterized by its read and write rates, its access time
+and its seek time. For common file sizes, seek time is negligible
+compared to read or write time as typical seek times range from about 0.1~ms
+for Solid-State Drives (SSD) to 10~ms for Hard-Disk
+Drives (HDD). However, as we will shown later, naive algorithms might
+seek up to $10^7$ times to merge a high-resolution image, which
+renders seek time comparable to read and write times. In addition,
+extensive seeking also has an effect on read and write rates, as these
+are typically increasing with the duration of uninterrupted reads or
+writes.
 
-\todo{Add ballpark values of seek time (5ms) and read/write rates.}
+In our complexity analysis, we do not distinguish between access time
+and seek time.  We also assume that seeks require a constant amount of
+time, regardless of the position seeked to. That is, we focus on the
+average seek time. In practice, large variations would be expected
+depending on the seek distance, but modeling such variations would
+inevitably lead to models specific to the hardware, file system or
+operating system, which we intentionally avoid here. Likewise, in
+contemporary systems, read and write times are greatly impacted by caches
+operating at several levels, which we do not model here. Thus, our
+goal is to find algorithms that minimize the \emph{number} of seek and
+file access operations, which we denote ``number of seeks'' in the
+remainder.
 

### 2017-08-11 paper.tex
Updates in Sections II and III

@@
       \STATE slice = sorted\_slices[i]
-      \IF{\texttt{sizeof}(buffer)+\texttt{sizeof}(slice) $\geq$ m}
+      \IF{sizeof(buffer)+sizeof(slice) $\geq$ m}
         \STATE write buffer in reconstructed image
@@
 extension to block merging is not easy. The remainder of this Section
-presents our attempts for such a generalization.
+presents our attempts for such an extension.
 
@@
 buffer will be written without seeking. A given
-block is accessed only once.
+block is accessed only once during the whole merging process.
 
@@
   their rows ($d^2$ seeks). Blue blocks need seeking before each of
-  their slices ($d$ seeks in total). Green blocks need only a single
+  their slices ($d$ seeks). Green blocks need only a single
   seek. Grey, dashed, transparent blocks represent the contiguous
@@
   reads. Left: configuration with incomplete block rows (multiplies
-  number of seeks by $d$). Right: configuration with block rows that
-  overlap multiple block slices (multiplies number of seeks by $2$).}
+  the number of seeks by $d$ compared to complete block rows). Right:
+  configuration with block rows that overlap multiple block slices
+  (multiplies the number of seeks by $2$ compared to non-overlapping
+  configurations).}
 \label{fig:avoided-configurations}
@@
 \end{equation*}
-where $x_i$ is the number of memory loads required to reconstruct the
-image and $b_i$ is the number of seeks required to write the memory
-load. The first $n$ seeks in the equation are required to read all the
-blocks once. According to Figure~\ref{fig:cluster-reads}, we have:
+where $x_i$ is the number of required memory loads and $b_i$ is the
+number of seeks required to write a memory load. The first $n$ seeks
+in the equation correspond to the reading of all the blocks. According
+to Figure~\ref{fig:cluster-reads}, we have:
 \begin{equation*}
@@
 proportional to the total number of block rows in the image
-($\sqrt[3]{n}^2$) and $x_2$ is proportional to the total number of
-block slices ($\sqrt[3]{n}$).
-
+$\left( \sqrt[3]{n}^2 \right)$ and $x_2$ is proportional to the total number of
+block slices $\left( \sqrt[3]{n} \right)$.
 Finally, the total number of seeks
@@
     \STATE initialize buffer
-    \FOR{split in sorted\_blocks}
-      \IF{split has voxels in write\_range}
-        \STATE split\_data = read split
-        \STATE in split\_data, extract the rows in write\_range
+    \FOR{block in sorted\_blocks}
+      \IF{block has voxels in write\_range}
+        \STATE block\_data = read block
+        \STATE in block\_data, extract the rows in write\_range
         \STATE insert rows in buffer
@@
 
-As illustrated in Figure~\ref{fig:multiple-reads-cases}, we assume
-that $m'$ represents an integer number $k$ of sub-rows (case 1,
-$k_1<\sqrt[3]{n}$), of complete rows (case 2, $k_2<d$), of tile rows
-(case 3, $k_3<\sqrt[3]{n}$), of slices (case 4, $k_4<d$) or of block
-slices (case 5, $k_5<\sqrt[3]{n}$). In each of these 5 cases, we
-define $v_i$ as follows:
+In the complexity analysis, we assume that $m'$ represents an integer
+number $k$ of sub-rows (case 1, $k_1<\sqrt[3]{n}$), of complete rows
+(case 2, $k_2<d$), of tile rows (case 3, $k_3<\sqrt[3]{n}$), of slices
+(case 4, $k_4<d$) or of block slices (case 5, $k_5<\sqrt[3]{n}$), as
+illustrated in Figure~\ref{fig:multiple-reads-cases}. In each of these
+5 cases, we define $v_i$ as follows:
 \begin{equation*}
@@
 where $x_i$ is the total number of memory loads, $b_i$ is the
-number of blocks accessed by the first ($x_i$-1) memory loads and
+number of blocks accessed by the first $(x_i-1)$ memory loads and
 $b_i'$ is the number of blocks access by the last memory load. The
-first $x_i$ seeks in the equation are required to write each memory
-load (1 seek per memory load). We have:
+first $x_i$ seeks in the equation correspond to the writing of all memory
+loads (1 seek per memory load). We have:
 \begin{equation*}
@@
 \begin{equation*}
-b_1=k_1 ; \quad b_2=\sqrt[3]{n} ; \quad b_3 = k_3\sqrt[3]{n} ; \quad b_4=\sqrt[3]{n}^2 ; \quad b_5=k_5\sqrt[3]{n}^2
+b_1=k_1 ; b_2=\sqrt[3]{n} ; b_3 = k_3\sqrt[3]{n} ; b_4=\sqrt[3]{n}^2 ; b_5=k_5\sqrt[3]{n}^2
 \end{equation*}
@@
 only for large values of $n$. When Cluster reads are in case 3, they
-are equivalent to Multiple reads; indeed, assuming that $Rb$ is an
-exact multiple of $m$, Equations~\ref{eq:seeks-cluster-reads-1}
+are equivalent to Multiple reads: assuming that $Rb$ is an exact
+multiple of $m$, Equations~\ref{eq:seeks-cluster-reads-1}
 and~\ref{eq:seeks-multiple-reads} both boil down to $n+\frac{Rb}{m}$.
+\todo{We could be a bit more rigorous and detail the cases where
+  cluster reads of multiple reads will be better.}
 
@@
 
-The 5 algorithms presented earlier are implemented in a Python library
-called \texttt{sam} (for ``split and merge''). It uses Nibabel~\cite{matthew_brett_2016_60808} for image I/O
-and NumPy for array manipulations.
+We implemented the 5 algorithms presented earlier in a
+Python library called \texttt{sam} (for ``split and merge''). It uses
+Nibabel~\cite{matthew_brett_2016_60808} for image I/O and NumPy for
+array manipulations.
 
@@
 specific position in a NumPy array copies the data in memory, which
-increases the runtime and the peak memory consumption.
+increases both the execution time and the peak memory consumption.
 
@@
 The implementation of splitting algorithms was greatly facilitated by
-the availability of so-called array proxies in Nibabel, which help to
-read specific sub-parts of large images. Nibabel's array proxies
+the availability of so-called array proxies in Nibabel, which help
+reading specific sub-parts of large images. Nibabel's array proxies
 essentially provided the buffer implementation for splitting
 algorithms. Unfortunately, they are not available to write
-data. \todo{Yongping, Valerie: any other comment on split
-  implementations?}
+data.
 
 % Additional optimizations.
-In Multiple reads, for faster reading all the blocks, before the
-merging process start, we read all the block headers to get their
-position information and calculate each corresponding row index in the
-reconstructed image. We only read the block when there is a row in
-that writing range (available memory), otherwise, we skip it.
-\todo{Valerie, Yongping, any other implementation trick to report?}
+In Multiple reads, block headers are read in a first pass where row
+indices in the reconstructed image are stored in memory. Those indices
+are then processed in each memory load, to identify the blocks that
+contribute to it.
 

### 2017-08-11 paper.tex

@@
   analyze sequential algorithms that leverage memory buffering, to
-  address this issue. Cluster reads and Cluster writes access image
+  address this issue. Clustered reads and Clustered writes access image
   chunks only once, but they have to seek in the reconstructed
@@
 
-\subsection{Buffered blocks: Cluster reads}
+\subsection{Buffered blocks: Clustered reads}
 
-Cluster reads are the more direct extension of buffered slices to
+Clustered reads are the more direct extension of buffered slices to
 blocks: they load multiple blocks in memory, concatenate them in a
@@
 the amount of memory $m'$ used by the algorithm is rounded down to the
-closest number of complete blocks (case 1), of complete block rows (case
-2) or of complete block slices (case 3). This is in general reasonable
+closest number of complete blocks (Case 1), of complete block rows (Case
+2) or of complete block slices (Case 3). This is in general reasonable
 since adding an incomplete row to a set of complete ones multiplies
@@
 Our algorithm also avoids configurations where the memory load
-overlaps multiple block slices in case 2 or multiple block rows in
-case 1, as such overlaps multiply the number of required seeks (see
+overlaps multiple block slices in Case 2 or multiple block rows in
+Case 1, as such overlaps multiply the number of required seeks (see
 Figure~\ref{fig:avoided-configurations}-Right).
 
-Cluster reads are described in
+Clustered reads are described in
 Algorithm~\ref{algo:cluster-reads}. Function \texttt{switch} (line 3)
@@
 (line 7) determines whether two blocks overlap multiple block slices
-(case 2) or multiple block rows (case 1). For case 3 it always returns
+(case 2) or multiple block rows (Case 1). For case 3 it always returns
 false.  Function \texttt{sizeof} (line 8) returns the actual memory
@@
 In the complexity analysis, we assume that $m'$ represents an integer
-number $k$ of sub-rows (case 1, $k_1<\sqrt[3]{n}$), of complete rows
-(case 2, $k_2<d$), of tile rows (case 3, $k_3<\sqrt[3]{n}$), of slices
-(case 4, $k_4<d$) or of block slices (case 5, $k_5<\sqrt[3]{n}$), as
+number $k$ of sub-rows (Case 1, $k_1<\sqrt[3]{n}$), of complete rows
+(Case 2, $k_2<d$), of tile rows (Case 3, $k_3<\sqrt[3]{n}$), of slices
+(Case 4, $k_4<d$) or of block slices (Case 5, $k_5<\sqrt[3]{n}$), as
 illustrated in Figure~\ref{fig:multiple-reads-cases}. In each of these
@@
 and~\ref{eq:seeks-multiple-reads} for different values of $n$. When
-Cluster reads are in Case 1 or 2, they may outperform Multiple reads
-only for large values of $n$. When Cluster reads are in case 3, they
+Clustered reads are in Case 1 or 2, they may outperform Multiple reads
+only for large values of $n$. When Clustered reads are in case 3, they
 are equivalent to Multiple reads: assuming that $Rb$ is an exact
@@
   \includegraphics[width=0.45\columnwidth]{figures/model-big-brain-comparison-moreblocks.pdf}
-  \caption{Number of seeks for Cluster reads vs Multiple reads, for
+  \caption{Number of seeks for Clustered reads vs Multiple reads, for
     D=3458 and b=2. Left: n=125; Right: n=64,000. Case 1 and Case 2
-    denote Cluster read configurations.}
+    denote Clustered read configurations.}
   % Gnuplot file: scripts/model/model.gnplt
@@
 % buffer implementation
-The data buffer used in Cluster reads and Multiple reads is
+The data buffer used in Clustered reads and Multiple reads is
 implemented as a Python dictionary where the keys are offsets in the
@@
 dictionary entries are written sequentially to the reconstructed
-image. In Cluster reads, some seeking might be required between
+image. In Clustered reads, some seeking might be required between
 writes. We implemented a defragmentation procedure for the dictionary
@@
 
+We also implemented the splitting algorithms corresponding to
+Clustered reads and Multiple reads, called Clustered writes and
+Multiple writes.
+
 %% \subsection{Lossless compression}
@@
 non-overlapping chunks of size 770x605x700 with 2 bytes per voxel
-(total size, uncompressed is 75.92~GB). Big~Brain is a reference brain
+(total size uncompressed is 75.92~GB). Big~Brain is a reference brain
 based on the reconstruction of 7404 histological sections at nearly
@@
 image, using our naive split algorithm. Slices are available at
-\todo{\url{x}}.
+\todo{Upload them (compressed) to figshare.}.
 
@@
 We used a Dell Precision Tower 3620 workstation with CentOS Linux
-release 7.3.1611, 32~GB of RAM and two disks. (1) a Hard disk drive
+release 7.3.1611, 32~GB of RAM and two disks: (1) a Hard Disk Drive
 (HDD): HGST Travelstar 7K1000, 7200~rpm, 931GiB (1TB), firmware
@@
 
-We used Git tag 0.1 of our \texttt{sam} library. Our experiment
-scripts are available at
+We used Git tag 0.1 of our \texttt{sam} library \todo{tag it}. Our
+experiment scripts are available at
 \url{https://github.com/big-data-lab-team/paper-sequential-split-merge/blob/master/scripts/experiment}. We
-used them to split and merge using Buffered slices, Cluster reads and
+used them to split and merge using Buffered slices, Clustered reads and
 Multiple reads, with 3~GB, 6~GB, 9~GB, 12~GB and 16~GB of
-memory. Table~\ref{table:configs} shows the configurations of Cluster
+memory. Table~\ref{table:configs} shows the configurations of Clustered
 reads and Multiple reads for each memory value, according to
 Equations~\ref{eq:seeks-cluster-reads-1}
-and~\ref{eq:seeks-multiple-reads}. For instance, Cluster reads are in
-case 1 for 3~GB. We also did 0~GB of memory for Buffered slices and
-Cluster reads, which triggered naive slice and block merging. We did 5
+and~\ref{eq:seeks-multiple-reads}. For instance, at 3~GB, Clustered reads were in
+Case 1. We also did a run with 0~GB of memory for Buffered slices and
+Clustered reads, which triggered naive slice and block merging. We did 5
 repetitions for each memory value. Memory values were shuffled in each
-repetition, to correct for potential biases coming from ordering, such
-as caching effects. To improve reproducibility, we dropped the kernel
+repetition, to avoid potential ordering biases such
+as caching effects. To ensure equal conditions, we dropped the kernel
 page, dentry and inode caches before each run (\texttt{echo 3 | sudo
@@
   \hline
-  Cluster  reads &  1 &  2 &  2 &  2 &  3 \\
+  Clustered  reads &  1 &  2 &  2 &  2 &  3 \\
   Multiple reads &  4 &  4 &  4 &  4 &  5\\
@@
 
+All the experimental data and scripts used to generate the figures in
+this Section are available at
+\url{https://github.com/big-data-lab-team/paper-sequential-split-merge}
+under GPLv3 license.
+
 \subsubsection{Seeks}
@@
 bars are not reported as numbers of seeks were constant across all
-repetitions. The mean model error is \todo{x\%}, explained by the fact
+repetitions. The average relative model errors are 12.7\% (Naive
+blocks), 0\% (Naive slices), 3.3\% (Clustered reads), 26.8\% (Multiple
+reads) and 0.9\% (Buffered slices), presumably explained by the fact
 that the model assumes cubic blocks while we used non-cubic ones in
-the experiment. Overall, the model is validated as it correctly
-explains the observations.  
+the experiment. For Multiple reads, our complexity analysis also
+assumed that one of the 5 Cases in
+Figure~\ref{fig:multiple-reads-cases} was used while they are usually
+blended in practice. Overall, the model correctly explains the
+observations.
 \begin{figure}[h]
@@
 tremendous, in the order of 50 million seeks. For 3~GB, the number of
-seeks in Cluster reads is 4 orders of magnitude higher than for
+seeks in Clustered reads is 4 orders of magnitude higher than for
 Multiple reads, and 5 orders of magnitude higher than naive
-slices. This huge difference comes from the fact that at 3~GB Cluster
-reads are in case 1. For 6~GB, 9~GB and 12~GB, Cluster reads are in
+slices. This huge difference comes from the fact that at 3~GB Clustered
+reads are in Case 1. For 6~GB, 9~GB and 12~GB, Clustered reads are in
 case 2 and the difference with Multiple reads and Buffered slices
@@
 
-\todo{ The measured seek times correlated with
-the measured merging times ($\rho$=x), which suggest that the
-observed time reductions were indeed caused by reduced seek times.}
-
 % Slices and blocks
 Figure~\ref{fig:merge-time} shows the merge time for all algorithms by
-memory values. Naive blocks take about 5 times longer than naive
-slices to merge the data, which quantifies the effect of the problem
-targeted by our algorithms. This effect appears both on HDD and SSD,
-which shows that SSDs might not be the only solution to reduce seek
-time. In the remainder we use naive blocks and naive slices and
-references to evaluate our algorithms.
+memory values. Naive blocks are 9.5 times slower than naive slices on
+HDD (6.7 times on SSD), which quantifies the effect of the targeted
+problem. It also shows that SSDs might not be the only solution to
+reduce the impact of seek time. In the remainder we use naive blocks
+and naive slices as references to evaluate our algorithms.
 
@@
 
-% Cluster reads
-Cluster reads provide important speed ups compared to naive blocks,
-both on HDD and on SSD. On SSD, they are up to 5 times faster than
-naive blocks at 16~GB, and 3 on average \todo{compute the actual
-  values}. \todo{On HDD, \ldots} Surprisingly, they perform
-substantially faster than naive blocks even at 3~GB, while in case
-1. This may be explained by the fact that the seeks required to write
-incomplete block rows to the reconstructed image are shorter than the
-ones for naive blocks.
+% Clustered reads
+Clustered reads provide important speed ups compared to naive blocks,
+both on HDD and on SSD. On HDD, they are up to 9.7 times faster than
+naive blocks at 16~GB and 6.8 times faster on average. On SSD, they
+are up to 5.1 times faster than naive blocks at 16~GB, and 3.1 times
+faster on average.  Surprisingly, they perform substantially faster
+than naive blocks even at 3~GB, while in Case 1. This may be explained
+by the fact that the seeks required to write incomplete block rows to
+the reconstructed image are shorter than the ones for naive blocks.
 
 % Multiple reads
-Multiple reads are even faster than Cluster reads. On SSD, they are 5
-times faster than naive blocks on average \todo{compute the actual
-  values} ( \todo{On HDD, \ldots}). At 12~GB, the total merging time
-is only 1.2 times slower than naive slices, which solves our initial
-problem to a large extent. Even at 3~GB the speed-up is considerable
-compared to naive blocks.
+Multiple reads are even faster than Clustered reads on this dataset. On
+HDD, they are 8.4 times faster than naive blocks on average (5.3 times
+on SSD). At 16~GB, they are 9.0 times faster than naive blocks on
+average (5.8 times faster on SSD).
 
@@
 write rate importantly in addition to introducing extra delays.
-Cluster reads reduce the seek time and thus the read time very
+Clustered reads reduce the seek time and thus the read time very
 substantially. Multiple reads almost anihilate the seek time and bring
@@
 behavior is observed on HDD and on SSD, although the effect of seeking
-is slightly lower on SSD, as expected. Read times are consistently and substantially lower than write times\todo{explain why}. The overhead time is small for
-both Cluster reads and Multiple reads, although it is not negligible
-for Multiple reads on HDD.
+is slightly smaller on SSD, as expected. Read times are consistently
+and substantially lower than write times. This may be a result of
+reading data using Python's NumPy package, which is more efficient
+than using native Python - as is the case with our writes. The
+overhead time is small for both Clustered reads and Multiple reads.
 \begin{figure}
@@
     \includegraphics[width=0.45\columnwidth]{figures/benchmark-creads/creads-breakdown-ssd.pdf}
-    \caption{Cluster reads}
+    \caption{Clustered reads}
   \end{subfigure}
@@
 
-\todo{define cluster writes and multiple writes somewhere.}
-
-The number of seeks for Cluster writes and Multiple writes were
-identical to the ones in Cluster reads and Multiple reads. \todo{Check
+The number of seeks for Clustered writes and Multiple writes were
+identical to the ones in Clustered reads and Multiple reads. \todo{Check
   that.} The total split time by algorithm is shown in
@@
     \includegraphics[width=0.45\columnwidth]{figures/benchmark-cwrites/cwrites-breakdown-ssd.pdf}
-    \caption{Cluster writes}
+    \caption{Clustered writes}
   \end{subfigure}
@@
 % Problem solved!
-Cluster reads and multiple reads are able to reduce to a negligible
+Clustered reads and multiple reads are able to reduce to a negligible
 amount the overall seek time required to split or merge 3D blocks in a
@@
 % overall, multiple reads is recommended as default parameters for the
-% library. Cluster reads could be useful in some cases.
+% library. Clustered reads could be useful in some cases.
 
 %Same algo works for blocks or slices.
-
-% Parallel algorithms
-In a real system, high-resolution images are likely to be processed on
-computing clusters, for instance using software from the Hadoop
-project, in particular the Hadoop Distributed File
-System~\cite{shvachko2010hadoop}. In this context parallel split-and-merge algorithms
-will be required, since the various blocks of a large image would be
-uploaded to different disks concurrently. In the same vein,
-``re-spliting'' algorithms would have to be devised in case a split
-image need to be split in a different geometry. Cluster reads and
-multiple reads could be used as starting points for such algorithms.
 
@@
 In terms of file formats, our results demonstrate that simple imaging
-formats might be used for the split-and-merge problem without
-performance loss compared to more complex formats that try to preserve
-spatial locality on disk, for instance MINC 2.0~\cite{vincent2016minc}
-or the format based on space-filling curves mentioned
-in~\cite{burns2013open}. This is of major interest in the current
-open-science context since simpler formats favor data-sharing and
-interoperability.  Moreover, our algorithms could potentially be
+formats may be split and merged without performance loss compared to
+more complex formats that try to preserve spatial locality on disk,
+for instance MINC 2.0 or the format based on space-filling curves
+mentioned in~\cite{burns2013open}. This is of major interest in the
+current open-science context since simpler formats favor data-sharing
+and interoperability.  Moreover, our algorithms could potentially be
 adapted to any split geometry, even though we demonstrated them on
@@
 particular geometry. For instance, the format in~\cite{burns2013open}
-wouldn't perform nicely with a naive splitting algorithm if slices had
-to be extracted instead of blocks.  Two points need to be made to be
-fair though. First, we aimed at extracting all the blocks while the
-problem targeted by the format in~\cite{burns2013open} was to extract
-a single block from a large image. Second, we have not considered
+is not designed to naively split slices. However, we aimed to extract
+all the blocks, whereas, ~\cite{burns2013open} aimed to extract a
+single block from a large image. Furthermore, we have not considered
 on-the-fly data compression yet, which is a unique feature of MINC
@@
 
-For more investigation and reproducibility purposes, the complete \LaTeX
-sources of this paper, including the collected data and Gnuplot
-scripts used to generate the results figures are available at
-\url{https://github.com/big-data-lab-team/paper-sequential-split-merge}
-under GPLv3 license. Our \texttt{sam} library is at
+% Parallel algorithms
+High-resolution images are likely to be processed on
+computing clusters, for instance using software from the Hadoop
+project, in particular the Hadoop Distributed File
+System~\cite{shvachko2010hadoop}. In this context parallel
+split-and-merge algorithms would be beneficial, since the various
+blocks of a large image could be uploaded to different disks
+concurrently. In the same vein, ``re-spliting'' algorithms would be
+beneficial in case an image already split needs to be split in a
+different geometry. Clustered reads and multiple reads could be used
+as starting points for such algorithms.
+
+Our \texttt{sam} library is available at
 \url{https://github.com/big-data-lab-team/sam}, under MIT license.
@@
 We warmly thank Lindsay B. Lewis and Claude Lepage for helping us with
-Big~Brain, Greg Kiar for useful discussion about the Open
-Connectome Data Cluster and Pierre Bellec for discussions about MINC.
+Big~Brain, Greg Kiar for useful discussion about the Open Connectome
+Data Cluster and Pierre Bellec for discussions about MINC.
 

### 2017-08-11 paper.tex
Updates following discussion with @Yongpinggao

@@
 
+implemnted on-the-fly compression using Python's gzip library. Use
+gzip open file. We don't need on-the-fly compression for multiple
+reads. gzip can find the seek point but it still has to decompress
+everything before the seek point. Cluster reads would be very
+penalized by that. Multiple reads wouldn't care. Same conclusions for
+splitting.
+
+indexed gzip would seek faster (improve the seekk time) but
+decompression time would remain (still have to decompress from
+begining of file).
+
+if we wanted cluster reads to work well with on the fly compression,
+we would need to investigate techniques like~\cite{rajna2015speeding}
+
 \section{Experiments}
@@
 
+\subsection{Compression}
+
+Naive blocks, naive slices, cluster reads and multiple reads at 3, 6,
+9, 12, 16GB, on SSD, with on-the-fly compression (graph will also show
+without compression (previous results)) for comparison.
+
 \section{Discussion}

### 2017-08-14 scripts/experiment/benchmark_split.py
minor fix

@@ -17,11 +17,12 @@ import os
 ori_image_hdd = "/data/bigbrain_40microns.nii"
 out_dir_hdd = "/data/gao/blocks_split"
 
-ori_image_ssd = "/home/gao/new_image.nii"
+#ori_image_ssd = "/home/gao/new_image.nii"
+ori_image_ssd = "/data/bigbrain_40microns.nii"
 out_dir_ssd = "/home/gao/blocks125"
 
-csv_file_hdd = "./hdd_mwrites.dat"
-csv_file_ssd = "./ssd_mwrites.dat"
+csv_file_hdd = "./mwrites_hdd.dat"
+csv_file_ssd = "./mwrites_ssd.dat"
 
 
 Y_splits=5

### 2017-08-14 scripts/experiment/generate_total_avg_var_data.py

@@ -132,10 +132,10 @@ def main():
     generate(input_file_mreads_ssd, input_file_creads_ssd, input_file_buff_slices_ssd, output_file=output_file)
     output_file = "./data/total-merge-time-hdd.dat"
     generate(input_file_mreads_hdd, input_file_creads_hdd, input_file_buff_slices_hdd, output_file=output_file)
-    #crash - add split data in cwrites/buff-slices
-    #output_file = "./data/total-split-time-ssd.dat"
-    #generate(input_file_mwrites_ssd, input_file_cwrites_ssd, input_file_buff_slices_split_ssd, output_file=output_file)
-    #output_file = "./data/total-split-time-hdd.dat"
-    #generate(input_file_mwrites_hdd, input_file_cwrites_hdd, input_file_buff_slices_split_hdd, output_file=output_file)
+    # crash
+    # output_file = "./data/total-split-time-ssd.dat"
+    # generate(input_file_mwrites_ssd, input_file_cwrites_ssd, input_file_buff_slices_split_ssd, output_file=output_file)
+    # output_file = "./data/total-split-time-hdd.dat"
+    # generate(input_file_mwrites_hdd, input_file_cwrites_hdd, input_file_buff_slices_split_hdd, output_file=output_file)
 if __name__ == '__main__':
     main()

### 2017-08-14 paper.tex

--- 2017-08-11 paper.tex
@@
 
-\author{Final ordering
-  to be discussed: Yongping Gao/Val\'erie Hayot-Sasson,
-  Yuhong Yan, Tristan Glatard\\
+\author{Val\'erie Hayot-Sasson$^*$, Yongping Gao$^*$, Yuhong Yan, Tristan Glatard\\
   Department of Computer Science and Software Engineering\\ Concordia University, Montreal, Quebec, Canada\\
-  {first.last}@concordia.ca
+  {first.last}@concordia.ca\\
+  $^*$ These authors have contributed equally
+
 }
@@
 
-%% \subsection{Lossless compression}
+\subsection{Lossless compression}
+
+We implemented lossless compression for all algorithms using Python's
+gzip library. Compression is done on the fly, that is, while the data
+is being read or written. On-the-fly compression of large datasets is
+a challenge when extensive seeking is involved as the gzip library has
+to decompress all the data until the seek position to read from it,
+making access time a linear function of the seek position.
 
@@
 Clustered reads provide important speed ups compared to naive blocks,
-both on HDD and on SSD. On HDD, they are up to 9.7 times faster than
-naive blocks at 16~GB and 6.8 times faster on average. On SSD, they
-are up to 5.1 times faster than naive blocks at 16~GB, and 3.1 times
-faster on average.  Surprisingly, they perform substantially faster
-than naive blocks even at 3~GB, while in Case 1. This may be explained
-by the fact that the seeks required to write incomplete block rows to
-the reconstructed image are shorter than the ones for naive blocks.
+both on HDD and on SSD. They are 6.8 times faster than naive blocks on
+HDD and 5.1 times on SSD (average accross all repetitions, all memory
+values). Surprisingly, they perform substantially faster than naive
+blocks even at 3~GB, while in Case 1. This may be explained by the
+fact that the seeks required to write incomplete block rows to the
+reconstructed image are shorter than the ones for naive blocks.
 
 % Multiple reads
-Multiple reads are even faster than Clustered reads on this dataset. On
-HDD, they are 8.4 times faster than naive blocks on average (5.3 times
-on SSD). At 16~GB, they are 9.0 times faster than naive blocks on
-average (5.8 times faster on SSD).
+Multiple reads are even faster than Clustered reads on this
+dataset. They are 8.4 times faster than naive blocks on HDD and 5.3
+times on SSD (average accross all repetitions, all memory values).
 
@@
 
+\subsection{Compression}
+
+Naive blocks, naive slices, cluster reads and multiple reads at 3, 6,
+9, 12, 16GB, on SSD, with on-the-fly compression (graph will also show
+without compression (previous results)) for comparison.
+
+Expected result: Compression speeds up Multiple reads. It speeds up
+Clustered reads to a smaller extent since Cluster reads still have to
+seek in the reconstructed image. \todo{finish that up when the data is
+  available}
+
 \section{Discussion}
@@
 all the blocks, whereas, ~\cite{burns2013open} aimed to extract a
-single block from a large image. Furthermore, we have not considered
-on-the-fly data compression yet, which is a unique feature of MINC
-2.0.
+single block from a large image. MINC 2.0 might also help with
+on-the-fly compression.
+
+% Compression
+Due to the minimal number of seeks in the reconstructed image,
+Multiple reads greatly benefit from on-the-fly compression. This is
+not the case of Clustered reads, where compression has an important
+overhead due to residual seeking in the reconstructed
+image. Techniques such as z-indexing have been used to address this
+limitation of the NiFTI format~\cite{rajna2015speeding} and could be
+employed here too.
 
 % Parallel algorithms
-High-resolution images are likely to be processed on
-computing clusters, for instance using software from the Hadoop
-project, in particular the Hadoop Distributed File
+High-resolution images are likely to be processed on computing
+clusters, for instance using software from the Hadoop project, in
+particular the Hadoop Distributed File
 System~\cite{shvachko2010hadoop}. In this context parallel

### 2017-08-14 paper.tex
Included cwrites data

@@
 Figure~\ref{fig:split-time} and the breakdown by read, write, seek and
-overhead time is in Figure~\ref{fig:breakdowns-ssd-split}. \todo{comment on the figures when results are available.}
+overhead time is in Figure~\ref{fig:breakdowns-ssd-split}. \todo{comment on the figures when results are available. Split time looks slightly longer than merge time.}
 
@@
   \begin{subfigure}[b]{\columnwidth}
-    \includegraphics[width=0.45\columnwidth]{figures/benchmark-cwrite/cwrites-breakdown-hdd.pdf}
+    \includegraphics[width=0.45\columnwidth]{figures/benchmark-cwrites/cwrites-breakdown-hdd.pdf}
     \includegraphics[width=0.45\columnwidth]{figures/benchmark-cwrites/cwrites-breakdown-ssd.pdf}

### 2017-08-14 scripts/experiment/generate_avg_var_data.py

@@ -10,6 +10,11 @@ output_file_creads_ssd = "./data/creads/creads_ssd_avg_var.dat"
 input_file_creads_hdd = "./data/creads/creads_hdd.dat"
 output_file_creads_hdd = "./data/creads/creads_hdd_avg_var.dat"
 
+input_file_cwrites_ssd = "./data/cwrites/cwrites_ssd.dat"
+output_file_cwrites_ssd = "./data/cwrites/cwrites_ssd_avg_var.dat"
+input_file_cwrites_hdd = "./data/cwrites/cwrites_hdd.dat"
+output_file_cwrites_hdd = "./data/cwrites/cwrites_hdd_avg_var.dat"
+
 input_file_buff_slices_ssd = "./data/buff-slices/buff-slices_ssd.dat"
 output_file_buff_slices_ssd = "./data/buff-slices/buff-slices_ssd_avg_var.dat"
 input_file_buff_slices_hdd = "./data/buff-slices/buff-slices_hdd.dat"
@@ -205,6 +210,9 @@ def main():
     # creads
     get_avg_var_naive(input_file_creads_ssd, output_file_creads_ssd, input_file_creads_ssd, input_file_buff_slices_ssd, hasnaive=True, disk="ssd")
     get_avg_var_naive(input_file_creads_hdd, output_file_creads_hdd, input_file_creads_hdd, input_file_buff_slices_hdd, hasnaive=True, disk="hdd")
+    # cwrites
+    get_avg_var_naive(input_file_cwrites_ssd, output_file_cwrites_ssd, input_file_cwrites_ssd, input_file_buff_slices_ssd, hasnaive=True, disk="ssd")
+    get_avg_var_naive(input_file_cwrites_hdd, output_file_cwrites_hdd, input_file_cwrites_hdd, input_file_buff_slices_hdd, hasnaive=True, disk="hdd")
     # buffered slicse
     get_avg_var_naive(input_file_buff_slices_ssd, output_file_buff_slices_ssd, input_file_creads_ssd, input_file_buff_slices_ssd, hasnaive=True, disk="ssd")
     get_avg_var_naive(input_file_buff_slices_hdd, output_file_buff_slices_hdd, input_file_creads_hdd, input_file_buff_slices_hdd, hasnaive=True, disk="hdd")

### 2017-08-14 scripts/experiment/generate_total_avg_var_data.py

@@ -79,7 +79,8 @@ def generate(input_file_mreads, input_file_creads, input_file_buff_slices, outpu
                 # 12g - mr
                 total_time_12g.append(float(items[19]))
                 # 16g - mr
-                total_time_16g.append(float(items[24]))
+                if len(items) >= 25:
+                  total_time_16g.append(float(items[24]))
 
         mr_total = [0, avg(total_time_3g), avg(total_time_6g), avg(total_time_9g), avg(total_time_12g), avg(total_time_16g)]
         mr_total_err = [0, var(total_time_3g), var(total_time_6g), var(total_time_9g), var(total_time_12g), var(total_time_16g)]
@@ -133,9 +134,9 @@ def main():
     output_file = "./data/total-merge-time-hdd.dat"
     generate(input_file_mreads_hdd, input_file_creads_hdd, input_file_buff_slices_hdd, output_file=output_file)
     #crash - add split data in cwrites/buff-slices
-    #output_file = "./data/total-split-time-ssd.dat"
-    #generate(input_file_mwrites_ssd, input_file_cwrites_ssd, input_file_buff_slices_split_ssd, output_file=output_file)
-    #output_file = "./data/total-split-time-hdd.dat"
-    #generate(input_file_mwrites_hdd, input_file_cwrites_hdd, input_file_buff_slices_split_hdd, output_file=output_file)
+    output_file = "./data/total-split-time-ssd.dat"
+    generate(input_file_mwrites_ssd, input_file_cwrites_ssd, input_file_buff_slices_split_ssd, output_file=output_file)
+    output_file = "./data/total-split-time-hdd.dat"
+    generate(input_file_mwrites_hdd, input_file_cwrites_hdd, input_file_buff_slices_split_hdd, output_file=output_file)
 if __name__ == '__main__':
     main()

### 2017-08-15 paper.tex
Integrated split results for buffered slices on SSD

@@
 and substantially lower than write times. This may be a result of
-reading data using Python's NumPy package, which is more efficient
-than using native Python - as is the case with our writes. The
-overhead time is small for both Clustered reads and Multiple reads.
+discrepancies between disk read and writes rates, or of reading data using Python's NumPy
+package, which is more efficient than using native Python - as is the
+case with our writes. The overhead time is small for both Clustered
+reads and Multiple reads.
 \begin{figure}
   \begin{subfigure}[b]{\columnwidth}
-    \includegraphics[width=0.45\columnwidth]{figures/benchmark-buff-slices/buff-slices-breakdown-hdd.pdf}
-    \includegraphics[width=0.45\columnwidth]{figures/benchmark-buff-slices/buff-slices-breakdown-ssd.pdf}
+    \includegraphics[width=0.45\columnwidth]{figures/benchmark-buff-slices/buff-slices-reads-breakdown-hdd.pdf}
+    \includegraphics[width=0.45\columnwidth]{figures/benchmark-buff-slices/buff-slices-reads-breakdown-ssd.pdf}
     \caption{Buffered slices}
@@
   \begin{subfigure}[b]{\columnwidth}
-    \includegraphics[width=0.45\columnwidth]{figures/benchmark-buff-slices/buff-slices-breakdown-hdd-split.pdf}
-    \includegraphics[width=0.45\columnwidth]{figures/benchmark-buff-slices/buff-slices-breakdown-ssd-split.pdf}
+    \includegraphics[width=0.45\columnwidth]{figures/benchmark-buff-slices/buff-slices-writes-breakdown-hdd.pdf}
+    \includegraphics[width=0.45\columnwidth]{figures/benchmark-buff-slices/buff-slices-writes-breakdown-ssd.pdf}
     \caption{Buffered slices}

### 2017-08-15 scripts/experiment/generate_total_avg_var_data.py

@@ -18,12 +18,12 @@ def var(l):
 #ssd
 input_file_mreads_ssd = "./data/mreads/mreads_ssd.dat"
 input_file_creads_ssd = "./data/creads/creads_ssd.dat"
-input_file_buff_slices_ssd = "./data/buff-slices/buff-slices_reads_ssd.dat"
+input_file_buff_slices_reads_ssd = "./data/buff-slices/buff-slices_reads_ssd.dat"
 
 #hdd
 input_file_mreads_hdd = "./data/mreads/mreads_hdd.dat"
 input_file_creads_hdd = "./data/creads/creads_hdd.dat"
-input_file_buff_slices_hdd = "./data/buff-slices/buff-slices_reads_hdd.dat"
+input_file_buff_slices_reads_hdd = "./data/buff-slices/buff-slices_reads_hdd.dat"
 
 # split
 #ssd
@@ -35,9 +35,6 @@ input_file_buff_slices_writes_ssd = "./data/buff-slices/buff-slices_writes_ssd.d
 input_file_mwrites_hdd = "./data/mwrites/mwrites_hdd.dat"
 input_file_cwrites_hdd = "./data/cwrites/cwrites_hdd.dat"
 input_file_buff_slices_writes_hdd = "./data/buff-slices/buff-slices_writes_hdd.dat"
-
-
-
 
 
 def generate(input_file_mreads, input_file_creads, input_file_buff_slices, output_file):
@@ -79,8 +76,7 @@ def generate(input_file_mreads, input_file_creads, input_file_buff_slices, outpu
                 # 12g - mr
                 total_time_12g.append(float(items[19]))
                 # 16g - mr
-                if len(items) >= 25:
-                  total_time_16g.append(float(items[24]))
+                total_time_16g.append(float(items[24]))
 
         mr_total = [0, avg(total_time_3g), avg(total_time_6g), avg(total_time_9g), avg(total_time_12g), avg(total_time_16g)]
         mr_total_err = [0, var(total_time_3g), var(total_time_6g), var(total_time_9g), var(total_time_12g), var(total_time_16g)]
@@ -101,13 +97,13 @@ def generate(input_file_mreads, input_file_creads, input_file_buff_slices, outpu
                 # 12g - bs
                 total_time_12g.append(float(items[24]))
                 # 16g - bs
-                total_time_16g.append(float(items[24]))
+                total_time_16g.append(float(items[29]))
 
         ns_total = [avg(total_time_06g), 0, 0, 0, 0, 0]
[…]

### 2017-08-16 paper.tex
Added split slices on HDD (@ValHayot's data).

@@
   n + \ceil*{\frac{Rb}{\sqrt[3]{n}m_2'}}\sqrt[3]{R}        & \text{if } \frac{Rb}{\sqrt[3]{n}^2} \leq m < \frac{Rb}{\sqrt[3]{n}}\\[0.4em]
-  n + \ceil*{\frac{Rb}{m_3'}}                              & \text{if } \frac{Rb}{\sqrt[3]{n}} \leq Rb < n\\[0.4em]
-  n + 1                                                          & \text{otherwise}
+  n + \ceil*{\frac{Rb}{m_3'}}                              & \text{if } \frac{Rb}{\sqrt[3]{n}} \leq m < Rb
 \end{cases} \label{eq:seeks-cluster-reads-1}
@@
 
-The number of seeks for Clustered writes and Multiple writes were
-identical to the ones in Clustered reads and Multiple reads. \todo{Check
-  that.} The total split time by algorithm is shown in
-Figure~\ref{fig:split-time} and the breakdown by read, write, seek and
-overhead time is in Figure~\ref{fig:breakdowns-ssd-split}. \todo{comment on the figures when results are available. Split time looks slightly longer than merge time.}
+The total split time by algorithm is shown in
+Figure~\ref{fig:split-time}. The difference between naive blocks and
+naive slices is still significant (average ratio is 1.4 on SSD,
+\todo{x} on HDD) although less than on HDD. Clustered writes and
+Multiple writes both perform similarly to naive slices. Buffered
+slices don't provide any significant speed-up compared to naive
+slices.
+
+The breakdown by read, write and overhead time is shown in
+Figure~\ref{fig:breakdowns-ssd-split}. We were not able to measure
+seek time for splitting algorithms as it was blended in read time by the Nibabel library. \todo{comment on the figures when results are
+  available. Split time looks slightly longer than merge time.}
 

### 2017-08-17 paper.tex
Included compression results to the text

@@
 buffer and write the buffer in the reconstructed image. Seeking is
-reduced compared to naive block merging since contiguous parts of the
+reduced compared to Naive blocks since contiguous parts of the
 buffer will be written without seeking. A given
@@
 \caption{Memory-load configurations in multiple reads (d=4, D=16,
-  n=64, $k_1=k_2=k_3=k_4=k_5=1$). Red colour shows the content of the
+  n=64, $k_1=k_2=k_3=k_4=k_5=1$). Red color shows the content of the
   memory load. Small dots depict block frontiers and long dashes
@@
 and~\ref{eq:seeks-multiple-reads} both boil down to $n+\frac{Rb}{m}$.
-\todo{We could be a bit more rigorous and detail the cases where
-  cluster reads of multiple reads will be better.}
+%\todo{We could be a bit more rigorous and detail the cases where
+%  cluster reads of multiple reads will be better.}
 
@@
 to decompress all the data until the seek position to read from it,
-making access time a linear function of the seek position.
-
-%% Lossless compression might reduce overall read and write times by
-%% reducing seek time and data sizes. However, it needs to be implemented
-%% on-the-fly, i.e., in chunks. See the work in~\cite{rajna2015speeding}
-%% and
-%% \href{https://github.com/pauldmccarthy/indexed\_gzip}{corresponding
-%%   implementation}. This is straightforward to apply for random reads
-%% but not for random writes. Multiple reads is the only merging strategy
-%% that works with literally no seek in the reconstructed image which
-%% could therefore potentially use in-memory compression.
-
-%% \todo{could we merge without de-compressing?}
+making access time a linear function of the seek position. This is
+potentially an issue for Clustered writes, which could be addressed by
+indexing techniques such as described in~\cite{rajna2015speeding} for
+the NiFTI format and implemented in
+\url{https://github.com/pauldmccarthy/indexed\_gzip}. Multiple writes
+are not impacted since they do not seek in the large image.
+
+In addition, ``negative'' seeking (seeking to a position that precedes
+the current one) is not possible while writing a compressed file,
+which renders Naive blocks unusable with compressed data and Clustered
+reads usable only in Case 3.  Again, Multiple reads are not impacted
+since they do not seek in the large image.
 
@@
 We generated NiFTI slices of size 3850x3025x28 from the reconstructed
-image, using our naive split algorithm. Slices are available at
-\todo{Upload them (compressed) to figshare.}.
+image, using our naive split algorithm.
 
 \subsection{Hardware}
-
 We used a Dell Precision Tower 3620 workstation with CentOS Linux
@@
 Case 1. We also did a run with 0~GB of memory for Buffered slices and
-Clustered reads, which triggered naive slice and block merging. We did 5
+Clustered reads, which triggered Naive slices and Naive blocks. We did 5
 repetitions for each memory value. Memory values were shuffled in each
@@
 
-As expected, the difference between naive blocks and naive slices is
+As expected, the difference between Naive blocks and Naive slices is
 tremendous, in the order of 50 million seeks. For 3~GB, the number of
 seeks in Clustered reads is 4 orders of magnitude higher than for
-Multiple reads, and 5 orders of magnitude higher than naive
+Multiple reads, and 5 orders of magnitude higher than Naive
 slices. This huge difference comes from the fact that at 3~GB Clustered
@@
 Figure~\ref{fig:merge-time} shows the merge time for all algorithms by
-memory values. Naive blocks are 9.5 times slower than naive slices on
+memory values. Naive blocks are 9.5 times slower than Naive slices on
 HDD (6.7 times on SSD), which quantifies the effect of the targeted
 problem. It also shows that SSDs might not be the only solution to
-reduce the impact of seek time. In the remainder we use naive blocks
-and naive slices as references to evaluate our algorithms.
+reduce the impact of seek time. In the remainder we use Naive blocks
+and Naive slices as references to evaluate our algorithms.
 
 % Buffered slices
-Buffered slices provide a negligible speed-up compared to naive
+Buffered slices provide a negligible speed-up compared to Naive
 slices. In most cases, their memory overhead would not be worth the
@@
 % Clustered reads
-Clustered reads provide important speed ups compared to naive blocks,
-both on HDD and on SSD. They are 6.8 times faster than naive blocks on
+Clustered reads provide important speed ups compared to Naive blocks,
+both on HDD and on SSD. They are 6.8 times faster than Naive blocks on
 HDD and 5.1 times on SSD (average accross all repetitions, all memory
-values). Surprisingly, they perform substantially faster than naive
+values). Surprisingly, they perform substantially faster than Naive
 blocks even at 3~GB, while in Case 1. This may be explained by the
 fact that the seeks required to write incomplete block rows to the
-reconstructed image are shorter than the ones for naive blocks.
+reconstructed image are shorter than the ones for Naive blocks.
 
@@
 Multiple reads are even faster than Clustered reads on this
-dataset. They are 8.4 times faster than naive blocks on HDD and 5.3
+dataset. They are 8.4 times faster than Naive blocks on HDD and 5.3
 times on SSD (average accross all repetitions, all memory values).
@@
 down to read, write, seek and overhead time for our algorithms. Naive
-blocks and naive slices are shown as references. The huge difference
-between naive blocks and naive slices is coming from both the seek
+blocks and Naive slices are shown as references. The huge difference
+between Naive blocks and Naive slices is coming from both the seek
 time and the write time, which suggests that seeking degrades the
@@
 substantially. Multiple reads almost anihilate the seek time and bring
-the read time to a value comparable to naive slices.  The same
+the read time to a value comparable to Naive slices.  The same
 behavior is observed on HDD and on SSD, although the effect of seeking
@@
 The total split time by algorithm is shown in
-Figure~\ref{fig:split-time}. The difference between naive blocks and
-naive slices is still significant (average ratio is 1.4 on SSD,
-\todo{x} on HDD) although less than on HDD. Clustered writes and
-Multiple writes both perform similarly to naive slices. Buffered
-slices don't provide any significant speed-up compared to naive
-slices.
+Figure~\ref{fig:split-time}. The difference between Naive blocks and
+Naive slices is still significant (average ratio is 1.4 on SSD, 2.0 on
+HDD) although less than for merging. On SSD, Clustered writes and
+Multiple writes both perform similarly to Naive slices. On HDD,
+Clustered writes are slightly slower than Multiple writes until
+16~GB. Buffered slices don't provide any significant speed-up compared
+to Naive slices.
 
@@
 Figure~\ref{fig:breakdowns-ssd-split}. We were not able to measure
-seek time for splitting algorithms as it was blended in read time by the Nibabel library. \todo{comment on the figures when results are
-  available. Split time looks slightly longer than merge time.}
+seek time for splitting algorithms as it was blended in read time by
+the Nibabel library. Naive blocks are strongly penalized by important
+read times coming from extensive seeking. Buffered slices, Clustered
+writes and multiple writes all reduce the read time compared to Naive
+blocks, but they also increase the write time, most likely due to
+caching effects in Naive blocks and slices.
 
@@
 
-Naive blocks, naive slices, cluster reads and multiple reads at 3, 6,
-9, 12, 16GB, on SSD, with on-the-fly compression (graph will also show
-without compression (previous results)) for comparison.
-
-Expected result: Compression speeds up Multiple reads. It speeds up
-Clustered reads to a smaller extent since Cluster reads still have to
-seek in the reconstructed image. \todo{finish that up when the data is
-  available}
+Figure~\ref{fig:compression} shows the impact of gzip compression on
+the merging time, for Multiple reads and Naive slices. As explained
+before, Clustered reads could only be applied to compressed data while
+in Case 3, i.e., for 16~GB, and Naive blocks could not be
+used. Compression dramatically slows down merging time for all
+algorithms \todo{Check that once all results are available}.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=\columnwidth]{figures/total-merge-time-ssd-compressed.pdf}
+  \hfill
+  \caption{Merging time by algorithm, using compression (on SSD). Each
+    algorithm is represented with a different color. Hatching
+    represents compression. Averages over 5 repetitions. Error bars
+    show $\pm$ 1 standard deviation. }
+  % Gnuplot file: scripts/experiment/split_and_merge_total_time.gnuplot
+\label{fig:compression}
+\end{figure}
 
@@
 single block from a large image. MINC 2.0 might also help with
-on-the-fly compression.
+on-the-fly compression for Clustered reads.
 
 % Compression
-Due to the minimal number of seeks in the reconstructed image,
-Multiple reads greatly benefit from on-the-fly compression. This is
-not the case of Clustered reads, where compression has an important
-overhead due to residual seeking in the reconstructed
-image. Techniques such as z-indexing have been used to address this
-limitation of the NiFTI format~\cite{rajna2015speeding} and could be
-employed here too.
+On-the-fly compression, indeed, could not be used with Clustered reads
+due to their use of negative seeking. This was not a problem for
+Multiple reads since Multiple reads completely remove seeking in the
+large image. Likewise, Multiple writes would not benefit from
+techniques aiming at accelerating random access reads in compressed
+images since they do not need to seek in it. Nonetheless, in our
+experimental setup compression drastically slowed down merging
+algorithms and proved of little use unless storage was limited.
 

### 2017-08-17 scripts/experiment/generate_total_avg_var_data.py

@@ -20,6 +20,11 @@ input_file_mreads_ssd = "./data/mreads/mreads_ssd.dat"
 input_file_creads_ssd = "./data/creads/creads_ssd.dat"
 input_file_buff_slices_reads_ssd = "./data/buff-slices/buff-slices_reads_ssd.dat"
 
+#ssd-compressed
+input_file_mreads_ssd_compressed = "./data/mreads/mreads_ssd_compressed.dat"
+input_file_creads_ssd_compressed = "./data/creads/creads_ssd_compressed.dat"
+input_file_buff_slices_reads_ssd_compressed = "./data/buff-slices/buff-slices_reads_ssd_compressed.dat"
+
 #hdd
 input_file_mreads_hdd = "./data/mreads/mreads_hdd.dat"
 input_file_creads_hdd = "./data/creads/creads_hdd.dat"
@@ -37,7 +42,7 @@ input_file_cwrites_hdd = "./data/cwrites/cwrites_hdd.dat"
 input_file_buff_slices_writes_hdd = "./data/buff-slices/buff-slices_writes_hdd.dat"
 
 
-def generate(input_file_mreads, input_file_creads, input_file_buff_slices, output_file):
+def generate(input_file_mreads, input_file_creads, input_file_buff_slices, output_file, compressed=False):
     total_time_06g, total_time_3g, total_time_6g, total_time_9g, total_time_12g, total_time_16g = [], [], [], [], [], []
 
     with open(input_file_creads, 'r') as f:
@@ -107,10 +112,16 @@ def generate(input_file_mreads, input_file_creads, input_file_buff_slices, outpu
     mem = ["0.6", "3", "6", "9", "12", "16"]
     with open(output_file, 'w+') as f:
         if "merge" in output_file:
-            f.write("mem \"Naive blocks\" NBR \"Naive slices\" NSR \"Cluster reads\" CRR \"Multiple reads\" MRR \"Buffered slices \" BSR")
+            if compressed:
+                f.write("mem \"Naive blocks (compressed)\" NBR \"Naive slices (compressed)\" NSR \"Cluster reads (compressed)\" CRR \"Multiple reads (compressed)\" MRR \"Buffered slices (compressed) \" BSR")
+            else:
+                f.write("mem \"Naive blocks\" NBR \"Naive slices\" NSR \"Cluster reads\" CRR \"Multiple reads\" MRR \"Buffered slices \" BSR")
             f.write('\n')
         elif "split" in output_file:
[…]

### 2017-08-17 scripts/experiment/compression/compressed.py
add compression for creads_16g, naive slices, buffered slices

@@ -1,0 +1,111 @@
+#!/usr/bin/env python
+# Ref: imageutils.py from
+# https://github.com/big-data-lab-team/sam/blob/master/imageutils.py
+import imageutils as img_utils
+import numpy as np
+from time import time
+import argparse
+import random
+import os
+
+# example
+# ./compressed.py -m 3221225472 6442450944 9663676416 12884901888 17179869184 -r 2 -d ssd
+# ./compressed.py -m 3221225472 9663676416 -r 5 -d hdd
+
+
+# on the consider
+reconstructed_hdd = "/data/gao/new_image.nii.gz"
+legend_hdd = "/data/gao/blocks_split/legend.txt"
+
+reconstructed_ssd = "/home/gao/new_image.nii.gz"
+legend_ssd = "/home/gao/blocks125/legend.txt"
+
+
+csv_file_hdd_mreads = "./mreads_hdd_compressed.dat"
+csv_file_hdd_creads = "./creads_hdd_compressed.dat"
+csv_file_ssd_mreads = "./mreads_ssd_compressed.dat"
+csv_file_ssd_creads = "./creads_ssd_compressed.dat"
+
+
+first_dim=3850
+second_dim=3025
+third_dim=3500
+
+files = {
+    "hdd": (reconstructed_hdd, legend_hdd, csv_file_hdd_mreads, csv_file_hdd_creads),
+    "ssd": (reconstructed_ssd, legend_ssd, csv_file_ssd_mreads, csv_file_ssd_creads)
+}
+
+def benchmark_mreads(mem, reconstructed, legend):
+    img = img_utils.ImageUtils(reconstructed, first_dim, second_dim, third_dim, np.uint16)
+    s_time = time()
+    total_read_time, total_write_time, total_seek_time, total_seek_number = img.reconstruct_img(legend, "multiple", mem, input_compressed=True, benchmark=True)
+    total_time = time() - s_time
+    print total_read_time, total_write_time, total_seek_time, total_seek_number, total_time
+    return (total_read_time, total_write_time, total_seek_time, total_seek_number, total_time)
+
+
+def benchmark_creads(mem, reconstructed, legend):
+    img = img_utils.ImageUtils(reconstructed, first_dim, second_dim, third_dim, np.uint16)
+    s_time = time()
+    total_read_time, total_write_time, total_seek_time, total_seek_number = img.reconstruct_img(legend, "clustered", mem, input_compressed=True, benchmark=True)
+    total_time = time() - s_time
[…]

### 2017-08-17 scripts/experiment/compression/compressed.py
minor fix

@@ -74,23 +74,18 @@ def main():
     rep = args.rep
     disk = args.disk
 
-
-
-
-
-    # print "mreads..."
-    # ## MREADS:
-    # for i in range(0, rep):
-    #     data_dict = {}
-    #     print "Repetition: {}".format(i)
-    #     random.shuffle(mem_list)
-    #     for mem in mem_list:
-    #         print "mem = {}".format(mem)
-    #         os.system("echo 3 | sudo tee /proc/sys/vm/drop_caches")
-    #         os.system("rm {}".format(files[disk][0]))
-    #         data = benchmark_mreads(mem=mem, reconstructed=files[disk][0], legend=files[disk][1])
-    #         data_dict[mem] = data
-    #     write_to_file(data_dict, files[disk][2])
+    ## MREADS:
+    for i in range(0, rep):
+        data_dict = {}
+        print "Repetition: {}".format(i)
+        random.shuffle(mem_list)
+        for mem in mem_list:
+            print "mem = {}".format(mem)
+            os.system("echo 3 | sudo tee /proc/sys/vm/drop_caches")
+            os.system("rm {}".format(files[disk][0]))
+            data = benchmark_mreads(mem=mem, reconstructed=files[disk][0], legend=files[disk][1])
+            data_dict[mem] = data
+        write_to_file(data_dict, files[disk][2])
 
 
 

### 2017-08-18 paper.tex
Integrated compression data

@@
 
-Clustered reads are the more direct extension of buffered slices to
+Clustered reads are the more direct extension of Buffered slices to
 blocks: they load multiple blocks in memory, concatenate them in a
@@
 as the total time minus the sum of all other times.
+
+Compressed blocks and slices were also merged into a compressed
+image. On-the-fly gzip compression was used for Naive slices, Buffered
+slices and Multiple reads. As explained before, Clustered reads could
+only be applied to compressed data while in Case 3, i.e., for 16~GB,
+and Naive blocks could not be used at all with on-the-fly
+compression. To give a reference, with used Naive blocks with
+\emph{offline} compression, that it, we wrote an uncompressed image
+and compressed it sequentially afterwards. \todo{confirm that.}
+
 \begin{table}
@@
 Figure~\ref{fig:compression} shows the impact of gzip compression on
-the merging time, for Multiple reads and Naive slices. As explained
-before, Clustered reads could only be applied to compressed data while
-in Case 3, i.e., for 16~GB, and Naive blocks could not be
-used. Compression dramatically slows down merging time for all
-algorithms \todo{Check that once all results are available}.
+the merging time, for Multiple reads and Naive slices.
+Unsurprisingly, compression dramatically slows down merging time for
+all algorithms. Multiple reads, however, still provide important
+speed-up compared to Naive blocks \todo{check that}, although they do
+not completely reach the performance of Buffered slices and Naive
+slices.
 
@@
 components might also result in performance differences particular to
-a particular algorithm. To ensure the portability of our library
+a specific algorithm. To ensure the portability of our library
 across systems and configurations, we focused on reducing the overall
@@
 formats may be split and merged without performance loss compared to
-more complex formats that try to preserve spatial locality on disk,
+more complex formats that can preserve spatial locality on disk,
 for instance MINC 2.0 or the format based on space-filling curves

### 2017-08-18 paper.tex
Minor fixes

@@
   analyze sequential algorithms that leverage memory buffering, to
-  address this issue. Clustered reads and Clustered writes access image
-  chunks only once, but they have to seek in the reconstructed
+  address this issue. Clustered reads and Clustered writes access
+  image chunks only once, but they have to seek in the reconstructed
   image. Multiple reads and Multiple writes minimize seeks in the
@@
   merging) provided that enough memory is available, which solves our
-  initial problem to a large extent.  We conclude that splitting and
-  merging large 3D images can be done efficiently without the need for
-  complex data formats.
+  initial problem to a large extent. Besides, Multiple reads
+  support on-the-fly compression of the merged image transparently but
+  Cluster reads do not due to their use of negative seeking. We
+  conclude that splitting and merging large 3D images can be done
+  efficiently without relying on complex data formats.
 \end{abstract}
@@
 increasingly found in a variety of disciplines. Big~Brain, for
-instance, is a 3D histological image of the
-human brain that represents 1~TB of raw data organized in 3600 planes
-at full resolution and 76~GB at a 40-micrometer isotropic resolution
-commonly used in neurosciences~\cite{amunts2013bigbrain}. Future
-versions of this dataset will be acquired at an even higher resolution
-and will represent 100~TB of raw data. Other examples found in medical
-imaging, our primary domain of interest, include high-resolution 3D
-electron microscopy (see, e.g., ~\cite{bock2011network}) or micro- and
+instance, is a 3D histological image of the human brain that
+represents 1~TB of raw data organized in 3600 planes at full
+resolution and 76~GB at a 40-micrometer isotropic resolution commonly
+used in neurosciences~\cite{amunts2013bigbrain}. Other examples found
+in medical imaging, our primary domain of interest, include
+high-resolution 3D electron microscopy (see, e.g.,
+~\cite{bock2011network}) or micro- and
 nano-tomography~\cite{10.1371/journal.pone.0035691}. As such images
@@
 seeks. In this paper we introduce and compare a family of sequential
-algorithms to split and merge images with limited numbers of seeks.
+algorithms to split and merge images with reduced seeking.
 
@@
 platforms~\cite{miguet1991elastic,tang2007eman2,yang2008parallel,braunl2013parallel,moise2013terabyte,bajcsy2013terabyte},
-but methods have focused on geometrical approaches to partition images
-and on load-balancing and on task scheduling techniques. Instead, we
+but methods have focused on geometrical approaches to partition images,
+and on load-balancing or task scheduling techniques. Instead, we
 aim at algorithms to efficiently split or merge images regardless of
@@
 To summarize, our paper makes the following contributions. (1) We
-propose a set of algorithms to split and merge terabyte-scale 3D
-images from 3D blocks or 2D slices. (2) We determine the complexity of
-those algorithms in terms of numbers of seeks, as a function of the image size, splitting scheme and available memory.
-(3)We evaluate our algorithms using the Big~Brain dataset and two storage drives with different characteristics.
-
-Section~\ref{sec:algos} presents our algorithms,
-Section~\ref{sec:implementation} describes their implementation,
-Section~\ref{sec:experiments} reports experimental results and
-Section~\ref{sec:discussion} concludes the paper.
+propose a set of algorithms to split and merge large 3D images from 3D
+blocks or 2D slices. (2) We determine the complexity of those
+algorithms in terms of numbers of seeks, as a function of the image
+size, splitting scheme and available memory. (3) We evaluate our
+algorithms using the Big~Brain dataset and two storage drives with
+different characteristics.  Section~\ref{sec:algos} presents our
+algorithms, Section~\ref{sec:implementation} describes their
+implementation, Section~\ref{sec:experiments} reports experimental
+results and Section~\ref{sec:discussion} concludes the paper.
 
@@
 and its seek time. For common file sizes, seek time is negligible
-compared to read or write time as typical seek times range from about 0.1~ms
-for Solid-State Drives (SSD) to 10~ms for Hard-Disk
-Drives (HDD). However, as we will shown later, naive algorithms might
-seek up to $10^7$ times to merge a high-resolution image, which
-renders seek time comparable to read and write times. In addition,
-extensive seeking also has an effect on read and write rates, as these
-are typically increasing with the duration of uninterrupted reads or
+compared to read or write time as typical seek times range from about
+0.1~ms for Solid-State Drives (SSD) to 10~ms for Hard-Disk Drives
+(HDD). However, as we will shown later, naive algorithms might seek up
+to $10^7$ times to merge a high-resolution image, which renders total
+seek time comparable to read and write times. In addition, extensive
+seeking also has an effect on read and write rates, as these are
+typically increasing with the duration of uninterrupted reads or
 writes.
 
-In our complexity analysis, we do not distinguish between access time
+In our analysis, we do not distinguish between access time
 and seek time.  We also assume that seeks require a constant amount of
@@
     \input{figures/svg/buffer.pdf_tex}
-    \caption{Buffer used in cluster reads (d=4).  White portions in
+    \caption{Buffer used in Clustered reads (d=4).  White portions in
       the buffer are not allocated. }
@@
 
-The number of seeks performed by cluster reads depends on how blocks
+The number of seeks performed by Clustered reads depends on how blocks
 loaded in memory arrange in the reconstructed image. In the best case,
@@
 \input{figures/svg/case3.pdf_tex}
-\caption{Memory-load configurations in cluster reads, leading to
+\caption{Memory-load configurations in Clustered reads, leading to
   different number of seeks. Red blocks need seeking before each of
@@
 \end{figure}
-Our cluster reads algorithm focuses on the three memory load
+Clustered reads focus on the three memory load
 configurations represented in Figure~\ref{fig:cluster-reads}, that is,
@@
 \caption{Configurations that increase the number of seeks per
-  memory-load and are thus deliberately avoided by cluster
+  memory-load and are thus deliberately avoided by Clustered
   reads. Left: configuration with incomplete block rows (multiplies
@@
 \begin{algorithm}[h]
-  \caption{Buffered merging of blocks with cluster reads}
+  \caption{Buffered merging of blocks with Clustered reads}
   \label{algo:cluster-reads}
@@
 \end{equation*}
-The number of seeks performed by cluster reads in each of the three cases is:
+The number of seeks performed by Clustered reads in each of the three cases is:
 \begin{equation*}
@@
 Finally, the total number of seeks
-performed by cluster reads to reconstruct the image is:
+performed by Clustered reads to reconstruct the image is:
 \begin{equation}
@@
 
+\todo{writing in this section could be improved.}
 We implemented the 5 algorithms presented earlier in a
@@
 As expected, the difference between Naive blocks and Naive slices is
-tremendous, in the order of 50 million seeks. For 3~GB, the number of
-seeks in Clustered reads is 4 orders of magnitude higher than for
-Multiple reads, and 5 orders of magnitude higher than Naive
-slices. This huge difference comes from the fact that at 3~GB Clustered
-reads are in Case 1. For 6~GB, 9~GB and 12~GB, Clustered reads are in
-case 2 and the difference with Multiple reads and Buffered slices
-reduces. At 16~GB, all algorithms perform the same.
+tremendous, in the order of 50 million seeks. Regarding the difference between
+Clustered reads and Multiple reads, it is consistent with our analysis. For
+3~GB, the number of seeks in Clustered reads is 4 orders of magnitude
+higher than for Multiple reads, and 5 orders of magnitude higher than
+Naive slices. This huge difference comes from the fact that at 3~GB
+Clustered reads are in Case 1. For 6~GB, 9~GB and 12~GB, Clustered
+reads are in Case 2 and the difference with Multiple reads and
+Buffered slices reduces. At 16~GB, all algorithms perform the same.
 
@@
 % Problem solved!
-Clustered reads and multiple reads are able to reduce to a negligible
+Clustered reads and Multiple reads reduce to a negligible
 amount the overall seek time required to split or merge 3D blocks in a
 high-resolution image where data is stored linearly. Both algorithms
-performed equivalently on our benchmark and compared to the reference
+performed equivalently and compared to the reference
 configuration where slices are merged or split without seeking. Our
@@
 %Same algo works for blocks or slices.
+
+% File formats
+Our results demonstrate that simple imaging
+formats may be split and merged without performance loss compared to
+more complex formats that can preserve spatial locality on disk,
+for instance MINC 2.0 or the format based on space-filling curves
+mentioned in~\cite{burns2013open}. This is of major interest in the
+current open-science context since simpler formats favor data-sharing
+and interoperability.  Moreover, our algorithms could potentially be
+adapted to any split geometry, even though we demonstrated them on
+slices and blocks only, while file formats inevitably assume a
+particular geometry. For instance, the format in~\cite{burns2013open}
+is not designed to naively split slices. However, we aimed to extract
+all the blocks, whereas, ~\cite{burns2013open} aimed to extract a
+single block from a large image. MINC 2.0 might also help with
+on-the-fly compression for Clustered reads.
+
+% Compression
+On-the-fly compression, indeed, could not be used with Clustered reads
+due to their use of negative seeking. This was not a problem for
+Multiple reads since Multiple reads completely remove seeking in the
+large image. Likewise, Multiple writes would not benefit from
+techniques aiming at accelerating random access reads in compressed
+images since they do not need to seek in it.
 
@@
 
-% File formats
-In terms of file formats, our results demonstrate that simple imaging
-formats may be split and merged without performance loss compared to
-more complex formats that can preserve spatial locality on disk,
-for instance MINC 2.0 or the format based on space-filling curves
-mentioned in~\cite{burns2013open}. This is of major interest in the
-current open-science context since simpler formats favor data-sharing
-and interoperability.  Moreover, our algorithms could potentially be
-adapted to any split geometry, even though we demonstrated them on
-slices and blocks only, while file formats inevitably assume a
-particular geometry. For instance, the format in~\cite{burns2013open}
-is not designed to naively split slices. However, we aimed to extract
-all the blocks, whereas, ~\cite{burns2013open} aimed to extract a
-single block from a large image. MINC 2.0 might also help with
-on-the-fly compression for Clustered reads.
-
-% Compression
-On-the-fly compression, indeed, could not be used with Clustered reads
-due to their use of negative seeking. This was not a problem for
-Multiple reads since Multiple reads completely remove seeking in the
-large image. Likewise, Multiple writes would not benefit from
-techniques aiming at accelerating random access reads in compressed
-images since they do not need to seek in it. Nonetheless, in our
-experimental setup compression drastically slowed down merging
-algorithms and proved of little use unless storage was limited.
-
 % Parallel algorithms
@@
 beneficial in case an image already split needs to be split in a
-different geometry. Clustered reads and multiple reads could be used
-as starting points for such algorithms.
+different geometry. Designing such algorithms is part of our future
+work. Clustered reads and multiple reads could be used as starting
+points.
 
 Our \texttt{sam} library is available at
-\url{https://github.com/big-data-lab-team/sam}, under MIT license.
-
-% Ideas for the future: measure the actual amount of consumed memory.
+\url{https://github.com/big-data-lab-team/sam} under MIT license.
 

### 2017-08-18 scripts/experiment/compression/compressed_blocks.py
naive block offline compression

@@ -1,0 +1,93 @@
+#!/usr/bin/env python
+# Ref: imageutils.py from
+# https://github.com/big-data-lab-team/sam/blob/master/imageutils.py
+import imageutils as img_utils
+import numpy as np
+from time import time
+import argparse
+import random
+import os
+
+# example
+# ./compressed_blocks.py -m 0  -r 2 -d ssd
+
+
+
+# on the consider
+reconstructed_hdd = "/data/gao/new_image.nii"
+legend_hdd = "/data/gao/blocks125/legend.txt"
+
+reconstructed_ssd = "/home/gao/new_image.nii"
+legend_ssd = "/home/gao/blocks125/legend.txt"
+
+reconstructed_ssd_compressed = "/home/gao/new_image.nii.gz"
+
+csv_file_ssd_slices = "./blocks_ssd_compressed.dat"
+csv_file_hdd_slices = "./blocks_hdd_compressed.dat"
+
+first_dim=3850
+second_dim=3025
+third_dim=3500
+
+files = {
+    "hdd": (reconstructed_hdd, legend_hdd, csv_file_hdd_slices),
+    "ssd": (reconstructed_ssd, legend_ssd, csv_file_ssd_slices)
+}
+
+
+def benchmark_creads(mem, reconstructed, legend):
+    img = img_utils.ImageUtils(reconstructed, first_dim, second_dim, third_dim, np.uint16)
+    s_time = time()
+    total_read_time, total_write_time, total_seek_time, total_seek_number = img.reconstruct_img(legend, "clustered", mem, input_compressed=True, benchmark=True)
+    e_time = time()
+    print "before compressing takes ", e_time - s_time
+    print "offline compressing..."
+    os.system("gzip {}".format(reconstructed))
+    compressed_e_time = time()
+    compressing_time = compressed_e_time - e_time
+    print "compressing_time", compressing_time
+    total_time = compressed_e_time - s_time
+    print total_read_time, total_write_time, total_seek_time, total_seek_number, compressing_time, total_time
+    return (total_read_time, total_write_time, total_seek_time, total_seek_number, compressing_time, total_time)
+
+def write_to_file(data_dict, dat_file):
+    # (total_read_time, total_write_time, total_seek_time, total_seek_number, total_time)
+    print "saved to ", dat_file
+    with open(dat_file, "a") as f:
+        for k in sorted(data_dict.keys()):
+            for e in data_dict[k]:
+                f.write(str(e) + " ")
[…]

### 2017-08-18 scripts/experiment/compression/compressed_blocks.py
add 4 more repetition on naive block offline compression

@@ -81,11 +81,12 @@ def main():
             print "---------------------- mem = {} ----------------------".format(mem)
             os.system("echo 3 | sudo tee /proc/sys/vm/drop_caches")
             os.system("rm {}".format(files[disk][0]))
+            os.system("rm {}".format(reconstructed_ssd_compressed))
 
             data = benchmark_creads(mem=mem, reconstructed=files[disk][0], legend=files[disk][1])
             data_dict[mem] = data
 
-            os.system("rm {}".format(reconstructed_ssd_compressed))
+
 
         write_to_file(data_dict, files[disk][2])
 

### 2017-08-19 paper.tex
removed todo on sam tagging
Updates on compressed data

@@
   perform very poorly to split or merge 3D blocks from ultra-high
-  resolution images, due to important seek times. We introduce and
+  resolution images, due to seek times. We introduce and
   analyze sequential algorithms that leverage memory buffering, to
@@
   support on-the-fly compression of the merged image transparently but
-  Cluster reads do not due to their use of negative seeking. We
+  Clustered reads do not, due to their use of negative seeking. We
   conclude that splitting and merging large 3D images can be done
@@
 aim at algorithms to efficiently split or merge images regardless of
-the geometry of the chunks. Although important seek times are often
+the geometry of the chunks. Although seek times are often
 identified as an issue, the preferred solution is usually to optimize
@@
 false.  Function \texttt{sizeof} (line 8) returns the actual memory
-used by its argument, including only its allocated segments in case
-the argument is a buffer.
+used by its argument, including only its allocated segments in the
+case that the argument is a buffer.
 \begin{algorithm}[h]
@@
 
-\todo{writing in this section could be improved.}
 We implemented the 5 algorithms presented earlier in a
@@
 image. In Clustered reads, some seeking might be required between
-writes. We implemented a defragmentation procedure for the dictionary
-that merges contiguous dictionary entries in a single one, but we
-abandoned it as it proved more time-consuming than going through all
-the initial entries, due to the overhead of resizing NumPy arrays to
-merge entries. In Multiple reads, dictionary entries are always
-contiguous in the reconstructed image.  We tried to use a single NumPy
-array as a buffer, but we finally abandoned it as inserting data at a
-specific position in a NumPy array copies the data in memory, which
-increases both the execution time and the peak memory consumption.
+writes. In Multiple reads, dictionary entries are always contiguous in
+the reconstructed image.  We tried to use a single NumPy array as a
+buffer, but we finally abandoned it as inserting data at a specific
+position in a NumPy array copies the data in memory, which increases
+both the execution time and the peak memory consumption.  We also
+implemented a defragmentation procedure for the dictionary that merges
+contiguous dictionary entries in a single one, but we abandoned it as
+it proved more time-consuming than going through all the initial
+entries, due to the overhead of resizing NumPy arrays to merge
+entries.
 
@@
 converted them to NiFTI 1.0 using Nibabel and left them uncompressed.
-We generated NiFTI slices of size 3850x3025x28 from the reconstructed
+We generated NiFTI slabs of size 3850x3025x28 from the reconstructed
 image, using our naive split algorithm.
@@
 
-We used Git tag 0.1 of our \texttt{sam} library \todo{tag it}. Our
+We used Git tag 0.1.1 of our \texttt{sam} library. Our
 experiment scripts are available at
@@
 \emph{offline} compression, that it, we wrote an uncompressed image
-and compressed it sequentially afterwards. \todo{confirm that.}
+and compressed it sequentially afterwards.
 
@@
 % Clustered reads
-Clustered reads provide important speed ups compared to Naive blocks,
+Clustered reads provide substantial speed ups compared to Naive blocks,
 both on HDD and on SSD. They are 6.8 times faster than Naive blocks on
@@
 time and the write time, which suggests that seeking degrades the
-write rate importantly in addition to introducing extra delays.
+write rate in addition to introducing extra delays.
 Clustered reads reduce the seek time and thus the read time very
@@
 Unsurprisingly, compression dramatically slows down merging time for
-all algorithms. Multiple reads, however, still provide important
-speed-up compared to Naive blocks \todo{check that}, although they do
+all algorithms. Multiple reads, however, still provide 
+speed-up compared to Naive blocks, although they do
 not completely reach the performance of Buffered slices and Naive

### 2017-08-20 paper.tex
Minor fixes

@@
 blocks), 0\% (Naive slices), 3.3\% (Clustered reads), 26.8\% (Multiple
-reads) and 0.9\% (Buffered slices), presumably explained by the fact
-that the model assumes cubic blocks while we used non-cubic ones in
+reads) and 0.9\% (Buffered slices), explained by the fact
+that the model assumes cuboid blocks while we used non-cuboid ones in
 the experiment. For Multiple reads, our complexity analysis also
 assumed that one of the 5 Cases in
-Figure~\ref{fig:multiple-reads-cases} was used while they are usually
+Figure~\ref{fig:multiple-reads-cases} was used while they are 
 blended in practice. Overall, the model correctly explains the
@@
 As expected, the difference between Naive blocks and Naive slices is
-tremendous, in the order of 50 million seeks. Regarding the difference between
-Clustered reads and Multiple reads, it is consistent with our analysis. For
+tremendous, in the order of 50 million seeks. The difference between
+Clustered reads and Multiple reads is consistent with our analysis. For
 3~GB, the number of seeks in Clustered reads is 4 orders of magnitude
@@
 HDD (6.7 times on SSD), which quantifies the effect of the targeted
-problem. It also shows that SSDs might not be the only solution to
-reduce the impact of seek time. In the remainder we use Naive blocks
+problem.  In the remainder we use Naive blocks
 and Naive slices as references to evaluate our algorithms.
@@
 Clustered writes are slightly slower than Multiple writes until
-16~GB. Buffered slices don't provide any significant speed-up compared
+16~GB. Buffered slices provide no speed-up compared
 to Naive slices.
@@
 % File formats
-Our results demonstrate that simple imaging
+Our results demonstrate that large images stored in simple 
 formats may be split and merged without performance loss compared to
@@
 different geometry. Designing such algorithms is part of our future
-work. Clustered reads and multiple reads could be used as starting
-points.
+work, in which Clustered reads and Multiple reads will be used as
+starting points.
 

### 2017-08-20 scripts/bigbrain/minc2nifti.py
added scripts used for reconstructing bigbrain

@@ -1,0 +1,97 @@
+import nibabel as nib
+import h5py
+import os
+import numpy as np
+import sys
+import argparse
+
+
+def get_dimensions(var):
+    try:
+        dimorder = var.attrs['dimorder']
+    except KeyError:  # No specified dimensions
+        return []
+    return dimorder.split(',')
+
+def convert2nifti(in_folder, out_folder, dtype, gzip=False):
+    for mnc_block in os.listdir(in_folder):
+
+        if mnc_block[-4:] != ".mnc":
+            continue
+
+        filepath = os.path.join(in_folder, mnc_block)
+        block_header = h5py.File(filepath, 'r')
+
+        minc_part = block_header['minc-2.0']
+        # The whole image is the first of the entries in 'image'
+        image = minc_part['image']['0']
+        image_data = image['image']
+        dim_names = get_dimensions(image_data)
+        dimensions = minc_part['dimensions']
+
+        dims = [dimensions[s].attrs.items() for s in dim_names]
+
+        ydim = dims[0][0][1]
+        zdim = dims[1][0][1]
+        xdim = dims[2][0][1]
+
+        ystep = dims[0][11][1]
+        zstep = dims[1][10][1]
+        xstep = dims[2][10][1]
+
+        ystart = dims[0][12][1]
+        zstart = dims[1][11][1]
+        xstart = dims[2][11][1]
+
+
+        data = nib.load(filepath).get_data().astype(dtype) 
+        nifti = nib.Nifti1Image(data, np.eye(4))
+        
+        nifti.header['descrip'] =  '{0} {1} {2}'.format(ystart, zstart, xstart)
+        
+        nifti.header['pixdim'][1] = ystep
+        nifti.header['pixdim'][2] = zstep
+        nifti.header['pixdim'][3] = xstep
+
+        print nifti.header['pixdim']
+
+        nii_file = mnc_block[:-4] + '.nii'
+        
+        if gzip:
+            nii_file += '.gz'
+        
+        nib.save(nifti, os.path.join(out_folder, nii_file))
+
+
+if __name__ == '__main__':
+    parser = argparse.ArgumentParser(description='Convert minc images to nifti images')
+
+    parser.add_argument('inpt', type=str, help="Input folder")
+    parser.add_argument('outpt', type=str, help="Output folder")
+    parser.add_argument('dtype', type=str, help="Numpy datatype \
+                                                            (np.int16, np.uint16, np.float32,\
[…]

### 2017-08-20 scripts/bigbrain/reconstruct_bb.py

@@ -1,0 +1,125 @@
+import nibabel as nib
+import os
+import h5py
+import numpy as np
+import math
+import sys
+import argparse
+from time import time
+
+
+def reconstruct(legend_fn, reconstructed_fn, block_folder, block_prefix, block_suffix, bytes_per_voxel):
+
+    legend = nib.load(legend_fn).get_data()
+    reconstructed_img = nib.load(reconstructed_fn)
+
+
+    ystart_0 = 0
+    zstart_0 = 0
+    xstart_0 = 0
+
+    blocks_copied =  {}
+
+    bb_header = reconstructed_img.header
+
+    try:
+      header_size = bb_header.single_vox_offset
+    except:
+      print 'ERROR: File not a NIfTI image'
+      sys.exit(1)
+
+    bb_ydim = bb_header.get_data_shape()[0]
+    bb_zdim = bb_header.get_data_shape()[1]
+    bb_xdim = bb_header.get_data_shape()[2]
+
+    with open(reconstructed_fn, "r+b") as reconstructed:
+        for x in range(0, legend.shape[2]):
+            for y in range(0, legend.shape[0]):
+                for z in range(0, legend.shape[1]):
+                   
+                    block_num = str(int(legend[y][z][x])).zfill(3) 
+                    
+                    block_filename = '{0}-0{1}-{2}'.format(os.path.join(block_folder,block_prefix), block_num, block_suffix)
+                    
+                    if block_num in blocks_copied:
+                        continue
+                    else:
+                        blocks_copied[block_num] = 1
+
+                    block_img = nib.load(block_filename)
+                    header = block_img.header
+                    shape = header.get_data_shape()
+                    ydim = shape[0]
+                    zdim = shape[1]
+                    xdim = shape[2]
+
+                    block_data = block_img.get_data()
+
+                    start = header['descrip'].tostring().strip('\x00').split()
+                    step = header['pixdim']
+
+                    ystep = round(float(step[1]), 2)
+                    zstep = round(float(step[2]), 2)
+                    xstep = round(float(step[3]), 2)
+                   
+
+                    ystart = float(start[0].strip())
+                    zstart = float(start[1].strip())
+                    xstart = float(start[2].strip())
+
+                    #get first block's start values to compare position with other blocks
+                    if y == 0 and z == 0 and x == 0 :
+                        ystart_0 = ystart
[…]

### 2017-08-20 scripts/bigbrain/minc2nifti.py
added reference

@@ -5,7 +5,8 @@ import numpy as np
 import sys
 import argparse
 
-
+# slightly modified version of _get_dimensions function 
+# obtained from  https://github.com/nipy/nibabel/blob/master/nibabel/minc2.py
 def get_dimensions(var):
     try:
         dimorder = var.attrs['dimorder']

### 2017-11-05 paper.tex
minor fixes

@@
   Splitting and merging data is a requirement for most parallel or
-  distributed processing operations. Naive algorithms, however,
-  perform very poorly to split or merge 3D blocks from ultra-high
-  resolution images, due to seek times. We introduce and
-  analyze sequential algorithms that leverage memory buffering, to
-  address this issue. Clustered reads and Clustered writes access
+  distributed processing operations. Naive algorithms to split and
+  merge 3D blocks from ultra-high resolution images perform very poorly,
+  as a result of seek times. In contrast, naive algorithms to split and 
+  merge 2D slices or 3D slabs perform optimally as seek time is significantly minimized.
+  We introduce and analyze sequential algorithms (Clustered reads, Multiple reads, Clustered write and Multiple writes)
+  that leverage memory buffering to
+  address this issue. Clustered reads and Clustered writes, access
   image chunks only once, but they have to seek in the reconstructed
@@
   times. Evaluation on a 3850x3025x3500 brain image shows that our
-  algorithms perform similarly to the optimal configuration (slice
-  merging) provided that enough memory is available, which solves our
-  initial problem to a large extent. Besides, Multiple reads
-  support on-the-fly compression of the merged image transparently but
-  Clustered reads do not, due to their use of negative seeking. We
+  algorithms perform similarly to the optimal configuration provided that enough memory is available. 
+  Additionally, Multiple reads support on-the-fly compression of the 
+  merged image transparently but Clustered reads do not, due to their use of negative seeking. We
   conclude that splitting and merging large 3D images can be done
@@
 decision to split an image into slices or blocks, and the size of the
-chunks, is up to the application. Some applications, for instance spatial filtering, would
-commonly require blocks while other
-ones, such as acquisition artifact removal, would rather work on
-slices. Applications that process voxels individually, for instance
+chunks, is up to the application. For instance, spatial filtering would
+commonly require blocks, whereas slices would be preferable for acquisition artifact removal. 
+Applications that process voxels individually, for instance
 histogram computation or k-means clustering, could work on either
@@
 -- \url{https://nifti.nimh.nih.gov}) store the complete image in
-column-major order, that is, for an image of dimensions (i, j, k), i
-is the fastest changing dimension, followed by dimension j, followed
-by dimension k. Other formats, for instance HDF5-based MINC
+column-major order, that is, elements belonging to the same column are
+stored in a contiguous order. Other formats, such as HDF5-based MINC
 2.0~\cite{vincent2016minc}, provide more flexibility by allowing data
@@
 proximity on disk. On the contrary, we are searching for algorithms
-that would reduce the seek time regardless of the data format, so that
-applications with flexible splitting schemes can be served
+that would reduce the seek time regardless of the data format, such that
+applications with flexible splitting schemes can be served.
 
@@
 
-The buffer might be a slightly complex data structure, for instance an
-associative array or a Python dictionary, capable of storing multiple
+The buffer, capable of storing multiple
 disjoint sequences of contiguous bytes without having to allocate
 memory for the bytes between such
-sequences. Figure~\ref{fig:cluster-reads-buffer} illustrates how the
+sequences, can be represented by an
+associative array or a Python dictionary.
+Figure~\ref{fig:cluster-reads-buffer} illustrates how the
 buffer would fill up for the two first blocks in a reconstructed
-image, assuming that blocks are of size 4x4x4.
+image, assuming, for the sake of this particular example, that blocks are of size 4x4x4.
 
@@
 Clustered reads focus on the three memory load
-configurations represented in Figure~\ref{fig:cluster-reads}, that is,
+configurations represented in Figure~\ref{fig:cluster-reads}:
 the amount of memory $m'$ used by the algorithm is rounded down to the
@@
 gzip library. Compression is done on the fly, that is, while the data
-is being read or written. On-the-fly compression of large datasets is
+is being read or written during splitting and merging. 
+On-the-fly compression of large datasets is
 a challenge when extensive seeking is involved as the gzip library has
@@
 \subsection{Data}
-We used the 3850x3025x3500 Big~Brain image split in 125
-non-overlapping chunks of size 770x605x700 with 2 bytes per voxel
-(total size uncompressed is 75.92~GB). Big~Brain is a reference brain
+We used the 3850x3025x3500 Big~Brain image split into 125
+non-overlapping chunks of size 770x605x700, with 2 bytes per voxel
+(total size uncompressed is 75.92~GB). The Big~Brain image was also split
+into 125 non-overlapping slabs of size 3850x3025x28 for our experiments.
+
+Big~Brain is a reference brain
 based on the reconstruction of 7404 histological sections at nearly
@@
 converted them to NiFTI 1.0 using Nibabel and left them uncompressed.
-We generated NiFTI slabs of size 3850x3025x28 from the reconstructed
-image, using our naive split algorithm.
+These blocks were then used to reconstruct the Big~Brain that was split into our desired block and slab
+configurations using our naive splitting algorithms.
 
@@
 
-Compressed blocks and slices were also merged into a compressed
+Compressed blocks and slabs were also merged into a compressed
 image. On-the-fly gzip compression was used for Naive slices, Buffered
@@
 and Naive blocks could not be used at all with on-the-fly
-compression. To give a reference, with used Naive blocks with
-\emph{offline} compression, that it, we wrote an uncompressed image
+compression. To give a reference, we used Naive blocks with
+\emph{offline} compression, that is, we wrote an uncompressed image
 and compressed it sequentially afterwards.
@@
 
+It should be noted that in some cases, splitting and merging can be considered to be an independent process
+from the other tasks. As such, the performance impact splitting and 
+merging has on the main tasks depends largely on the duration of the main tasks.
+
 % Several other factors might influence performance

### 2017-11-09 paper.tex
Removed paragraph on task duration
convert slice to slab

@@
 We assume that the high-resolution image is split into chunks
-representing 3D blocks or 2D slices that fit in memory. A dataset such
+representing 3D blocks, 3D slabs, or  2D slices that fit in memory. A dataset such
 as Big~Brain would perhaps be split into 125 chunks of 600~MB. The
-decision to split an image into slices or blocks, and the size of the
+decision to split an image into slices, slabs or blocks, and the size of the
 chunks, is up to the application. For instance, spatial filtering would
@@
 propose a set of algorithms to split and merge large 3D images from 3D
-blocks or 2D slices. (2) We determine the complexity of those
+blocks, 3D slabs or 2D slices. (2) We determine the complexity of those
 algorithms in terms of numbers of seeks, as a function of the image
@@
 \item $b$: number of bytes per voxel (in B).
-\item $n$: number of chunks (blocks or slices).
+\item $n$: number of chunks (blocks or slabs).
 \item $m$: amount of available memory (in B).
@@
 
-\subsection{Slices vs blocks}
+\subsection{Slabs vs blocks}
 
@@
     \caption{Notations. A \emph{block row} is shown in red. A
-      \emph{block slice} is shown in blue.}
+      \emph{block slab} is shown in blue.}
     \label{fig:notations}
@@
 \end{figure}
-Algorithms~\ref{algo:naive-slices} and~\ref{algo:naive-blocks} show
-the naive merging methods for slices and blocks. These algorithms
+Algorithms~\ref{algo:naive-slabs} and~\ref{algo:naive-blocks} show
+the naive merging methods for slabs and blocks. These algorithms
 actually have very different complexities even though blocks and
-slices have identical sizes. Since slices are stored contiguously in
+slabs have identical sizes. Since slabs are stored contiguously in
 the reconstructed image, the number of seeks in
-Algorithm~\ref{algo:naive-slices} is only $2n$ as $n$ seeks are
-required to read the slices and $n$ seeks are required to write them:
+Algorithm~\ref{algo:naive-slabs} is only $2n$ as $n$ seeks are
+required to read the slabs and $n$ seeks are required to write them:
 \begin{equation}
-  N_\mathrm{slices} = 2n \label{eq:naive-blocks}
+  N_\mathrm{slabs} = 2n \label{eq:naive-blocks}
 \end{equation}
@@
 \begin{equation}
-N_\mathrm{blocks} = n+n\left(\sqrt[3]{\frac{R}{n}}\right)^2 \label{eq:naive-slices}
+N_\mathrm{blocks} = n+n\left(\sqrt[3]{\frac{R}{n}}\right)^2 \label{eq:naive-slabs}
 \end{equation}
@@
 \begin{algorithm}
-\caption{Naive merging from slices}
-\label{algo:naive-slices} 
+\caption{Naive merging from slabs}
+\label{algo:naive-slabs} 
 \begin{algorithmic}
-  \FOR{each slice}
-    \STATE read slice
-    \STATE write slice in reconstructed image
+  \FOR{each slab}
+    \STATE read slab
+    \STATE write slab in reconstructed image
   \ENDFOR      
@@
 \end{algorithm}
-%%  For slices, the total merge time is modeled as follows:
+%%  For slabs, the total merge time is modeled as follows:
 %% \begin{equation}
-%% T_\mathrm{slices} = \frac{bR}{\alpha_r}+\frac{bR}{\alpha_w}+2\beta n
+%% T_\mathrm{slabs} = \frac{bR}{\alpha_r}+\frac{bR}{\alpha_w}+2\beta n
 %% \end{equation}
@@
 
-\subsection{Buffered slices}
-
-Algorithm~\ref{algo:naive-slices} is a particular case of memory
+\subsection{Buffered slabs}
+
+Algorithm~\ref{algo:naive-slabs} is a particular case of memory
 buffering where the amount of available memory equals the maximum size
@@
 available memory increases, as shown in
-Algorithm~\ref{algo:buffered-slices}.
+Algorithm~\ref{algo:buffered-slabs}.
 \begin{algorithm}[h]
-  \caption{Buffered merging from slices}
-  \label{algo:buffered-slices}
+  \caption{Buffered merging from slabs}
+  \label{algo:buffered-slabs}
   \begin{algorithmic}[1]
-    \STATE sorted\_slices = sort slices by increasing k values
+    \STATE sorted\_slabs = sort slabs by increasing k values
     \STATE initialize buffer
     \FOR{i = 0 ; i \textless n ; i+=1 }
-      \STATE slice = sorted\_slices[i]
-      \IF{sizeof(buffer)+sizeof(slice) $\geq$ m}
+      \STATE slab = sorted\_slabs[i]
+      \IF{sizeof(buffer)+sizeof(slab) $\geq$ m}
         \STATE write buffer in reconstructed image
@@
       \ENDIF
-      \STATE read slice and append it to buffer
+      \STATE read slab and append it to buffer
     \ENDFOR
@@
 \begin{equation}
-N_\mathrm{buff\_slices} =  n + \ceil*{\frac{bR}{m}} \label{eq:buff-slices}
+N_\mathrm{buff\_slabs} =  n + \ceil*{\frac{bR}{m}} \label{eq:buff-slabs}
 \end{equation}
-Buffered slices are straightforward to implement, however, their
+Buffered slabs are straightforward to implement, however, their
 extension to block merging is not easy. The remainder of this Section
@@
 
-Clustered reads are the more direct extension of Buffered slices to
+Clustered reads are the more direct extension of Buffered slabs to
 blocks: they load multiple blocks in memory, concatenate them in a
@@
 loaded in memory arrange in the reconstructed image. In the best case,
-complete contiguous slices of the reconstructed image can be assembled
+complete contiguous slabs of the reconstructed image can be assembled
 in memory and written in a single seek. In the worst case, the
@@
 O($d^2$) seeks are then required during writing, one for every partial
-row in every partial slice. In the intermediary case, rows are
-complete but some slices can only be partially reconstructed: O($d$)
+row in every partial slabs. In the intermediary case, rows are
+complete but some slabs can only be partially reconstructed: O($d$)
 seeks are then required.
@@
 closest number of complete blocks (Case 1), of complete block rows (Case
-2) or of complete block slices (Case 3). This is in general reasonable
+2) or of complete block slabs (Case 3). This is in general reasonable
 since adding an incomplete row to a set of complete ones multiplies
@@
   the number of seeks by $d$ compared to complete block rows). Right:
-  configuration with block rows that overlap multiple block slices
+  configuration with block rows that overlap multiple block slabs
   (multiplies the number of seeks by $2$ compared to non-overlapping
@@
 Our algorithm also avoids configurations where the memory load
-overlaps multiple block slices in Case 2 or multiple block rows in
+overlaps multiple block slabs in Case 2 or multiple block rows in
 Case 1, as such overlaps multiply the number of required seeks (see
@@
 the identifier of the selected case. Function \texttt{check\_overlap}
-(line 7) determines whether two blocks overlap multiple block slices
+(line 7) determines whether two blocks overlap multiple block slabs
 (case 2) or multiple block rows (Case 1). For case 3 it always returns
@@
 $\left( \sqrt[3]{n}^2 \right)$ and $x_2$ is proportional to the total number of
-block slices $\left( \sqrt[3]{n} \right)$.
+block slabs $\left( \sqrt[3]{n} \right)$.
 Finally, the total number of seeks
@@
 %% In case (1) on
-%% this Figure, the memory load completely fills rows and slices in the
+%% this Figure, the memory load completely fills rows and slabs in the
 %% reconstructed image. Thus it can be written with a single seek,
@@
 number $k$ of sub-rows (Case 1, $k_1<\sqrt[3]{n}$), of complete rows
-(Case 2, $k_2<d$), of tile rows (Case 3, $k_3<\sqrt[3]{n}$), of slices
-(Case 4, $k_4<d$) or of block slices (Case 5, $k_5<\sqrt[3]{n}$), as
+(Case 2, $k_2<d$), of tile rows (Case 3, $k_3<\sqrt[3]{n}$), of slabs
+(Case 4, $k_4<d$) or of block slabs (Case 5, $k_5<\sqrt[3]{n}$), as
 illustrated in Figure~\ref{fig:multiple-reads-cases}. In each of these
@@
 \url{https://github.com/big-data-lab-team/paper-sequential-split-merge/blob/master/scripts/experiment}. We
-used them to split and merge using Buffered slices, Clustered reads and
+used them to split and merge using Buffered slabs, Clustered reads and
 Multiple reads, with 3~GB, 6~GB, 9~GB, 12~GB and 16~GB of
@@
 and~\ref{eq:seeks-multiple-reads}. For instance, at 3~GB, Clustered reads were in
-Case 1. We also did a run with 0~GB of memory for Buffered slices and
-Clustered reads, which triggered Naive slices and Naive blocks. We did 5
+Case 1. We also did a run with 0~GB of memory for Buffered slabs and
+Clustered reads, which triggered Naive slabs and Naive blocks. We did 5
 repetitions for each memory value. Memory values were shuffled in each
@@
 Compressed blocks and slabs were also merged into a compressed
-image. On-the-fly gzip compression was used for Naive slices, Buffered
-slices and Multiple reads. As explained before, Clustered reads could
+image. On-the-fly gzip compression was used for Naive slabs, Buffered
+slabs and Multiple reads. As explained before, Clustered reads could
 only be applied to compressed data while in Case 3, i.e., for 16~GB,
@@
 repetitions. The average relative model errors are 12.7\% (Naive
-blocks), 0\% (Naive slices), 3.3\% (Clustered reads), 26.8\% (Multiple
-reads) and 0.9\% (Buffered slices), explained by the fact
+blocks), 0\% (Naive slabs), 3.3\% (Clustered reads), 26.8\% (Multiple
+reads) and 0.9\% (Buffered slabs), explained by the fact
 that the model assumes cuboid blocks while we used non-cuboid ones in
@@
 
-As expected, the difference between Naive blocks and Naive slices is
+As expected, the difference between Naive blocks and Naive slabs is
 tremendous, in the order of 50 million seeks. The difference between
@@
 higher than for Multiple reads, and 5 orders of magnitude higher than
-Naive slices. This huge difference comes from the fact that at 3~GB
+Naive slabs. This huge difference comes from the fact that at 3~GB
 Clustered reads are in Case 1. For 6~GB, 9~GB and 12~GB, Clustered
 reads are in Case 2 and the difference with Multiple reads and
-Buffered slices reduces. At 16~GB, all algorithms perform the same.
+Buffered slabs reduces. At 16~GB, all algorithms perform the same.
 
@@
 Figure~\ref{fig:merge-time} shows the merge time for all algorithms by
-memory values. Naive blocks are 9.5 times slower than Naive slices on
+memory values. Naive blocks are 9.5 times slower than Naive slabs on
 HDD (6.7 times on SSD), which quantifies the effect of the targeted
 problem.  In the remainder we use Naive blocks
-and Naive slices as references to evaluate our algorithms.
-
-% Buffered slices
-Buffered slices provide a negligible speed-up compared to Naive
-slices. In most cases, their memory overhead would not be worth the
+and Naive slabs as references to evaluate our algorithms.
+
+% Buffered slabs
+Buffered slabs provide a negligible speed-up compared to Naive
+slabs. In most cases, their memory overhead would not be worth the
 time gain.
@@
 down to read, write, seek and overhead time for our algorithms. Naive
-blocks and Naive slices are shown as references. The huge difference
-between Naive blocks and Naive slices is coming from both the seek
+blocks and Naive slabs are shown as references. The huge difference
+between Naive blocks and Naive slabs is coming from both the seek
 time and the write time, which suggests that seeking degrades the
@@
 substantially. Multiple reads almost anihilate the seek time and bring
-the read time to a value comparable to Naive slices.  The same
+the read time to a value comparable to Naive slabs.  The same
 behavior is observed on HDD and on SSD, although the effect of seeking
@@
     \includegraphics[width=0.45\columnwidth]{figures/benchmark-buff-slices/buff-slices-reads-breakdown-ssd.pdf}
-    \caption{Buffered slices}
+    \caption{Buffered slabs}
   \end{subfigure}
@@
 Figure~\ref{fig:split-time}. The difference between Naive blocks and
-Naive slices is still significant (average ratio is 1.4 on SSD, 2.0 on
+Naive slabs is still significant (average ratio is 1.4 on SSD, 2.0 on
 HDD) although less than for merging. On SSD, Clustered writes and
-Multiple writes both perform similarly to Naive slices. On HDD,
+Multiple writes both perform similarly to Naive slabs. On HDD,
 Clustered writes are slightly slower than Multiple writes until
-16~GB. Buffered slices provide no speed-up compared
-to Naive slices.
+16~GB. Buffered slabs provide no speed-up compared
+to Naive slabs.
 
@@
 the Nibabel library. Naive blocks are strongly penalized by important
-read times coming from extensive seeking. Buffered slices, Clustered
+read times coming from extensive seeking. Buffered slabs, Clustered
 writes and multiple writes all reduce the read time compared to Naive
 blocks, but they also increase the write time, most likely due to
-caching effects in Naive blocks and slices.
+caching effects in Naive blocks and slabs.
 
@@
     \includegraphics[width=0.45\columnwidth]{figures/benchmark-buff-slices/buff-slices-writes-breakdown-ssd.pdf}
-    \caption{Buffered slices}
+    \caption{Buffered slabs}
   \end{subfigure}
@@
 Figure~\ref{fig:compression} shows the impact of gzip compression on
-the merging time, for Multiple reads and Naive slices.
+the merging time, for Multiple reads and Naive slabs.
 Unsurprisingly, compression dramatically slows down merging time for
@@
 speed-up compared to Naive blocks, although they do
-not completely reach the performance of Buffered slices and Naive
-slices.
+not completely reach the performance of Buffered slabs and Naive
+slabs.
 
@@
 performed equivalently and compared to the reference
-configuration where slices are merged or split without seeking. Our
+configuration where slabs are merged or split without seeking. Our
 initial problem is solved.
@@
 adapted to any split geometry, even though we demonstrated them on
-slices and blocks only, while file formats inevitably assume a
+slabs and blocks only, while file formats inevitably assume a
 particular geometry. For instance, the format in~\cite{burns2013open}
-is not designed to naively split slices. However, we aimed to extract
+is not designed to naively split slabs. However, we aimed to extract
 all the blocks, whereas, ~\cite{burns2013open} aimed to extract a

### 2017-11-09 paper.tex
fix slab/slice more

@@
   as a result of seek times. In contrast, naive algorithms to split and 
-  merge 2D slices or 3D slabs perform optimally as seek time is significantly minimized.
-  We introduce and analyze sequential algorithms (Clustered reads, Multiple reads, Clustered write and Multiple writes)
+  merge 3D slabs perform optimally as seek time is significantly minimized.
+  We introduce and analyze sequential algorithms (Clustered reads, Multiple reads, Clustered writes, and Multiple writes)
   that leverage memory buffering to
@@
 We assume that the high-resolution image is split into chunks
-representing 3D blocks, 3D slabs, or  2D slices that fit in memory. A dataset such
+representing 3D blocks or 3D slabs that fit in memory. A 3D block
+consists of a stack of one or more 2D tiles (incomplete slices) spanning sequentially occuring slices, whereas a 3D slab is 
+a stack of one of more contiguous 2D slices. A dataset such
 as Big~Brain would perhaps be split into 125 chunks of 600~MB. The
-decision to split an image into slices, slabs or blocks, and the size of the
+decision to split an image into slabs or blocks, and the size of the
 chunks, is up to the application. For instance, spatial filtering would
@@
 closest number of complete blocks (Case 1), of complete block rows (Case
-2) or of complete block slabs (Case 3). This is in general reasonable
+2) or of complete block slices (Case 3). This is in general reasonable
 since adding an incomplete row to a set of complete ones multiplies
@@
   the number of seeks by $d$ compared to complete block rows). Right:
-  configuration with block rows that overlap multiple block slabs
+  configuration with block rows that overlap multiple block slices
   (multiplies the number of seeks by $2$ compared to non-overlapping
@@
 $\left( \sqrt[3]{n}^2 \right)$ and $x_2$ is proportional to the total number of
-block slabs $\left( \sqrt[3]{n} \right)$.
+block slices $\left( \sqrt[3]{n} \right)$.
 Finally, the total number of seeks
@@
 number $k$ of sub-rows (Case 1, $k_1<\sqrt[3]{n}$), of complete rows
-(Case 2, $k_2<d$), of tile rows (Case 3, $k_3<\sqrt[3]{n}$), of slabs
-(Case 4, $k_4<d$) or of block slabs (Case 5, $k_5<\sqrt[3]{n}$), as
+(Case 2, $k_2<d$), of tile rows (Case 3, $k_3<\sqrt[3]{n}$), of slices
+(Case 4, $k_4<d$) or of block slices (Case 5, $k_5<\sqrt[3]{n}$), as
 illustrated in Figure~\ref{fig:multiple-reads-cases}. In each of these
@@
 Clustered reads reduce the seek time and thus the read time very
-substantially. Multiple reads almost anihilate the seek time and bring
+substantially. Multiple reads almost annihilate the seek time and bring
 the read time to a value comparable to Naive slabs.  The same
@@
 
-It should be noted that in some cases, splitting and merging can be considered to be an independent process
-from the other tasks. As such, the performance impact splitting and 
-merging has on the main tasks depends largely on the duration of the main tasks.
-
 % Several other factors might influence performance

### 2017-11-10 figures/svg/export_svg_files.sh
fixed figures to be column-major

@@ -4,7 +4,7 @@ for name in buffer case1-a case2-a case3 incomplete-rows overlap mreads-case1 \
         mreads-case2 mreads-case3 mreads-case4 mreads-case5 Notations
 do
   #echo "Exporting ${name}.svg"
-  inkscape -D -z --file=figures/svg/${name}.svg --export-pdf=figures/svg/${name}.pdf --export-latex
-  sed s,${name}\.pdf,figures/svg/${name}\.pdf,g figures/svg/${name}.pdf_tex > figures/svg/${name}.pdf_tex.new
-  \mv figures/svg/${name}.pdf_tex.new figures/svg/${name}.pdf_tex
+  inkscape -D -z --file=$PWD/figures/svg/${name}.svg --export-pdf=$PWD/figures/svg/${name}.pdf --export-latex
+  sed s,${name}\.pdf,$PWD/figures/svg/${name}\.pdf,g $PWD/figures/svg/${name}.pdf_tex > $PWD/figures/svg/${name}.pdf_tex.new
+  \mv $PWD/figures/svg/${name}.pdf_tex.new $PWD/figures/svg/${name}.pdf_tex
 done

### 2017-11-10 paper.tex
minor fixes

@@
   algorithms perform similarly to the optimal configuration provided that enough memory is available. 
-  Additionally, Multiple reads support on-the-fly compression of the 
+  Additionally, Multiple reads supports on-the-fly compression of the 
   merged image transparently but Clustered reads do not, due to their use of negative seeking. We
@@
 representing 3D blocks or 3D slabs that fit in memory. A 3D block
-consists of a stack of one or more 2D tiles (incomplete slices) spanning sequentially occuring slices, whereas a 3D slab is 
-a stack of one of more contiguous 2D slices. A dataset such
+consists of a stack of one or more 2D tiles (incomplete slices and/or columns) spanning sequentially occuring slices, whereas a 3D slab is 
+a series of one of more contiguous 2D slices. A dataset such
 as Big~Brain would perhaps be split into 125 chunks of 600~MB. The
@@
 propose a set of algorithms to split and merge large 3D images from 3D
-blocks, 3D slabs or 2D slices. (2) We determine the complexity of those
+blocks or 3D slabs. (2) We determine the complexity of those
 algorithms in terms of numbers of seeks, as a function of the image
@@
 \item Number of slices/rows/columns in a block: $\sqrt[3]{\frac{R}{n}}=d$.
-\item Number of blocks in a block row: $\sqrt[3]{n}$.
+\item Number of blocks in a block column: $\sqrt[3]{n}$.
 \end{itemize}
@@
     \input{figures/svg/Notations.pdf_tex}
-    \caption{Notations. A \emph{block row} is shown in red. A
+    \caption{Notations. A \emph{block column} is shown in red. A
       \emph{block slab} is shown in blue.}
@@
 However,
-Algorithm~\ref{algo:naive-blocks} has to do extra seeks for each row
+Algorithm~\ref{algo:naive-blocks} has to do extra seeks for each column
 in each slice of each block:
@@
 \end{equation}
-Buffered slabs are straightforward to implement, however, their
+Buffered slabs is straightforward to implement, however, its
 extension to block merging is not easy. The remainder of this Section
@@
 
-Clustered reads are the more direct extension of Buffered slabs to
-blocks: they load multiple blocks in memory, concatenate them in a
-buffer and write the buffer in the reconstructed image. Seeking is
+Clustered reads is the more direct extension of Buffered slabs to
+blocks: it loads multiple blocks in memory, concatenates them in a
+buffer and writes the buffer in the reconstructed image. Seeking is
 reduced compared to Naive blocks since contiguous parts of the
@@
 in memory and written in a single seek. In the worst case, the
-memory load only partially covers rows in the reconstructed image:
+memory load only partially covers columns in the reconstructed image:
 O($d^2$) seeks are then required during writing, one for every partial
-row in every partial slabs. In the intermediary case, rows are
+column in every partial slabs. In the intermediary case, columns are
 complete but some slabs can only be partially reconstructed: O($d$)
@@
   different number of seeks. Red blocks need seeking before each of
-  their rows ($d^2$ seeks). Blue blocks need seeking before each of
+  their columns ($d^2$ seeks). Blue blocks need seeking before each of
   their slices ($d$ seeks). Green blocks need only a single
@@
 \end{figure}
-Clustered reads focus on the three memory load
+Clustered reads focuses on the three memory load
 configurations represented in Figure~\ref{fig:cluster-reads}:
 the amount of memory $m'$ used by the algorithm is rounded down to the
-closest number of complete blocks (Case 1), of complete block rows (Case
+closest number of complete blocks (Case 1), of complete block columns (Case
 2) or of complete block slices (Case 3). This is in general reasonable
-since adding an incomplete row to a set of complete ones multiplies
+since adding an incomplete columns to a set of complete ones multiplies
 the number of required seeks by $d$, as illustrated in
@@
   memory-load and are thus deliberately avoided by Clustered
-  reads. Left: configuration with incomplete block rows (multiplies
+  reads. Left: configuration with incomplete block columns (multiplies
   the number of seeks by $d$ compared to complete block rows). Right:
-  configuration with block rows that overlap multiple block slices
+  configuration with block columns that overlap multiple block slices
   (multiplies the number of seeks by $2$ compared to non-overlapping
@@
 Our algorithm also avoids configurations where the memory load
-overlaps multiple block slabs in Case 2 or multiple block rows in
+overlaps multiple block slabs in Case 2 or multiple block columns in
 Case 1, as such overlaps multiply the number of required seeks (see
@@
 
-Clustered reads are described in
+Clustered reads is described in
 Algorithm~\ref{algo:cluster-reads}. Function \texttt{switch} (line 3)
@@
 
-Multiple reads are shown in Algorithm~\ref{algo:multiple-reads}.  The
+Multiple reads is shown in Algorithm~\ref{algo:multiple-reads}.  The
 main idea of this algorithm is that blocks are read partially (line 9)
@@
 In the complexity analysis, we assume that $m'$ represents an integer
-number $k$ of sub-rows (Case 1, $k_1<\sqrt[3]{n}$), of complete rows
-(Case 2, $k_2<d$), of tile rows (Case 3, $k_3<\sqrt[3]{n}$), of slices
+number $k$ of sub-columns (Case 1, $k_1<\sqrt[3]{n}$), of complete columns
+(Case 2, $k_2<d$), of tile columns (Case 3, $k_3<\sqrt[3]{n}$), of slices
 (Case 4, $k_4<d$) or of block slices (Case 5, $k_5<\sqrt[3]{n}$), as
@@
 and~\ref{eq:seeks-multiple-reads} for different values of $n$. When
-Clustered reads are in Case 1 or 2, they may outperform Multiple reads
-only for large values of $n$. When Clustered reads are in case 3, they
-are equivalent to Multiple reads: assuming that $Rb$ is an exact
+Clustered reads is in Case 1 or 2, it may outperform Multiple reads
+only for large values of $n$. When Clustered reads is in case 3, it
+is equivalent to Multiple reads: assuming that $Rb$ is an exact
 multiple of $m$, Equations~\ref{eq:seeks-cluster-reads-1}
@@
 % Additional optimizations.
-In Multiple reads, block headers are read in a first pass where row
+In Multiple reads, block headers are read in a first pass where column
 indices in the reconstructed image are stored in memory. Those indices
@@
 \url{https://github.com/pauldmccarthy/indexed\_gzip}. Multiple writes
-are not impacted since they do not seek in the large image.
+is not impacted since it does not seek in the large image.
 
@@
 which renders Naive blocks unusable with compressed data and Clustered
-reads usable only in Case 3.  Again, Multiple reads are not impacted
-since they do not seek in the large image.
+reads usable only in Case 3.  Again, Multiple reads is not impacted
+since it does not seek in the large image.
 
@@
 % Buffered slabs
-Buffered slabs provide a negligible speed-up compared to Naive
+Buffered slabs provides a negligible speed-up compared to Naive
 slabs. In most cases, their memory overhead would not be worth the
@@
 % Clustered reads
-Clustered reads provide substantial speed ups compared to Naive blocks,
-both on HDD and on SSD. They are 6.8 times faster than Naive blocks on
+Clustered reads provides substantial speed ups compared to Naive blocks,
+both on HDD and on SSD. It is 6.8 times faster than Naive blocks on
 HDD and 5.1 times on SSD (average accross all repetitions, all memory
-values). Surprisingly, they perform substantially faster than Naive
+values). Surprisingly, it performs substantially faster than Naive
 blocks even at 3~GB, while in Case 1. This may be explained by the
@@
 % Multiple reads
-Multiple reads are even faster than Clustered reads on this
+Multiple reads is even faster than Clustered reads on this
 dataset. They are 8.4 times faster than Naive blocks on HDD and 5.3
@@
 write rate in addition to introducing extra delays.
-Clustered reads reduce the seek time and thus the read time very
-substantially. Multiple reads almost annihilate the seek time and bring
+Clustered reads reduces the seek time and thus the read time very
+substantially. Multiple reads almost annihilates the seek time and brings
 the read time to a value comparable to Naive slabs.  The same
@@
 Multiple writes both perform similarly to Naive slabs. On HDD,
-Clustered writes are slightly slower than Multiple writes until
-16~GB. Buffered slabs provide no speed-up compared
+Clustered writes is slightly slower than Multiple writes until
+16~GB. Buffered slabs provides no speed-up compared
 to Naive slabs.
@@
 read times coming from extensive seeking. Buffered slabs, Clustered
-writes and multiple writes all reduce the read time compared to Naive
+writes and Multiple writes all reduce the read time compared to Naive
 blocks, but they also increase the write time, most likely due to
@@
 Unsurprisingly, compression dramatically slows down merging time for
-all algorithms. Multiple reads, however, still provide 
-speed-up compared to Naive blocks, although they do
+all algorithms. Multiple reads, however, still provides 
+speed-up compared to Naive blocks, although it does
 not completely reach the performance of Buffered slabs and Naive
@@
 On-the-fly compression, indeed, could not be used with Clustered reads
-due to their use of negative seeking. This was not a problem for
-Multiple reads since Multiple reads completely remove seeking in the
+due to its use of negative seeking. This was not a problem for
+Multiple reads since Multiple reads completely removes seeking in the
 large image. Likewise, Multiple writes would not benefit from

### 2017-11-10 paper.tex
Small fixes in abstract and introduction

@@
 \begin{abstract}
-  Splitting and merging data is a requirement for most parallel or
+  Splitting and merging data is a requirement for many parallel or
   distributed processing operations. Naive algorithms to split and
@@
   Additionally, Multiple reads supports on-the-fly compression of the 
-  merged image transparently but Clustered reads do not, due to their use of negative seeking. We
+  merged image transparently but Clustered reads does not, due to its use of negative seeking. We
   conclude that splitting and merging large 3D images can be done
@@
 split and merge them efficiently, in particular to limit file
-seeks. In this paper we introduce and compare a family of sequential
+seeks. In this paper we introduce and compare sequential
 algorithms to split and merge images with reduced seeking.
@@
 representing 3D blocks or 3D slabs that fit in memory. A 3D block
-consists of a stack of one or more 2D tiles (incomplete slices and/or columns) spanning sequentially occuring slices, whereas a 3D slab is 
-a series of one of more contiguous 2D slices. A dataset such
-as Big~Brain would perhaps be split into 125 chunks of 600~MB. The
+consists of a stack of one or more 2D tiles (incomplete slices or
+incomplete columns) spanning contiguous slices, whereas a 3D slab is a
+series of one of more complete contiguous 2D slices. A dataset such as
+Big~Brain would perhaps be split into 125 chunks of 600~MB. The
 decision to split an image into slabs or blocks, and the size of the
-chunks, is up to the application. For instance, spatial filtering would
-commonly require blocks, whereas slices would be preferable for acquisition artifact removal. 
-Applications that process voxels individually, for instance
-histogram computation or k-means clustering, could work on either
-slices or blocks. Flexibility is thus required in the splitting
-scheme.
+chunks, is up to the application. For instance, spatial filtering
+would commonly require blocks, whereas slabs might be preferable for
+acquisition artifact removal.  Applications that process voxels
+individually, for instance histogram computation or k-means
+clustering, could work on either slabs or blocks. Flexibility is thus
+required in the splitting scheme.
 

### 2017-11-10 paper.tex

--- 2017-11-10 paper.tex
@@
 representing 3D blocks or 3D slabs that fit in memory. A 3D block
-consists of a stack of one or more 2D tiles (incomplete slices and/or columns) spanning sequentially occuring slices, whereas a 3D slab is 
+consists of a stack of one or more 2D tiles (incomplete slices or incomplete columns) spanning sequentially occuring slices, whereas a 3D slab is 
 a series of one of more contiguous 2D slices. A dataset such

### 2017-11-10 paper.tex
Final fixes for camera-ready version

--- 2017-11-10 paper.tex
@@
 order. All voxels in a \emph{slice} have the same $k$ and all voxels
-in a \emph{row} have the same $j$.
+in a \emph{column} have the same $j$.
 
@@
 \end{figure}
-Algorithms~\ref{algo:naive-slabs} and~\ref{algo:naive-blocks} show
-the naive merging methods for slabs and blocks. These algorithms
-actually have very different complexities even though blocks and
-slabs have identical sizes. Since slabs are stored contiguously in
-the reconstructed image, the number of seeks in
-Algorithm~\ref{algo:naive-slabs} is only $2n$ as $n$ seeks are
-required to read the slabs and $n$ seeks are required to write them:
+Algorithms~\ref{algo:naive-slabs} and~\ref{algo:naive-blocks} show the
+naive merging methods for slabs and blocks. These algorithms have very
+different complexities even though blocks and slabs have identical
+sizes. Since slabs are stored contiguously in the reconstructed image,
+the number of seeks in Algorithm~\ref{algo:naive-slabs} is only $2n$
+as $n$ seeks are required to read the slabs and $n$ seeks are required
+to write them:
 \begin{equation}
@@
 \end{algorithm}
-This algorithm writes in the reconstructed image using a single seek
+
+\noindent This algorithm writes in the reconstructed image using a single seek
 per memory load. Therefore:
@@
 reduced compared to Naive blocks since contiguous parts of the
-buffer will be written without seeking. A given
+buffer are written without seeking. A given
 block is accessed only once during the whole merging process.
@@
 complete contiguous slabs of the reconstructed image can be assembled
-in memory and written in a single seek. In the worst case, the
-memory load only partially covers columns in the reconstructed image:
-O($d^2$) seeks are then required during writing, one for every partial
-column in every partial slabs. In the intermediary case, columns are
-complete but some slabs can only be partially reconstructed: O($d$)
-seeks are then required.
+in memory and written in a single seek. In the worst case, the memory
+load only partially covers columns in the reconstructed image:
+O($d^2$) seeks are then required during writing, one for every column
+of every tile. In the intermediary case, columns are complete but some
+slices can only be partially reconstructed: O($d$) seeks are then
+required.
 
@@
   their columns ($d^2$ seeks). Blue blocks need seeking before each of
-  their slices ($d$ seeks). Green blocks need only a single
+  their tiles ($d$ seeks). Green blocks need only a single
   seek. Grey, dashed, transparent blocks represent the contiguous
@@
 \def\svgwidth{0.3\columnwidth}
-\input{figures/svg/incomplete-rows.pdf_tex}
+\input{figures/svg/incomplete-columns.pdf_tex}
 \quad \quad \quad
@@
   reads. Left: configuration with incomplete block columns (multiplies
-  the number of seeks by $d$ compared to complete block rows). Right:
+  the number of seeks by $d$ compared to complete block columns). Right:
   configuration with block columns that overlap multiple block slices
@@
 Our algorithm also avoids configurations where the memory load
-overlaps multiple block slabs in Case 2 or multiple block columns in
+overlaps multiple block slices in Case 2 or multiple block columns in
 Case 1, as such overlaps multiply the number of required seeks (see
@@
 (line 7) determines whether two blocks overlap multiple block slabs
-(case 2) or multiple block rows (Case 1). For case 3 it always returns
+(Case 2) or multiple block columns (Case 1). For Case 3 it always returns
 false.  Function \texttt{sizeof} (line 8) returns the actual memory
@@
 Because our algorithm avoids overlapping configurations, $x_1$ is
-proportional to the total number of block rows in the image
+proportional to the total number of block columns in the image
 $\left( \sqrt[3]{n}^2 \right)$ and $x_2$ is proportional to the total number of
@@
   memory load. Small dots depict block frontiers and long dashes
-  depict rows and slices within blocks.}
+  depict columns and slices within blocks.}
 \label{fig:multiple-reads-cases}
@@
         \STATE block\_data = read block
-        \STATE in block\_data, extract the rows in write\_range
-        \STATE insert rows in buffer
+        \STATE in block\_data, extract the columns in write\_range
+        \STATE insert columns in buffer
       \ENDIF
@@
 
-The total number of seeks performed by multiple reads in case $i$ is:
+The total number of seeks performed by multiple reads in Case $i$ is:
 \begin{eqnarray*}
@@
 Clustered reads is in Case 1 or 2, it may outperform Multiple reads
-only for large values of $n$. When Clustered reads is in case 3, it
+only for large values of $n$. When Clustered reads is in Case 3, it
 is equivalent to Multiple reads: assuming that $Rb$ is an exact
@@
 memory values. Naive blocks are 9.5 times slower than Naive slabs on
-HDD (6.7 times on SSD), which quantifies the effect of the targeted
-problem.  In the remainder we use Naive blocks
-and Naive slabs as references to evaluate our algorithms.
+HDD (6.7 times on SSD), which quantifies the effect of the problem
+targeted by our algorithms.  In the remainder we use Naive blocks and
+Naive slabs as references to evaluate our algorithms.
 
@@
 blocks even at 3~GB, while in Case 1. This may be explained by the
-fact that the seeks required to write incomplete block rows to the
+fact that the seeks required to write incomplete block columns to the
 reconstructed image are shorter than the ones for Naive blocks.
@@
 discrepancies between disk read and writes rates, or of reading data using Python's NumPy
-package, which is more efficient than using native Python - as is the
+package, which is more efficient than using native Python as is the
 case with our writes. The overhead time is small for both Clustered
@@
 % File formats
-Our results demonstrate that large images stored in simple 
-formats may be split and merged without performance loss compared to
-more complex formats that can preserve spatial locality on disk,
-for instance MINC 2.0 or the format based on space-filling curves
-mentioned in~\cite{burns2013open}. This is of major interest in the
-current open-science context since simpler formats favor data-sharing
-and interoperability.  Moreover, our algorithms could potentially be
-adapted to any split geometry, even though we demonstrated them on
-slabs and blocks only, while file formats inevitably assume a
-particular geometry. For instance, the format in~\cite{burns2013open}
-is not designed to naively split slabs. However, we aimed to extract
-all the blocks, whereas, ~\cite{burns2013open} aimed to extract a
-single block from a large image. MINC 2.0 might also help with
-on-the-fly compression for Clustered reads.
+Our results demonstrate that large images stored in simple formats may
+be split and merged without performance loss compared to more complex
+formats, for instance MINC 2.0 or the format based on space-filling
+curves mentioned in~\cite{burns2013open}. This is of major interest in
+the current open-science context since simpler formats favor
+data-sharing and interoperability.  Moreover, our algorithms could
+potentially be adapted to any chunk geometry, even though we
+demonstrated them on slabs and blocks only, while file formats
+inevitably assume a particular geometry. For instance, the format
+in~\cite{burns2013open} is not designed to naively split
+slabs. However, we aimed to extract all the blocks, whereas
+~\cite{burns2013open} aimed to extract a single block from a large
+image. MINC 2.0 might also help with on-the-fly compression for
+Clustered reads.
 

### 2017-11-10 figures/svg/export_svg_files.sh

@@ -1,6 +1,6 @@
 #!/usr/bin/env bash
 
-for name in buffer case1-a case2-a case3 incomplete-rows overlap mreads-case1 \
+for name in buffer case1-a case2-a case3 incomplete-columns overlap mreads-case1 \
         mreads-case2 mreads-case3 mreads-case4 mreads-case5 Notations
 do
   #echo "Exporting ${name}.svg"
